---
title: "01-Plot_PreHarmony_BanksyPCAs_by_spe_variables"
format: html
author: "Bernie Mulvey"
date: "06-19-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(ggrastr)
library(Cairo)
library(scater)
library(uwot)
library(RcppHNSW)
library(SpatialExperiment)
# library(SpatialFeatureExperiment)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
precision <- sqrt(.Machine$double.eps)
(x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
precision <- sqrt(.Machine$double.eps)
(x <= y) | (abs(x-y) <= precision)}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost='localhost')


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = '#000000'), axis.title.x = element_text(size = 9,color = '#000000'), axis.text.y = element_text(size = 8,color = '#000000'), axis.title.y = element_text(size =9,color = '#000000'), plot.title = element_markdown(size = 10,hjust=0.5,color = '#000000'), strip.text = element_text(size=11), legend.text = element_text(size=7,color = '#000000'), legend.title = element_text(size=8,hjust=0.5,color = '#000000'),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position='plot'))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

```{r}
## get metadata
lcx <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

lcxcd <- as.data.table(colData(lcx),keep.rownames=T)

## append slide/kit info
xensl <- fread("raw-data/donor_sample_slide_kit.txt")
lcxcd <- merge.data.table(lcxcd,xensl,by.x="brnum",by.y="Donor_sample")
stopifnot(nrow(lcxcd)==ncol(lcx))
rm(lcx,xensl)

### select 6.25k cells per sample (comes out to 40% of dataset) and use them for all of the UMAPs and PCA plots
set.seed(42)
pltcells <- unlist(lapply(unique(lcxcd$brnum),function(x){sample(lcxcd[brnum==x,rn],size = 6250,replace = FALSE)}))
```

## we only saved BanksyPCA matrices -- will need to calculate the corresponding UMAPs on the fly here for plotting purposes.
```{r}
pcaouts <- list.files("processed-data_newseg/11_Banksy_clustering/01-banksy_clustruns_outs/banksyPCAs",pattern = "RDS",full.names=T)

pcas <- lapply(X = pcaouts,readRDS)
names(pcas) <- gsub(pcaouts,pattern="^processed-data_newseg/11_Banksy_clustering/01-banksy_clustruns_outs/banksyPCAs/PCA_(M.*)\\.RDS$",replacement="\\1")

## downsample cells to 6.25k/sample
pcas <- lapply(pcas,function(x){
    x[pltcells,]
})
rm(pltcells)
gc(full=T)
```

make UMAPs for plotting purposes ONLY--output will not be reproducible!
```{r}
umaps <- lapply(pcas,function(x){
    ## fast_sgd does not yield reproducible UMAPs even with set.seed.
    ## however, uwot notes that this method is perfectly servicable for plotting
    y <- uwot::umap(x,nn_method = "hnsw",fast_sgd = TRUE,verbose=TRUE)
    y <- as.data.table(as.data.frame(y),keep.rownames=T)
    setnames(y,c(2:3),c("UMAP1","UMAP2"))
    return(y)
})
names(umaps) <- names(pcas)

## pare down PCAs to cell id, pc1, p2 for memory effish
pcas <- lapply(pcas,as.data.table,keep.rownames=T)

pcas <- lapply(pcas,function(x){
    y <- x[,.(rn,PC1,PC2)]
    return(y)
})
```

PCA, UMAP plots for each banksy parameter set by tech vars of interest.
```{r}
covs <- c("brnum", "transcript_counts", "cell_area", "nucleus_area", "Age","Sex","Rin","APOE","YRI","nuc_frac","subsets_any_neg_percent","Xenium_slide","Kit")
keepn <- c("rn",covs)
lcxcd <- lcxcd[,..keepn]
rm(keepn)
```

### PCA plots
```{r}
bpmapply(d=pcas,f=names(pcas),SIMPLIFY=FALSE,BPPARAM = MulticoreParam(8),FUN=function(d,f){
    tmp <- copy(d)
    
    mdcol <- c("rn",covs)
    tmp <- merge.data.table(tmp,lcxcd[,..mdcol],by="rn")
    
    plts <- lapply(covs,function(x){
        if(x %in% c("transcript_counts", "cell_area", "nucleus_area", "Age","Rin","YRI","nuc_frac","subsets_any_neg_percent")){
            xv <- names(tmp)[2]
            yv <- names(tmp)[3]
            p <- ggplot(tmp,aes(x=.data[[xv]],y=.data[[yv]],color=as.numeric(.data[[x]])))+
                geom_point(size=0.1,stroke = 0)+
                labs(title=paste0("Dimred by ",x),color="none")
        } else {
            xv <- names(tmp)[2]
            yv <- names(tmp)[3]
            p <- ggplot(tmp,aes(x=.data[[xv]],y=.data[[yv]],color=.data[[x]]))+
                geom_point(size=0.1,stroke = 0)+
                labs(title=paste0("Dimred by ",x))+
                guides(color="none")
        }
        
        
        q <- rasterize(p,layers="Points",dev="cairo_png",dpi=200)
        return(q)
    })
    Cairo(paste0("plots_newseg/11_Banksy_Clustering/01_NoHarmony_PCAs_UMAPS/",f,"_PCA.png"),width=18,height=15,units="in",type="png",dpi=200)
    do.call("grid.arrange",c(plts,ncol=4))
    dev.off()
})

```

UMAP plots:
```{r}
bpmapply(d=umaps,f=names(umaps),SIMPLIFY=FALSE,BPPARAM = MulticoreParam(8),FUN=function(d,f){
    tmp <- copy(d)
    
    mdcol <- c("rn",covs)
    tmp <- merge.data.table(tmp,lcxcd[,..mdcol],by="rn")
    
    plts <- lapply(covs,function(x){
        if(x %in% c("transcript_counts", "cell_area", "nucleus_area", "Age","Rin","YRI","nuc_frac","subsets_any_neg_percent")){
            xv <- names(tmp)[2]
            yv <- names(tmp)[3]
            p <- ggplot(tmp,aes(x=.data[[xv]],y=.data[[yv]],color=as.numeric(.data[[x]])))+
                geom_point(size=0.1,stroke = 0)+
                labs(title=paste0("UMAP by ",x),color="none")
        } else {
            xv <- names(tmp)[2]
            yv <- names(tmp)[3]
            p <- ggplot(tmp,aes(x=.data[[xv]],y=.data[[yv]],color=.data[[x]]))+
                geom_point(size=0.1,stroke = 0)+
                labs(title=paste0("UMAP by ",x))+
                guides(color="none")
        }
        
        
        q <- rasterize(p,layers="Points",dev="cairo_png",dpi=200)
        return(q)
    })
    Cairo(paste0("plots_newseg/11_Banksy_Clustering/01_NoHarmony_PCAs_UMAPS/",f,"_UMAP.png"),width=18,height=15,units="in",type="png",dpi=200)
    do.call("grid.arrange",c(plts,ncol=4))
    dev.off()
})



```

### three of these look bad / have kit or slide specific areas of UMAP space: m0_lam1_kgeom30; m1_lam0.8_kgeom30; m1_lam1_kgeom30. plot each clustering for the remaining 21 UMAPs.

```{r}
# don't need the pcas anymore
rm(pcas)

# load clusters
bclus <- fread("processed-data_newseg/11_Banksy_clustering/01c-AllBanksyClusteringAssignments.txt.gz")
keepn <- grep(names(bclus),pattern="M0_lam0.8_.*kgeom30|M1_lam0.8_.*kgeom30|M1_lam1_.*kgeom30",invert = T,value=T)
bclus <- bclus[,..keepn]

# also get rid of the corresp. umaps
umaps <- umaps[!(names(umaps) %in% c("M0_lam1_kgeom30","M1_lam1_kgeom30","M1_lam0.8_kgeom30"))]

# split into list of tables by banksy parameter set, in same order as in umaps. we need to strip out the "clust_..._k25_" in each name to match
setnames(bclus,gsub(names(bclus),pattern="^clust_(.*)$",replacement="\\1"))
setnames(bclus,gsub(names(bclus),pattern="_k25",replacement=""))
setnames(bclus,gsub(names(bclus),pattern="__",replacement=""))
# ugh
setnames(bclus,gsub(names(bclus),pattern="^(.*[[:alnum:]]{1})(kgeom.*$)",replacement=paste0("\\1","_","\\2")))

clusls <- lapply(names(umaps),function(x){
    keepcl <- c("rn",grep(names(bclus),pattern=x,value = T))
    out <- copy(bclus)[,..keepcl]
    return(out)
})
names(clusls) <- names(umaps)
``` 

## cluster sets per dimred
```{r}
bpmapply(d=umaps,f=names(umaps),clus=clusls,SIMPLIFY=FALSE,BPPARAM = MulticoreParam(8),FUN=function(d,f,clus){
    tmp <- copy(d)
    tmp <- merge.data.table(tmp,clus,by="rn")
    
    plts <- lapply(names(clus)[2:4],function(x){
        p <- ggplot(tmp,aes(x=UMAP1,y=UMAP2,color=.data[[x]]))+
            geom_point(size=0.1,alpha=0.55,stroke = 0)+
            labs(title=paste0("UMAP by ",x))+
            guides(color="none")
        
        
        q <- rasterize(p,layers="Points",dev="cairo_png",dpi=200)
        return(q)
    })
    Cairo(paste0("plots_newseg/11_Banksy_Clustering/01_NoHarmony_PCAs_UMAPS/",f,"_UMAPclusters.png"),width=23,height=7,units="in",type="png",dpi=200)
    do.call("grid.arrange",c(plts,ncol=3))
    dev.off()
})
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0         parallelly_1.47.0          
 [3] SpatialExperiment_1.20.0    RcppHNSW_0.6.0             
 [5] uwot_0.2.4                  Matrix_1.7-4               
 [7] scater_1.38.1               scuttle_1.20.0             
 [9] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
[11] Biobase_2.70.0              GenomicRanges_1.62.1       
[13] Seqinfo_1.0.0               IRanges_2.44.0             
[15] S4Vectors_0.48.1            BiocGenerics_0.56.0        
[17] generics_0.1.4              MatrixGenerics_1.22.0      
[19] matrixStats_1.5.0           Cairo_1.7-0                
[21] ggrastr_1.0.2               gridExtra_2.3              
[23] ggtext_0.1.2                ggplot2_4.0.2              
[25] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] tidyselect_1.2.1    viridisLite_0.4.3   dplyr_1.2.0        
 [4] vipor_0.4.7         farver_2.1.2        viridis_0.6.5      
 [7] S7_0.2.1            fastmap_1.2.0       digest_0.6.39      
[10] rsvd_1.0.5          lifecycle_1.0.5     magrittr_2.0.4     
[13] compiler_4.5.2      rlang_1.1.7         tools_4.5.2        
[16] yaml_2.3.12         knitr_1.51          S4Arrays_1.10.1    
[19] here_1.0.2          DelayedArray_0.36.1 xml2_1.5.2         
[22] RColorBrewer_1.1-3  abind_1.4-8         rsconnect_1.7.0    
[25] withr_3.0.2         grid_4.5.2          beachmat_2.26.0    
[28] scales_1.4.0        dichromat_2.0-0.1   cli_3.6.5          
[31] rmarkdown_2.30      otel_0.2.0          rstudioapi_0.18.0  
[34] RSpectra_0.16-2     rjson_0.2.23        sessioninfo_1.2.3  
[37] ggbeeswarm_0.7.3    parallel_4.5.2      XVector_0.50.0     
[40] vctrs_0.7.1         BiocSingular_1.26.0 BiocNeighbors_2.4.0
[43] ggrepel_0.9.6       irlba_2.3.7         beeswarm_0.4.0     
[46] magick_2.9.0        glue_1.8.0          codetools_0.2-20   
[49] gtable_0.3.6        ScaledMatrix_1.18.0 tibble_3.3.1       
[52] pillar_1.11.1       htmltools_0.5.9     R6_2.6.1           
[55] rprojroot_2.1.1     evaluate_1.0.5      lattice_0.22-7     
[58] gridtext_0.1.5      Rcpp_1.1.1          SparseArray_1.10.8 
[61] xfun_0.56           pkgconfig_2.0.3    
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-19
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 beachmat               2.26.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocNeighbors          2.4.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Cairo                * 1.7-0    2025-10-29 [1] CRAN (R 4.5.0)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.1   2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 ggbeeswarm             0.7.3    2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggrastr              * 1.0.2    2023-06-01 [1] CRAN (R 4.5.0)
 ggrepel                0.9.6    2024-09-07 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7    2026-01-30 [1] CRAN (R 4.5.2)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix               * 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.47.0   2026-04-17 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 RcppHNSW             * 0.6.0    2024-02-04 [1] CRAN (R 4.5.0)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 RSpectra               0.16-2   2024-07-18 [1] CRAN (R 4.5.0)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 rsvd                   1.0.5    2021-04-16 [1] CRAN (R 4.5.0)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1   2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 ScaledMatrix           1.18.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 scater               * 1.38.1   2026-03-20 [1] Bioconductor 3.22 (R 4.5.2)
 scuttle              * 1.20.0   2025-10-30 [1] Bioconductor 3.22 (R 4.5.2)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 uwot                 * 0.2.4    2025-11-10 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.5.0)
 viridis                0.6.5    2024-01-29 [1] CRAN (R 4.5.0)
 viridisLite            0.4.3    2026-02-04 [1] CRAN (R 4.5.2)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────