---
title: "01-Get_IntraLCdomain_Transcripts_allLCdomain_definitions"
author: "Bernie Mulvey"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(arrow)
library(sf)
library(sfheaders)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##


## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

### get coordinates of LC domains for each of the 6 definitions (as we did in code 14 for calculating LC domain area in xenium samples). while these are cell-level annotations, we can use this information to extract the coordinate info about those cells and the areas the domain covers (we also saved a table of cell coordinates in code 14 which we'll use for this.)
```{r}
wdlcs <- fread("processed-data_newseg/04_snSeqLabelTransfer/06-wdLCsmoothings_LUT.txt")
wdlcs <- wdlcs[,.(rn,domain12412,domain8492)] # indicates whether a given cell is in LC domain
setnames(wdlcs,c(2,3),c("wd_domain12412","wd_domain8492"))

srtlcs <- fread("processed-data_newseg/07_visLabelTransfer/05-srtLCsmoothDoms_and_NMFpreds_lut.txt")
srtlcs <- srtlcs[,.(rn,domain8426,domain12199,domain16119,domain20079)]
setnames(srtlcs,c(2:5),c("srt_domain8426","srt_domain12199","srt_domain16119","srt_domain20079"))

domassigns <- merge.data.table(wdlcs,srtlcs,by="rn")
## append brnum using the info in the rowname (cell_id)
domassigns[,brnum:=gsub(rn,pattern="^(Br.*)_.*$",replacement="\\1")]
rm(srtlcs,wdlcs)
gc(full=T)

### now read in aforementioned cell polygonal boundary info
allpoly <- fread("processed-data_newseg/14_CellTypeProximity/retainedCellBoundaries.txt.gz")

## for efficiency's sake, only keep cells and the corresponding polygon info for those cells  assigned to at least one of the LC domains
domassigns[,everLC:=apply(.SD,MARGIN=1,function(x){sum(x=="LC")}),.SDcols=c(2:7)]
domassigns <- domassigns[everLC>0]
allpoly <- allpoly[cell_id %in% domassigns$rn]

gc(full=T)
```

```{r}
xscripfiles <- list.files("processed-data/xenium_imageProcessing_new/xeniumRanger",recursive=T,pattern="transcripts.parquet",full.names=T)

## crashed partway. pick up at sample number 7 
xscripfiles <- xscripfiles[c(7:length(xscripfiles))]

arrow::set_cpu_count(4)
setDTthreads(2)
## load one sample, calculate totals for all 6 domain defs, save, move onto next sample

i <- 1
for (i in c(1:length(xscripfiles))){
  p <- xscripfiles[i]
  message(paste0("starting ",p))
  xscriploc <- read_parquet(p,as_data_frame=T)
  xscriploc <- as.data.table(xscriploc)
  
  # retain qv≥ 20, as is done automatically for transcript filtering when xeniumranger filters
  # transcripts in cell-level data assembly:
  xscriploc <- xscriploc[greaterOrEqual(qv,20),.(cell_id,qv,x_location,y_location,feature_name)]
  
  ## append brnum (sample)
  xscriploc[,brnum:=gsub(p,pattern="^.*/xeniumRanger/xeniumranger_NM_DAPI_(Br.*)/transcripts.parquet$",replacement="\\1")]
  xscriploc[,brnum:=gsub(brnum,pattern="-L",replacement="L")]
  
  ## skip the sample that was not analyzed:
  if(!(unique(xscriploc$brnum) %in% unique(domassigns$brnum))){
    rm(p,xscriploc)
    gc(full=T)
    next(i)
  }
  
  # convert to sf format for intersecting transcript coords to LC domain areas
  xscriploc <- st_as_sf(xscriploc,coords=c("x_location","y_location"),remove=FALSE)
  
  ## get counts for this sample falling within each domain definition
  domdefs <- c("wd_domain12412","wd_domain8492","srt_domain8426","srt_domain12199","srt_domain16119","srt_domain20079")
  
    ## narrow fix: if sample Br5854 or 1556, don't use srt_domain20079 
  ## because there are no such cells for these samples
  if(unique(xscriploc$brnum) %in% c("Br5854","Br1556")){
     domdefs <- domdefs[1:5]
  }
  
  domxscripts <- list()
  for (domdef in domdefs){
    tmp <- copy(domassigns)[get(domdef)=="LC"]
    tmp <- merge.data.table(tmp,allpoly,by.x="rn",by.y="cell_id")
    ## get the coordinates of intra-LC cells for the current sample and domain definition
    tmp <- tmp[brnum==unique(xscriploc$brnum)]
    
    # subset your data to the cells that make up your zone (already done)
    # Define the gap distance-this should be large enough to bridge the space between dots in the same domain zone, but small enough not to merge distinct, separate ones.
    # test values using br5415 -- should get 11 separate LC domain areas. testing this on several addnl brains with multiple discontig LC domain pieces returned the correct number of domains in each case.
    gap_distance <- 30 ## 
    # Then, apply st_union() to merge touching geometries, and st_cast() to break apart the unconnected bodies.
    lccellpolygons_sf <- sf_polygon(
      obj = tmp,
      x = "vertex_x",
      y = "vertex_y",
      polygon_id = "cell_id",
      keep = TRUE # Retains the label and other attributes
    )
    
    ## clean up the preceding step's objects for memory effish
    buffered_sf <- st_buffer(lccellpolygons_sf, dist = gap_distance)
    rm(lccellpolygons_sf)
    gc(full=T)
    
    merged_sf <- st_union(buffered_sf)
    rm(buffered_sf)
    
    lczonepolys <- st_cast(merged_sf, "POLYGON")
    rm(merged_sf)
    
    # retract the the buffer area we applied in Step 1 to make cells touch
    # This pulls the boundary back to meet the original cell polygons
    tight_zones_sfc <- st_buffer(lczonepolys, dist = -gap_distance)
    rm(lczonepolys)
    
    zones_sf <- st_sf(geometry = tight_zones_sfc)
    rm(tight_zones_sfc)
    ## now filter the corresponding sample's transcripts to those transcripts found within our LC domain boundaries:
    
    matched_points_sf <- st_join(
      xscriploc, 
      zones_sf, 
      join = st_intersects, 
      left = FALSE
    )
    rm(zones_sf)
    
    # tabulate total transcripts for each gene in the intersected objects
    counttab <- table(matched_points_sf$feature_name)
    counttab <- as.data.table(counttab)
    counttab[,brnum:=unique(xscriploc$brnum)]
    ## append to list for this brain
    domxscripts[[(length(domxscripts)+1)]] <- counttab
    
    ## clean up
    rm(counttab,matched_points_sf)
    gc(full=T)
  }
  names(domxscripts) <- domdefs
  message(paste0("finished ",p))
  
  saveRDS(domxscripts,paste0("processed-data_newseg/15_LCdomain_AllTranscripts_DE/01tmp/",unique(xscriploc$brnum),".RDS"))
  rm(domxscripts,xscriploc)
  gc(full=T)
}
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] colorout_1.3-3    sfheaders_0.4.5   sf_1.0-24         arrow_24.0.0      data.table_1.18.4

loaded via a namespace (and not attached):
 [1] bit_4.6.0          compiler_4.5.2     tidyselect_1.2.1   Rcpp_1.1.1        
 [5] assertthat_0.2.1   yaml_2.3.12        fastmap_1.2.0      here_1.0.2        
 [9] R6_2.6.1           classInt_0.4-11    knitr_1.51         units_1.0-0       
[13] rprojroot_2.1.1    DBI_1.2.3          rlang_1.1.7        xfun_0.56         
[17] bit64_4.6.0-1      otel_0.2.0         cli_3.6.5          magrittr_2.0.4    
[21] class_7.3-23       digest_0.6.39      grid_4.5.2         rstudioapi_0.18.0 
[25] lifecycle_1.0.5    vctrs_0.7.1        KernSmooth_2.23-26 proxy_0.4-29      
[29] evaluate_1.0.5     glue_1.8.0         sessioninfo_1.2.3  rsconnect_1.7.0   
[33] e1071_1.7-17       rmarkdown_2.30     purrr_1.2.1        tools_4.5.2       
[37] htmltools_0.5.9   
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-24
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────────────────
 package     * version date (UTC) lib source
 arrow       * 24.0.0  2026-04-29 [1] CRAN (R 4.5.2)
 assertthat    0.2.1   2019-03-21 [1] CRAN (R 4.5.0)
 bit           4.6.0   2025-03-06 [1] CRAN (R 4.5.0)
 bit64         4.6.0-1 2025-01-16 [1] CRAN (R 4.5.0)
 class         7.3-23  2025-01-01 [1] CRAN (R 4.5.2)
 classInt      0.4-11  2025-01-08 [1] CRAN (R 4.5.0)
 cli           3.6.5   2025-04-23 [1] CRAN (R 4.5.0)
 colorout    * 1.3-3   2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table  * 1.18.4  2026-05-06 [1] CRAN (R 4.5.2)
 DBI           1.2.3   2024-06-02 [1] CRAN (R 4.5.0)
 digest        0.6.39  2025-11-19 [1] CRAN (R 4.5.2)
 e1071         1.7-17  2025-12-18 [1] CRAN (R 4.5.2)
 evaluate      1.0.5   2025-08-27 [1] CRAN (R 4.5.0)
 fastmap       1.2.0   2024-05-15 [1] CRAN (R 4.5.0)
 glue          1.8.0   2024-09-30 [1] CRAN (R 4.5.0)
 here          1.0.2   2025-09-15 [1] CRAN (R 4.5.0)
 htmltools     0.5.9   2025-12-04 [1] CRAN (R 4.5.2)
 KernSmooth    2.23-26 2025-01-01 [1] CRAN (R 4.5.2)
 knitr         1.51    2025-12-20 [1] CRAN (R 4.5.2)
 lifecycle     1.0.5   2026-01-08 [1] CRAN (R 4.5.2)
 magrittr      2.0.4   2025-09-12 [1] CRAN (R 4.5.0)
 otel          0.2.0   2025-08-29 [1] CRAN (R 4.5.0)
 proxy         0.4-29  2025-12-29 [1] CRAN (R 4.5.2)
 purrr         1.2.1   2026-01-09 [1] CRAN (R 4.5.2)
 R6            2.6.1   2025-02-15 [1] CRAN (R 4.5.0)
 Rcpp          1.1.1   2026-01-10 [1] CRAN (R 4.5.2)
 rlang         1.1.7   2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown     2.30    2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot     2.1.1   2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect     1.7.0   2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi    0.18.0  2026-01-16 [1] CRAN (R 4.5.2)
 sessioninfo   1.2.3   2025-02-05 [1] CRAN (R 4.5.0)
 sf          * 1.0-24  2026-01-13 [1] CRAN (R 4.5.2)
 sfheaders   * 0.4.5   2025-11-25 [1] CRAN (R 4.5.2)
 tidyselect    1.2.1   2024-03-11 [1] CRAN (R 4.5.0)
 units         1.0-0   2025-10-09 [1] CRAN (R 4.5.0)
 vctrs         0.7.1   2026-01-23 [1] CRAN (R 4.5.2)
 xfun          0.56    2026-01-18 [1] CRAN (R 4.5.2)
 yaml          2.3.12  2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────────────────