---
title: "03-OneInteractionPerRun_AllLCdomXscripts_DE.Rmd"
author: "Bernie Mulvey"
date: "06-24-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(edgeR)
library(limma)
## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
dgels <- readRDS("processed-data_newseg/15_LCdomain_AllTranscripts_DE/02-Allxscript_pbulkDGEls_per_domaindef.RDS")

## these already have genes, donor info, but need simple APOE column
dgels <- lapply(dgels,function(x){
  x$samples$APOEsimple <- ifelse(x$samples$APOE %in% c("E2hom","E2het"),"E2","E4")
  return(x)
})
```


##### In the next SEVERAL blocks, we will define a function that runs each DE analysis so that we can parallelize across dgelists (categorical or continuous YRI) and analysis parameters (defined functions). #####

##### MODEL 1: SEX DE (Sex_cluster "interaction" edgeR style; APOE by carrier type #######
####################################################################
```{r}
sexfun <- function(dgel,n){
    dgelsex <- dgel
    dgelsex$samples$clus_sex <- paste0(dgelsex$samples$clus, "_", dgelsex$samples$Sex)
    
    # design mat
    des <- model.matrix(~0 + Sex + APOEsimple + Age + YRI, data = dgelsex$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "Sex", replacement = "")
    
    
    # filter using design
    dgelsex <- edgeR::normLibSizes(dgelsex)
    keep <- filterByExpr.DGEList(dgelsex, design = des)
    dgelsex <- dgelsex[keep, , keep.lib.sizes = FALSE]
    dgelsex <- edgeR::normLibSizes(dgelsex)
    
    # define contrasts
    trastmat <- makeContrasts("LCdom_sex"=M-F,levels=des)

    ## fit and fit contrasts
    v.swt <- voomLmFit(dgelsex, design = des, block = as.factor(dgelsex$samples$donor), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    ## get top tables, save an RDS with the DE results and other parts
    sexlabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
    })
    names(sexlabel_detabs) <- colnames(trastmat)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelsex, trasts = trastmat, detabs = sexlabel_detabs)
    
    saveRDS(keeplist, paste0("processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp/03bTMP_keeplist_sexlabel_noMixClusts_", n, ".RDS"))
    gc(full = TRUE)
}
```

#####################################################################
##### MODEL 2: E2/E4 carrier status "interaction" edgeR style #######
#####################################################################
```{r}
e2e4fun <- function(dgel,n){
    dgelE2E4 <- dgel
    
    des <- model.matrix(~0 + APOEsimple + Sex + Age + YRI, data = dgelE2E4$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "APOEsimple", replacement = "")

    # Filter using the design specified
    dgelE2E4 <- edgeR::normLibSizes(dgelE2E4)
    keep <- filterByExpr.DGEList(dgelE2E4, design = des)
    dgelE2E4 <- dgelE2E4[keep, , keep.lib.sizes = FALSE]
    dgelE2E4 <- edgeR::normLibSizes(dgelE2E4)
    
    # define contrasts to pass to makeContrasts (we have to do it this way due to environment weirdness with parallelization)
    trastmat <- makeContrasts("LCdom_E4vE2"=E4-E2,levels=des)

    ## fit and fit contrasts
    v.swt <- voomLmFit(dgelE2E4, design = des, block = as.factor(dgelE2E4$samples$donor), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    ## get top tables, save an RDS with the DE results and other parts
    E2E4label_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
    })
    names(E2E4label_detabs) <- colnames(trastmat)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelE2E4, trasts = trastmat, detabs = E2E4label_detabs)
                     
    saveRDS(keeplist, paste0("processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp/03bTMP_keeplist_E2E4label_", n, ".RDS"))
    gc(full = TRUE)}
```

##########################################################################
#### MODEL 3: APOE alleles (E2hom/het, 4hom/het) "intxns" edgeR style; APOE by diplotype ####
##########################################################################
```{r}
apogenofun <- function(dgel,n){
    dgelApoAlleles <- dgel
    
    des <- model.matrix(~0 + APOE + Sex + Age + YRI, data = dgelApoAlleles$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "APOE", replacement = "")
    
    # filter using the design specified
    dgelApoAlleles <- edgeR::normLibSizes(dgelApoAlleles)
    keep <- filterByExpr.DGEList(dgelApoAlleles, design = des)
    dgelApoAlleles <- dgelApoAlleles[keep, , keep.lib.sizes = FALSE]
    dgelApoAlleles <- edgeR::normLibSizes(dgelApoAlleles)

    # Create a named vector of contrast formulas for the current label
    formulas <- c("E2hom - (0.5*E4het + 0.5*E4hom)",
                  "E2het - (0.5*E4het + 0.5*E4hom)",
                  "E4hom - (0.5*E2het + 0.5*E2hom)",
                  "E4het - (0.5*E2het + 0.5*E2hom)",
                  "E4hom - E4het",
                  "E4hom - ((E4het + E2het + E2hom)/3)")
    names(formulas) <- c("LCdom_E2hom_vs_anyE4", "LCdom_E2het_vs_anyE4", "LCdom_E4hom_vs_anyE2", "LCdom_E4het_vs_anyE2", "LCdom_E4hom_vs_E4het", "LCdom_E4hom_vs_others")
    contrasts_list <- as.list(formulas)
    trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))

    # fit and fit contrasts
    v.swt <- voomLmFit(dgelApoAlleles, design = des, block = as.factor(dgelApoAlleles$samples$donor), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    # get toptabs and other components, save
    ApoAllelesLabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
    })
    names(ApoAllelesLabel_detabs) <- colnames(trastmat)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelApoAlleles, trasts = trastmat, detabs = ApoAllelesLabel_detabs)
                     
    saveRDS(keeplist, paste0("processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp/03bTMP_keeplist_ApoAllelesLabel_", n, ".RDS"))
    gc(full = TRUE)
}
```

#############################################################################
#### MODEL 4: label-by-YRI (continuous, categorical) styles, with APOEsimple control ####
#############################################################################
```{r}
ancesfunCon <- function(dgel,n){
    dgelances <- dgel
    
    # Continuous YRI setup
    des <- model.matrix(~YRI + Sex + APOEsimple + Age, data = dgelances$samples)
    trastmat <- "YRI" # These are coefficients, not contrasts
    stopifnot(all(trastmat %in% colnames(des)))
    
    # Filtering using appropriate design from above
    dgelances <- edgeR::normLibSizes(dgelances)
    keep <- filterByExpr.DGEList(dgelances, design = des)
    dgelances <- dgelances[keep, , keep.lib.sizes = FALSE]
    dgelances <- edgeR::normLibSizes(dgelances)

    # Fit
    v.swt <- voomLmFit(dgelances, design = des, block = as.factor(dgelances$samples$donor), adaptive.span = TRUE, sample.weights = TRUE)
    
    # Fit contrasts or coefficients depending on YRI type
    v.swt.fit <- contrasts.fit(v.swt, coefficients = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)
    
    # Get results and other components, save
    yrilabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
    })
    names(yrilabel_detabs) <- colnames(v.swt.fit)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelances, trasts = trastmat, detabs = yrilabel_detabs)

    saveRDS(keeplist, paste0("processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp/03bTMP_keeplist_AncesContinuous_", n, ".RDS"))
    gc(full = TRUE)
}


###
###


ancesfunCat <- function(dgel,n){
    # make design matrices based on YRI as categorical
    # Categorical YRI setup
    message("Subsetting to 75%+ ancestry groups", n)
    newsamp <- as.data.table(dgel$samples,keep.rownames=T)
    newct <- dgel$counts
    
    newsamp <- newsamp[YRI>0.75|YRI<0.25]
    newsamp[,YRIcat:=ifelse(YRI>0.75,"YRI","CEU")]
    newct <- newct[,colnames(newct) %in% newsamp$rn]
    newsamp <- as.data.frame(newsamp,row.names=newsamp$rn)[colnames(newct),]
    newg <- as.data.frame(dgel$genes,row.names=rownames(dgel$genes))[rownames(newct),]
    newdgel <- DGEList(counts=newct,samples=newsamp,genes=newg)
  
    des <- model.matrix(~0 + YRIcat + APOEsimple + Sex + Age, data = newdgel$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "^YRIcat(.*)$", replacement = "\\1")
    trastmat <- makeContrasts("LCdom_YRIvCEU"=YRI-CEU,levels=des)

    
    # Filtering using appropriate design from above
    newdgel <- edgeR::normLibSizes(newdgel)
    keep <- filterByExpr.DGEList(newdgel, design = des)
    newdgel <- newdgel[keep, , keep.lib.sizes = FALSE]
    newdgel <- edgeR::normLibSizes(newdgel)

    # Fit
    v.swt <- voomLmFit(newdgel, design = des, block = as.factor(newdgel$samples$donor), adaptive.span = TRUE, sample.weights = TRUE)
    
    # Fit contrasts or coefficients depending on YRI type
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)
    
    # Get results and other components, save
    yrilabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
    })
    names(yrilabel_detabs) <- colnames(v.swt.fit)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = newdgel, trasts = trastmat, detabs = yrilabel_detabs)

    saveRDS(keeplist, paste0("processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp/03bTMP_keeplist_AncesCategorical_", n, ".RDS"))
    gc(full = TRUE)
}
```

#############################################################################
####### MODEL 5: label*YRI*E4orE2 (cat-cat-cat) "intxns" edgeR style ########
#############################################################################
```{r}
ancesapofun <- function(dgel,n){
    message("Subsetting to 75%+ ancestry groups", n)
    newsamp <- as.data.table(dgel$samples,keep.rownames=T)
    newct <- dgel$counts
    
    newsamp <- newsamp[YRI>0.75|YRI<0.25]
    newsamp[,YRIcat:=ifelse(YRI>0.75,"YRI","CEU")]
    newct <- newct[,colnames(newct) %in% newsamp$rn]
    newsamp <- as.data.frame(newsamp,row.names=newsamp$rn)[colnames(newct),]
    newg <- as.data.frame(dgel$genes,row.names=rownames(dgel$genes))[rownames(newct),]
    newdgel <- DGEList(counts=newct,samples=newsamp,genes=newg)

    # with the now categorical data, set it up and go
    newdgel$samples$APOE_PredomAnces <- paste0(newdgel$samples$APOEsimple, "_", newdgel$samples$YRIcat)
    
    # design matrix
    des <- model.matrix(~0 + APOE_PredomAnces + Sex + Age, data = newdgel$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "^APOE_PredomAnces(.*)$", replacement = "\\1")
    
    # filter using the design
    newdgel <- edgeR::normLibSizes(newdgel)
    keep <- filterByExpr.DGEList(newdgel, design = des)
    newdgel <- newdgel[keep, , keep.lib.sizes = FALSE]
    newdgel <- edgeR::normLibSizes(newdgel)

    # define contrasts
    formulas <- c(
      "E4_YRI - E4_CEU",
      "E2_YRI - E2_CEU",
      "E4_CEU - E2_CEU",
      "E4_YRI - E2_YRI")
    names(formulas) <- c("LCdom_E4YRI_vs_E4CEU", "LCdom_E2YRI_vs_E2CEU", "LCdom_E4CEU_vs_E2CEU", "LCdom_E4YRI_vs_E2YRI")
    contrasts_list <- as.list(formulas)
    trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))

    # fit, fit contrasts
    v.swt <- voomLmFit(newdgel, design = des, block = as.factor(newdgel$samples$donor), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    # get toptabs, other results, and save
    yriapolabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        o <- as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
        #setnames(o,"genes","gene_name") ## force uniformimty with other results
        return(o)
    })
    names(yriapolabel_detabs) <- colnames(v.swt.fit)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = newdgel, trasts = trastmat, detabs = yriapolabel_detabs)
                     
    saveRDS(keeplist, paste0("processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp/03bTMP_keeplist_YRIAPO_", n, ".RDS"))
    gc(full = TRUE)
}
```


## now run all of these * serially -- parallelization throwing an error now even though ive used an internal mclapply here before...
```{r}
funs <- list(sexfun,e2e4fun,apogenofun,ancesfunCon,ancesfunCat,ancesapofun)

mapply(dgel=dgels,n=names(dgels),SIMPLIFY = FALSE,FUN=function(dgel,n){
    lapply(X=funs,FUN=function(f){
        f(dgel,n)
    })
})
```

## clean up, load the lists and concatenate into one giant result obj per domain definition set
```{r}
rm(list=ls())
gc(full=T)
```

## load the tmp RDSes
```{r}
srt8426 <- list.files("processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp",pattern="srt_domain8426",full.names = TRUE)
srt12412 <- list.files("processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp",pattern="srt_domain12199",full.names = TRUE)
srt16119 <- list.files("processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp",pattern="srt_domain16119",full.names = TRUE)
srt20079 <- list.files("processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp",pattern="srt_domain20079",full.names = TRUE)
wd_domain8492 <- list.files("processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp",pattern="wd_domain8492",full.names = TRUE)
wd_domain12412 <- list.files("processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp",pattern="wd_domain12412",full.names = TRUE)

## combine into one metalist
srt8o <- lapply(srt8426,readRDS)
names(srt8o) <- gsub(srt8426,pattern="processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp/03bTMP_keeplist_(.*)_srt.*$",replacement="\\1")

srt12o <- lapply(srt12412,readRDS)
names(srt12o) <- gsub(srt12412,pattern="processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp/03bTMP_keeplist_(.*)_srt.*$",replacement="\\1")

srt16o <- lapply(srt16119,readRDS)
names(srt16o) <- gsub(srt16119,pattern="processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp/03bTMP_keeplist_(.*)_srt.*$",replacement="\\1")

srt20o <- lapply(srt20079,readRDS)
names(srt20o) <- gsub(srt20079,pattern="processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp/03bTMP_keeplist_(.*)_srt.*$",replacement="\\1")

wd8o <- lapply(wd_domain8492,readRDS)
names(wd8o) <- gsub(wd_domain8492,pattern="processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp/03bTMP_keeplist_(.*)_wd.*$",replacement="\\1")

wd12o <- lapply(wd_domain12412,readRDS)
names(wd12o) <- gsub(wd_domain12412,pattern="processed-data_newseg/15_LCdomain_AllTranscripts_DE/03tmp/03bTMP_keeplist_(.*)_wd.*$",replacement="\\1")

outlist <- list(srt8426=srt8o, srt12412=srt12o, srt16119=srt16o,srt20079=srt20o,wd8492=wd8o,wd12412=wd12o)

saveRDS(outlist,"processed-data_newseg/15_LCdomain_AllTranscripts_DE/03-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_LCdomaindefs_allXenTranscripts.RDS")
### 
### find nominal/near nominal DE events that have a stable effect direction across domain definitions (same sign and P<0.1 in 4 or more of the 6 definitions):
stablenomde <- lapply(outlist,function(a){
  b <- lapply(a,function(h){
    j <- h$detabs
    d <- mapply(x=j,n=names(j),SIMPLIFY=FALSE,FUN=function(x,n){
      z <- x[P.Value<0.05]
      z[,contrast:=n]
      return(z)
    })
    e <- rbindlist(d,idcol="something")
    return(e)
  })
  k <- rbindlist(b,fill=T)
  return(k)
})


stablenomde2 <- stablenomde[something=="LCdom_E4vE2"][P.Value<0.1]
stablenomde2[,sign:=ifelse(logFC>0,"e4up","e4down")]
stablenomde2[,.N,by=c("rn","sign")][N>3]

## only PLPP3 is a stable hit across domain definitions even when we take ALL transcripts in the domain. interestingly though (given NM-gene model results), both GPX3 and GPX4 are persistently E4-upregulated.
#           rn   sign     N
#       <char> <char> <int>
#  1:     RYR3 e4down     6
#  2:   P2RY12 e4down     6
#  3:    PLPP3 e4down     6
#  4:    GRID2 e4down     6
#  5:     GPX4   e4up     6
#  6:   CALCRL e4down     6
#  7:   STK32B e4down     6
#  8:   P2RY13 e4down     6
#  9:     ENC1   e4up     6
# 10:      ALK   e4up     6
# 11:     PENK e4down     6
# 12:    NPY1R e4down     6
# 13:  RNASET2 e4down     5
# 14:     GPX3   e4up     4
# 15:   ELOVL2 e4down     6
# 16:     PAX6 e4down     4
# 17:    ITGAM e4down     4
# 18:  CABLES1 e4down     5
# 19:    THSD4   e4up     6
# 20: B4GALNT1   e4up     4
# 21:   CEMIP2   e4up     4
# 22:   PTPRZ1 e4down     4
# 23:   CARTPT   e4up     5
# 24:   CTNNA3   e4up     4
# 25:    LRRK1 e4down     4
# 26:    RGS10 e4down     4
# 27:    NXPH2 e4down     5
# 28:    NDST4 e4down     5
# 29:   CLDN11   e4up     4
```
## can now delete 03bTMP RDSes.

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
 [4] Biobase_2.70.0              GenomicRanges_1.62.1        Seqinfo_1.0.0              
 [7] IRanges_2.44.0              S4Vectors_0.48.1            BiocGenerics_0.56.0        
[10] generics_0.1.4              MatrixGenerics_1.22.0       matrixStats_1.5.0          
[13] BiocParallel_1.44.0         parallelly_1.47.0           colorout_1.3-3             
[16] edgeR_4.8.2                 limma_3.66.0                data.table_1.18.4          

loaded via a namespace (and not attached):
 [1] Matrix_1.7-4        rjson_0.2.23        compiler_4.5.2      Rcpp_1.1.1         
 [5] magick_2.9.0        yaml_2.3.12         fastmap_1.2.0       statmod_1.5.1      
 [9] lattice_0.22-7      here_1.0.2          XVector_0.50.0      S4Arrays_1.10.1    
[13] knitr_1.51          DelayedArray_0.36.1 rprojroot_2.1.1     rlang_1.1.7        
[17] xfun_0.56           otel_0.2.0          SparseArray_1.10.8  cli_3.6.5          
[21] magrittr_2.0.4      digest_0.6.39       grid_4.5.2          locfit_1.5-9.12    
[25] rstudioapi_0.18.0   lifecycle_1.0.5     evaluate_1.0.5      codetools_0.2-20   
[29] abind_1.4-8         rsconnect_1.7.0     rmarkdown_2.30      tools_4.5.2        
[33] htmltools_0.5.9    
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-24
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.4   2026-05-06 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.1   2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 edgeR                * 4.8.2    2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 limma                * 3.66.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 locfit                 1.5-9.12 2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.47.0   2026-04-17 [1] CRAN (R 4.5.2)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1   2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 statmod                1.5.1    2025-10-09 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────────────────