---
title: "02-Register_NonHarmony_Banksys_to_Manual_xLC_and_vLC"
author: "Bernie Mulvey"
date: "03/15/26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)
library(bettermc) # for troubleshooting mclapply

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

## one-time preflight: concatenate all the banksy clustering output tables into one big fella
```{r}
banksysout <- list.files("processed-data/04_Banksy_clusterings/01-banksy_clustruns_outs",pattern=".txt")

## the column names didn't save anyhow, so this is also an opportunity to fix those...
megabank <- lapply(banksysout,function(b){
  curb <- fread(paste0("processed-data/04_Banksy_clusterings/01-banksy_clustruns_outs/", b
  ),header=F)
  setnames(curb,c("rn",
      paste0(
        gsub(b,pattern="^(.*)_louv05.*$",replacement="\\1"),
        c("_louv05","_louv1","_louv2"))))
  return(curb)
})
megabank2 <- megabank[[1]]
for (i in c(2:length(megabank))){
  megabank2 <- merge.data.table(megabank2,megabank[[i]],by="rn")
}
saveRDS(megabank2,"processed-data/04_Banksy_clusterings/01-AllBanksyClustsAssigns.RDS")
rm(i,banksysout,megabank)
gc()
```


load xenium data, manual xenium LC assignments; Visium NMF label transfers; add the latter 2 to coldata
```{r}
lcx <- readRDS("processed-data/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")
## remove slots we don't need
assays(lcx) <- assays(lcx)[1]

manlc <- fread("processed-data/03_ManualLC/01-XprBasedLClabels_and_cellSegsGainedorAlteredByNM.txt")
manlc <- manlc[,.(rn,LCloose,LCstrict)]

vispreds <- fread("processed-data/05_LabelTransfer/05b-LCsmoothings_LUT.txt")
vispreds <- vispreds[,.(rn,nmfpred)]

megabank2 <- readRDS("processed-data/04_Banksy_clusterings/01-AllBanksyClustsAssigns.RDS")

megabank2 <- merge.data.table(megabank2,manlc,by="rn")
megabank2 <- merge.data.table(megabank2,vispreds,by="rn")
stopifnot(nrow(megabank2)==ncol(lcx))

## reorder megabank2 to match lcx
megabank2 <- DataFrame(megabank2,row.names=megabank2$rn)[colnames(lcx),]
megabank2$rn  <- NULL
colData(lcx) <- cbind(colData(lcx),megabank2)

bcols <- grep("clust_M",colnames(colData(lcx)),value=T)
reglabs <- c("LCloose","LCstrict","nmfpred")
 
rm(manlc,megabank2,vispreds)
gc(full=T)
```

## get rid of unnecess parts of SPE for RAM purposes
```{r}
## remove unnecess columns from colData
tmpcd <- as.data.table(colData(lcx),keep.rownames=T)
keepcols <- c("rn","brnum","Age","Sex","APOE","Ancestry","YRI","CEU","LCloose","LCstrict","nmfpred",grep("clust_M",names(tmpcd),value=T))
tmpcd <- tmpcd[,..keepcols]
setnames(tmpcd,"brnum","sample_id") # we're using brnum as sample_id but SPE requires a sample_id column explicitly
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(lcx),]
tmpcd$rn <- NULL
colData(lcx) <- tmpcd

rm(tmpcd,keepcols)
```

## get the LC visium data registration wrapper stats themselves (already run including duplicateCorrelation due to multiple sections per donor in that dataset)
```{r}
visreg <- readRDS("../LFF_spatial_LC/processed-data/13_markers_and_registration_final/01-HDGSVG2575louv1_regWrapOut.RDS")
```

iterate thru spatial registration runs. remember, these just get us the marker stats, and then we have to use some manually written code to actually test the marker correlations (because the built-in command for it has never worked in my hands)
```{r}
setDTthreads(1,restore_after_fork = FALSE)

mannmf.regs <- mclapply(reglabs,mc.cores=3,mc.share.copy = FALSE,function(m){
    manout <- suppressWarnings(registration_wrapper(sce = lcx,var_registration = m,var_sample_id = "sample_id",covars = c("Age","Sex","APOE","YRI"),gene_ensembl = "ID",gene_name = "Symbol",min_ncells = 20))
    return(manout)
})
names(mannmf.regs) <- reglabs

## append the visium registration stats to "manual" LCs list (we want to compare each banksy run to the two manual xenium labelings and the final visium labeling)
mannmf.regs[[length(mannmf.regs)+1]] <- visreg
names(mannmf.regs)[length(mannmf.regs)] <- "LFF_Visium_9clust"

###

bankregs <- mclapply(bcols,mc.cores=3,mc.preschedule = TRUE,mc.share.copy = FALSE,function(b){
    manout <- registration_wrapper(sce = lcx,var_registration = b,var_sample_id = "sample_id",covars = c("Age","Sex","APOE","YRI"),gene_ensembl = "ID",gene_name = "Symbol",min_ncells = 20)
})
names(bankregs) <- bcols

gc(full=T)

## save these so we can load them later without re-running
saveRDS(mannmf.regs,"processed-data/04_Banksy_clusterings/02tmpfile-regwrapper_outs_manual_xLCs_Visdata_and_Vislabelxfers.RDS")
saveRDS(bankregs,"processed-data/04_Banksy_clusterings/02tmpfile-regwrapper_outs_nonHarmoBanksys.RDS")
```

```{r}
### I wrote this code to use inputs already subsetted to just t-statistics from the enrichment test tables. we can put these together in the following way. the xenium's simpler
bankregs2 <- lapply(bankregs,function(x){
  z <- as.data.table(x$enrichment)
  keepnames <- c("ensembl",grep(names(z),pattern="t_stat",value=T))
  z <- z[,..keepnames]
  # take the name down to just the generic cluster id
  setnames(z,
           c(2:ncol(z)),
           gsub(names(z)[c(2:ncol(z))],pattern="t_stat_",replacement=""))
  return(z)
})

mannmf.regs2 <- lapply(mannmf.regs,function(x){
  z <- as.data.table(x$enrichment)
  keepnames <- c("ensembl",grep(names(z),pattern="t_stat",value=T))
  z <- z[,..keepnames]
  # take the name down to just the generic cluster id
  setnames(z,
           c(2:ncol(z)),
           gsub(names(z)[c(2:ncol(z))],pattern="t_stat_",replacement=""))
  return(z)
})
```

### Define a function to run the registration over a series of N top genes, with option to go by top query markers 
```{r}
setDTthreads(1,restore_after_fork = FALSE)
# putting this ^ at every level as belt+susp
megatabulate.regist <- function(ref=list(),query=list(),orderbytop=c("query","ref")){
   regres <- lapply(ref,function(m){
      # putting this  at every level as belt+susp
      setDTthreads(1,restore_after_fork = FALSE)
      
      subres <- mcmapply(b=query,n=names(query),mc.cores=4,SIMPLIFY=FALSE,FUN=function(b,n){
       setDTthreads(1,restore_after_fork = FALSE) # put it at every level as belt+susp
       j<-seq(10,300,by=10)
       i<-1
       outlist <- list()
       cortab <- merge.data.table(m,b,by="ensembl")
       cortab$ensembl <- NULL
       for (i in c(1:(length(j)+1))){
           if(i<=length(j)){
               ngene <- j[i]
           }
           else{
               # no value for top_n --> all genes used
               ngene <- nrow(cortab)
           }
           # initialize cormat
           nclxen <- ncol(b)
           nclref <- ncol(m)
           cormat <- as.data.frame(matrix(nrow=nclxen-1,ncol=nclref-1))
           # skim the first column, the gene id
           k <- 2
           for (k in c(2:nclxen)){
               rownames(cormat)[k-1] <- colnames(b)[k]
               l <- 2
               for (l in c(2:nclref)){
                   colnames(cormat)[l-1] <- colnames(m)[l]
                   
                   curxen <- colnames(b)[k]
                   curref <- colnames(m)[l]
                   if(orderbytop=="query"){
                     setorderv(cortab,curxen,-1) # order the rows by current xenium cluster's top marker t stats in descending order, which we get the top N of below
                   } else {
                      setorderv(cortab,curref,-1)
                   }
                   
                   cormat[(k-1),(l-1)] <- cor(cortab[c(1:ngene),..curref],cortab[c(1:ngene),..curxen],use="pairwise.complete.obs",method = "spearman")
               }
           }
           plttab <- as.data.table(cormat,keep.rownames=T)
           setnames(plttab,"rn","queryclust")
           outlist[[i]] <- plttab
           names(outlist)[i] <- paste0(n,"_top_",ngene)
       }
       rm(cortab)
       return(outlist)
     })
   })

   regres2 <- lapply(regres,function(x){
      z <- lapply(x,function(y){
         a <- mapply(b=y,n=names(y),SIMPLIFY=FALSE,FUN=function(b,n){
            b[,newcol:=as.numeric(gsub(n,pattern="^clust.*_top_(.*)$",replacement="\\1"))]
            setnames(b,"newcol",paste0("topN_",orderbytop,"_mks"))
            return(b)
         })
         z2 <- rbindlist(a)
         return(z2)
      })
      z3 <- rbindlist(z, idcol="banksy_run")
      return(z3)
   })
   return(regres2)
}
```

```{r}
regres2 <- megatabulate.regist(ref=mannmf.regs2,query=bankregs2,orderbytop = "query")
saveRDS(regres2,"processed-data/04_Banksy_clusterings/02-BanksyToManualLCs_VisiumData_and_VisLabelXfer_regist_spears_byTopQueryMks.RDS")

regres3 <- megatabulate.regist(ref=mannmf.regs2,query=bankregs2,orderbytop="ref")
saveRDS(regres2,"processed-data/04_Banksy_clusterings/02-BanksyToManualLCs_VisiumData_and_VisLabelXfer_regist_spears_byTopRefMks.RDS")

```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0         parallelly_1.46.1           colorout_1.3-3             
 [4] spatialLIBD_1.22.0          SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
 [7] SummarizedExperiment_1.40.0 Biobase_2.70.0              GenomicRanges_1.62.1       
[10] Seqinfo_1.0.0               IRanges_2.44.0              S4Vectors_0.48.0           
[13] BiocGenerics_0.56.0         generics_0.1.4              MatrixGenerics_1.22.0      
[16] matrixStats_1.5.0           gridExtra_2.3               ggtext_0.1.2               
[19] ggplot2_4.0.2               data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3       rstudioapi_0.18.0        jsonlite_2.0.0          
  [4] shape_1.4.6.1            magrittr_2.0.4           ggbeeswarm_0.7.3        
  [7] magick_2.9.0             farver_2.1.2             rmarkdown_2.30          
 [10] BiocIO_1.20.0            GlobalOptions_0.1.3      vctrs_0.7.1             
 [13] Rsamtools_2.26.0         memoise_2.0.1            config_0.3.2            
 [16] RCurl_1.98-1.17          paletteer_1.7.0          benchmarkme_1.0.8       
 [19] htmltools_0.5.9          S4Arrays_1.10.1          AnnotationHub_4.0.0     
 [22] curl_7.0.0               BiocNeighbors_2.4.0      SparseArray_1.10.8      
 [25] sass_0.4.10              bslib_0.10.0             htmlwidgets_1.6.4       
 [28] httr2_1.2.2              plotly_4.12.0            cachem_1.1.0            
 [31] GenomicAlignments_1.46.0 mime_0.13                lifecycle_1.0.5         
 [34] iterators_1.0.14         pkgconfig_2.0.3          rsvd_1.0.5              
 [37] Matrix_1.7-4             R6_2.6.1                 fastmap_1.2.0           
 [40] shiny_1.12.1             clue_0.3-66              digest_0.6.39           
 [43] colorspace_2.1-2         rematch2_2.1.2           AnnotationDbi_1.72.0    
 [46] scater_1.38.0            rprojroot_2.1.1          irlba_2.3.7             
 [49] ExperimentHub_3.0.0      RSQLite_2.4.5            beachmat_2.26.0         
 [52] filelock_1.0.3           httr_1.4.7               abind_1.4-8             
 [55] compiler_4.5.2           here_1.0.2               bit64_4.6.0-1           
 [58] withr_3.0.2              doParallel_1.0.17        attempt_0.3.1           
 [61] S7_0.2.1                 viridis_0.6.5            DBI_1.2.3               
 [64] sessioninfo_1.2.3        rappdirs_0.3.4           DelayedArray_0.36.0     
 [67] rjson_0.2.23             tools_4.5.2              vipor_0.4.7             
 [70] otel_0.2.0               beeswarm_0.4.0           httpuv_1.6.16           
 [73] glue_1.8.0               restfulr_0.0.16          promises_1.5.0          
 [76] gridtext_0.1.5           grid_4.5.2               cluster_2.1.8.1         
 [79] gtable_0.3.6             tidyr_1.3.2              ScaledMatrix_1.18.0     
 [82] BiocSingular_1.26.0      xml2_1.5.2               XVector_0.50.0          
 [85] ggrepel_0.9.6            BiocVersion_3.22.0       foreach_1.5.2           
 [88] pillar_1.11.1            limma_3.66.0             later_1.4.5             
 [91] circlize_0.4.17          benchmarkmeData_2.0.0    dplyr_1.2.0             
 [94] BiocFileCache_3.0.0      lattice_0.22-7           rtracklayer_1.70.1      
 [97] bit_4.6.0                tidyselect_1.2.1         ComplexHeatmap_2.26.1   
[100] locfit_1.5-9.12          scuttle_1.20.0           Biostrings_2.78.0       
[103] knitr_1.51               edgeR_4.8.2              xfun_0.56               
[106] statmod_1.5.1            DT_0.34.0                lazyeval_0.2.2          
[109] yaml_2.3.12              shinyWidgets_0.9.0       cigarillo_1.0.0         
[112] evaluate_1.0.5           codetools_0.2-20         tibble_3.3.1            
[115] BiocManager_1.30.27      cli_3.6.5                xtable_1.8-4            
[118] jquerylib_0.1.4          golem_0.5.1              dichromat_2.0-0.1       
[121] Rcpp_1.1.1               dbplyr_2.5.1             png_0.1-8               
[124] XML_3.99-0.20            blob_1.3.0               bitops_1.0-9            
[127] viridisLite_0.4.3        scales_1.4.0             purrr_1.2.1             
[130] crayon_1.5.3             GetoptLong_1.1.0         rlang_1.1.7.9000        
[133] cowplot_1.2.0            KEGGREST_1.50.0         
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-03-16
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [2] CRAN (R 4.5.0)
 AnnotationDbi          1.72.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 AnnotationHub          4.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 attempt                0.3.1      2020-05-03 [2] CRAN (R 4.5.0)
 beachmat               2.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0      2021-06-01 [2] CRAN (R 4.5.0)
 benchmarkme            1.0.8      2022-06-12 [2] CRAN (R 4.5.0)
 benchmarkmeData        2.0.0      2026-01-19 [2] CRAN (R 4.5.2)
 Biobase              * 2.70.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocFileCache          3.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocIO                 1.20.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocManager            1.30.27    2025-11-14 [2] CRAN (R 4.5.2)
 BiocNeighbors          2.4.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocVersion            3.22.0     2025-10-07 [2] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0      2025-03-06 [2] CRAN (R 4.5.0)
 bit64                  4.6.0-1    2025-01-16 [2] CRAN (R 4.5.0)
 bitops                 1.0-9      2024-10-03 [2] CRAN (R 4.5.0)
 blob                   1.3.0      2026-01-14 [2] CRAN (R 4.5.2)
 bslib                  0.10.0     2026-01-26 [2] CRAN (R 4.5.2)
 cachem                 1.1.0      2024-05-16 [2] CRAN (R 4.5.0)
 cigarillo              1.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 circlize               0.4.17     2025-12-08 [2] CRAN (R 4.5.2)
 cli                    3.6.5      2025-04-23 [2] CRAN (R 4.5.0)
 clue                   0.3-66     2024-11-13 [2] CRAN (R 4.5.0)
 cluster                2.1.8.1    2025-03-12 [2] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [2] CRAN (R 4.5.2)
 colorout             * 1.3-3      2026-01-31 [2] Github (jalvesaq/colorout@64863bb)
 colorspace             2.1-2      2025-09-22 [2] CRAN (R 4.5.0)
 ComplexHeatmap         2.26.1     2026-01-30 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 config                 0.3.2      2023-08-30 [2] CRAN (R 4.5.0)
 cowplot                1.2.0      2025-07-07 [2] CRAN (R 4.5.0)
 crayon                 1.5.3      2024-06-20 [2] CRAN (R 4.5.0)
 curl                   7.0.0      2025-08-19 [2] CRAN (R 4.5.0)
 data.table           * 1.18.2.1   2026-01-27 [2] CRAN (R 4.5.2)
 DBI                    1.2.3      2024-06-02 [2] CRAN (R 4.5.0)
 dbplyr                 2.5.1      2025-09-10 [2] CRAN (R 4.5.0)
 DelayedArray           0.36.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [2] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [2] CRAN (R 4.5.2)
 doParallel             1.0.17     2022-02-07 [2] CRAN (R 4.5.0)
 dplyr                  1.2.0      2026-02-03 [2] CRAN (R 4.5.2)
 DT                     0.34.0     2025-09-02 [2] CRAN (R 4.5.0)
 edgeR                  4.8.2      2025-12-23 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate               1.0.5      2025-08-27 [2] CRAN (R 4.5.0)
 ExperimentHub          3.0.0      2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 farver                 2.1.2      2024-05-13 [2] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [2] CRAN (R 4.5.0)
 filelock               1.0.3      2023-12-11 [2] CRAN (R 4.5.0)
 foreach                1.5.2      2022-02-02 [2] CRAN (R 4.5.0)
 generics             * 0.1.4      2025-05-09 [2] CRAN (R 4.5.0)
 GenomicAlignments      1.46.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 GenomicRanges        * 1.62.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 GetoptLong             1.1.0      2025-11-28 [2] CRAN (R 4.5.2)
 ggbeeswarm             0.7.3      2025-11-29 [2] CRAN (R 4.5.2)
 ggplot2              * 4.0.2      2026-02-03 [2] CRAN (R 4.5.2)
 ggrepel                0.9.6      2024-09-07 [2] CRAN (R 4.5.0)
 ggtext               * 0.1.2      2022-09-16 [2] CRAN (R 4.5.0)
 GlobalOptions          0.1.3      2025-11-28 [2] CRAN (R 4.5.2)
 glue                   1.8.0      2024-09-30 [2] CRAN (R 4.5.0)
 golem                  0.5.1      2024-08-27 [2] CRAN (R 4.5.0)
 gridExtra            * 2.3        2017-09-09 [2] CRAN (R 4.5.0)
 gridtext               0.1.5      2022-09-16 [2] CRAN (R 4.5.0)
 gtable                 0.3.6      2024-10-25 [2] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [2] CRAN (R 4.5.0)
 htmltools              0.5.9      2025-12-04 [2] CRAN (R 4.5.2)
 htmlwidgets            1.6.4      2023-12-06 [2] CRAN (R 4.5.0)
 httpuv                 1.6.16     2025-04-16 [2] CRAN (R 4.5.0)
 httr                   1.4.7      2023-08-15 [2] CRAN (R 4.5.0)
 httr2                  1.2.2      2025-12-08 [2] CRAN (R 4.5.2)
 IRanges              * 2.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [2] CRAN (R 4.5.2)
 iterators              1.0.14     2022-02-05 [2] CRAN (R 4.5.0)
 jquerylib              0.1.4      2021-04-26 [2] CRAN (R 4.5.0)
 jsonlite               2.0.0      2025-03-27 [2] CRAN (R 4.5.0)
 KEGGREST               1.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51       2025-12-20 [2] CRAN (R 4.5.2)
 later                  1.4.5      2026-01-08 [2] CRAN (R 4.5.2)
 lattice                0.22-7     2025-04-02 [2] CRAN (R 4.5.2)
 lazyeval               0.2.2      2019-03-15 [2] CRAN (R 4.5.0)
 lifecycle              1.0.5      2026-01-08 [2] CRAN (R 4.5.2)
 limma                  3.66.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 locfit                 1.5-9.12   2025-03-05 [2] CRAN (R 4.5.0)
 magick                 2.9.0      2025-09-08 [2] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [2] CRAN (R 4.5.0)
 Matrix                 1.7-4      2025-08-28 [2] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [2] CRAN (R 4.5.0)
 memoise                2.0.1      2021-11-26 [2] CRAN (R 4.5.0)
 mime                   0.13       2025-03-17 [2] CRAN (R 4.5.0)
 otel                   0.2.0      2025-08-29 [2] CRAN (R 4.5.0)
 paletteer              1.7.0      2026-01-08 [2] CRAN (R 4.5.2)
 parallelly           * 1.46.1     2026-01-08 [2] CRAN (R 4.5.2)
 pillar                 1.11.1     2025-09-17 [2] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [2] CRAN (R 4.5.0)
 plotly                 4.12.0     2026-01-24 [2] CRAN (R 4.5.2)
 png                    0.1-8      2022-11-29 [2] CRAN (R 4.5.0)
 promises               1.5.0      2025-11-01 [2] CRAN (R 4.5.0)
 purrr                  1.2.1      2026-01-09 [2] CRAN (R 4.5.2)
 R6                     2.6.1      2025-02-15 [2] CRAN (R 4.5.0)
 rappdirs               0.3.4      2026-01-17 [2] CRAN (R 4.5.2)
 RColorBrewer           1.1-3      2022-04-03 [2] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [2] CRAN (R 4.5.2)
 RCurl                  1.98-1.17  2025-03-22 [2] CRAN (R 4.5.0)
 rematch2               2.1.2      2020-05-01 [2] CRAN (R 4.5.0)
 restfulr               0.0.16     2025-06-27 [2] CRAN (R 4.5.0)
 rjson                  0.2.23     2024-09-16 [2] CRAN (R 4.5.0)
 rlang                  1.1.7.9000 2026-02-04 [1] Github (r-lib/rlang@bae1fe1)
 rmarkdown              2.30       2025-09-28 [2] CRAN (R 4.5.0)
 rprojroot              2.1.1      2025-08-26 [2] CRAN (R 4.5.0)
 Rsamtools              2.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 RSQLite                2.4.5      2025-11-30 [2] CRAN (R 4.5.2)
 rstudioapi             0.18.0     2026-01-16 [2] CRAN (R 4.5.2)
 rsvd                   1.0.5      2021-04-16 [2] CRAN (R 4.5.0)
 rtracklayer            1.70.1     2025-12-20 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 S4Arrays               1.10.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1      2025-11-14 [2] CRAN (R 4.5.2)
 sass                   0.4.10     2025-04-11 [2] CRAN (R 4.5.0)
 ScaledMatrix           1.18.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0      2025-04-24 [2] CRAN (R 4.5.0)
 scater                 1.38.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 scuttle                1.20.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [2] CRAN (R 4.5.0)
 shape                  1.4.6.1    2024-02-23 [2] CRAN (R 4.5.0)
 shiny                  1.12.1     2025-12-09 [2] CRAN (R 4.5.2)
 shinyWidgets           0.9.0      2025-02-21 [2] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [2] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 spatialLIBD          * 1.22.0     2025-11-04 [2] Bioconductor 3.22 (R 4.5.2)
 statmod                1.5.1      2025-10-09 [2] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1      2026-01-11 [2] CRAN (R 4.5.2)
 tidyr                  1.3.2      2025-12-19 [2] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [2] CRAN (R 4.5.0)
 vctrs                  0.7.1      2026-01-23 [2] CRAN (R 4.5.2)
 vipor                  0.4.7      2023-12-18 [2] CRAN (R 4.5.0)
 viridis                0.6.5      2024-01-29 [2] CRAN (R 4.5.0)
 viridisLite            0.4.3      2026-02-04 [2] CRAN (R 4.5.2)
 withr                  3.0.2      2024-10-28 [2] CRAN (R 4.5.0)
 xfun                   0.56       2026-01-18 [2] CRAN (R 4.5.2)
 XML                    3.99-0.20  2025-11-08 [2] CRAN (R 4.5.0)
 xml2                   1.5.2      2026-01-17 [2] CRAN (R 4.5.2)
 xtable                 1.8-4      2019-04-21 [2] CRAN (R 4.5.0)
 XVector                0.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12     2025-12-10 [2] CRAN (R 4.5.2)

 [1] /Users/bmulvey/Library/R/arm64/4.5/library
 [2] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────────