---
title: "03-Register_Banksys_to_webdivSnseq"
author: "Bernie Mulvey"
date: "04-12-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(WeberDivechaLCdata)
library(spatialLIBD)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)
library(bettermc) # for troubleshooting mclapply/better functionality

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```


## the interesting RCTD findings came out at the "label" level, so we want to recreate those and use that here, including the sort of if-it-was-ne-by-any-definition-its-NE approach to maximize NE cells 
```{r}
#######
lcsn <- WeberDivechaLCdata_singleNucleus()
tmpcd <- as.data.table(colData(lcsn),keep.rownames=T)
## we will make a cluster hierarchy as follows: cluster_merged > label > label or inhibitory subcluster
# to ensure there's enough LCNE cells to work with, change label_merged and then label itself to NE for any cell classified as supervised NE OR unsup. NE
tmpcd[,label_merged:=as.character(label_merged)]
tmpcd[unsupervised_NE==TRUE,label_merged:="NE"]
tmpcd[supervised_NE==TRUE,label_merged:="NE"]

## then, make label a character column of merged_label (parent) and label (granular clust #)
tmpcd[,label:=as.character(label)]
tmpcd[,label:=paste0(label_merged,"_",label)]
## for NE neurons, disregard the granular cluster of origin since we're integrating several definition types now
tmpcd[label_merged=="NE",label:="NE"]
## should be 30 unique labels.
stopifnot(length(unique(tmpcd$label))>length(unique(tmpcd$label_merged)))
stopifnot(length(unique(tmpcd$label))==30)
colData(lcsn) <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(lcsn),]

rm(tmpcd)
gc(full=T)
```

```{r}
lcx <- readRDS("processed-data/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")
## remove slots we don't need
assays(lcx) <- assays(lcx)[1]
## banksy clusterings
megabank2 <- readRDS("processed-data/04_Banksy_clusterings/01-AllBanksyClustsAssigns.RDS")

## also load web-div NMF projections from granular labels (for comparison) 
wdpreds <- fread("processed-data/09_snSeqLabelTransfer/07-wdLCsmoothings_LUT.txt")

megabank2 <- merge.data.table(megabank2,wdpreds[,.(rn,nmfpred)],by="rn")
stopifnot(nrow(megabank2)==ncol(lcx))

## reorder megabank2 to match lcx
megabank2 <- DataFrame(megabank2,row.names=megabank2$rn)[colnames(lcx),]
megabank2$rn  <- NULL
colData(lcx) <- cbind(colData(lcx),megabank2)

## we're getting nmfpred and banksy cluster regist stats from the xenium obj
bcols <- c("nmfpred",grep("clust_M",colnames(colData(lcx)),value=T)) 
## and snseq labels (coarse and granular) from the snseq
scols <- c("label_merged","label")


rm(megabank2,wdpreds)
gc(full=T)
```

```{r}
### single nuc clustering stats
setDTthreads(1,restore_after_fork = FALSE)
snstats <- mclapply(scols,mc.cores=2,mc.share.copy = FALSE,FUN=function(s){ ## s is the column name
  regout <- registration_wrapper(sce = lcsn,var_registration = s,var_sample_id = "Sample",gene_ensembl = "gene_id",gene_name = "gene_name",min_ncells = 10)
  x <- as.data.table(regout$enrichment)
  y <- melt.data.table(x,id.vars=c("ensembl","gene"))
  y[,stat:=gsub(variable,pattern="^(t_stat|p_value|logFC|fdr)_.*$",replacement="\\1")]
  y[,clust:=gsub(variable,pattern="^(t_stat|p_value|logFC|fdr)_(.*)$",replacement="\\2")]
  y[,variable:=NULL]
  z <- dcast(y,ensembl+gene+clust~stat,value.var="value")
  return(z)
})
names(snstats) <- scols
rm(lcsn)
gc(full=T)

### banksy and nmfpred registration stats
bankregs <- mclapply(bcols,mc.cores=4,mc.share.vectors = TRUE,function(b){
    manout <- registration_wrapper(sce = lcx,var_registration = b,var_sample_id = "sample_id",covars = c("Age","Sex","APOE","YRI"),gene_ensembl = "ID",gene_name = "Symbol",min_ncells = 20)
    x <- as.data.table(manout$enrichment)
    y <- melt.data.table(x,id.vars=c("ensembl","gene"))
    y[,stat:=gsub(variable,pattern="^(t_stat|p_value|logFC|fdr)_.*$",replacement="\\1")]
    y[,clust:=gsub(variable,pattern="^(t_stat|p_value|logFC|fdr)_(.*)$",replacement="\\2")]
    y[,variable:=NULL]
    z <- dcast(y,ensembl+gene+clust~stat,value.var="value")
    return(z)
})
names(bankregs) <- bcols


## append the nmfpred registration stats to snseq LCs stats list (we want to compare each banksy run to the two manual xenium labelings and the final visium labeling)
snstats[[(length(snstats)+1)]] <- bankregs[["nmfpred"]]
names(snstats)[length(snstats)] <- "asNMFprojd_to_xen"

bankregs$nmfpred <- NULL
### save the banksy output itself since these are marker tables
saveRDS(bankregs,"processed-data/04_Banksy_clusterings/03-noHarmo_banksys_markertables.RDS")
gc(full=T)
```

extract putative LC clusters from each cluster set
```{r}
bsy.lcs <- lapply(bankregs,function(x){
  y <- x[gene %in% c("DBH","TH","SLC6A2","PHOX2A","PHOX2B","MAOA","MAOB") & fdr<0.1 & logFC>1.5]
  return(y)
})

saveRDS(bsy.lcs,"processed-data/04_Banksy_clusterings/03-noHarmo_banksys_putativeLCclusters.RDS")
```

how many LC clusters per clustering?
```{r}
nlc <- as.data.frame(unlist(lapply(bsy.lcs,function(x){length(unique(x$clust))})))
ntot <- as.data.frame(unlist(lapply(bankregs,function(x){length(unique(x$clust))})))
nlc <- as.data.table(nlc,keep.rownames=T)
ntot <- as.data.table(ntot,keep.rownames=T)
setnames(nlc,2,"n_putative_LC_clusts")
setnames(ntot,2,"n_total_clusts")
nlc <- merge.data.table(nlc,ntot,by="rn")
nlc[,n_putative_LC_clusts:=as.numeric(n_putative_LC_clusts)]
nlc[,n_total_clusts:=as.numeric(n_total_clusts)]
## well now. anywhere from 1-11 clusters enriched for these (?!)

png("plots/04_Banksy_clusterings/03-nPutativeLCclusts_vs_nTotal_clusts.png",width=4,height=4,res=300,units = "in")
ggplot(nlc,aes(x=n_total_clusts,y=n_putative_LC_clusts))+geom_point()
dev.off()
```

```{r}
### anyhow, re-arrange the outputs to a gene id column and column with enrichment t-stat per cluster 
bankregs2 <- lapply(bankregs,function(x){
  z <- dcast(x,ensembl~clust,value.var="t_stat")
  return(z)
})

snstats2 <- lapply(snstats,function(x){
  z <- dcast(x,ensembl~clust,value.var="t_stat")
  return(z)
})
```

### Define a function to run the registration over a series of N top genes, with option to go by top query markers 
```{r}
setDTthreads(1,restore_after_fork = FALSE)
# putting this ^ at every level as belt+susp

## we need actual parallel for mcmapply
library(parallel)
megatabulate.regist <- function(ref=list(),query=list(),orderbytop=c("query","ref")){
   regres <- lapply(ref,function(m){
      # putting this  at every level as belt+susp
      setDTthreads(1,restore_after_fork = FALSE)
      
      subres <- mcmapply(b=query,n=names(query),mc.cores=4,SIMPLIFY=FALSE,FUN=function(b,n){
       setDTthreads(1,restore_after_fork = FALSE) # put it at every level as belt+susp
       j<-seq(10,300,by=10)
       i<-1
       outlist <- list()
       cortab <- merge.data.table(m,b,by="ensembl")
       cortab$ensembl <- NULL
       for (i in c(1:(length(j)+1))){
           if(i<=length(j)){
               ngene <- j[i]
           }
           else{
               # no value for top_n --> all genes used
               ngene <- nrow(cortab)
           }
           # initialize cormat
           nclxen <- ncol(b)
           nclref <- ncol(m)
           cormat <- as.data.frame(matrix(nrow=nclxen-1,ncol=nclref-1))
           # skim the first column, the gene id
           k <- 2
           for (k in c(2:nclxen)){
               rownames(cormat)[k-1] <- colnames(b)[k]
               l <- 2
               for (l in c(2:nclref)){
                   colnames(cormat)[l-1] <- colnames(m)[l]
                   
                   curxen <- colnames(b)[k]
                   curref <- colnames(m)[l]
                   if(orderbytop=="query"){
                     setorderv(cortab,curxen,-1) # order the rows by current xenium cluster's top marker t stats in descending order, which we get the top N of below
                   } else {
                      setorderv(cortab,curref,-1)
                   }
                   
                   cormat[(k-1),(l-1)] <- cor(cortab[c(1:ngene),..curref],cortab[c(1:ngene),..curxen],use="pairwise.complete.obs",method = "spearman")
               }
           }
           plttab <- as.data.table(cormat,keep.rownames=T)
           setnames(plttab,"rn","queryclust")
           outlist[[i]] <- plttab
           names(outlist)[i] <- paste0(n,"_top_",ngene)
       }
       rm(cortab)
       return(outlist)
     })
   })

   regres2 <- lapply(regres,function(x){
      z <- lapply(x,function(y){
         a <- mapply(b=y,n=names(y),SIMPLIFY=FALSE,FUN=function(b,n){
            b[,newcol:=as.numeric(gsub(n,pattern="^clust.*_top_(.*)$",replacement="\\1"))]
            setnames(b,"newcol",paste0("topN_",orderbytop,"_mks"))
            return(b)
         })
         z2 <- rbindlist(a)
         return(z2)
      })
      z3 <- rbindlist(z, idcol="banksy_run")
      return(z3)
   })
   return(regres2)
}
```

```{r}
rm(lcx,bsy.lcs,snstats,bankregs,ntot,nlc,bcols,scols)
gc(full=T)

regres2 <- megatabulate.regist(ref=snstats2,query=bankregs2,orderbytop = "query")
saveRDS(regres2,"processed-data/04_Banksy_clusterings/03-BanksyToWebDivsnSeq_coarse_granular_and_wdSnGranularNMFprojToXen_regist_spears_byTopQueryMks.RDS")

regres3 <- megatabulate.regist(ref=snstats2,query=bankregs2,orderbytop="ref")
saveRDS(regres3,"processed-data/04_Banksy_clusterings/03-BanksyToWebDivsnSeq_coarse_granular_and_wdSnGranularNMFprojToXen_regist_spears_byTopRefMks.RDS")
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics 
[5] grDevices utils     methods   base     

other attached packages:
 [1] bettermc_1.2.2.9000        
 [2] BiocParallel_1.44.0        
 [3] parallelly_1.46.1          
 [4] colorout_1.3-3             
 [5] spatialLIBD_1.22.0         
 [6] WeberDivechaLCdata_1.12.0  
 [7] ExperimentHub_3.0.0        
 [8] AnnotationHub_4.0.0        
 [9] BiocFileCache_3.0.0        
[10] dbplyr_2.5.1               
[11] SpatialExperiment_1.20.0   
[12] SingleCellExperiment_1.32.0
[13] SummarizedExperiment_1.40.0
[14] Biobase_2.70.0             
[15] GenomicRanges_1.62.1       
[16] Seqinfo_1.0.0              
[17] IRanges_2.44.0             
[18] S4Vectors_0.48.0           
[19] BiocGenerics_0.56.0        
[20] generics_0.1.4             
[21] MatrixGenerics_1.22.0      
[22] matrixStats_1.5.0          
[23] gridExtra_2.3              
[24] ggtext_0.1.2               
[25] ggplot2_4.0.2              
[26] data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] later_1.4.5             
  [2] BiocIO_1.20.0           
  [3] bitops_1.0-9            
  [4] filelock_1.0.3          
  [5] tibble_3.3.1            
  [6] XML_3.99-0.20           
  [7] lifecycle_1.0.5         
  [8] httr2_1.2.2             
  [9] edgeR_4.8.2             
 [10] doParallel_1.0.17       
 [11] rprojroot_2.1.1         
 [12] lattice_0.22-7          
 [13] backports_1.5.0         
 [14] magrittr_2.0.4          
 [15] limma_3.66.0            
 [16] plotly_4.12.0           
 [17] sass_0.4.10             
 [18] jquerylib_0.1.4         
 [19] yaml_2.3.12             
 [20] httpuv_1.6.16           
 [21] otel_0.2.0              
 [22] sessioninfo_1.2.3       
 [23] cowplot_1.2.0           
 [24] DBI_1.2.3               
 [25] RColorBrewer_1.1-3      
 [26] golem_0.5.1             
 [27] abind_1.4-8             
 [28] purrr_1.2.1             
 [29] RCurl_1.98-1.17         
 [30] rappdirs_0.3.4          
 [31] circlize_0.4.17         
 [32] ggrepel_0.9.6           
 [33] irlba_2.3.7             
 [34] codetools_0.2-20        
 [35] DelayedArray_0.36.0     
 [36] DT_0.34.0               
 [37] scuttle_1.20.0          
 [38] xml2_1.5.2              
 [39] tidyselect_1.2.1        
 [40] shape_1.4.6.1           
 [41] farver_2.1.2            
 [42] ScaledMatrix_1.18.0     
 [43] viridis_0.6.5           
 [44] shinyWidgets_0.9.0      
 [45] GenomicAlignments_1.46.0
 [46] jsonlite_2.0.0          
 [47] GetoptLong_1.1.0        
 [48] BiocNeighbors_2.4.0     
 [49] scater_1.38.0           
 [50] iterators_1.0.14        
 [51] systemfonts_1.3.1       
 [52] foreach_1.5.2           
 [53] progress_1.2.3          
 [54] tools_4.5.2             
 [55] ragg_1.5.0              
 [56] Rcpp_1.1.1              
 [57] glue_1.8.0              
 [58] SparseArray_1.10.8      
 [59] xfun_0.56               
 [60] here_1.0.2              
 [61] dplyr_1.2.0             
 [62] withr_3.0.2             
 [63] BiocManager_1.30.27     
 [64] fastmap_1.2.0           
 [65] digest_0.6.39           
 [66] rsvd_1.0.5              
 [67] R6_2.6.1                
 [68] mime_0.13               
 [69] textshaping_1.0.4       
 [70] colorspace_2.1-2        
 [71] dichromat_2.0-0.1       
 [72] RSQLite_2.4.5           
 [73] cigarillo_1.0.0         
 [74] config_0.3.2            
 [75] tidyr_1.3.2             
 [76] rtracklayer_1.70.1      
 [77] prettyunits_1.2.0       
 [78] httr_1.4.7              
 [79] htmlwidgets_1.6.4       
 [80] S4Arrays_1.10.1         
 [81] pkgconfig_2.0.3         
 [82] gtable_0.3.6            
 [83] blob_1.3.0              
 [84] ComplexHeatmap_2.26.1   
 [85] S7_0.2.1                
 [86] XVector_0.50.0          
 [87] htmltools_0.5.9         
 [88] clue_0.3-66             
 [89] scales_1.4.0            
 [90] png_0.1-8               
 [91] attempt_0.3.1           
 [92] knitr_1.51              
 [93] rjson_0.2.23            
 [94] checkmate_2.3.4         
 [95] curl_7.0.0              
 [96] cachem_1.1.0            
 [97] GlobalOptions_0.1.3     
 [98] BiocVersion_3.22.0      
 [99] vipor_0.4.7             
[100] AnnotationDbi_1.72.0    
[101] restfulr_0.0.16         
[102] pillar_1.11.1           
[103] grid_4.5.2              
[104] vctrs_0.7.1             
[105] promises_1.5.0          
[106] BiocSingular_1.26.0     
[107] beachmat_2.26.0         
[108] xtable_1.8-4            
[109] cluster_2.1.8.1         
[110] beeswarm_0.4.0          
[111] paletteer_1.7.0         
[112] evaluate_1.0.5          
[113] magick_2.9.0            
[114] cli_3.6.5               
[115] locfit_1.5-9.12         
[116] compiler_4.5.2          
[117] Rsamtools_2.26.0        
[118] rlang_1.1.7             
[119] crayon_1.5.3            
[120] labeling_0.4.3          
[121] rematch2_2.1.2          
[122] ggbeeswarm_0.7.3        
[123] viridisLite_0.4.3       
[124] Biostrings_2.78.0       
[125] lazyeval_0.2.2          
[126] Matrix_1.7-4            
[127] benchmarkme_1.0.8       
[128] hms_1.1.4               
[129] bit64_4.6.0-1           
[130] KEGGREST_1.50.0         
[131] statmod_1.5.1           
[132] shiny_1.12.1            
[133] gridtext_0.1.5          
[134] memoise_2.0.1           
[135] bslib_0.10.0            
[136] benchmarkmeData_2.0.0   
[137] bit_4.6.0               
> sessioninfo::session_info()
─ Session info ──────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       Positron
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-04-12
 pandoc   NA
 quarto   1.8.27 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [1] CRAN (R 4.5.0)
 AnnotationDbi          1.72.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationHub        * 4.0.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 attempt                0.3.1      2020-05-03 [1] CRAN (R 4.5.0)
 backports              1.5.0      2024-05-23 [1] CRAN (R 4.5.0)
 beachmat               2.26.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0      2021-06-01 [1] CRAN (R 4.5.0)
 benchmarkme            1.0.8      2022-06-12 [1] CRAN (R 4.5.0)
 benchmarkmeData        2.0.0      2026-01-19 [1] CRAN (R 4.5.2)
 bettermc             * 1.2.2.9000 2026-03-16 [1] Github (akersting/bettermc@d6ec1aa)
 Biobase              * 2.70.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocFileCache        * 3.0.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocIO                 1.20.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocManager            1.30.27    2025-11-14 [1] CRAN (R 4.5.2)
 BiocNeighbors          2.4.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocVersion            3.22.0     2025-10-07 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0      2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1    2025-01-16 [1] CRAN (R 4.5.0)
 bitops                 1.0-9      2024-10-03 [1] CRAN (R 4.5.0)
 blob                   1.3.0      2026-01-14 [1] CRAN (R 4.5.2)
 bslib                  0.10.0     2026-01-26 [1] CRAN (R 4.5.2)
 cachem                 1.1.0      2024-05-16 [1] CRAN (R 4.5.0)
 checkmate              2.3.4      2026-02-03 [1] CRAN (R 4.5.2)
 cigarillo              1.0.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 circlize               0.4.17     2025-12-08 [1] CRAN (R 4.5.2)
 cli                    3.6.5      2025-04-23 [1] CRAN (R 4.5.0)
 clue                   0.3-66     2024-11-13 [1] CRAN (R 4.5.0)
 cluster                2.1.8.1    2025-03-12 [1] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3      2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 colorspace             2.1-2      2025-09-22 [1] CRAN (R 4.5.0)
 ComplexHeatmap         2.26.1     2026-01-30 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 config                 0.3.2      2023-08-30 [1] CRAN (R 4.5.0)
 cowplot                1.2.0      2025-07-07 [1] CRAN (R 4.5.0)
 crayon                 1.5.3      2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0      2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1   2026-01-27 [1] CRAN (R 4.5.2)
 DBI                    1.2.3      2024-06-02 [1] CRAN (R 4.5.0)
 dbplyr               * 2.5.1      2025-09-10 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [1] CRAN (R 4.5.2)
 doParallel             1.0.17     2022-02-07 [1] CRAN (R 4.5.0)
 dplyr                  1.2.0      2026-02-03 [1] CRAN (R 4.5.2)
 DT                     0.34.0     2025-09-02 [1] CRAN (R 4.5.0)
 edgeR                  4.8.2      2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate               1.0.5      2025-08-27 [1] CRAN (R 4.5.0)
 ExperimentHub        * 3.0.0      2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 farver                 2.1.2      2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [1] CRAN (R 4.5.0)
 filelock               1.0.3      2023-12-11 [1] CRAN (R 4.5.0)
 foreach                1.5.2      2022-02-02 [1] CRAN (R 4.5.0)
 generics             * 0.1.4      2025-05-09 [1] CRAN (R 4.5.0)
 GenomicAlignments      1.46.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 GenomicRanges        * 1.62.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 GetoptLong             1.1.0      2025-11-28 [1] CRAN (R 4.5.2)
 ggbeeswarm             0.7.3      2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2      2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6      2024-09-07 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2      2022-09-16 [1] CRAN (R 4.5.0)
 GlobalOptions          0.1.3      2025-11-28 [1] CRAN (R 4.5.2)
 glue                   1.8.0      2024-09-30 [1] CRAN (R 4.5.0)
 golem                  0.5.1      2024-08-27 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3        2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5      2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6      2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [1] CRAN (R 4.5.0)
 hms                    1.1.4      2025-10-17 [1] CRAN (R 4.5.0)
 htmltools              0.5.9      2025-12-04 [1] CRAN (R 4.5.2)
 htmlwidgets            1.6.4      2023-12-06 [1] CRAN (R 4.5.0)
 httpuv                 1.6.16     2025-04-16 [1] CRAN (R 4.5.0)
 httr                   1.4.7      2023-08-15 [1] CRAN (R 4.5.0)
 httr2                  1.2.2      2025-12-08 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [1] CRAN (R 4.5.2)
 iterators              1.0.14     2022-02-05 [1] CRAN (R 4.5.0)
 jquerylib              0.1.4      2021-04-26 [1] CRAN (R 4.5.0)
 jsonlite               2.0.0      2025-03-27 [1] CRAN (R 4.5.0)
 KEGGREST               1.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51       2025-12-20 [1] CRAN (R 4.5.2)
 labeling               0.4.3      2023-08-29 [1] CRAN (R 4.5.0)
 later                  1.4.5      2026-01-08 [1] CRAN (R 4.5.2)
 lattice                0.22-7     2025-04-02 [1] CRAN (R 4.5.2)
 lazyeval               0.2.2      2019-03-15 [1] CRAN (R 4.5.0)
 lifecycle              1.0.5      2026-01-08 [1] CRAN (R 4.5.2)
 limma                  3.66.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 locfit                 1.5-9.12   2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0      2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4      2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1      2021-11-26 [1] CRAN (R 4.5.0)
 mime                   0.13       2025-03-17 [1] CRAN (R 4.5.0)
 otel                   0.2.0      2025-08-29 [1] CRAN (R 4.5.0)
 paletteer              1.7.0      2026-01-08 [1] CRAN (R 4.5.2)
 parallelly           * 1.46.1     2026-01-08 [1] CRAN (R 4.5.2)
 pillar                 1.11.1     2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [1] CRAN (R 4.5.0)
 plotly                 4.12.0     2026-01-24 [1] CRAN (R 4.5.2)
 png                    0.1-8      2022-11-29 [1] CRAN (R 4.5.0)
 prettyunits            1.2.0      2023-09-24 [1] CRAN (R 4.5.0)
 progress               1.2.3      2023-12-06 [1] CRAN (R 4.5.0)
 promises               1.5.0      2025-11-01 [1] CRAN (R 4.5.0)
 purrr                  1.2.1      2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1      2025-02-15 [1] CRAN (R 4.5.0)
 ragg                   1.5.0      2025-09-02 [1] CRAN (R 4.5.0)
 rappdirs               0.3.4      2026-01-17 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3      2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [1] CRAN (R 4.5.2)
 RCurl                  1.98-1.17  2025-03-22 [1] CRAN (R 4.5.0)
 rematch2               2.1.2      2020-05-01 [1] CRAN (R 4.5.0)
 restfulr               0.0.16     2025-06-27 [1] CRAN (R 4.5.0)
 rjson                  0.2.23     2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7      2026-01-09 [1] CRAN (R 4.5.2)
 rprojroot              2.1.1      2025-08-26 [1] CRAN (R 4.5.0)
 Rsamtools              2.26.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 RSQLite                2.4.5      2025-11-30 [1] CRAN (R 4.5.2)
 rsvd                   1.0.5      2021-04-16 [1] CRAN (R 4.5.0)
 rtracklayer            1.70.1     2025-12-20 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 S4Arrays               1.10.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1      2025-11-14 [1] CRAN (R 4.5.2)
 sass                   0.4.10     2025-04-11 [1] CRAN (R 4.5.0)
 ScaledMatrix           1.18.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0      2025-04-24 [1] CRAN (R 4.5.0)
 scater                 1.38.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 scuttle                1.20.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [1] CRAN (R 4.5.0)
 shape                  1.4.6.1    2024-02-23 [1] CRAN (R 4.5.0)
 shiny                  1.12.1     2025-12-09 [1] CRAN (R 4.5.2)
 shinyWidgets           0.9.0      2025-02-21 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 spatialLIBD          * 1.22.0     2025-11-04 [1] Bioconductor 3.22 (R 4.5.2)
 statmod                1.5.1      2025-10-09 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 systemfonts            1.3.1      2025-10-01 [1] CRAN (R 4.5.0)
 textshaping            1.0.4      2025-10-10 [1] CRAN (R 4.5.0)
 tibble                 3.3.1      2026-01-11 [1] CRAN (R 4.5.2)
 tidyr                  1.3.2      2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1      2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7      2023-12-18 [1] CRAN (R 4.5.0)
 viridis                0.6.5      2024-01-29 [1] CRAN (R 4.5.0)
 viridisLite            0.4.3      2026-02-04 [1] CRAN (R 4.5.2)
 WeberDivechaLCdata   * 1.12.0     2025-11-04 [1] Bioconductor 3.22 (R 4.5.2)
 withr                  3.0.2      2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56       2026-01-18 [1] CRAN (R 4.5.2)
 XML                    3.99-0.20  2025-11-08 [1] CRAN (R 4.5.0)
 xml2                   1.5.2      2026-01-17 [1] CRAN (R 4.5.2)
 xtable                 1.8-4      2019-04-21 [1] CRAN (R 4.5.0)
 XVector                0.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12     2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────
