---
title: "05-DAPIonly_PreBanksy spatCoord staggering"
author: "Bernie Mulvey"
date: "2026-04-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(scuttle)
library(Banksy)
library(BiocParallel)
library(RcppHNSW) ## faster UMAP in uwot 0.2 (used by banksy UMAP)


## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
```

## load QCed, genes only, NON-log normalized data
```{r}
lcx <- readRDS("processed-data/02_QC/07b-DAPIonly_spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

## re normalization: Banksy uses  NON-log normcounts (per its documentation)
## the loaded SPE already contains the nonlog normaliezd count matrix and is subsetted to gene-targeting probes, so nothing to do on these fronts. lets get rid of the logcounts matrix just to be safe.

assays(lcx)["logcounts"] <- NULL
assayNames(lcx)
# "counts"  "normcounts"
# we just tell banksy in the actual call to use normcounts.
```


## to run Banksy in multisample mode, adjust coordinates so that
## every sample has a unique, non-overlapping set of coords (i.e. to prevent
## cells from a given area on two different slides from colliding)
## code adapted from: https://prabhakarlab.github.io/Banksy/articles/batch-correction.html
```{r}
# get the spatial coords matrix and their rownames to make sure everything matches back up in the end
locs <- cbind(rownames(spatialCoords(lcx)),spatialCoords(lcx)) 
colnames(locs)[1] <- "rn"

# get sample IDs (each brnum entry (brnum+L if left hemi) is a sample id. we have to make into an integer factor here though, which makes the spacing math simpler. we can't do that with a data.table though for some reason
locs <- cbind(locs,factor(lcx$sample_id))
locs <- as.data.table(locs)
setnames(locs,4,"group") # the sample-as-integer-factor column.

locs[,x_centroid:=as.numeric(x_centroid)]
locs[,y_centroid:=as.numeric(y_centroid)]
locs[,group:=as.numeric(group)] # for mathing
setnames(locs,c(2:3),c("sdimx", "sdimy")) # names expected by banksy

# now spread the x coords out so that no two samples overlap in space, even from different slides         
locs[, sdimx := sdimx - min(sdimx), by = group]
global_max <- max(locs$sdimx) * 1.5 # 1.5x the widest sample's width
locs[, sdimx := sdimx + group * global_max] # put that much space between each sample
locs <- as.matrix(as.data.frame(locs[, .(sdimx,sdimy)],row.names=locs$rn))
spatialCoords(lcx) <- locs[colnames(lcx),] # ensuring that ordering of spatialCoords matches everything else

## finally, banksy can pull these from the colData instead of spatialcoords if we're having trouble, so put them in coldata just in case.
colData(lcx)$sdimx <- spatialCoords(lcx)[,1]
colData(lcx)$sdimy <- spatialCoords(lcx)[,2]

# save this SPE for inputting into banksy
saveRDS(lcx,"processed-data/04_Banksy_clusterings/05-DAPIonly_staggeredCoord_spe_genetargOnly_NONlogNorm.RDS")
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] parallelly_1.46.1           colorout_1.3-3             
 [3] RcppHNSW_0.6.0              BiocParallel_1.44.0        
 [5] Banksy_1.6.0                scuttle_1.20.0             
 [7] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
 [9] SummarizedExperiment_1.40.0 Biobase_2.70.0             
[11] GenomicRanges_1.62.1        Seqinfo_1.0.0              
[13] IRanges_2.44.0              S4Vectors_0.48.0           
[15] BiocGenerics_0.56.0         generics_0.1.4             
[17] MatrixGenerics_1.22.0       matrixStats_1.5.0          
[19] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] gtable_0.3.6        rjson_0.2.23        xfun_0.56           ggplot2_4.0.2      
 [5] lattice_0.22-7      vctrs_0.7.1         tools_4.5.2         parallel_4.5.2     
 [9] tibble_3.3.1        sccore_1.0.6        pkgconfig_2.0.3     Matrix_1.7-4       
[13] RColorBrewer_1.1-3  S7_0.2.1            lifecycle_1.0.5     compiler_4.5.2     
[17] farver_2.1.2        aricode_1.0.3       codetools_0.2-20    htmltools_0.5.9    
[21] yaml_2.3.12         pillar_1.11.1       uwot_0.2.4          DelayedArray_0.36.0
[25] dbscan_1.2.4        magick_2.9.0        abind_1.4-8         mclust_6.1.2       
[29] tidyselect_1.2.1    digest_0.6.39       dplyr_1.2.0         rprojroot_2.1.1    
[33] fastmap_1.2.0       grid_4.5.2          here_1.0.2          cli_3.6.5          
[37] SparseArray_1.10.8  magrittr_2.0.4      leidenAlg_1.1.5     S4Arrays_1.10.1    
[41] dichromat_2.0-0.1   scales_1.4.0        rmarkdown_2.30      XVector_0.50.0     
[45] igraph_2.2.1        otel_0.2.0          beachmat_2.26.0     evaluate_1.0.5     
[49] RcppHungarian_0.3   knitr_1.51          irlba_2.3.7         rlang_1.1.7.9000   
[53] Rcpp_1.1.1          glue_1.8.0          rstudioapi_0.18.0   R6_2.6.1           
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-04-16
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [2] CRAN (R 4.5.0)
 aricode                1.0.3      2023-10-20 [2] CRAN (R 4.5.0)
 Banksy               * 1.6.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 beachmat               2.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 Biobase              * 2.70.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5      2025-04-23 [2] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [2] CRAN (R 4.5.2)
 colorout             * 1.3-3      2026-01-31 [2] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.2.1   2026-01-27 [2] CRAN (R 4.5.2)
 dbscan                 1.2.4      2025-12-19 [2] CRAN (R 4.5.2)
 DelayedArray           0.36.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [2] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [2] CRAN (R 4.5.2)
 dplyr                  1.2.0      2026-02-03 [2] CRAN (R 4.5.2)
 evaluate               1.0.5      2025-08-27 [2] CRAN (R 4.5.0)
 farver                 2.1.2      2024-05-13 [2] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [2] CRAN (R 4.5.0)
 generics             * 0.1.4      2025-05-09 [2] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 ggplot2                4.0.2      2026-02-03 [2] CRAN (R 4.5.2)
 glue                   1.8.0      2024-09-30 [2] CRAN (R 4.5.0)
 gtable                 0.3.6      2024-10-25 [2] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [2] CRAN (R 4.5.0)
 htmltools              0.5.9      2025-12-04 [2] CRAN (R 4.5.2)
 igraph                 2.2.1      2025-10-27 [2] CRAN (R 4.5.0)
 IRanges              * 2.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [2] CRAN (R 4.5.2)
 knitr                  1.51       2025-12-20 [2] CRAN (R 4.5.2)
 lattice                0.22-7     2025-04-02 [2] CRAN (R 4.5.2)
 leidenAlg              1.1.5      2025-04-19 [2] CRAN (R 4.5.0)
 lifecycle              1.0.5      2026-01-08 [2] CRAN (R 4.5.2)
 magick                 2.9.0      2025-09-08 [2] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [2] CRAN (R 4.5.0)
 Matrix                 1.7-4      2025-08-28 [2] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [2] CRAN (R 4.5.0)
 mclust                 6.1.2      2025-10-31 [2] CRAN (R 4.5.0)
 otel                   0.2.0      2025-08-29 [2] CRAN (R 4.5.0)
 parallelly           * 1.46.1     2026-01-08 [2] CRAN (R 4.5.2)
 pillar                 1.11.1     2025-09-17 [2] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [2] CRAN (R 4.5.0)
 R6                     2.6.1      2025-02-15 [2] CRAN (R 4.5.0)
 RColorBrewer           1.1-3      2022-04-03 [2] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [2] CRAN (R 4.5.2)
 RcppHNSW             * 0.6.0      2024-02-04 [2] CRAN (R 4.5.0)
 RcppHungarian          0.3        2023-09-05 [2] CRAN (R 4.5.0)
 rjson                  0.2.23     2024-09-16 [2] CRAN (R 4.5.0)
 rlang                  1.1.7.9000 2026-02-04 [1] Github (r-lib/rlang@bae1fe1)
 rmarkdown              2.30       2025-09-28 [2] CRAN (R 4.5.0)
 rprojroot              2.1.1      2025-08-26 [2] CRAN (R 4.5.0)
 rstudioapi             0.18.0     2026-01-16 [2] CRAN (R 4.5.2)
 S4Arrays               1.10.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1      2025-11-14 [2] CRAN (R 4.5.2)
 scales                 1.4.0      2025-04-24 [2] CRAN (R 4.5.0)
 sccore                 1.0.6      2025-04-13 [2] CRAN (R 4.5.0)
 scuttle              * 1.20.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [2] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [2] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1      2026-01-11 [2] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [2] CRAN (R 4.5.0)
 uwot                   0.2.4      2025-11-10 [2] CRAN (R 4.5.0)
 vctrs                  0.7.1      2026-01-23 [2] CRAN (R 4.5.2)
 xfun                   0.56       2026-01-18 [2] CRAN (R 4.5.2)
 XVector                0.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12     2025-12-10 [2] CRAN (R 4.5.2)

 [1] /Users/bmulvey/Library/R/arm64/4.5/library
 [2] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────────────