---
title: "06-DAPIonly_Banksy_M0l0.2k25kg30louv1"
author: "Bernie Mulvey"
date: "04-16-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(scuttle)
library(Banksy)
library(BiocParallel)
## faster UMAP in uwot 0.2 (used by banksy UMAP)
library(RcppHNSW)
library(here)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
lcx <- readRDS("processed-data/04_Banksy_clusterings/05-DAPIonly_staggeredCoord_spe_genetargOnly_NONlogNorm.RDS")

gabor=FALSE
kgeom=30
lam=0.2
useassay="normcounts"
monocot=42 # seed

# aaand run Banksy. this would be the same as NOT rearranging coords, then running on an SPE for each sample individually and joining them all back together after this step into one  SPE again. I think.

lcx <- computeBanksy(lcx, 
                      assay_name = useassay,
                      compute_agf = gabor,
                      k_geom = kgeom,
                      seed=monocot,
                      parallel = TRUE,
                      num_cores = 10)

## run Banksy PCA for subseequent input to clustering
## CRITICAL HERE: specify "group" as the sample_id -- i.e., each
## sample's Banksy PCA will be computed separately, making this
## equivalent to running Banksy on a list of SPEs with each entry as
## one sample, the other approach their vignettes uses
lcx <- runBanksyPCA(lcx,
                         assay_name=useassay,
                         use_agf=gabor,
                         lambda=lam,
                         group="sample_id",
                         seed=monocot)

## make sure to grab the BanksyPCA in case we want to subsequently Harmony it.
## we only need to save this once since we can run louvain at all desired
## resolutions on this. clusterBanksy does not parallelize so we run all resolutions
## in parallel instead.

pcaoutname <- paste0("PCA_M",ifelse(gabor,1,0),"_lam",lam)
pcaout <- reducedDims(lcx)[[pcaoutname]]
pcaoutpath <- paste0(
        "processed-data/04_Banksy_clusterings/06-DAPIonly_banksyPCA_",
        pcaoutname,"_kgeom",kgeom,".RDS")

saveRDS(pcaout,pcaoutpath)
rm(pcaout,pcaoutpath)
gc()
```
## run Banksy clustering -- louvain clustering (default method is leiden).
## several unwritten parameters will default to their defaults here. 
## and this will take a longgggg time, though louvain seems snappier.
## note that by default clusterBanksy would use k=50 for the 
## knn graph upstream of louv/leid which is a silly big number here.
## a two sample test run with k=50 at resolution 1 -> 19 clusters;
## k_neighb=25, res 1 -> 23 clusters, so similar # of clusters. we'll use
## k_neighbors=25, resolutions 0.5, 1, and 2.

### since louvain clustering itself cannot be parallelized, we can run a couple different k_neighbors in parallel. to avoid stupid collisions in memory, make copies of the SPE (its not THAT big) to ensure each process has its own copy. do this with base parallel:: , i don't trust biocp with SPE objects
```{r}
lcx3res <- list(lcx,lcx,lcx)
library(parallel)
bclusts <- mcmapply(s=lcx3res,kn=c(25,50,100),mc.cores=3,SIMPLIFY=FALSE,function(s,kn){
        o <- Banksy::clusterBanksy(s,
              assay_name=useassay,
              use_agf = gabor,
              lambda = lam,
              k_neighbors = kn,
              resolution = 1,
              group="sample_id",
              seed = monocot,algo = "louvain")

        ## make the clustering result non-numeric for ease of plotting later;
        ## extract the clustering result, rather than saving a whole SPE for this one clustering result of manyy being tested.
        # the clustering result is a new column tacked onto colData.

        keepname <- c("rn",colnames(colData(o))[ncol(colData(o))])
        outtab <- as.data.table(colData(o),keep.rownames=T)[,..keepname]
        setnames(outtab,2,"clusts")
        outtab[,clusts:=paste0("X",as.character(clusts))]
        # append res (its already there, we fix it in the final table)

        setnames(outtab,2,paste0(keepname[2],"_kgeom",kgeom,"_res1"))
        
        # this column name will tell us each of the clustering parameters used for this run
        # e.g., clust_M0_lam0_k25_kgeom30_res1 means
        # 1. M=0 (no gabor), 
        # 2. lambda = 0,
        # 3. k25 means k neighbors for clustering = 25
        # 4. kgeom=30 here, but is a parameter we've varied in full banksy param sweeps
        # 5. res2 means clustering resolution 1, 
        
return(outtab)})

# concat and save the cluster assignments
stopifnot(nrow(bclusts[[1]])==nrow(bclusts[[2]])&nrow(bclusts[[1]])==nrow(bclusts[[3]]))
outtab <- merge.data.table(bclusts[[1]],bclusts[[2]],by="rn")
outtab <- merge.data.table(outtab,bclusts[[3]],by="rn")

## get rid of the redundant, less visible duplication of the resolution info
names(outtab)
setnames(outtab,gsub(names(outtab),pattern="^(clust_M0_lam0.2_k.*_)res1_(kgeom.*_res1)$",replacement=paste0("\\1","\\2")))

fwrite(outtab,
       "processed-data/04_Banksy_clusterings/06-DAPIonly_M0l0.2kgeom30louv1_with_kneighb25-50-100.txt",
       sep='\t',
       quote=F,
       row.names=F,
       col.names=T)
```
seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] parallelly_1.46.1           colorout_1.3-3             
 [3] here_1.0.2                  RcppHNSW_0.6.0             
 [5] BiocParallel_1.44.0         Banksy_1.6.0               
 [7] scuttle_1.20.0              SpatialExperiment_1.20.0   
 [9] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
[11] Biobase_2.70.0              GenomicRanges_1.62.1       
[13] Seqinfo_1.0.0               IRanges_2.44.0             
[15] S4Vectors_0.48.0            BiocGenerics_0.56.0        
[17] generics_0.1.4              MatrixGenerics_1.22.0      
[19] matrixStats_1.5.0           data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] gtable_0.3.6        rjson_0.2.23        xfun_0.56           ggplot2_4.0.2      
 [5] lattice_0.22-7      vctrs_0.7.1         tools_4.5.2         tibble_3.3.1       
 [9] sccore_1.0.6        pkgconfig_2.0.3     Matrix_1.7-4        RColorBrewer_1.1-3 
[13] S7_0.2.1            lifecycle_1.0.5     compiler_4.5.2      farver_2.1.2       
[17] aricode_1.0.3       codetools_0.2-20    htmltools_0.5.9     yaml_2.3.12        
[21] pillar_1.11.1       uwot_0.2.4          DelayedArray_0.36.0 dbscan_1.2.4       
[25] magick_2.9.0        abind_1.4-8         mclust_6.1.2        tidyselect_1.2.1   
[29] digest_0.6.39       dplyr_1.2.0         rprojroot_2.1.1     fastmap_1.2.0      
[33] grid_4.5.2          cli_3.6.5           SparseArray_1.10.8  magrittr_2.0.4     
[37] leidenAlg_1.1.5     S4Arrays_1.10.1     dichromat_2.0-0.1   scales_1.4.0       
[41] rmarkdown_2.30      XVector_0.50.0      igraph_2.2.1        otel_0.2.0         
[45] beachmat_2.26.0     evaluate_1.0.5      RcppHungarian_0.3   knitr_1.51         
[49] irlba_2.3.7         rlang_1.1.7.9000    Rcpp_1.1.1          glue_1.8.0         
[53] rstudioapi_0.18.0   R6_2.6.1           
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-04-16
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [2] CRAN (R 4.5.0)
 aricode                1.0.3      2023-10-20 [2] CRAN (R 4.5.0)
 Banksy               * 1.6.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 beachmat               2.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 Biobase              * 2.70.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5      2025-04-23 [2] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [2] CRAN (R 4.5.2)
 colorout             * 1.3-3      2026-01-31 [2] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.2.1   2026-01-27 [2] CRAN (R 4.5.2)
 dbscan                 1.2.4      2025-12-19 [2] CRAN (R 4.5.2)
 DelayedArray           0.36.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [2] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [2] CRAN (R 4.5.2)
 dplyr                  1.2.0      2026-02-03 [2] CRAN (R 4.5.2)
 evaluate               1.0.5      2025-08-27 [2] CRAN (R 4.5.0)
 farver                 2.1.2      2024-05-13 [2] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [2] CRAN (R 4.5.0)
 generics             * 0.1.4      2025-05-09 [2] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 ggplot2                4.0.2      2026-02-03 [2] CRAN (R 4.5.2)
 glue                   1.8.0      2024-09-30 [2] CRAN (R 4.5.0)
 gtable                 0.3.6      2024-10-25 [2] CRAN (R 4.5.0)
 here                 * 1.0.2      2025-09-15 [2] CRAN (R 4.5.0)
 htmltools              0.5.9      2025-12-04 [2] CRAN (R 4.5.2)
 igraph                 2.2.1      2025-10-27 [2] CRAN (R 4.5.0)
 IRanges              * 2.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [2] CRAN (R 4.5.2)
 knitr                  1.51       2025-12-20 [2] CRAN (R 4.5.2)
 lattice                0.22-7     2025-04-02 [2] CRAN (R 4.5.2)
 leidenAlg              1.1.5      2025-04-19 [2] CRAN (R 4.5.0)
 lifecycle              1.0.5      2026-01-08 [2] CRAN (R 4.5.2)
 magick                 2.9.0      2025-09-08 [2] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [2] CRAN (R 4.5.0)
 Matrix                 1.7-4      2025-08-28 [2] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [2] CRAN (R 4.5.0)
 mclust                 6.1.2      2025-10-31 [2] CRAN (R 4.5.0)
 otel                   0.2.0      2025-08-29 [2] CRAN (R 4.5.0)
 parallelly           * 1.46.1     2026-01-08 [2] CRAN (R 4.5.2)
 pillar                 1.11.1     2025-09-17 [2] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [2] CRAN (R 4.5.0)
 R6                     2.6.1      2025-02-15 [2] CRAN (R 4.5.0)
 RColorBrewer           1.1-3      2022-04-03 [2] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [2] CRAN (R 4.5.2)
 RcppHNSW             * 0.6.0      2024-02-04 [2] CRAN (R 4.5.0)
 RcppHungarian          0.3        2023-09-05 [2] CRAN (R 4.5.0)
 rjson                  0.2.23     2024-09-16 [2] CRAN (R 4.5.0)
 rlang                  1.1.7.9000 2026-02-04 [1] Github (r-lib/rlang@bae1fe1)
 rmarkdown              2.30       2025-09-28 [2] CRAN (R 4.5.0)
 rprojroot              2.1.1      2025-08-26 [2] CRAN (R 4.5.0)
 rstudioapi             0.18.0     2026-01-16 [2] CRAN (R 4.5.2)
 S4Arrays               1.10.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1      2025-11-14 [2] CRAN (R 4.5.2)
 scales                 1.4.0      2025-04-24 [2] CRAN (R 4.5.0)
 sccore                 1.0.6      2025-04-13 [2] CRAN (R 4.5.0)
 scuttle              * 1.20.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [2] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [2] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1      2026-01-11 [2] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [2] CRAN (R 4.5.0)
 uwot                   0.2.4      2025-11-10 [2] CRAN (R 4.5.0)
 vctrs                  0.7.1      2026-01-23 [2] CRAN (R 4.5.2)
 xfun                   0.56       2026-01-18 [2] CRAN (R 4.5.2)
 XVector                0.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12     2025-12-10 [2] CRAN (R 4.5.2)

 [1] /Users/bmulvey/Library/R/arm64/4.5/library
 [2] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────────────