---
title: "07-DAPIonly_M0l0.2k25kg30louv1_markers"
author: "Bernie Mulvey"
date: "04-16-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(spatialLIBD)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(bettermc)


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

filtered DAPI only spe and labels
```{r}
lcx <- readRDS("processed-data/02_QC/07b-DAPIonly_spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

bs <- fread("processed-data/04_Banksy_clusterings/06-DAPIonly_M0l0.2kgeom30louv1_with_kneighb25-50-100.txt")

bs <- DataFrame(bs,row.names=bs$rn)
colData(lcx) <- cbind(colData(lcx),bs[colnames(lcx),])

```

registwrapper
```{r}
mktabs <- mclapply(colnames(bs)[2:4],mc.cores=3,function(b){
  regs <- registration_wrapper(lcx,var_registration=b,var_sample_id = "brnum",covars = c("APOE","Sex","Age","YRI"),gene_ensembl = "ID",gene_name = "Symbol",min_ncells = 10)
  regsdt <- as.data.table(regs$enrichment)
  regsdt <- melt.data.table(regsdt,id.vars=c("gene","ensembl"))
  regsdt[,variable:=as.character(variable)]
  regsdt[,clus:=gsub(variable,pattern="^.*_(X.*)$",replacement="\\1")]
  regsdt[,stat:=gsub(variable,pattern="^(.*)_X.*$",replacement="\\1")]
  regsdt2 <- dcast(regsdt,ensembl+gene+clus~stat,value.var="value")
  return(regsdt2)
})

putativelcs <- lapply(mktabs,function(x){
  putcluss <- x[(gene=="DBH"|gene=="TH")&lessOrEqual(fdr,0.1)&greaterOrEqual(logFC,0),unique(clus)]
  y <- x[clus %in% putcluss & ((lessOrEqual(fdr,0.1) & greaterOrEqual(logFC,0))|gene %in% c("DBH","TH"))]
  return(y)
})


## get top 5 markers and DBH TH stats per cluster enriched for DBH andor TH:
putativelcstop <- lapply(putativelcs,function(x){
  topsandne <- rbind(
  x[gene %in% c("DBH","TH"),.(clus,gene,round(logFC,2),round(fdr,2))],
  x[,head(.SD[order(-logFC)],5),by="clus"][,.(clus,gene,round(logFC,2),round(fdr,2))])
  topsandne <- as.data.table(topsandne)
  setorderv(topsandne,"clus")
  topsandne <- unique(topsandne)
  setnames(topsandne,c(3,4),c("logFC","fdr"))
  return(topsandne)
})

## peep the ones for exact parameter set we've been tracking,  "clust_M0_lam0.2_k25_kgeom30_res1"

putativelcstop[["clust_M0_lam0.2_k25_kgeom30_res1"]]

```
Hmm...the same GENERAL pattern is still here (all clusters but one being off-target with secondary DBH/TH enrichment) when clustering DAPI-only cells (compare to the table of top markers + DBH + TH for NMDAPI nuclei, code 04/04). BUT the DBH/TH logFC is lower both in the one like pure LC cluster, as well as in the secondarily enriched clusters (more on the order of logFC 2-3 in off-target clusters in the NMDAPI nuclear counts).

So, we go one level further down next...DAPI only nuclear counts only.
      clus    gene logFC   fdr
    <char>  <char> <num> <num>
 1:    X15     DBH  0.77  0.03
 2:    X15      TH  0.31  0.28
 3:    X15    CCL5  8.51  0.00
 4:    X15    NKG7  8.18  0.00
 5:    X15    IL7R  7.64  0.00
 6:    X15    CD3G  7.32  0.00
 7:    X15    CCL4  7.12  0.00
 8:     X2     DBH  1.27  0.00
 9:     X2      TH  0.70  0.02
10:     X2    FLT1  6.22  0.00
11:     X2    KLF4  5.86  0.00
12:     X2  APOLD1  5.83  0.00
13:     X2    KLF2  5.76  0.00
14:     X2  PECAM1  5.53  0.00
15:    X24     DBH  3.16  0.00
16:    X24      TH  2.12  0.00
17:    X24   FGFR3  3.17  0.00
18:    X24   SPON1  2.94  0.00
19:    X24    MT1X  2.92  0.00
20:    X24    RGMA  2.87  0.00
21:    X25     DBH  7.16  0.00
22:    X25      TH  5.82  0.00
23:    X25  SLC6A2  6.61  0.00
24:    X25 SLC18A2  6.59  0.00
25:    X25     DDC  5.46  0.00
26:     X3     DBH  0.18  0.68
27:     X3      TH  0.91  0.00
28:     X3  SLC6A3  2.01  0.00
29:     X3   TSHZ2  1.93  0.00
30:     X3   CNDP1  1.88  0.00
31:     X3    ROS1  1.81  0.00
32:     X3    KLK6  1.80  0.00
33:     X5     DBH  1.21  0.00
34:     X5      TH  0.67  0.03
35:     X5  ATP2C2  5.65  0.00
36:     X5  PDGFRA  5.58  0.00
37:     X5   OLIG2  4.90  0.00
38:     X5    VCAN  4.82  0.00
39:     X5   CSPG4  4.51  0.00
      clus    gene logFC   fdr
    <char>  <char> <num> <num>

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] bettermc_1.2.2.9000         BiocParallel_1.44.0        
 [3] parallelly_1.46.1           colorout_1.3-3             
 [5] spatialLIBD_1.22.0          SpatialExperiment_1.20.0   
 [7] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
 [9] Biobase_2.70.0              GenomicRanges_1.62.1       
[11] Seqinfo_1.0.0               IRanges_2.44.0             
[13] S4Vectors_0.48.0            BiocGenerics_0.56.0        
[15] generics_0.1.4              MatrixGenerics_1.22.0      
[17] matrixStats_1.5.0           data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] later_1.4.5              BiocIO_1.20.0            bitops_1.0-9            
  [4] filelock_1.0.3           tibble_3.3.1             XML_3.99-0.20           
  [7] lifecycle_1.0.5          httr2_1.2.2              edgeR_4.8.2             
 [10] doParallel_1.0.17        rprojroot_2.1.1          lattice_0.22-7          
 [13] backports_1.5.0          magrittr_2.0.4           limma_3.66.0            
 [16] plotly_4.12.0            sass_0.4.10              rmarkdown_2.30          
 [19] jquerylib_0.1.4          yaml_2.3.12              httpuv_1.6.16           
 [22] otel_0.2.0               sessioninfo_1.2.3        cowplot_1.2.0           
 [25] DBI_1.2.3                RColorBrewer_1.1-3       golem_0.5.1             
 [28] abind_1.4-8              purrr_1.2.1              RCurl_1.98-1.17         
 [31] rappdirs_0.3.4           circlize_0.4.17          ggrepel_0.9.6           
 [34] irlba_2.3.7              codetools_0.2-20         DelayedArray_0.36.0     
 [37] DT_0.34.0                scuttle_1.20.0           tidyselect_1.2.1        
 [40] shape_1.4.6.1            farver_2.1.2             ScaledMatrix_1.18.0     
 [43] viridis_0.6.5            shinyWidgets_0.9.0       BiocFileCache_3.0.0     
 [46] GenomicAlignments_1.46.0 jsonlite_2.0.0           GetoptLong_1.1.0        
 [49] BiocNeighbors_2.4.0      scater_1.38.0            iterators_1.0.14        
 [52] foreach_1.5.2            tools_4.5.2              progress_1.2.3          
 [55] Rcpp_1.1.1               glue_1.8.0               gridExtra_2.3           
 [58] SparseArray_1.10.8       xfun_0.56                here_1.0.2              
 [61] dplyr_1.2.0              BiocManager_1.30.27      fastmap_1.2.0           
 [64] digest_0.6.39            rsvd_1.0.5               R6_2.6.1                
 [67] mime_0.13                colorspace_2.1-2         dichromat_2.0-0.1       
 [70] RSQLite_2.4.5            cigarillo_1.0.0          config_0.3.2            
 [73] tidyr_1.3.2              rtracklayer_1.70.1       prettyunits_1.2.0       
 [76] httr_1.4.7               htmlwidgets_1.6.4        S4Arrays_1.10.1         
 [79] pkgconfig_2.0.3          gtable_0.3.6             blob_1.3.0              
 [82] ComplexHeatmap_2.26.1    S7_0.2.1                 XVector_0.50.0          
 [85] htmltools_0.5.9          clue_0.3-66              scales_1.4.0            
 [88] png_0.1-8                attempt_0.3.1            knitr_1.51              
 [91] rstudioapi_0.18.0        rjson_0.2.23             checkmate_2.3.4         
 [94] curl_7.0.0               cachem_1.1.0             GlobalOptions_0.1.3     
 [97] BiocVersion_3.22.0       parallel_4.5.2           vipor_0.4.7             
[100] AnnotationDbi_1.72.0     restfulr_0.0.16          pillar_1.11.1           
[103] grid_4.5.2               vctrs_0.7.1              promises_1.5.0          
[106] BiocSingular_1.26.0      dbplyr_2.5.1             beachmat_2.26.0         
[109] xtable_1.8-4             cluster_2.1.8.1          beeswarm_0.4.0          
[112] paletteer_1.7.0          evaluate_1.0.5           magick_2.9.0            
[115] cli_3.6.5                locfit_1.5-9.12          compiler_4.5.2          
[118] Rsamtools_2.26.0         rlang_1.1.7.9000         crayon_1.5.3            
[121] rematch2_2.1.2           ggbeeswarm_0.7.3         viridisLite_0.4.3       
[124] Biostrings_2.78.0        lazyeval_0.2.2           Matrix_1.7-4            
[127] ExperimentHub_3.0.0      benchmarkme_1.0.8        hms_1.1.4               
[130] bit64_4.6.0-1            ggplot2_4.0.2            KEGGREST_1.50.0         
[133] statmod_1.5.1            shiny_1.12.1             AnnotationHub_4.0.0     
[136] memoise_2.0.1            bslib_0.10.0             benchmarkmeData_2.0.0   
[139] bit_4.6.0               
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-04-16
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [2] CRAN (R 4.5.0)
 AnnotationDbi          1.72.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 AnnotationHub          4.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 attempt                0.3.1      2020-05-03 [2] CRAN (R 4.5.0)
 backports              1.5.0      2024-05-23 [2] CRAN (R 4.5.0)
 beachmat               2.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0      2021-06-01 [2] CRAN (R 4.5.0)
 benchmarkme            1.0.8      2022-06-12 [2] CRAN (R 4.5.0)
 benchmarkmeData        2.0.0      2026-01-19 [2] CRAN (R 4.5.2)
 bettermc             * 1.2.2.9000 2026-03-16 [2] Github (akersting/bettermc@d6ec1aa)
 Biobase              * 2.70.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocFileCache          3.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocIO                 1.20.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocManager            1.30.27    2025-11-14 [2] CRAN (R 4.5.2)
 BiocNeighbors          2.4.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocVersion            3.22.0     2025-10-07 [2] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0      2025-03-06 [2] CRAN (R 4.5.0)
 bit64                  4.6.0-1    2025-01-16 [2] CRAN (R 4.5.0)
 bitops                 1.0-9      2024-10-03 [2] CRAN (R 4.5.0)
 blob                   1.3.0      2026-01-14 [2] CRAN (R 4.5.2)
 bslib                  0.10.0     2026-01-26 [2] CRAN (R 4.5.2)
 cachem                 1.1.0      2024-05-16 [2] CRAN (R 4.5.0)
 checkmate              2.3.4      2026-02-03 [2] CRAN (R 4.5.2)
 cigarillo              1.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 circlize               0.4.17     2025-12-08 [2] CRAN (R 4.5.2)
 cli                    3.6.5      2025-04-23 [2] CRAN (R 4.5.0)
 clue                   0.3-66     2024-11-13 [2] CRAN (R 4.5.0)
 cluster                2.1.8.1    2025-03-12 [2] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [2] CRAN (R 4.5.2)
 colorout             * 1.3-3      2026-01-31 [2] Github (jalvesaq/colorout@64863bb)
 colorspace             2.1-2      2025-09-22 [2] CRAN (R 4.5.0)
 ComplexHeatmap         2.26.1     2026-01-30 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 config                 0.3.2      2023-08-30 [2] CRAN (R 4.5.0)
 cowplot                1.2.0      2025-07-07 [2] CRAN (R 4.5.0)
 crayon                 1.5.3      2024-06-20 [2] CRAN (R 4.5.0)
 curl                   7.0.0      2025-08-19 [2] CRAN (R 4.5.0)
 data.table           * 1.18.2.1   2026-01-27 [2] CRAN (R 4.5.2)
 DBI                    1.2.3      2024-06-02 [2] CRAN (R 4.5.0)
 dbplyr                 2.5.1      2025-09-10 [2] CRAN (R 4.5.0)
 DelayedArray           0.36.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [2] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [2] CRAN (R 4.5.2)
 doParallel             1.0.17     2022-02-07 [2] CRAN (R 4.5.0)
 dplyr                  1.2.0      2026-02-03 [2] CRAN (R 4.5.2)
 DT                     0.34.0     2025-09-02 [2] CRAN (R 4.5.0)
 edgeR                  4.8.2      2025-12-23 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate               1.0.5      2025-08-27 [2] CRAN (R 4.5.0)
 ExperimentHub          3.0.0      2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 farver                 2.1.2      2024-05-13 [2] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [2] CRAN (R 4.5.0)
 filelock               1.0.3      2023-12-11 [2] CRAN (R 4.5.0)
 foreach                1.5.2      2022-02-02 [2] CRAN (R 4.5.0)
 generics             * 0.1.4      2025-05-09 [2] CRAN (R 4.5.0)
 GenomicAlignments      1.46.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 GenomicRanges        * 1.62.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 GetoptLong             1.1.0      2025-11-28 [2] CRAN (R 4.5.2)
 ggbeeswarm             0.7.3      2025-11-29 [2] CRAN (R 4.5.2)
 ggplot2                4.0.2      2026-02-03 [2] CRAN (R 4.5.2)
 ggrepel                0.9.6      2024-09-07 [2] CRAN (R 4.5.0)
 GlobalOptions          0.1.3      2025-11-28 [2] CRAN (R 4.5.2)
 glue                   1.8.0      2024-09-30 [2] CRAN (R 4.5.0)
 golem                  0.5.1      2024-08-27 [2] CRAN (R 4.5.0)
 gridExtra              2.3        2017-09-09 [2] CRAN (R 4.5.0)
 gtable                 0.3.6      2024-10-25 [2] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [2] CRAN (R 4.5.0)
 hms                    1.1.4      2025-10-17 [2] CRAN (R 4.5.0)
 htmltools              0.5.9      2025-12-04 [2] CRAN (R 4.5.2)
 htmlwidgets            1.6.4      2023-12-06 [2] CRAN (R 4.5.0)
 httpuv                 1.6.16     2025-04-16 [2] CRAN (R 4.5.0)
 httr                   1.4.7      2023-08-15 [2] CRAN (R 4.5.0)
 httr2                  1.2.2      2025-12-08 [2] CRAN (R 4.5.2)
 IRanges              * 2.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [2] CRAN (R 4.5.2)
 iterators              1.0.14     2022-02-05 [2] CRAN (R 4.5.0)
 jquerylib              0.1.4      2021-04-26 [2] CRAN (R 4.5.0)
 jsonlite               2.0.0      2025-03-27 [2] CRAN (R 4.5.0)
 KEGGREST               1.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51       2025-12-20 [2] CRAN (R 4.5.2)
 later                  1.4.5      2026-01-08 [2] CRAN (R 4.5.2)
 lattice                0.22-7     2025-04-02 [2] CRAN (R 4.5.2)
 lazyeval               0.2.2      2019-03-15 [2] CRAN (R 4.5.0)
 lifecycle              1.0.5      2026-01-08 [2] CRAN (R 4.5.2)
 limma                  3.66.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 locfit                 1.5-9.12   2025-03-05 [2] CRAN (R 4.5.0)
 magick                 2.9.0      2025-09-08 [2] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [2] CRAN (R 4.5.0)
 Matrix                 1.7-4      2025-08-28 [2] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [2] CRAN (R 4.5.0)
 memoise                2.0.1      2021-11-26 [2] CRAN (R 4.5.0)
 mime                   0.13       2025-03-17 [2] CRAN (R 4.5.0)
 otel                   0.2.0      2025-08-29 [2] CRAN (R 4.5.0)
 paletteer              1.7.0      2026-01-08 [2] CRAN (R 4.5.2)
 parallelly           * 1.46.1     2026-01-08 [2] CRAN (R 4.5.2)
 pillar                 1.11.1     2025-09-17 [2] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [2] CRAN (R 4.5.0)
 plotly                 4.12.0     2026-01-24 [2] CRAN (R 4.5.2)
 png                    0.1-8      2022-11-29 [2] CRAN (R 4.5.0)
 prettyunits            1.2.0      2023-09-24 [2] CRAN (R 4.5.0)
 progress               1.2.3      2023-12-06 [2] CRAN (R 4.5.0)
 promises               1.5.0      2025-11-01 [2] CRAN (R 4.5.0)
 purrr                  1.2.1      2026-01-09 [2] CRAN (R 4.5.2)
 R6                     2.6.1      2025-02-15 [2] CRAN (R 4.5.0)
 rappdirs               0.3.4      2026-01-17 [2] CRAN (R 4.5.2)
 RColorBrewer           1.1-3      2022-04-03 [2] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [2] CRAN (R 4.5.2)
 RCurl                  1.98-1.17  2025-03-22 [2] CRAN (R 4.5.0)
 rematch2               2.1.2      2020-05-01 [2] CRAN (R 4.5.0)
 restfulr               0.0.16     2025-06-27 [2] CRAN (R 4.5.0)
 rjson                  0.2.23     2024-09-16 [2] CRAN (R 4.5.0)
 rlang                  1.1.7.9000 2026-02-04 [1] Github (r-lib/rlang@bae1fe1)
 rmarkdown              2.30       2025-09-28 [2] CRAN (R 4.5.0)
 rprojroot              2.1.1      2025-08-26 [2] CRAN (R 4.5.0)
 Rsamtools              2.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 RSQLite                2.4.5      2025-11-30 [2] CRAN (R 4.5.2)
 rstudioapi             0.18.0     2026-01-16 [2] CRAN (R 4.5.2)
 rsvd                   1.0.5      2021-04-16 [2] CRAN (R 4.5.0)
 rtracklayer            1.70.1     2025-12-20 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 S4Arrays               1.10.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1      2025-11-14 [2] CRAN (R 4.5.2)
 sass                   0.4.10     2025-04-11 [2] CRAN (R 4.5.0)
 ScaledMatrix           1.18.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0      2025-04-24 [2] CRAN (R 4.5.0)
 scater                 1.38.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 scuttle                1.20.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [2] CRAN (R 4.5.0)
 shape                  1.4.6.1    2024-02-23 [2] CRAN (R 4.5.0)
 shiny                  1.12.1     2025-12-09 [2] CRAN (R 4.5.2)
 shinyWidgets           0.9.0      2025-02-21 [2] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [2] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 spatialLIBD          * 1.22.0     2025-11-04 [2] Bioconductor 3.22 (R 4.5.2)
 statmod                1.5.1      2025-10-09 [2] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1      2026-01-11 [2] CRAN (R 4.5.2)
 tidyr                  1.3.2      2025-12-19 [2] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [2] CRAN (R 4.5.0)
 vctrs                  0.7.1      2026-01-23 [2] CRAN (R 4.5.2)
 vipor                  0.4.7      2023-12-18 [2] CRAN (R 4.5.0)
 viridis                0.6.5      2024-01-29 [2] CRAN (R 4.5.0)
 viridisLite            0.4.3      2026-02-04 [2] CRAN (R 4.5.2)
 xfun                   0.56       2026-01-18 [2] CRAN (R 4.5.2)
 XML                    3.99-0.20  2025-11-08 [2] CRAN (R 4.5.0)
 xtable                 1.8-4      2019-04-21 [2] CRAN (R 4.5.0)
 XVector                0.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12     2025-12-10 [2] CRAN (R 4.5.2)

 [1] /Users/bmulvey/Library/R/arm64/4.5/library
 [2] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────────────