---
title: "08-Examine_DAPIonly_multiLC_cluster_in_DAPIonlyNucOnlySPE"
author: "Bernie Mulvey"
date: "04-16-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(spatialLIBD)
# library(SpatialFeatureExperiment)


## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}


## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)
library(bettermc) # for troubleshooting mclapply/better functionality

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

tldr: many of the banksy clusterings give >>1 DBH-enriched cluster; inspection reveals that usually there's 1 super-strong DBH enriched cluster and then the others are enriched for other cell type markers. this remains the case when using NMDAPI "nuclear" counts. this further remains (albeit somewhat weaker DBH enrichment) even with DAPI-only cells. so, does this go all the way down to the transcripts within the DAPI itself? (i.e., DAPI nuclear transcripts only)? here, we have the nuclei for the DAPI-only cells previously clustered and can run the same marker analyses to see whether the multi-type contamination is any better. 

a handy little shortcut we can use, thanks to nuclei and cells sharing identifiers, is taking the banksy results we got from cell level clustering and assigning those same labels to nucleus-counts-only data. we can then check our working example banksy analysis, label the nuclei the same way we did the cells, and run the same marker analysis again.

preliminary poked around in a mid-granularity result that gave 5 DBH clusters, clust_M0_lam0.2_k25_kgeom30_louv1 , 4 of which were enriched for other markers besides LCNE. in the DAPIonly cells, there were 6 clusters total, 5 with weird not LCNE cells.
```{r}
nucspe <- readRDS("processed-data/01_spe/05b_DAPIonly_nuclearCounts_SPE_retainedDAPIcellsOnly.RDS")

## banksy results from the parameter set we're following, performed on DAPI-only cell level data:

testbanks <- fread("processed-data/04_Banksy_clusterings/06-DAPIonly_M0l0.2kgeom30louv1_with_kneighb25-50-100.txt")

## we varied k neighbors there, just because I was curious, but we want the 25 neighbors for comparability here.
testbanks <- DataFrame(testbanks,row.names=testbanks$rn)[colnames(nucspe),]
colData(nucspe) <- cbind(colData(nucspe),testbanks[,c(2:4)])

## run registration wrapper and loook at the marker enrichment stats for these clusters now that we have "nucleus" only counts
## first we need something in rowdata though
newrd <- cbind(rownames(nucspe),rownames(nucspe))
## spatialLIBD is going to want an ensembl column but we'll just make it a second column of the symbols
colnames(newrd) <- c("gene_name","gene_id")
rownames(newrd) <- newrd[,1]
newrd <- newrd[rownames(nucspe),]
rowData(nucspe) <- newrd
```

## marker analyses
```{r}
mktabs <- mclapply(colnames(testbanks)[2:4],mc.cores=3,function(b){
  regs <- registration_wrapper(nucspe,var_registration=b,var_sample_id = "brnum",covars = c("APOE","Sex","Age","YRI"),gene_ensembl = "gene_id",gene_name = "gene_name",min_ncells = 10)
  regsdt <- as.data.table(regs$enrichment)
  regsdt <- melt.data.table(regsdt,id.vars=c("gene","ensembl"))
  regsdt[,variable:=as.character(variable)]
  regsdt[,clus:=gsub(variable,pattern="^.*_(X.*)$",replacement="\\1")]
  regsdt[,stat:=gsub(variable,pattern="^(.*)_X.*$",replacement="\\1")]
  regsdt2 <- dcast(regsdt,ensembl+gene+clus~stat,value.var="value")
  return(regsdt2)
})

putativelcs <- lapply(mktabs,function(x){
  putcluss <- x[(gene=="DBH"|gene=="TH")&lessOrEqual(fdr,0.1)&greaterOrEqual(logFC,0),unique(clus)]
  y <- x[clus %in% putcluss & ((lessOrEqual(fdr,0.1) & greaterOrEqual(logFC,0))|gene %in% c("DBH","TH"))]
  return(y)
})

putativelcstop <- lapply(putativelcs,function(x){
  topsandne <- rbind(
  x[gene %in% c("DBH","TH"),.(clus,gene,round(logFC,2),round(fdr,2))],
  x[,head(.SD[order(-logFC)],5),by="clus"][,.(clus,gene,round(logFC,2),round(fdr,2))])
  topsandne <- as.data.table(topsandne)
  setorderv(topsandne,"clus")
  topsandne <- unique(topsandne)
  setnames(topsandne,c(3,4),c("logFC","fdr"))
  return(topsandne)
})

## peep the ones for exact parameter set we've been tracking,  "clust_M0_lam0.2_k25_kgeom30_res1"

putativelcstop[["clust_M0_lam0.2_k25_kgeom30_res1"]]
```

Well slap my ass and call me Sally, this is practically UNCHANGED relative to DAPI with cell expansion! All 6 are enriched for DBH at FDR < 0.1:
      clus    gene logFC   fdr
    <char>  <char> <num> <num>
 1:     X1     DBH  0.97  0.01
 2:     X1      TH  0.85  0.01
 3:     X1  FCGR1A  6.55  0.00
 4:     X1  P2RY12  6.37  0.00
 5:     X1   ITGAX  6.30  0.00
 6:     X1   GPR34  6.28  0.00
 7:     X1  P2RY13  6.13  0.00
 8:    X10     DBH  0.38  0.38
 9:    X10      TH  0.59  0.07
10:    X10     MOG  5.86  0.00
11:    X10     MAG  5.75  0.00
12:    X10    MYRF  5.61  0.00
13:    X10    ST18  5.59  0.00
14:    X10   ERBB3  5.52  0.00
15:     X2     DBH  0.91  0.02
16:     X2      TH  0.51  0.12
17:     X2    FLT1  7.77  0.00
18:     X2  APOLD1  7.24  0.00
19:     X2    KLF2  7.09  0.00
20:     X2    KLF4  6.87  0.00
21:     X2  PECAM1  6.75  0.00
22:    X24     DBH  2.40  0.00
23:    X24      TH  1.18  0.00
24:    X24    MT1X  3.83  0.00
25:    X24   FGFR3  3.79  0.00
26:    X24   NTSR2  3.57  0.00
27:    X24   MGST1  3.57  0.00
28:    X24   SPON1  3.54  0.00
29:    X25     DBH  7.36  0.00
30:    X25      TH  5.64  0.00
31:    X25  SLC6A2  6.62  0.00
32:    X25 SLC18A2  6.60  0.00
33:    X25     DDC  5.39  0.00
34:     X3     DBH  0.95  0.01
35:     X3      TH  1.64  0.00
36:     X3   CNDP1  2.55  0.00
37:     X3  SLC6A3  2.50  0.00
38:     X3    KLK6  2.37  0.00
39:     X3   TSHZ2  2.29  0.00
40:     X3     MAG  2.29  0.00
41:     X5     DBH  0.83  0.04
42:     X5      TH  0.37  0.29
43:     X5  PDGFRA  7.17  0.00
44:     X5  ATP2C2  6.79  0.00
45:     X5    VCAN  5.78  0.00
46:     X5   OLIG2  5.76  0.00
47:     X5   CSPG4  5.08  0.00
      clus    gene logFC   fdr
    <char>  <char> <num> <num>

So, is this because of cells being "stacked" over nuclei in the sections? And/or is this just the so-called phenomenon of "transcript diffusion"? And most importantly, are these phenomenon (impactfully) more marked when using the NM-DAPI joint segmentations (whether its to produce "nuclei" and nuclear counts or those "nuclei" further expanded to cells)?

That'll be the question for discussion with SCP, MT, and probably others.

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] bettermc_1.2.2.9000         BiocParallel_1.44.0        
 [3] parallelly_1.46.1           colorout_1.3-3             
 [5] spatialLIBD_1.22.0          SpatialExperiment_1.20.0   
 [7] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
 [9] Biobase_2.70.0              GenomicRanges_1.62.1       
[11] Seqinfo_1.0.0               IRanges_2.44.0             
[13] S4Vectors_0.48.0            BiocGenerics_0.56.0        
[15] generics_0.1.4              MatrixGenerics_1.22.0      
[17] matrixStats_1.5.0           data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3       rstudioapi_0.18.0        jsonlite_2.0.0          
  [4] shape_1.4.6.1            magrittr_2.0.4           ggbeeswarm_0.7.3        
  [7] magick_2.9.0             farver_2.1.2             rmarkdown_2.30          
 [10] GlobalOptions_0.1.3      BiocIO_1.20.0            vctrs_0.7.1             
 [13] memoise_2.0.1            config_0.3.2             Rsamtools_2.26.0        
 [16] paletteer_1.7.0          RCurl_1.98-1.17          benchmarkme_1.0.8       
 [19] htmltools_0.5.9          S4Arrays_1.10.1          AnnotationHub_4.0.0     
 [22] curl_7.0.0               BiocNeighbors_2.4.0      SparseArray_1.10.8      
 [25] sass_0.4.10              bslib_0.10.0             htmlwidgets_1.6.4       
 [28] httr2_1.2.2              plotly_4.12.0            cachem_1.1.0            
 [31] GenomicAlignments_1.46.0 mime_0.13                lifecycle_1.0.5         
 [34] iterators_1.0.14         pkgconfig_2.0.3          rsvd_1.0.5              
 [37] Matrix_1.7-4             R6_2.6.1                 fastmap_1.2.0           
 [40] shiny_1.12.1             clue_0.3-66              digest_0.6.39           
 [43] colorspace_2.1-2         rematch2_2.1.2           AnnotationDbi_1.72.0    
 [46] scater_1.38.0            rprojroot_2.1.1          irlba_2.3.7             
 [49] ExperimentHub_3.0.0      RSQLite_2.4.5            beachmat_2.26.0         
 [52] filelock_1.0.3           httr_1.4.7               abind_1.4-8             
 [55] compiler_4.5.2           here_1.0.2               bit64_4.6.0-1           
 [58] doParallel_1.0.17        attempt_0.3.1            S7_0.2.1                
 [61] viridis_0.6.5            DBI_1.2.3                sessioninfo_1.2.3       
 [64] rappdirs_0.3.4           DelayedArray_0.36.0      rjson_0.2.23            
 [67] tools_4.5.2              vipor_0.4.7              otel_0.2.0              
 [70] beeswarm_0.4.0           httpuv_1.6.16            glue_1.8.0              
 [73] restfulr_0.0.16          promises_1.5.0           grid_4.5.2              
 [76] cluster_2.1.8.1          gtable_0.3.6             tidyr_1.3.2             
 [79] ScaledMatrix_1.18.0      BiocSingular_1.26.0      XVector_0.50.0          
 [82] ggrepel_0.9.6            BiocVersion_3.22.0       foreach_1.5.2           
 [85] pillar_1.11.1            limma_3.66.0             later_1.4.5             
 [88] circlize_0.4.17          benchmarkmeData_2.0.0    dplyr_1.2.0             
 [91] BiocFileCache_3.0.0      lattice_0.22-7           rtracklayer_1.70.1      
 [94] bit_4.6.0                tidyselect_1.2.1         ComplexHeatmap_2.26.1   
 [97] locfit_1.5-9.12          scuttle_1.20.0           Biostrings_2.78.0       
[100] knitr_1.51               gridExtra_2.3            edgeR_4.8.2             
[103] xfun_0.56                statmod_1.5.1            DT_0.34.0               
[106] lazyeval_0.2.2           yaml_2.3.12              shinyWidgets_0.9.0      
[109] cigarillo_1.0.0          evaluate_1.0.5           codetools_0.2-20        
[112] tibble_3.3.1             BiocManager_1.30.27      cli_3.6.5               
[115] xtable_1.8-4             jquerylib_0.1.4          golem_0.5.1             
[118] dichromat_2.0-0.1        Rcpp_1.1.1               dbplyr_2.5.1            
[121] png_0.1-8                XML_3.99-0.20            ggplot2_4.0.2           
[124] blob_1.3.0               bitops_1.0-9             viridisLite_0.4.3       
[127] scales_1.4.0             purrr_1.2.1              crayon_1.5.3            
[130] GetoptLong_1.1.0         rlang_1.1.7.9000         cowplot_1.2.0           
[133] KEGGREST_1.50.0         
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-04-15
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [2] CRAN (R 4.5.0)
 AnnotationDbi          1.72.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 AnnotationHub          4.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 attempt                0.3.1      2020-05-03 [2] CRAN (R 4.5.0)
 beachmat               2.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0      2021-06-01 [2] CRAN (R 4.5.0)
 benchmarkme            1.0.8      2022-06-12 [2] CRAN (R 4.5.0)
 benchmarkmeData        2.0.0      2026-01-19 [2] CRAN (R 4.5.2)
 bettermc             * 1.2.2.9000 2026-03-16 [2] Github (akersting/bettermc@d6ec1aa)
 Biobase              * 2.70.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocFileCache          3.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocIO                 1.20.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocManager            1.30.27    2025-11-14 [2] CRAN (R 4.5.2)
 BiocNeighbors          2.4.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocVersion            3.22.0     2025-10-07 [2] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0      2025-03-06 [2] CRAN (R 4.5.0)
 bit64                  4.6.0-1    2025-01-16 [2] CRAN (R 4.5.0)
 bitops                 1.0-9      2024-10-03 [2] CRAN (R 4.5.0)
 blob                   1.3.0      2026-01-14 [2] CRAN (R 4.5.2)
 bslib                  0.10.0     2026-01-26 [2] CRAN (R 4.5.2)
 cachem                 1.1.0      2024-05-16 [2] CRAN (R 4.5.0)
 cigarillo              1.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 circlize               0.4.17     2025-12-08 [2] CRAN (R 4.5.2)
 cli                    3.6.5      2025-04-23 [2] CRAN (R 4.5.0)
 clue                   0.3-66     2024-11-13 [2] CRAN (R 4.5.0)
 cluster                2.1.8.1    2025-03-12 [2] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [2] CRAN (R 4.5.2)
 colorout             * 1.3-3      2026-01-31 [2] Github (jalvesaq/colorout@64863bb)
 colorspace             2.1-2      2025-09-22 [2] CRAN (R 4.5.0)
 ComplexHeatmap         2.26.1     2026-01-30 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 config                 0.3.2      2023-08-30 [2] CRAN (R 4.5.0)
 cowplot                1.2.0      2025-07-07 [2] CRAN (R 4.5.0)
 crayon                 1.5.3      2024-06-20 [2] CRAN (R 4.5.0)
 curl                   7.0.0      2025-08-19 [2] CRAN (R 4.5.0)
 data.table           * 1.18.2.1   2026-01-27 [2] CRAN (R 4.5.2)
 DBI                    1.2.3      2024-06-02 [2] CRAN (R 4.5.0)
 dbplyr                 2.5.1      2025-09-10 [2] CRAN (R 4.5.0)
 DelayedArray           0.36.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [2] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [2] CRAN (R 4.5.2)
 doParallel             1.0.17     2022-02-07 [2] CRAN (R 4.5.0)
 dplyr                  1.2.0      2026-02-03 [2] CRAN (R 4.5.2)
 DT                     0.34.0     2025-09-02 [2] CRAN (R 4.5.0)
 edgeR                  4.8.2      2025-12-23 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate               1.0.5      2025-08-27 [2] CRAN (R 4.5.0)
 ExperimentHub          3.0.0      2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 farver                 2.1.2      2024-05-13 [2] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [2] CRAN (R 4.5.0)
 filelock               1.0.3      2023-12-11 [2] CRAN (R 4.5.0)
 foreach                1.5.2      2022-02-02 [2] CRAN (R 4.5.0)
 generics             * 0.1.4      2025-05-09 [2] CRAN (R 4.5.0)
 GenomicAlignments      1.46.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 GenomicRanges        * 1.62.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 GetoptLong             1.1.0      2025-11-28 [2] CRAN (R 4.5.2)
 ggbeeswarm             0.7.3      2025-11-29 [2] CRAN (R 4.5.2)
 ggplot2                4.0.2      2026-02-03 [2] CRAN (R 4.5.2)
 ggrepel                0.9.6      2024-09-07 [2] CRAN (R 4.5.0)
 GlobalOptions          0.1.3      2025-11-28 [2] CRAN (R 4.5.2)
 glue                   1.8.0      2024-09-30 [2] CRAN (R 4.5.0)
 golem                  0.5.1      2024-08-27 [2] CRAN (R 4.5.0)
 gridExtra              2.3        2017-09-09 [2] CRAN (R 4.5.0)
 gtable                 0.3.6      2024-10-25 [2] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [2] CRAN (R 4.5.0)
 htmltools              0.5.9      2025-12-04 [2] CRAN (R 4.5.2)
 htmlwidgets            1.6.4      2023-12-06 [2] CRAN (R 4.5.0)
 httpuv                 1.6.16     2025-04-16 [2] CRAN (R 4.5.0)
 httr                   1.4.7      2023-08-15 [2] CRAN (R 4.5.0)
 httr2                  1.2.2      2025-12-08 [2] CRAN (R 4.5.2)
 IRanges              * 2.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [2] CRAN (R 4.5.2)
 iterators              1.0.14     2022-02-05 [2] CRAN (R 4.5.0)
 jquerylib              0.1.4      2021-04-26 [2] CRAN (R 4.5.0)
 jsonlite               2.0.0      2025-03-27 [2] CRAN (R 4.5.0)
 KEGGREST               1.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51       2025-12-20 [2] CRAN (R 4.5.2)
 later                  1.4.5      2026-01-08 [2] CRAN (R 4.5.2)
 lattice                0.22-7     2025-04-02 [2] CRAN (R 4.5.2)
 lazyeval               0.2.2      2019-03-15 [2] CRAN (R 4.5.0)
 lifecycle              1.0.5      2026-01-08 [2] CRAN (R 4.5.2)
 limma                  3.66.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 locfit                 1.5-9.12   2025-03-05 [2] CRAN (R 4.5.0)
 magick                 2.9.0      2025-09-08 [2] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [2] CRAN (R 4.5.0)
 Matrix                 1.7-4      2025-08-28 [2] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [2] CRAN (R 4.5.0)
 memoise                2.0.1      2021-11-26 [2] CRAN (R 4.5.0)
 mime                   0.13       2025-03-17 [2] CRAN (R 4.5.0)
 otel                   0.2.0      2025-08-29 [2] CRAN (R 4.5.0)
 paletteer              1.7.0      2026-01-08 [2] CRAN (R 4.5.2)
 parallelly           * 1.46.1     2026-01-08 [2] CRAN (R 4.5.2)
 pillar                 1.11.1     2025-09-17 [2] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [2] CRAN (R 4.5.0)
 plotly                 4.12.0     2026-01-24 [2] CRAN (R 4.5.2)
 png                    0.1-8      2022-11-29 [2] CRAN (R 4.5.0)
 promises               1.5.0      2025-11-01 [2] CRAN (R 4.5.0)
 purrr                  1.2.1      2026-01-09 [2] CRAN (R 4.5.2)
 R6                     2.6.1      2025-02-15 [2] CRAN (R 4.5.0)
 rappdirs               0.3.4      2026-01-17 [2] CRAN (R 4.5.2)
 RColorBrewer           1.1-3      2022-04-03 [2] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [2] CRAN (R 4.5.2)
 RCurl                  1.98-1.17  2025-03-22 [2] CRAN (R 4.5.0)
 rematch2               2.1.2      2020-05-01 [2] CRAN (R 4.5.0)
 restfulr               0.0.16     2025-06-27 [2] CRAN (R 4.5.0)
 rjson                  0.2.23     2024-09-16 [2] CRAN (R 4.5.0)
 rlang                  1.1.7.9000 2026-02-04 [1] Github (r-lib/rlang@bae1fe1)
 rmarkdown              2.30       2025-09-28 [2] CRAN (R 4.5.0)
 rprojroot              2.1.1      2025-08-26 [2] CRAN (R 4.5.0)
 Rsamtools              2.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 RSQLite                2.4.5      2025-11-30 [2] CRAN (R 4.5.2)
 rstudioapi             0.18.0     2026-01-16 [2] CRAN (R 4.5.2)
 rsvd                   1.0.5      2021-04-16 [2] CRAN (R 4.5.0)
 rtracklayer            1.70.1     2025-12-20 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 S4Arrays               1.10.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1      2025-11-14 [2] CRAN (R 4.5.2)
 sass                   0.4.10     2025-04-11 [2] CRAN (R 4.5.0)
 ScaledMatrix           1.18.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0      2025-04-24 [2] CRAN (R 4.5.0)
 scater                 1.38.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 scuttle                1.20.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [2] CRAN (R 4.5.0)
 shape                  1.4.6.1    2024-02-23 [2] CRAN (R 4.5.0)
 shiny                  1.12.1     2025-12-09 [2] CRAN (R 4.5.2)
 shinyWidgets           0.9.0      2025-02-21 [2] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [2] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 spatialLIBD          * 1.22.0     2025-11-04 [2] Bioconductor 3.22 (R 4.5.2)
 statmod                1.5.1      2025-10-09 [2] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1      2026-01-11 [2] CRAN (R 4.5.2)
 tidyr                  1.3.2      2025-12-19 [2] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [2] CRAN (R 4.5.0)
 vctrs                  0.7.1      2026-01-23 [2] CRAN (R 4.5.2)
 vipor                  0.4.7      2023-12-18 [2] CRAN (R 4.5.0)
 viridis                0.6.5      2024-01-29 [2] CRAN (R 4.5.0)
 viridisLite            0.4.3      2026-02-04 [2] CRAN (R 4.5.2)
 xfun                   0.56       2026-01-18 [2] CRAN (R 4.5.2)
 XML                    3.99-0.20  2025-11-08 [2] CRAN (R 4.5.0)
 xtable                 1.8-4      2019-04-21 [2] CRAN (R 4.5.0)
 XVector                0.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12     2025-12-10 [2] CRAN (R 4.5.2)

 [1] /Users/bmulvey/Library/R/arm64/4.5/library
 [2] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────────────