---
title: "03-Concat_wdLabel_and_wdLabelMerged_DE_result_tables"
author: "Bernie Mulvey"
date: "06-10-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
deres <- readRDS("processed-data_newseg/06_snSeqxferedLabelsDE/02-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_XferedWdLabels_andParentMergedLabels.RDS")
deres <- lapply(deres,function(x){
  z <- lapply(x,function(y){y$detabs})})

wdLabelres <- c(deres$wdLabels$AncesByLabel,deres$wdLabels$ApoAllelesLabel,deres$wdLabels$E2E4label,deres$wdLabels$sexlabel,deres$wdLabels$YRIAPOlabel)
wdLabelres <- rbindlist(wdLabelres,idcol="contrast")
wdLabelres[,label:=gsub(contrast,pattern="^(.*)_E4YRI_vs_E2YRI|_E4CEU_vs_E4YRI|_E4YRI_vs_E4CEU|_E4CEU_vs_E2CEU|_E2YRI_vs_E2CEU|_Sex|_E4vE2|_E4hom_vs_anyE2|_E4hom_vs_E4het|_E4hom_vs_others|_E4het_vs_anyE2|_E2hom_vs_anyE4|_E2het_vs_anyE4|_by_YRI",replacement="\\1")]
wdLabelres[,contrast:=apply(.SD,MARGIN=1,function(x){gsub(x[1],pattern=paste0(x[2],"_"),replacement="")}),.SDcols=c("contrast","label")]
reordn <- c("label","contrast","gene_name","adj.P.Val","logFC","P.Value","t","AveExpr","B")
wdLabelres <- wdLabelres[,..reordn]

wdLabmerres <- c(deres$wdMergedLabels$AncesByLabel,deres$wdMergedLabels$ApoAllelesLabel,deres$wdMergedLabels$E2E4label,deres$wdMergedLabels$sexlabel,deres$wdMergedLabels$YRIAPOlabel)
wdLabmerres <- rbindlist(wdLabmerres,idcol="contrast")
wdLabmerres[,label_merged:=gsub(contrast,pattern="^(.*)_E4YRI_vs_E2YRI|_E4CEU_vs_E4YRI|_E4YRI_vs_E4CEU|_E4CEU_vs_E2CEU|_E2YRI_vs_E2CEU|_Sex|_E4vE2|_E4hom_vs_anyE2|_E4hom_vs_E4het|_E4hom_vs_others|_E4het_vs_anyE2|_E2hom_vs_anyE4|_E2het_vs_anyE4|_by_YRI",replacement="\\1")]
wdLabmerres[,contrast:=apply(.SD,MARGIN=1,function(x){gsub(x[1],pattern=paste0(x[2],"_"),replacement="")}),.SDcols=c("contrast","label_merged")]
reordn <- c("label_merged","contrast","gene_name","adj.P.Val","logFC","P.Value","t","AveExpr","B")
wdLabmerres <- wdLabmerres[,..reordn]

fwrite(wdLabelres,"processed-data_newseg/06_snSeqxferedLabelsDE/03-XenXfered_wdLabels_allDEresults.txt",sep='\t',quote=F,row.names=F,col.names=T)
fwrite(wdLabmerres,"processed-data_newseg/06_snSeqxferedLabelsDE/03-XenXfered_wdLabelsMerged_allDEresults.txt",sep='\t',quote=F,row.names=F,col.names=T)
```

explor
```{r}
wdLabelres[label %in% grep(label,pattern="ast",value=T)&contrast=="E4vE2"&adj.P.Val<0.1,.(gene_name,contrast,label,logFC,adj.P.Val)]

wdLabelres[gene_name %in% c("FGFR3","PLPP3","")]
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] colorout_1.3-3      data.table_1.18.2.1

loaded via a namespace (and not attached):
 [1] digest_0.6.39     fastmap_1.2.0     xfun_0.56         lattice_0.22-7   
 [5] limma_3.66.0      knitr_1.51        htmltools_0.5.9   rmarkdown_2.30   
 [9] lifecycle_1.0.5   cli_3.6.5         grid_4.5.2        rsconnect_1.7.0  
[13] statmod_1.5.1     compiler_4.5.2    rprojroot_2.1.1   here_1.0.2       
[17] rstudioapi_0.18.0 tools_4.5.2       evaluate_1.0.5    yaml_2.3.12      
[21] otel_0.2.0        locfit_1.5-9.12   rlang_1.1.7       edgeR_4.8.2      
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-10
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────
 package     * version  date (UTC) lib source
 cli           3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 colorout    * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table  * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 digest        0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 edgeR         4.8.2    2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate      1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 fastmap       1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 here          1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools     0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 knitr         1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice       0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle     1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 limma         3.66.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 locfit        1.5-9.12 2025-03-05 [1] CRAN (R 4.5.0)
 otel          0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 rlang         1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown     2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot     2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect     1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi    0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 sessioninfo   1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 statmod       1.5.1    2025-10-09 [1] CRAN (R 4.5.0)
 xfun          0.56     2026-01-18 [1] CRAN (R 4.5.2)
 yaml          2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────