```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(ggrepel)
library(gridExtra)
library(SpatialExperiment)
library(lme4)
# library(SpatialFeatureExperiment)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
precision <- sqrt(.Machine$double.eps)
(x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
precision <- sqrt(.Machine$double.eps)
(x <= y) | (abs(x-y) <= precision)}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost='localhost')


## ggplot defaults
theme_set(theme_bw()+
  theme(panel.grid = element_blank(),
    axis.text.x = element_text(size = 8,color = '#000000'), 
    axis.title.x = element_text(size = 9,color = '#000000'), 
    axis.text.y = element_text(size = 8,color = '#000000'), 
    axis.title.y = element_text(size =9,color = '#000000'), 
    plot.title = element_markdown(size = 10,hjust=0.5,color = '#000000'), 
    strip.text = element_text(size=11), 
    legend.text = element_text(size=7,color = '#000000'),
    legend.title = element_text(size=8,hjust=0.5,color = '#000000'),
    axis.ticks = element_line(linewidth=0.5),
    panel.grid.major = element_line(linewidth=0.25),
    panel.grid.minor=element_line(linewidth=0.125),
    plot.title.position='plot'))
## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

These panels extend the NM-gene expression association findings from Visium in Fig 4.

###### FIG 4D-E: Residualized quartile plots ########

read model results with all coefs; prepare data as done for running the LMMs
```{r}
nmlm <- readRDS("processed-data_newseg/10_NManalyses/05-EverInLCdomain_wdXfer-orSRTxfer_LClabeled_NonMultimapNM_meanMedIntensity_LMMed_by_geneLogcounts_2apoeFactorizations_withRanEfFits.RDS")
nmlm <- nmlm$E4carrierYN

## read in corresponding SPE so we can feed in the gene logcounts, which is one of our variables in the model;
## subset to the cells that were used for the LMM analysis--straight outta code 10/05
lcx <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

## label transfers and manual lcs
wdx <- fread("processed-data_newseg/04_snSeqLabelTransfer/06-wdLCsmoothings_LUT.txt")
srt <- fread("processed-data_newseg/07_visLabelTransfer/05-srtLCsmoothDoms_and_NMFpreds_lut.txt")

lccells <- merge.data.table(srt,wdx,by="rn")

bsy <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_assignmentsMergeClusts.txt")
bsyan <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_mergeClusts_Annots.txt")
bsy <- merge.data.table(bsy,bsyan,by="clusid")
bsy[,anno:=ifelse(anno=="LCNE","LC","Other")]

lccells <- merge.data.table(lccells,bsy[,.(rn,anno)],by="rn")

### subset to cells that are in ANY lc domain definition first...
lccells <- lccells[domain8426=="LC"|domain12199=="LC"|domain16119=="LC"|domain20079=="LC"|domain12412=="LC"|domain8492=="LC"]
lccells <- lccells[nmfpred=="LC"|SRTnmfpred=="LC"|anno=="LC",rn] # 18.2k intra-domain LC cells across all definitions, before considering NM seg regist.

## read in nm measures-retain ONLY cells uniquely mapped to NMmoment
lcxnm <- fread("processed-data_newseg/10_NManalyses/01-NMmetrics_per_QCretained_cell.txt")
lcxnm <- lcxnm[cell_id %in% lccells&match_status=="unique_match"]
## make sure that this left no ambiguous NM-to-cell assignments
stopifnot(nrow(lcxnm[NM_bbox_n_cells==1])==nrow(lcxnm))
## OK
lcxnm <- lcxnm[,.(cell_id,NM_mean_intensity,NM_median_intensity)] 
# ^ 2892 cells ever annotated as LC cells AND in 1+ LC domain definition AND uniquely registered to an NM seg

lcx <- lcx[,lcxnm$cell_id]

lcnm <- DataFrame(lcxnm,row.names=lcxnm$cell_id)[colnames(lcx),]

## assign (FLIP THE SIGN FOR CONSISTENCY WITH SRT)
colData(lcx)$NM_mean_intensity <- (1-lcnm$NM_mean_intensity)
colData(lcx)$NM_median_intensity <- (1-lcnm$NM_median_intensity)

### get metadata
flares <- fread("../LFF_spatial_LC/raw-data/ancestryAndHaplotypes/FLARE/global_ancestry_estimates.txt")

lcnm.cd <- as.data.table(colData(lcx),keep.rownames=T)
lcnm.cd <- lcnm.cd[,.(rn,NM_mean_intensity,NM_median_intensity,Age,Sex,APOE,donor,brnum)] # brnum = sample (donor with hemisphere); donor entries do not have hemisphere
setnames(lcnm.cd,"APOE","APOEgeno")

### append YRI
stopifnot(all(lcnm.cd$donor%in%flares$ID))
lcnm.cd <- merge.data.table(lcnm.cd,flares[,.(ID,YRI)],by.x="donor",by.y="ID")

### xenium slide info
techmet <- fread("raw-data/donor_sample_slide_kit.txt")
stopifnot(all(lcnm.cd$brnum %in% techmet$Donor_sample))
lcnm.cd <- merge.data.table(lcnm.cd,techmet[,.(Donor_sample,Xenium_slide)],by.x="brnum",by.y="Donor_sample")

## we'll use two models with different approaches for APOE: E4+/- or individual genotypes (already in APOEgeno as E4hom/E4het/E2hom/E2het)
lcnm.cd[,E4carrier:="No"]
lcnm.cd[APOEgeno %in% c("E4hom","E4het"),E4carrier:="Yes"]

lcnm.lct <- logcounts(lcx)
lcnm.lct <- as.data.frame(as.matrix(lcnm.lct))

### keep this subsetted SPE in session for later. can clean out some other stuff though
rm(techmet,wdx,srt,flares,bsy,bsyan,lccells,lcnm,lcxnm)
```

WHATS BEING DONE BELOW: getting the residuals (without a gene effect in the model).
these plots are a little confusing to look at / comprehend in terms of +/- values when the imaging data of course only can give values 0-1.
```{r}
## write as a function to make residualized plot for gene of interest
residplt <- function(g,nonzero_only=FALSE){
    exprtab <- as.data.table(t(lcnm.lct[g,]),keep.rownames=T)
    exprtab <- merge.data.table(exprtab,lcnm.cd[,.(rn,brnum,NM_mean_intensity,Age,Sex,APOEgeno,YRI,Xenium_slide,E4carrier)],by="rn")


    ### fit a model that doesn't have an expression term
    f2 <- formula(paste0("NM_mean_intensity ~ Age + Sex + YRI + E4carrier + (1|Xenium_slide) + (1|Xenium_slide:brnum)")) # no term for gene; E2/E4 model; sample within xenium slide
    noxprmod <- lmer(f2,data=exprtab)

    ### get fitted values -> residuals, and expression quantiles
    exprtab[,noxprfitted:=predict(noxprmod)]
    exprtab[,noxprresid:=NM_mean_intensity-noxprfitted]
    
    if(nonzero_only){
        exprtab <- exprtab[get(g)>0]
        xtitle <- paste0("Quartile *",g,"* Expression<br>(In Cells with Nonzero Counts)")
    } else{
        xtitle <- paste0("Quartile *",g,"* Expression")
    }
    quartile_breaks <- quantile(exprtab[,..g], probs = seq(0, 1, by = 0.25), na.rm = TRUE)
    exprtab[, x_quartile := cut(get(g), 
                        breaks = quartile_breaks, 
                        labels = 1:4,         # Label bins accoring to nth decile
                        include.lowest = TRUE)]
    exprtab[x_quartile==1,xax_quart:="0-25th"]
    exprtab[x_quartile==2,xax_quart:="25th-50th"]
    exprtab[x_quartile==3,xax_quart:="50th-75th"]
    exprtab[x_quartile==4,xax_quart:="75th-100th"]
    exprtab[,xax_quart:=factor(xax_quart,levels=c("0-25th","25th-50th","50th-75th","75th-100th"))]

    # calculate filll color as the mean (mean pixel intensity) value for that quartile
    # exprtab[,calcfill:=mean(NM_mean_intensity),by="xax_quart"]
    # sample mean per quartile
    exprtab[,calcfill:=mean(NM_mean_intensity),by=c("brnum","xax_quart")]
    # mean of sample means per quartile
    exprtab[,calcfill:=mean(calcfill),by="xax_quart"]
    
    p <- ggplot(exprtab,aes(x=xax_quart,y=noxprresid,fill=calcfill))+
        geom_boxplot(notch = T)+
        scale_fill_gradient(low="brown",high="beige",limits=c(0.975*min(exprtab$calcfill),1.025*max(exprtab$calcfill)))+
        labs(y="Residualized NM Intensity",x=xtitle)+
        guides(fill="none")+
        theme(axis.title.x=element_markdown(size=8))
    return(p)
}

## no, these are just confusing, especially since the SRT plots are just the straight-up values from samples.

# pdf("plots_newseg/manuscript/Fig4/4D-MAP1LC3B_NMresiduals_by_quartileXprReplicated_Xen_NMgeneXprAssns.pdf",height=2.2,width=2.25)
residplt("MAP1LC3B",nonzero_only = FALSE)
dev.off()

# pdf("plots_newseg/manuscript/Fig4/4E-SARAF_NMresiduals_by_quartileXprReplicated_Xen_NMgeneXprAssns.pdf",height=2.2,width=2.25)
residplt("SARAF",nonzero_only = TRUE)
dev.off()

```

## alternative: mean cellwsie NM pixel intensity per sample per quartile
```{r}

## write as a function to make residualized plot for gene of interest
qtileplt <- function(g,nonzero_only=FALSE){
    exprtab <- as.data.table(t(lcnm.lct[g,]),keep.rownames=T)
    exprtab <- merge.data.table(exprtab,lcnm.cd[,.(rn,brnum,NM_mean_intensity,Age,Sex,APOEgeno,YRI,Xenium_slide,E4carrier)],by="rn")

    if(nonzero_only){
        exprtab <- exprtab[get(g)>0]
        xtitle <- paste0("Quartile *",g,"* Expression<br>(In Cells with Nonzero Counts)")
    } else{
        xtitle <- paste0("Quartile *",g,"* Expression")
    }
    quartile_breaks <- quantile(exprtab[,..g], probs = seq(0, 1, by = 0.25), na.rm = TRUE)
    exprtab[, x_quartile := cut(get(g), 
                        breaks = quartile_breaks, 
                        labels = 1:4,         # Label bins accoring to nth decile
                        include.lowest = TRUE)]
    exprtab[x_quartile==1,xax_quart:="0-25th"]
    exprtab[x_quartile==2,xax_quart:="25th-50th"]
    exprtab[x_quartile==3,xax_quart:="50th-75th"]
    exprtab[x_quartile==4,xax_quart:="75th-100th"]
    exprtab[,xax_quart:=factor(xax_quart,levels=c("0-25th","25th-50th","50th-75th","75th-100th"))]

    # calculate filll color as the mean (mean pixel intensity) value for that quartile
    ## mean of means by sample
    exprtab[,calcfill:=mean(NM_mean_intensity),by=c("brnum","xax_quart")]
    ## mean of sample means
    exprtab[,calcfill:=mean(calcfill),by="xax_quart"]

    p <- ggplot(exprtab,aes(x=xax_quart,y=NM_mean_intensity,fill=calcfill))+
        geom_boxplot(notch = T)+
        scale_fill_gradient(low="brown",high="beige",limits=c(0.975*min(exprtab$calcfill),1.025*max(exprtab$calcfill)))+
        labs(y="Residualized NM Intensity",x=xtitle)+
        guides(fill="none")+
        theme(axis.title.x=element_markdown(size=8))
    return(p)
}

qtileplt("MAP1LC3B",nonzero_only = F)
```

```{r}
saveplt <- function(g){
    pltdat <- as.data.frame(t(as.matrix(logcounts(lc3[g,]))))
    pltdat <- as.data.table(pltdat,keep.rownames=T)
    setnames(pltdat,g,"g_xpr") # since the value of g will be a variable in coldata
    pltdat <- merge.data.table(metad,pltdat,by.x="rowname",by.y="rn")
    pltdat[,nspot:=.N,by="sample_id"] # determine the number of spots from each sample INCLUDING those with logcounts=0 (though we won't plot those)
    pltdat <- pltdat[nspot>19]
    pltdat <- pltdat[g_xpr!=0]
    
    noxprmod <- lmer(formula(Intensity_NM~Age+Sex+YRI+APOEgeno+(1|brnum)+(1|brnum:capture_id)+(1|brnum:capture_id:sample_id)),data=pltdat)
    
    pltdat[,noxprfitted:=predict(noxprmod)]
    pltdat[,noxprresid:=Intensity_NM-noxprfitted]

    quartile_breaks <- quantile(pltdat$g_xpr, probs = seq(0, 1, by = 0.25), na.rm = TRUE)
    
    pltdat[, x_quartile := cut(g_xpr, 
                     breaks = quartile_breaks, 
                     labels = 1:4,         # Label bins accoring to nth decile
                     include.lowest = TRUE)]
    
    pltdat[x_quartile==1,xvals:="0-25th %ile"]
    pltdat[x_quartile==2,xvals:="25-50th %ile"]
    pltdat[x_quartile==3,xvals:="50-75th %ile"]
    pltdat[x_quartile==4,xvals:="75-100th %ile"]
    pltdat[,xvals:=factor(xvals,levels=c("0-25th %ile","25-50th %ile","50-75th %ile","75-100th %ile"))]
    
    pltdat[,calcfill:=mean(Intensity_NM),by="x_quartile"]
    p <- ggplot(pltdat[,.(xvals,noxprresid,calcfill)],aes(x=xvals,group=xvals,y=noxprresid,fill=calcfill))+
        scale_fill_gradient(low="brown",high="beige")+
        guides(fill="none")+
        geom_boxplot()+
        labs(y=paste0("Spot NM Intensity Residual<br>(Model Without *",g,"* Term)"),x=paste0("Quartile logcounts *",g,"*"))+
        theme(axis.title.x=element_markdown(size=9),
              axis.title.y=element_markdown(size=9),
              axis.text.x=element_text(size=8),
              axis.text.y=element_text(size=8))

    # pdf(paste0("plots/manuscript/Fig4/4x-Spotlogcountquartiles_",g,"_spotNMintensity_nozeros.pdf"),height=2.5,width=2.5)
    # print(p)
    # dev.off()
    return(p)
}

saveplt("PGRMC1")
```





###### FIG 4F - Volcano of NM-gene expression associations from SRT replicated in ISS ######
First, get the visium and Xenium NM-gene expression associations using E2/E4 two-condition model and mean NM pixel intensity from the Xenium data. (This was the LC-domain-only + some sort of LC label assigned to cell only + NM containing (obviously) analysis, code 10/05) 
```{r}
vislmm <- readRDS("../LFF_spatial_LC/processed-data/12_DEanalyses_removedsampsAndFinalNMseg/11-LCNMspots_NMmetric_LMMed_by_geneLogcounts_2apoeFactorizations_withRanEfFits.RDS")

# get fixed effects tables for pixel intensity (no NM prop in xenium), with gene effects only
vislmm <- vislmm$E4carrierYN$Intensity_NM$fixed
setnames(vislmm,c("Pr(>Chisq)","coef"),c("pval_vis","coef_vis"))
vislmm[,fdr_vis:=p.adjust(pval_vis,method="BH")]
vislmm[,bonf_vis:=pval_vis*length(unique(vislmm$gene_name))]
vislmm[bonf_vis>1,bonf_vis:=1]

## then the same model and mean intensity from xenium:
##  remember, the sign convention flipped in xenium--positive means more NM with higher xpr in xenium
xenlmm <- readRDS("processed-data_newseg/10_NManalyses/05-EverInLCdomain_wdXfer-orSRTxfer_LClabeled_NonMultimapNM_meanMedIntensity_LMMed_by_geneLogcounts_2apoeFactorizations_withRanEfFits.RDS")

xenlmm <- xenlmm$E4carrierYN$NM_mean_intensity$fixed
setnames(xenlmm,c("Pr(>Chisq)","coef"),c("pval_xen","coef_xen"))
xenlmm[,coef_xen:=-1*coef_xen] # sign flip to match visium
xenlmm <- xenlmm[variable==gene_name] # subset to gene coefficients
xenlmm[,fdr_xen:=p.adjust(pval_xen,method="BH")]
xenlmm[,bonf_xen:=pval_xen*length(unique(xenlmm$gene_name))]
xenlmm[bonf_xen>1,bonf_xen:=1]

### combine
nmmerge <- merge.data.table(xenlmm[,.(gene_name,coef_xen,pval_xen,fdr_xen,bonf_xen)],vislmm[,.(gene_name,coef_vis,fdr_vis,bonf_vis)],by="gene_name")

nmmerge[,replic:="No"]
# replication: bonferonni on visium, nomsig on xenium
nmmerge[lessOrEqual(bonf_vis,0.05)&lessOrEqual(pval_xen,0.05)&(coef_vis*coef_xen)>0,replic:="Yes"]
# how many visium signif, xenium nonsignif?
nmmerge[lessOrEqual(bonf_vis,0.05)&greaterOrEqual(pval_xen,0.05)]
## none!
## how many vis signif, xen signif, discordant?
nmmerge[lessOrEqual(bonf_vis,0.05)&lessOrEqual(pval_xen,0.05)&(coef_vis*coef_xen)<0]
## also none!!

## prep data to make a volcano plot of the xenium genes that replicate
nmmerge <- nmmerge[lessOrEqual(bonf_vis,0.05)&lessOrEqual(pval_xen,0.05)&(coef_vis*coef_xen)>0]
## genes to label: "DBH", "MAP1LC3B", "ADRA2A", "SLC6A2", "PDGFD", "ENC1", "PGRMC1", "SARAF", "GPX3", "APOE", "MT1E"
nmmerge[gene_name %in% c("DBH", "MAP1LC3B", "ADRA2A", "SLC6A2", "PDGFD", "ENC1", "PGRMC1", "SARAF", "GPX3", "APOE", "MT1E"),genelab:=gene_name]
```

## get info on the number of cells tested expressing each gene
```{r}
lcx <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

## label transfers and manual lcs
wdx <- fread("processed-data_newseg/04_snSeqLabelTransfer/06-wdLCsmoothings_LUT.txt")
srt <- fread("processed-data_newseg/07_visLabelTransfer/05-srtLCsmoothDoms_and_NMFpreds_lut.txt")

lccells <- merge.data.table(srt,wdx,by="rn")

bsy <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_assignmentsMergeClusts.txt")
bsyan <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_mergeClusts_Annots.txt")
bsy <- merge.data.table(bsy,bsyan,by="clusid")
bsy[,anno:=ifelse(anno=="LCNE","LC","Other")]

lccells <- merge.data.table(lccells,bsy[,.(rn,anno)],by="rn")

### subset to cells that are in ANY lc domain definition first...
lccells <- lccells[domain8426=="LC"|domain12199=="LC"|domain16119=="LC"|domain20079=="LC"|domain12412=="LC"|domain8492=="LC"]
lccells <- lccells[nmfpred=="LC"|SRTnmfpred=="LC"|anno=="LC",rn] # 18.2k intra-domain LC cells across all definitions, before considering NM seg regist.

## read in nm measures-retain ONLY cells uniquely mapped to NMmoment
lcxnm <- fread("processed-data_newseg/10_NManalyses/01-NMmetrics_per_QCretained_cell.txt")
lcxnm <- lcxnm[cell_id %in% lccells&match_status=="unique_match",
              .(cell_id,NM_mean_intensity,NM_median_intensity)] 
# ^ 2892 LC cells ever annotated as LC cells AND in an LC domain AND registered to an NM seg

## subset to genes of nascent volcano, get counts table
lcx <- lcx[,lcxnm$cell_id]
lcx <- lcx[nmmerge$gene_name,]

lcct <- rowSums(counts(lcx)>0)
lcct <- as.data.table(lcct,keep.rownames=T)
setnames(lcct,2,"Nxpr")
lcct[,pctXpr:=Nxpr/2892]
lcct[,pctXpr:=pctXpr*100]
lcct[,pctXpr:=as.numeric(pctXpr)]
lcm <- cbind(rownames(lcx),rowMedians(as.matrix(counts(lcx))))
lcm <- as.data.table(lcm)
setnames(lcm,c("gene_name","medXpr"))
lcm[,medXpr:=as.numeric(medXpr)]
```

## we can fix the point sizes in illustrator later, ggplot just fuckign sucks at that bit.
```{r}
### append info on number of cells with each gene to our plot table, make and save the plot.
nmmerge <- merge.data.table(nmmerge,lcct,by.x="gene_name",by.y="rn")
nmmerge <- merge.data.table(nmmerge,lcm,by="gene_name")

# save it
pdf("plots_newseg/manuscript/Fig4/4F-Replicated_Xen_NMgeneXprAssns.pdf",height=4.5,width=4.4)
ggplot(nmmerge,aes(x=coef_xen,y=-log10(pval_xen),col=pctXpr,size=medXpr))+
    geom_point()+
    scale_color_viridis_c(limits=c(0,100),begin=1,end=0.4)+
    scale_size(limits = c(0,161),range=c(0.65,2.5),breaks=c(0,55,110,160))+
    geom_hline(yintercept = -log10(0.05), linetype="dashed", color="red",linewidth=0.5) +
    geom_text_repel(seed = 23942,aes(y = -log10(pval_xen),
            x = coef_xen,
            label = ifelse(is.na(genelab), "", genelab)),
        min.segment.length = 0,
        size = 2,
        show.legend=FALSE,
        max.iter=100000,
        max.overlaps=10000,
        max.time=3,
        segment.size=0.4,
        force_pull = -0.005,
        fontface = "italic")+
    labs(x="ISS gene coefficient (<0: Darker<br>NM with greater expression)",
            y= "-log10(ISS *p*)",
            color="Cell % with<br>>0 counts",
            size="Median counts")+
    scale_x_continuous(limits=c(min(nmmerge$coef_xen),-min(nmmerge$coef_xen)))+
    theme(axis.title.x=element_markdown(size=8),
        axis.title.y=element_markdown(size=8),
        legend.title=element_markdown(size=8))
dev.off()
```

```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices
[5] utils     methods   base     

other attached packages:
 [1] lme4_1.1-38                
 [2] Matrix_1.7-4               
 [3] ggrepel_0.9.6              
 [4] colorout_1.3-3             
 [5] SpatialExperiment_1.20.0   
 [6] SingleCellExperiment_1.32.0
 [7] SummarizedExperiment_1.40.0
 [8] Biobase_2.70.0             
 [9] GenomicRanges_1.62.1       
[10] Seqinfo_1.0.0              
[11] IRanges_2.44.0             
[12] S4Vectors_0.48.1           
[13] BiocGenerics_0.56.0        
[14] generics_0.1.4             
[15] MatrixGenerics_1.22.0      
[16] matrixStats_1.5.0          
[17] BiocParallel_1.44.0        
[18] parallelly_1.47.0          
[19] gridExtra_2.3              
[20] ggtext_0.1.2               
[21] ggplot2_4.0.2              
[22] data.table_1.18.4          

loaded via a namespace (and not attached):
 [1] tidyselect_1.2.1    viridisLite_0.4.3  
 [3] dplyr_1.2.0         farver_2.1.2       
 [5] S7_0.2.1            lifecycle_1.0.5    
 [7] magrittr_2.0.4      compiler_4.5.2     
 [9] rlang_1.1.7         tools_4.5.2        
[11] knitr_1.51          S4Arrays_1.10.1    
[13] labeling_0.4.3      here_1.0.2         
[15] DelayedArray_0.36.1 xml2_1.5.2         
[17] RColorBrewer_1.1-3  abind_1.4-8        
[19] withr_3.0.2         grid_4.5.2         
[21] scales_1.4.0        MASS_7.3-65        
[23] dichromat_2.0-0.1   cli_3.6.5          
[25] reformulas_0.4.4    ragg_1.5.0         
[27] otel_0.2.0          rjson_0.2.23       
[29] sessioninfo_1.2.3   commonmark_2.0.0   
[31] minqa_1.2.8         stringr_1.6.0      
[33] splines_4.5.2       parallel_4.5.2     
[35] XVector_0.50.0      vctrs_0.7.1        
[37] boot_1.3-32         litedown_0.9       
[39] systemfonts_1.3.1   magick_2.9.0       
[41] glue_1.8.0          nloptr_2.2.1       
[43] codetools_0.2-20    stringi_1.8.7      
[45] gtable_0.3.6        tibble_3.3.1       
[47] pillar_1.11.1       R6_2.6.1           
[49] textshaping_1.0.4   Rdpack_2.6.5       
[51] rprojroot_2.1.1     evaluate_1.0.5     
[53] lattice_0.22-7      markdown_2.0       
[55] rbibutils_2.4.1     gridtext_0.1.5     
[57] Rcpp_1.1.1          SparseArray_1.10.8 
[59] nlme_3.1-168        xfun_0.56          
[61] pkgconfig_2.0.3    
> sessioninfo::session_info()
─ Session info ──────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       Positron
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-07-09
 pandoc   NA
 quarto   1.9.38 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────
 package              * version date (UTC) lib source
 abind                  1.4-8   2024-09-12 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 boot                   1.3-32  2025-08-29 [1] CRAN (R 4.5.2)
 cli                    3.6.5   2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20  2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3   2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 commonmark             2.0.0   2025-07-07 [1] CRAN (R 4.5.0)
 data.table           * 1.18.4  2026-05-06 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.1  2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1 2022-05-02 [1] CRAN (R 4.5.0)
 dplyr                  1.2.0   2026-02-03 [1] CRAN (R 4.5.2)
 evaluate               1.0.5   2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2   2024-05-13 [1] CRAN (R 4.5.0)
 generics             * 0.1.4   2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1  2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 ggplot2              * 4.0.2   2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel              * 0.9.6   2024-09-07 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2   2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0   2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3     2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5   2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6   2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2   2025-09-15 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51    2025-12-20 [1] CRAN (R 4.5.2)
 labeling               0.4.3   2023-08-29 [1] CRAN (R 4.5.0)
 lattice                0.22-7  2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5   2026-01-08 [1] CRAN (R 4.5.2)
 litedown               0.9     2025-12-18 [1] CRAN (R 4.5.2)
 lme4                 * 1.1-38  2025-12-02 [1] CRAN (R 4.5.2)
 magick                 2.9.0   2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4   2025-09-12 [1] CRAN (R 4.5.0)
 markdown               2.0     2025-03-23 [1] CRAN (R 4.5.0)
 MASS                   7.3-65  2025-02-28 [1] CRAN (R 4.5.2)
 Matrix               * 1.7-4   2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0   2025-01-07 [1] CRAN (R 4.5.0)
 minqa                  1.2.8   2024-08-17 [1] CRAN (R 4.5.0)
 nlme                   3.1-168 2025-03-31 [1] CRAN (R 4.5.2)
 nloptr                 2.2.1   2025-03-17 [1] CRAN (R 4.5.0)
 otel                   0.2.0   2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.47.0  2026-04-17 [1] CRAN (R 4.5.2)
 pillar                 1.11.1  2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.5.0)
 R6                     2.6.1   2025-02-15 [1] CRAN (R 4.5.0)
 ragg                   1.5.0   2025-09-02 [1] CRAN (R 4.5.0)
 rbibutils              2.4.1   2026-01-21 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3   2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1   2026-01-10 [1] CRAN (R 4.5.2)
 Rdpack                 2.6.5   2026-01-23 [1] CRAN (R 4.5.2)
 reformulas             0.4.4   2026-02-02 [1] CRAN (R 4.5.2)
 rjson                  0.2.23  2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7   2026-01-09 [1] CRAN (R 4.5.2)
 rprojroot              2.1.1   2025-08-26 [1] CRAN (R 4.5.0)
 S4Arrays               1.10.1  2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1  2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1   2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0   2025-04-24 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3   2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0  2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8  2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0  2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 stringi                1.8.7   2025-03-27 [1] CRAN (R 4.5.0)
 stringr                1.6.0   2025-11-04 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0  2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 systemfonts            1.3.1   2025-10-01 [1] CRAN (R 4.5.0)
 textshaping            1.0.4   2025-10-10 [1] CRAN (R 4.5.0)
 tibble                 3.3.1   2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1   2026-01-23 [1] CRAN (R 4.5.2)
 viridisLite            0.4.3   2026-02-04 [1] CRAN (R 4.5.2)
 withr                  3.0.2   2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56    2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2   2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────