---
title: "Fig7-SmoothedLCDomains_AncestrySpecificDE"
format: html
date: "07-01-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(Polychrome)
library(SpatialExperiment)
library(ggrastr)
library(escheR)
library(jaffelab)
library(sf)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
##

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
precision <- sqrt(.Machine$double.eps)
(x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
precision <- sqrt(.Machine$double.eps)
(x <= y) | (abs(x-y) <= precision)}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost='localhost')


## ggplot defaults
theme_set(theme_bw()+
  theme(panel.grid = element_blank(),
    axis.text.x = element_text(size = 8,color = '#000000'), 
    axis.title.x = element_text(size = 9,color = '#000000'), 
    axis.text.y = element_text(size = 8,color = '#000000'), 
    axis.title.y = element_text(size =9,color = '#000000'), 
    plot.title = element_markdown(size = 10,hjust=0.5,color = '#000000'), 
    strip.text = element_text(size=11), 
    legend.text = element_text(size=7,color = '#000000'),
    legend.title = element_text(size=8,hjust=0.5,color = '#000000'),
    axis.ticks = element_line(linewidth=0.5),
    panel.grid.major = element_line(linewidth=0.25),
    panel.grid.minor=element_line(linewidth=0.125),
    plot.title.position='plot'))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

## Plot from a previously-plotted sample, showing transferred labels themselves, with similar anat to its visium counterpart
## poss brs: Fig 2 (Br5941, Br0946, Br6538, Br1105) or Fig 1 (Br3974)
## try Br5941, Br6538

# SRT dom12199 least stringent, 20079 most stringent (44189, 32865 cells)
```{r}

# helper function to extract a legend and pipe back to a diff plot: from https://statisticsglobe.com/add-common-legend-to-combined-ggplot2-plots-in-r/#example-2-add-shared-legend-to-ggplot2-plots-using-gridextra-package 

extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}
```

```{r}
pal <- readRDS("../LFF_spatial_LC/plots/finalpal_somuchfornorainbows.RDS")

xlc <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

lxf <- fread("processed-data_newseg/07_visLabelTransfer/05-srtLCsmoothDoms_and_NMFpreds_lut.txt")
## use a permissive and strict domain for plotting
lxf <- lxf[,.(rn,domain12199,domain20079,SRTnmfpred)]
## convert working names to plotting names for cluster labels
lxf[SRTnmfpred=="Astro_Oligo",SRTnmfpred:="Astro-Oligo Mixed"]
lxf[SRTnmfpred=="Oligo_Astro",SRTnmfpred:="Oligo-Astro Mixed"]
lxf[SRTnmfpred=="ACHE_SERT_Npep",SRTnmfpred:="*ACHE*, *SERT*,<br>Neuropept. Mixed"]
lxf[SRTnmfpred=="ACHE_SERT_LC_WM",SRTnmfpred:="*ACHE*, *SERT*,<br>LC, Oligo Mixed"]
lxf[SRTnmfpred=="Oligo_Complex",SRTnmfpred:="Oligo-Other<br>Types Mixed"]

names(pal)[2] <- "*ACHE*, *SERT*,<br>LC, Oligo Mixed"
names(pal)[3] <- "*ACHE*, *SERT*,<br>Neuropept. Mixed"
names(pal)[9] <- "Oligo-Other<br>Types Mixed"

stopifnot(all(lxf$SRTnmfpred %in% names(pal)))

## add to coldata
tmpcd <- as.data.table(colData(xlc),keep.rownames=T)
tmpcd <- merge.data.table(tmpcd,lxf,by="rn")
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(xlc),]
tmpcd$rn <- NULL
colData(xlc) <- tmpcd

rm(lxf)
```

### make the plots of Br5941 with cell annots, LC domain perimeter drawn on, and an inset that zooms to just the domain part. we'll do whatever rotating is necess. to match visium orientation in illustrator.

### first, determine the domain boundaries.
```{r}
pane <- xlc[,xlc$brnum=="Br5941"]

d12 <- st_as_sf(as.data.frame(spatialCoords(pane[,pane$domain12199=="LC"])),coords = c("x_centroid","y_centroid"))

d12hullpt <- st_concave_hull(st_union(d12),ratio = 0.2)
## create the data table we will use for overlaying the boundary polygons. geom_path will need the first point to also be popped into the last slot to close the shape
dompoly12 <- rbind(d12hullpt[[1]][[1]],
                  d12hullpt[[1]][[1]][1,])
dompoly12 <- cbind(dompoly12,rep("LC",nrow(dompoly12)))

dompoly12 <- as.data.table(dompoly12)

setnames(dompoly12,c("xpol","ypol","dompol"))

dompoly12[,xpol:=as.numeric(xpol)]
dompoly12[,ypol:=as.numeric(ypol)]

rm(d12,d12hullpt)
```

### then, make full-section escher plot with the domain perimeter overlaid...
(for main fig, dom12199 only)
```{r}
# sizepal <- c(0.5,rep(0.2,8)) <- we need to pass a vector of point sizes to have the polygon-over-spot plotting scheme work here:

sizesman <- ifelse(pane$SRTnmfpred=="LC",0.5,0.2)

plt <- make_escheR(pane,y_reverse=FALSE)
plt <- plt |> add_fill("SRTnmfpred",size=0.2,point_size = 0.2)


## assign spot colors based on our existing palette
p <- plt +
  labs(fill="Transferred\nVisium Label")+
  scale_fill_manual(values=pal)+
  guides(fill=guide_legend(override.aes=list(size=1.5)))

## override the escheR size settings with the spot sizes determined above:
p@layers[[1]]$aes_params$size <- sizesman


## draw LC domain boundary; drop guides for first panel; format for second to extract later
p12 <- p + 
  geom_path(data=dompoly12,aes(x=xpol,y=ypol,group=dompol,col=dompol))+
  scale_color_manual(values=pal)+guides(color="none",size="none")

## title
p12 <- p12 + ggtitle("Br5941 (E4/E4, AA, M) SRT12199")

### format for consistency/illustrator editing simplifying
p12 <- p12 + theme(panel.grid = element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      axis.ticks = element_blank(),
      legend.text = element_markdown(size=6,margin=margin(0,0,0,0,"in")),
      legend.key.size = ggplot2::unit(0.075,"in"),
      legend.title=element_text(size=7,hjust=0.5),
      legend.background = element_blank(),
      # legend.box = element_blank(),
      legend.box.spacing = ggplot2::unit(0,"in"),
      legend.key = element_blank(),
      legend.key.height=ggplot2::unit(0.0125,"in"),
      legend.key.width=ggplot2::unit(0.0125,"in"),
      legend.key.spacing.y = ggplot2::unit(0.04,"in"),
      title=element_text(size=9,hjust=0.5))
      # legend.position.inside = c(0.19,0.165),
      # legend.position="inside")

p12 <- rasterize(p12,layers="Points",dpi=600,dev="cairo_png")

# pdf(paste0("plots_newseg/manuscript/SuppFigs/Predlabels_SRT12199_SRT20079_wDomainBounds_Br5941.pdf"),height=2,width=6.5)
# do.call("grid.arrange",list(arrangeGrob(grobs=c(p12,p20),ncol=2),leg,ncol=2,widths=c(5,1.5)))
# dev.off()

# main: srtdom12
pdf(paste0("plots_newseg/manuscript/Fig7/Fig7A-Predlabels_SRT12199_wDomainBounds_Br5941.pdf"),height=2,width=2,onefile = F)
print(p12)
dev.off()

rm(p12,sizesman,plt,p)
```

#### and an inset that just spans the LC area
```{r}
### get the min and max x and y coords of the LC polygon, add a little buffer space, and get cell IDs for all cells that correspond
rownames(spatialCoords(pane)) <- colnames(pane)
zoomcoord <- as.data.table(spatialCoords(pane),keep.rownames=T)
range(dompoly12$xpol) # 88-1256
range(dompoly12$ypol) # 2177-4019

zoomcoord <- zoomcoord[greaterOrEqual(x_centroid,50)&lessOrEqual(x_centroid,1350)]
zoomcoord <- zoomcoord[greaterOrEqual(y_centroid,2050)&lessOrEqual(y_centroid,4150)]

pane2 <- pane[,zoomcoord$rn]


sizesman <- ifelse(pane2$SRTnmfpred=="LC",0.75,0.5)

plt <- make_escheR(pane2,y_reverse=FALSE)
plt <- plt |> add_fill("SRTnmfpred",size=0.2,point_size = 0.2)


## assign spot colors based on our existing palette; will use same legend
p <- plt + 
  scale_fill_manual(values=pal)

## override the escheR size settings with the spot sizes determined above:
p@layers[[1]]$aes_params$size <- sizesman


## draw LC domain boundary
p <- p + 
  geom_path(data=dompoly12,aes(x=xpol,y=ypol,group=dompol,col=dompol))+
  scale_color_manual(values=pal)

## no legends needed
p <- p +
  guides(fill="none",size="none",color="none")

## no title

## for consistency/illustrator editing simplifying
p <- p+ theme(panel.grid = element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      axis.ticks = element_blank(),
      legend.text = element_markdown(size=6,margin=margin(0,0,0,0,"in")),
      legend.key.size = ggplot2::unit(0.075,"in"),
      legend.title=element_text(size=7,hjust=0.5),
      legend.background = element_blank(),
      legend.box = element_blank(),
      legend.box.spacing = ggplot2::unit(0,"in"),
      legend.key = element_blank(),
      legend.key.height=ggplot2::unit(0.0125,"in"),
      legend.key.width=ggplot2::unit(0.0125,"in"),
      legend.key.spacing.y = ggplot2::unit(0.04,"in"),
      title=element_text(size=10,hjust=0.5))
      # legend.position.inside = c(0.19,0.165),
      # legend.position="inside")

p <- rasterize(p,layers="Points",dpi=600,dev="cairo_png")

pdf(paste0("plots_newseg/manuscript/Fig7/7Ainset-Predlabels_wSRT20079DomainBounds_Br5941.pdf"),height=2.25,width=1.5)
print(p)
dev.off()

### keep srt12199 domain polygon so we can show the banksy clusters within it too
rm(pane,pane2,p,sizesman)
```

### and the visium LC in the same donor
```{r}
vlc <- readRDS("../LFF_spatial_LC/processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

vclus <- readRDS("../LFF_spatial_LC/processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
vclus <- vclus$HARMONYlmbna_HDG_SVG_2575
vclus <- vclus[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(vclus,2,"clusid")

anno <- fread("../LFF_spatial_LC/processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")
vclus <- merge(vclus,anno[,.(clusid,plot_anno)],by="clusid")
vclus[plot_anno=="Oligo and Other Types Mixed",plot_anno:="Oligo-Other<br>Types Mixed"]
vclus[plot_anno=="*ACHE*, *SERT*, LC, Oligo Mixed",plot_anno:="*ACHE*, *SERT*,<br>LC, Oligo Mixed"]
vclus[plot_anno=="*ACHE*, *SERT*, Neuropeptide Mixed",plot_anno:="*ACHE*, *SERT*,<br>Neuropept. Mixed"]
stopifnot(all(vclus$plot_anno %in% names(pal)))

vclus <- DataFrame(vclus,row.names=vclus$rn)[colnames(vlc),]
# length(unique(vlc[,vlc$brnum=="Br5941"]$sample_id)) # 2
vlc$plot_anno <- vclus$plot_anno
vlc <- vlc[,vlc$brnum=="Br5941"]
vlc$pltalph <- ifelse(vlc$plot_anno=="LC",1,0.6)

samps <- unique(vlc$sample_id)
visplt <- lapply(samps,function(x){
    plt <- make_escheR(vlc[,vlc$sample_id==x],y_reverse=FALSE)
    plt <- add_fill(plt,"plot_anno",size=0.75,point_size = 0.75)
    plt <- plt + 
        scale_fill_manual(values=pal)+
        aes(alpha=pltalph)+
        guides(fill="none",size="none",color="none",alpha="none") # use the one from the xenium panel
    plt <- plt+ theme(panel.grid = element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      axis.ticks = element_blank(),
      legend.text = element_markdown(size=6,margin=margin(0,0,0,0,"in")),
      legend.key.size = ggplot2::unit(0.075,"in"),
      legend.title=element_text(size=7,hjust=0.5),
      legend.background = element_blank(),
      legend.box = element_blank(),
      legend.box.spacing = ggplot2::unit(0,"in"),
      legend.key = element_blank(),
      legend.key.height=ggplot2::unit(0.0125,"in"),
      legend.key.width=ggplot2::unit(0.0125,"in"),
      legend.key.spacing.y = ggplot2::unit(0.04,"in"),
      title=element_text(size=10,hjust=0.5))
    plt <- rasterize(plt,layers="Points",dpi=300,dev="cairo_png")
    
    return(plt)
})

pdf(paste0("plots_newseg/manuscript/Fig7/7B-VisiumLCs_Br5941.pdf"),height=2.25,width=3)
do.call("grid.arrange",c(visplt,nrow=1))
dev.off()
```

### next, we want to plot the pseudobulk results for AA, EA, and all ancestries in 12199
```{r}
srtdomde <- readRDS("processed-data_newseg/08_VisXferDomainDE/02-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_LCdoms_andOtherXferedVisLabels.RDS")

pbulk <- srtdomde$dom12199$YRIAPOlabel$voomLFt$EList$E
keepn <- grep(colnames(pbulk),pattern="^LC_Br",value=T)

sampinf <- srtdomde$dom12199$YRIAPOlabel$dgel$samples
keepn <- grep(rownames(sampinf),pattern="^LC_Br",value=T)
sampinf <- sampinf[keepn,]

### cleaning y for the pbulk
sampinf$lab_APOE_PredomAnces <- factor(sampinf$lab_APOE_PredomAnces,levels = unique(sampinf$lab_APOE_PredomAnces))
sampinf$Sex <- factor(sampinf$Sex)
sampinf$Age <- as.numeric(sampinf$Age)

pbulk <- pbulk[,rownames(sampinf)]

modmat <-  model.matrix(~0+lab_APOE_PredomAnces+Age+Sex,data = sampinf)
degs <- jaffelab::cleaningY(pbulk,modmat,5) # correct for terms that arent in the ancestry-label groupings (term 1 intercept, 2,3,4,5 levels of lab_APOE_ancest, 6 Age, 7 Sex)
rownames(degs) <- rownames(pbulk)
colnames(degs) <- colnames(pbulk)
degs <- as.data.table(degs,keep.rownames=T)
setnames(degs,"rn","gene")
```

### Main fig plot: dom12199 (srt)
```{r}
b <- melt.data.table(degs,id.vars="gene")
b[,variable:=as.character(variable)]
b <- b[gene %in% c("PLPP3","NTSR2","FGFR3")&variable %in% grep(variable,pattern="LC_Br",value=T)]
b[,titleg:=paste0("*",gene,"*")]
    
# drop existing rn from sample table if present
tmpsampinf <- sampinf
tmpsampinf$rn <- NULL
s2 <- as.data.table(tmpsampinf,keep.rownames=T)
s2[,YRIcat:=ifelse(YRIcat=="YRI","AA","EA")]
b <- merge.data.table(b,s2,by.x="variable",by.y="rn")

b[,lab_APOE_PredomAnces:=gsub(lab_APOE_PredomAnces,pattern="^LC_",replacement="")]
b[,lab_APOE_PredomAnces:=gsub(lab_APOE_PredomAnces,pattern="_",replacement=" ")]
b[,lab_APOE_PredomAnces:=gsub(lab_APOE_PredomAnces,pattern="YRI",replacement="AA")]
b[,lab_APOE_PredomAnces:=gsub(lab_APOE_PredomAnces,pattern="CEU",replacement="EA")]
b[,lab_APOE_PredomAnces:=factor(lab_APOE_PredomAnces,levels=c("E2 AA","E4 AA","E2 EA","E4 EA"))]

plt <- ggplot(b,aes(x=lab_APOE_PredomAnces,y=value,fill=YRIcat))+
  geom_boxplot(outlier.size = 0.4)+
  facet_wrap(~titleg,ncol=3)+
  labs(x="Ancestry",y=paste0("Expression in ISS LC Domain<br>(Age and Sex adjusted)"),fill="Predominant\nAncestry")+
  theme(
    axis.title.x=element_text(size=8),
    axis.text.x=element_text(size=7,angle=90,vjust=0.5,hjust=1),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.title.y=element_markdown(size=8),
    legend.text = element_markdown(size=7,margin=margin(0,0,0,0,"in")),
    legend.title=element_text(size=7,hjust=0.5),
    legend.background = element_blank(),
    legend.box = element_blank(),
    legend.box.spacing = ggplot2::unit(0,"in"),
    strip.text = element_markdown(size=9),
    strip.background = element_blank(),
    strip.background.y = element_blank(),
    strip.background.x=element_blank(),
    #legend.key = element_blank(),
    legend.key.height=ggplot2::unit(0.0125,"in"),
    legend.key.width=ggplot2::unit(0.0125,"in"),
    legend.key.spacing.y = ggplot2::unit(0.04,"in"),
    title=element_markdown(size=9,hjust=0.5))
    # legend.position.inside = c(0.19,0.165),
    # legend.position="inside")

pdf("plots_newseg/manuscript/Fig7/7B_LCpbulk_xpr_SRT12199.pdf",height=2.25,width=3.5)
print(plt)
dev.off()

rm(adjpbulks,domde,NTSRplts,pbulks,PLPplts,sampinf,srtdomde,wddomde)
gc(full=T)
```

## and then, banksy clusters in the same samples
```{r}
bcl <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_assignmentsMergeClusts.txt")
banno <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_mergeClusts_Annots.txt")
bcl <- merge.data.table(bcl,banno,by="clusid")

bcl[anno=="Astro_RYR3",anno:="Astro *RYR3*"]
bcl[anno=="Astro_FGFR3",anno:="Astro *FGFR3*"]
bcl[anno=="Astro_SERPINA3",anno:="Astro *SERPINA3*"]
bcl[anno=="Inhib_5HTR2A",anno:="GABA *5HTR2A*"]
bcl[,anno:=gsub(anno,pattern="_",replacement=" ")]
bcl[anno=="Endothelium",anno:="Endothel."]
bcl[anno=="Oligodendrocytes",anno:="Oligodend."]
bcl[anno=="NK and T cells",anno:="NK/T Cells"]
bcl[anno=="Serotonergic",anno:="5HTergic"]
bcl[anno=="Mixed Neuropept.",anno:="Mixed Npept."]

bankpan <- xlc[,xlc$brnum=="Br5941"]
bcl <- DataFrame(bcl,row.names=bcl$rn)[colnames(bankpan),]
bankpan$anno <- bcl$anno
bankpan <- bankpan[,bankpan$anno!="DISCARD"]
### palette:
# 1 - blood
# 2 - OPC
# 3 - Mixed Npept.
# 4 Astro_FGFR3
# 5 - Microglia
# 6 NK/T
# 7 Oligo (sub gray)
# 8 GABA 5htr2a
# 9 Astro RYR3
# 10 astro serpina3
# 11 ependyma
# 12 - vasc smc
# 13 - lcne
# 14 - endothel
# 15 - 5HTergic

pal <- Polychrome::palette36.colors(n=22)[c(3:17)]
names(pal) <- c("Blood","OPC","Mixed Npept.","Astro *FGFR3*",
                "Microglia","NK/T Cells","Oligodend.","GABA *5HTR2A*",
                "Astro *RYR3*","Astro *SERPINA3*","Ependyma","Vasc. SMC",
                "LCNE","Endothel.","5HTergic")

        

pal[which(names(pal)=="Oligodend.")] <- "#BCBCBC"

bankpan$anno <- factor(bankpan$anno,levels=c("LCNE","Astro *FGFR3*","Astro *RYR3*",
                                             "Astro *SERPINA3*","Ependyma","Microglia",
                                             "5HTergic","GABA *5HTR2A*","Mixed Npept.",
                                             "Oligodend.", "OPC",
                                             "Blood","Endothel.","Vasc. SMC","NK/T Cells"))


## plot it
dompoly12[dompol=="LC",dompol:="LCNE"] # to use the banksy palette for LC domain boundary color

plt <- make_escheR(bankpan,y_reverse=FALSE)
plt <- plt |> add_fill("anno",size=0.2375,point_size = 0.2375)
plt <- plt + 
  geom_path(data=dompoly12,aes(x=xpol,y=ypol,group=dompol,col=dompol))+
  scale_color_manual(values=pal)

## assign spot colors based on our existing palette
p <- plt +
  labs(fill="Banksy Cluster")+
  scale_fill_manual(values=pal)+
  guides(fill=guide_legend(override.aes=list(size=2)),color="none")+
  theme(legend.text=element_markdown(size=7),
        legend.title=element_text(size=7),
        legend.key.size=unit(0.2,"cm"))

pdf("plots_newseg/manuscript/Fig7/7C-BanksyClusters_Br5941.pdf",width=2.5,height=1.5)
print(p)
dev.off()

## an inset of above with LC domain boundary and cluster labels
plt <- make_escheR(bankpan[,colnames(bankpan) %in% zoomcoord$rn],y_reverse=FALSE)
plt <- plt |> add_fill("anno",size=0.325,point_size = 0.325)
plt <- plt + 
  geom_path(data=dompoly12,aes(x=xpol,y=ypol,group=dompol,col=dompol))+
  scale_color_manual(values=pal)+
  scale_fill_manual(values=pal)+
  guides(color="none",fill=guide_legend(override.aes=list(size=2)))+
  theme(legend.text=element_markdown(size=7),
        legend.title=element_text(size=7),
        legend.key.size=unit(0.2,"cm"))

pdf("plots_newseg/manuscript/Fig7/7C-BanksyClusters_inSRT12199inset_Br5941.pdf",width=2.5,height=1.5)
print(plt)
dev.off()

rm(plt)
```

### now, what about DE in astrocyte clusters themselves? First, all astrocytes combined, within domain:
```{r}
## single astro cluster DE, cross-anc and single-anc
inlc.1astde <- readRDS("processed-data_newseg/13_BanksyDE/06-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_BksyClus_collapsedAstros_inSRTlcDoms_or_wholetiss.RDS")

### E2-E4, Age Sex YRI controlled ### 
inlc.1astde.alldon <- inlc.1astde$intraLC12199$E2E4label$voomLFt$EList$E
keepn <- grep(colnames(inlc.1astde.alldon),pattern="^All_Astrocytes_Br",value=T)
inlc.1astde.alldon <- inlc.1astde.alldon[,keepn]
inlc.1ast.alldon.samps <- inlc.1astde$intraLC12199$E2E4label$dgel$samples[keepn,]
## cleaning y

inlc.1ast.alldon.samps$clus_E2E4 <- factor(inlc.1ast.alldon.samps$clus_E2E4)
inlc.1ast.alldon.samps$Sex <- factor(inlc.1ast.alldon.samps$Sex)
inlc.1ast.alldon.samps$Age <- as.numeric(inlc.1ast.alldon.samps$Age)
inlc.1ast.alldon.samps$YRI <- as.numeric(inlc.1ast.alldon.samps$YRI)
mod.alldon <-  model.matrix(~0+clus_E2E4+Age+Sex+YRI,data = inlc.1ast.alldon.samps)
degs.alldon <- jaffelab::cleaningY(inlc.1astde.alldon,mod.alldon,3) # correct for terms that arent in the ancestry-label groupings (term 1 intercept, 2 APOE genos)

degs.alldon <- as.data.table(degs.alldon,keep.rownames=T)
setnames(degs.alldon,"rn","gene")
degs.alldon[,contrast:="Ancestry-Controlled"]

# organize for plotting
inlc.1ast.alldon.samps <- as.data.table(inlc.1ast.alldon.samps,keep.rownames=T)
setnames(inlc.1ast.alldon.samps,"rn","pbulkid")

degs.alldon <- melt.data.table(degs.alldon,id.vars = c("gene","contrast"))
degs.alldon[,variable:=as.character(variable)]
degs.alldon <- merge.data.table(degs.alldon,inlc.1ast.alldon.samps[,c("pbulkid","APOEsimple")],by.x="variable",by.y="pbulkid")
degs.alldon <- degs.alldon[gene %in% c("PLPP3","FGFR3","NTSR2")]
degs.alldon[,titleg:=paste0("*",gene,"*")]

allast.pal <- c(E2="#009E73",E4="#D55E00") # to match fig 6 E2/E4 palette

p.allast.alldon <- ggplot(degs.alldon,aes(x=APOEsimple,y=as.numeric(value),fill=APOEsimple))+
  geom_boxplot(outlier.size = 0.4)+
  facet_wrap(~titleg,ncol=3)+
  labs(x="APOE Genotype",y=paste0("Expression in All ISS LC Domain<br>Astrocytes (Age, Sex, %AA adjusted)"),fill="APOE Status")+
  scale_fill_manual(values=allast.pal)+
  guides(fill="none")+ ## unnecess
  theme(
    axis.title.x=element_text(size=8),
    axis.text.x=element_text(size=7,angle=90,vjust=0.5,hjust=1),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.title.y=element_markdown(size=8),
    legend.text = element_markdown(size=7,margin=margin(0,0,0,0,"in")),
    legend.title=element_text(size=7,hjust=0.5),
    legend.background = element_blank(),
    legend.box = element_blank(),
    legend.box.spacing = ggplot2::unit(0,"in"),
    strip.text = element_markdown(size=9),
    strip.background = element_blank(),
    strip.background.y = element_blank(),
    strip.background.x=element_blank(),
    #legend.key = element_blank(),
    legend.key.height=ggplot2::unit(0.0125,"in"),
    legend.key.width=ggplot2::unit(0.0125,"in"),
    legend.key.spacing.y = ggplot2::unit(0.04,"in"),
    title=element_markdown(size=9,hjust=0.5))

pdf("plots_newseg/manuscript/Fig7/7D_AllSRT12199_Astro_AllDonorE2E4_boxplot.pdf",height=2.25,width=2.25)
print(p.allast.alldon)
dev.off()

rm(degs.alldon,mod.alldon,inlc.1ast.alldon.samps,inlc.1astde.alldon,p.allast.alldon,allast.pal)
gc(full=T)
```


### All astrocyte, Ancestry-Stratified E2-E4, Age Sex YRI controlled ### 
```{r}
inlc.1astde.ancstrat <- inlc.1astde$intraLC12199$YRIAPOlabel$voomLFt$EList$E
keepn <- grep(colnames(inlc.1astde.ancstrat),pattern="^All_Astrocytes_Br",value=T)
inlc.1astde.ancstrat <- inlc.1astde.ancstrat[,keepn]
inlc.1ast.ancstrat.samps <- inlc.1astde$intraLC12199$YRIAPOlabel$dgel$samples[keepn,]

inlc.1ast.ancstrat.samps$lab_APOE_PredomAnces <- 
  factor(inlc.1ast.ancstrat.samps$lab_APOE_PredomAnces,
  levels = unique(inlc.1ast.ancstrat.samps$lab_APOE_PredomAnces))
inlc.1ast.ancstrat.samps$Sex <- factor(inlc.1ast.ancstrat.samps$Sex)
inlc.1ast.ancstrat.samps$Age <- as.numeric(inlc.1ast.ancstrat.samps$Age)
mod.ancstrat <-  model.matrix(~0+lab_APOE_PredomAnces+Age+Sex,data = inlc.1ast.ancstrat.samps)
degs.ancstrat <- jaffelab::cleaningY(inlc.1astde.ancstrat,mod.ancstrat,5) # correct for terms that arent in the ancestry-label groupings (term 1 intercept, 4 ancestry-APOE groups)


degs.ancstrat <- as.data.table(degs.ancstrat,keep.rownames=T)
setnames(degs.ancstrat,"rn","gene")
degs.ancstrat[,contrast:="Ancestry-Stratified"]

# organize for plotting
degs.ancstrat <- melt.data.table(degs.ancstrat,id.vars = c("gene","contrast"))
degs.ancstrat[,variable:=as.character(variable)]
inlc.1ast.ancstrat.samps <- as.data.table(inlc.1ast.ancstrat.samps,keep.rownames=T)
degs.ancstrat <- merge.data.table(degs.ancstrat,inlc.1ast.ancstrat.samps[,c("rn","lab_APOE_PredomAnces","YRIcat")],by.x="variable",by.y="rn")
degs.ancstrat[,lab_APOE_PredomAnces:=as.character(lab_APOE_PredomAnces)]
degs.ancstrat[,lab_APOE_PredomAnces:=gsub(lab_APOE_PredomAnces,pattern="All_Astrocytes_",replacement="")]
degs.ancstrat[,lab_APOE_PredomAnces:=gsub(lab_APOE_PredomAnces,pattern="_",replacement=" ")]
degs.ancstrat[,lab_APOE_PredomAnces:=gsub(lab_APOE_PredomAnces,pattern="YRI",replacement="AA")]
degs.ancstrat[,lab_APOE_PredomAnces:=gsub(lab_APOE_PredomAnces,pattern="CEU",replacement="EA")]
degs.ancstrat[,lab_APOE_PredomAnces:=factor(lab_APOE_PredomAnces,levels=c("E2 AA","E4 AA","E2 EA","E4 EA"))]
degs.ancstrat[,YRIcat:=ifelse(YRIcat=="CEU","EA","AA")]
degs.ancstrat <- degs.ancstrat[gene %in% c("FGFR3","PLPP3","NTSR2")]
degs.ancstrat[,titleg:=paste0("*",gene,"*")]
### plots ###
p.allast.ancstrat <- ggplot(data=degs.ancstrat,aes(x=lab_APOE_PredomAnces,y=value,fill=YRIcat))+
  geom_boxplot(outlier.size = 0.4)+
  facet_wrap(~titleg,nrow=1)+
  labs(x="Gene",y=paste0("Expression in All ISS LC Domain<br>Astrocytes (Age, Sex adjusted)"),fill="APOE-Ancestry Group")+
  guides(fill="none")+ # we already have a copy of this in another panel
  theme(
    axis.title.x=element_text(size=8),
    axis.text.x=element_text(size=7,angle=90,vjust=0.5,hjust=1),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.title.y=element_markdown(size=8),
    legend.text = element_markdown(size=7,margin=margin(0,0,0,0,"in")),
    legend.title=element_text(size=7,hjust=0.5),
    legend.background = element_blank(),
    legend.box = element_blank(),
    legend.box.spacing = ggplot2::unit(0,"in"),
    strip.text = element_markdown(size=9),
    strip.background = element_blank(),
    strip.background.y = element_blank(),
    strip.background.x=element_blank(),
    #legend.key = element_blank(),
    legend.key.height=ggplot2::unit(0.0125,"in"),
    legend.key.width=ggplot2::unit(0.0125,"in"),
    legend.key.spacing.y = ggplot2::unit(0.04,"in"),
    title=element_markdown(size=9,hjust=0.5))

pdf("plots_newseg/manuscript/Fig7/7D_AllAstr_intraSRT12199_AncestryStratDE.pdf",height=2.25,width=2.25)
print(p.allast.ancstrat)
dev.off()
```

## Aaaand within domain 12199, by astrocyte cluster (and ancestry)
```{r}
bsyde <- readRDS("processed-data_newseg/13_BanksyDE/04-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_BksyClus_inSRTlcDoms_or_wholetiss.RDS")

intra12detab <- bsyde$intraLC12199$YRIAPOlabel$detabs
wholetissdetab <- bsyde$wholetissue$YRIAPOlabel$detabs

emats <- list(intraLC12199=bsyde$intraLC12199$YRIAPOlabel$voomLFt$EList$E,
              wholetissue=bsyde$wholetissue$YRIAPOlabel$voomLFt$EList$E)
metas <- list(bsyde$intraLC12199$YRIAPOlabel$dgel$samples,
              bsyde$wholetissue$YRIAPOlabel$dgel$samples)

cleanys <- mapply(e=emats,m=metas,SIMPLIFY=FALSE,function(e,m){
  astrocols <- grep(colnames(e),pattern="Astro_FGFR3|Astro_RYR3",value=T)
  e <- e[,astrocols]
  m <- m[astrocols,]
  
  m$lab_APOE_PredomAnces <- factor(m$lab_APOE_PredomAnces,levels = unique(m$lab_APOE_PredomAnces))
  m$Sex <- factor(m$Sex)
  m$Age <- as.numeric(m$Age)

  mod <- model.matrix(~0+lab_APOE_PredomAnces+Age+Sex,data = m)
  cleannum <- length(unique(m$lab_APOE_PredomAnces))+1
  cleanexp <- jaffelab::cleaningY(e,mod,cleannum) # correct for terms that arent in the ancestry-label groupings (term 1 intercept + 4 levels of lab_APOE_ancest * 2 clusters, 6 Age, 7 Sex)

  cleanexp <- as.data.table(cleanexp,keep.rownames=T)
  cleanexp.melt <- melt.data.table(cleanexp,id.vars="rn")
  cleanexp.melt <- cleanexp.melt[rn %in% c("FGFR3","PLPP3","NTSR2")]
  cleanexp.melt[,brnum:=gsub(variable,pattern="^.*_(Br.*)$",replacement="\\1")]
  cleanexp.melt[,clus:=gsub(variable,pattern="^(.*)_Br.*$",replacement="\\1")]
  setnames(cleanexp.melt,"rn","gene")
  return(cleanexp.melt)
})
names(cleanys) <- names(emats)

cleanys <- rbindlist(cleanys,idcol="Tissue_Region")
sampmd <- as.data.table(bsyde$wholetissue$YRIAPOlabel$dgel$samples)
sampmd <- unique(sampmd[,.(brnum,APOEsimple,YRIcat)])
cleanys <- merge.data.table(cleanys,sampmd,by="brnum")
cleanys[YRIcat=="CEU",YRIcat:="EA"]
cleanys[YRIcat=="YRI",YRIcat:="AA"]
cleanys[,APO_Anc:=paste0(APOEsimple," ",YRIcat)]
cleanys[,APO_Anc:=factor(APO_Anc,levels=c("E2 AA","E4 AA","E2 EA","E4 EA"))]
cleanys[,titleg:=paste0("*",gene,"*")]

### intraLC
intraLC.ancstrat.byastclus <- lapply(unique(cleanys$clus),function(x){
  p <- ggplot(cleanys[Tissue_Region=="intraLC12199" & gene %in% c("FGFR3","PLPP3","NTSR2")&clus==x],aes(x=APO_Anc,y=as.numeric(value),fill=YRIcat))+
    facet_wrap(~titleg,nrow=1)+
    geom_boxplot(outlier.size = 0.4)+
    labs(x="APOE Genotype",
         y=paste0("Expression in ",
          gsub(x,pattern="Astro_",replacement="*"),"* Astrocytes<br>of ISS LC Domain (Age, Sex adjusted)"),fill="APO_Anc")+
    guides(fill="none")+ # use from preceding ancestry-strat plot
    theme(
    axis.title.x=element_text(size=8),
    axis.text.x=element_text(size=7,angle=90,vjust=0.5,hjust=1),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.title.y=element_markdown(size=8),
    legend.text = element_markdown(size=7,margin=margin(0,0,0,0,"in")),
    legend.title=element_text(size=7,hjust=0.5),
    legend.background = element_blank(),
    legend.box = element_blank(),
    legend.box.spacing = ggplot2::unit(0,"in"),
    strip.text = element_markdown(size=9),
    strip.background = element_blank(),
    strip.background.y = element_blank(),
    strip.background.x=element_blank(),
    #legend.key = element_blank(),
    legend.key.height=ggplot2::unit(0.0125,"in"),
    legend.key.width=ggplot2::unit(0.0125,"in"),
    legend.key.spacing.y = ggplot2::unit(0.04,"in"),
    title=element_markdown(size=9,hjust=0.5))

})

### save em
pdf("plots_newseg/manuscript/Fig7/7D_intraSRT12199_ancstrat_byAstroClus.pdf",height=4.75,width=2.25)
do.call("grid.arrange",c(intraLC.ancstrat.byastclus,nrow=2))
dev.off()
```

### last, not a cell type density effect:
```{r}
cpa <- readRDS("processed-data_newseg/14_CellTypeProximity/03_BsyLabelCounts_per_LCDomainArea.RDS")

cpa <- rbindlist(cpa,idcol="LCDomainDef")
## get metadata for all samples using the spe
tmpcd <- as.data.table(colData(xlc),keep.rownames=T)
tmpcd <- unique(tmpcd[,.(brnum,APOE,Sex,Age,YRI)])
tmpcd <- tmpcd[greaterOrEqual(YRI,0.75)|lessOrEqual(YRI,0.25)]
tmpcd[,catanc:=ifelse(YRI>0.5,"AA","EA")]
tmpcd[,e2e4:=ifelse(APOE %in% c("E4hom","E4het"),"E4","E2")]
tmpcd[,APOE_Anc:=paste0(e2e4," ",catanc)]
cpa <- merge.data.table(cpa,tmpcd,"brnum")
cpa[,APOE_Anc:=factor(APOE_Anc,levels=c("E2 AA","E4 AA","E2 EA","E4 EA"))]

# convert from per unit^2 to per 100 units^2
cpa[,annoPerLCdomArea:=100*annoPerLCdomArea]
cpa <- cpa[LCDomainDef=="srt_domain12199" & anno %in% c("Astro_RYR3","Astro_FGFR3","LCNE")]
cpa[,anno:=gsub(anno,pattern="^(.*)_(.*)$",replacement=paste0("\\1"," *","\\2","*"))]
## excluding LCNE cells, which are going to skew the y axis here


### version 1: boxplot by apoeanc group, colored by cell type
p2 <- ggplot(cpa[anno!="LCNE"],aes(x=APOE_Anc,y=annoPerLCdomArea,fill=anno))+
  geom_boxplot()+
  labs(y="Cells per 100 units^2<br>of ISS LC domain area",x="APOE-Ancestry Group",fill="Cell Type")+
  theme(axis.text.x=element_markdown(size=8,angle=90,hjust=0.5,vjust=1))+
  scale_fill_manual(values=pal)+
  theme(axis.text.x=element_markdown(size=7,angle=90,hjust=0.5,vjust=1),
    axis.title.x=element_text(size=8),
    axis.title.y=element_markdown(size=8),
    axis.text.y=element_text(size=7),
    legend.text = element_markdown(size=7,margin=margin(0,0,0,0,"in")),
    legend.title=element_markdown(size=7,hjust=0.5),
    legend.background = element_blank(),
    legend.box = element_blank(),
    legend.box.spacing = ggplot2::unit(0,"in"),
    title=element_markdown(size=9,hjust=0.5))

pdf("plots_newseg/manuscript/Fig7/7E_intraSRT12199_cellTypeDensity_by_Anc.pdf",width=2.5,height=2)
print(p2)
dev.off()


### version 2: boxplot by cell type
## excluding LCNE cells, which are going to skew the y axis here
cpa[,anno:=gsub(x=anno,pattern=" ",replacement="<br>")]

p <- ggplot(cpa[anno!="LCNE"],aes(x=anno,y=annoPerLCdomArea,fill=APOE_Anc))+
  geom_boxplot()+
  labs(y="Cells per 100 units^2<br>of ISS LC domain area",
    fill="APOE-Ancestry<br>Group",
    x="Cell Type")+
  theme(axis.text.x=element_markdown(size=7,angle=90,hjust=0.5,vjust=1),
    legend.text = element_markdown(size=7,margin=margin(0,0,0,0,"in")),
    axis.title.y=element_markdown(size=8),
    axis.text.y=element_text(size=7),
    axis.title.x=element_text(size=8),
    legend.title=element_markdown(size=7,hjust=0.5),
    legend.background = element_blank(),
    legend.box = element_blank(),
    legend.box.spacing = ggplot2::unit(0,"in"),
    title=element_markdown(size=9,hjust=0.5))
pdf("plots_newseg/manuscript/Fig7/7E_intraSRT12199_cellTypeDensity_by_Celltype.pdf",width=2.5,height=2)
print(p)
dev.off()

```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()

```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices
[5] utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0        
 [2] parallelly_1.47.0          
 [3] sf_1.0-24                  
 [4] jaffelab_0.99.34           
 [5] rafalib_1.0.4              
 [6] escheR_1.10.0              
 [7] ggrastr_1.0.2              
 [8] SpatialExperiment_1.20.0   
 [9] SingleCellExperiment_1.32.0
[10] SummarizedExperiment_1.40.0
[11] Biobase_2.70.0             
[12] GenomicRanges_1.62.1       
[13] Seqinfo_1.0.0              
[14] IRanges_2.44.0             
[15] S4Vectors_0.48.1           
[16] BiocGenerics_0.56.0        
[17] generics_0.1.4             
[18] MatrixGenerics_1.22.0      
[19] matrixStats_1.5.0          
[20] Polychrome_1.5.4           
[21] gridExtra_2.3              
[22] ggtext_0.1.2               
[23] ggplot2_4.0.2              
[24] data.table_1.18.4          

loaded via a namespace (and not attached):
 [1] DBI_1.2.3            rlang_1.1.7         
 [3] magrittr_2.0.4       otel_0.2.0          
 [5] e1071_1.7-17         compiler_4.5.2      
 [7] systemfonts_1.3.1    vctrs_0.7.1         
 [9] stringr_1.6.0        pkgconfig_2.0.3     
[11] magick_2.9.0         XVector_0.50.0      
[13] labeling_0.4.3       markdown_2.0        
[15] ggbeeswarm_0.7.3     ragg_1.5.0          
[17] purrr_1.2.1          xfun_0.56           
[19] litedown_0.9         DelayedArray_0.36.1 
[21] parallel_4.5.2       R6_2.6.1            
[23] stringi_1.8.7        RColorBrewer_1.1-3  
[25] limma_3.66.0         Rcpp_1.1.1          
[27] knitr_1.51           Matrix_1.7-4        
[29] splines_4.5.2        tidyselect_1.2.1    
[31] dichromat_2.0-0.1    abind_1.4-8         
[33] codetools_0.2-20     lattice_0.22-7      
[35] tibble_3.3.1         withr_3.0.2         
[37] S7_0.2.1             evaluate_1.0.5      
[39] units_1.0-0          proxy_0.4-29        
[41] xml2_1.5.2           pillar_1.11.1       
[43] KernSmooth_2.23-26   rprojroot_2.1.1     
[45] commonmark_2.0.0     scales_1.4.0        
[47] class_7.3-23         glue_1.8.0          
[49] scatterplot3d_0.3-44 tools_4.5.2         
[51] locfit_1.5-9.12      fs_1.6.6            
[53] grid_4.5.2           edgeR_4.8.2         
[55] colorspace_2.1-2     nlme_3.1-168        
[57] beeswarm_0.4.0       googledrive_2.1.2   
[59] vipor_0.4.7          cli_3.6.5           
[61] textshaping_1.0.4    gargle_1.6.1        
[63] segmented_2.2-1      S4Arrays_1.10.1     
[65] viridisLite_0.4.3    dplyr_1.2.0         
[67] gtable_0.3.6         classInt_0.4-11     
[69] SparseArray_1.10.8   rjson_0.2.23        
[71] farver_2.1.2         lifecycle_1.0.5     
[73] here_1.0.2           statmod_1.5.1       
[75] gridtext_0.1.5       MASS_7.3-65         
> sessioninfo::session_info()
─ Session info ──────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       Positron
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-07-11
 pandoc   NA
 quarto   1.9.38 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 class                  7.3-23   2025-01-01 [1] CRAN (R 4.5.2)
 classInt               0.4-11   2025-01-08 [1] CRAN (R 4.5.0)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorspace             2.1-2    2025-09-22 [1] CRAN (R 4.5.0)
 commonmark             2.0.0    2025-07-07 [1] CRAN (R 4.5.0)
 data.table           * 1.18.4   2026-05-06 [1] CRAN (R 4.5.2)
 DBI                    1.2.3    2024-06-02 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.1   2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 e1071                  1.7-17   2025-12-18 [1] CRAN (R 4.5.2)
 edgeR                  4.8.2    2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 escheR               * 1.10.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fs                     1.6.6    2025-04-12 [1] CRAN (R 4.5.0)
 gargle                 1.6.1    2026-01-29 [1] CRAN (R 4.5.2)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 ggbeeswarm             0.7.3    2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggrastr              * 1.0.2    2023-06-01 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 googledrive            2.1.2    2025-09-10 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 jaffelab             * 0.99.34  2026-02-24 [1] Github (LieberInstitute/jaffelab@e7a26ca)
 KernSmooth             2.23-26  2025-01-01 [1] CRAN (R 4.5.2)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 labeling               0.4.3    2023-08-29 [1] CRAN (R 4.5.0)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 limma                  3.66.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 litedown               0.9      2025-12-18 [1] CRAN (R 4.5.2)
 locfit                 1.5-9.12 2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 markdown               2.0      2025-03-23 [1] CRAN (R 4.5.0)
 MASS                   7.3-65   2025-02-28 [1] CRAN (R 4.5.2)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 nlme                   3.1-168  2025-03-31 [1] CRAN (R 4.5.2)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.47.0   2026-04-17 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 Polychrome           * 1.5.4    2025-04-06 [1] CRAN (R 4.5.0)
 proxy                  0.4-29   2025-12-29 [1] CRAN (R 4.5.2)
 purrr                  1.2.1    2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 rafalib              * 1.0.4    2025-04-08 [1] CRAN (R 4.5.0)
 ragg                   1.5.0    2025-09-02 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1   2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 scatterplot3d          0.3-44   2023-05-05 [1] CRAN (R 4.5.0)
 segmented              2.2-1    2026-01-29 [1] CRAN (R 4.5.2)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 sf                   * 1.0-24   2026-01-13 [1] CRAN (R 4.5.2)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 statmod                1.5.1    2025-10-09 [1] CRAN (R 4.5.0)
 stringi                1.8.7    2025-03-27 [1] CRAN (R 4.5.0)
 stringr                1.6.0    2025-11-04 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 systemfonts            1.3.1    2025-10-01 [1] CRAN (R 4.5.0)
 textshaping            1.0.4    2025-10-10 [1] CRAN (R 4.5.0)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 units                  1.0-0    2025-10-09 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.5.0)
 viridisLite            0.4.3    2026-02-04 [1] CRAN (R 4.5.2)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────