---
title: "03-TestLCsmoothing_knnstep2_k75"
author: "Bernie Mulvey"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(FastKNN)
library(caTools)
library(FNN)

# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallel)


## ggplot defaults
## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load KNN step 1 results
```{r}
labxf <- readRDS("processed-data/00_prelim_procdat_BJM/05_LabelTransfer/02-TestLCknnsmoothing_step1k20.RDS")

## split back to samplewise
labxf_list <- split(labxf,labxf$brnum)
rm(labxf)

## second smoothing
```

## step 2 knn smoothing of LC with k=75 (200 makes big blotches)
#### note that the data.table-internalized sapply call is not the absolute fastest way to run that step (but still only takes a matter of seconds at scale here).
#### however, this approach is more transparent in that the user can take an entry of xyspe_list into the environment as v and step through the bplapply to see what's being done here, and can see that cells are always matched in the process (as opposed to working with vectors where we don't have cell ids matched up anymore. which always makes me unccomfortable, however rationally/irrationally.)
```{r}
setDTthreads(1,restore_after_fork = FALSE)
thresh1 <- seq(0.05,0.5,by=0.05) # thresholds above 0.5 pretty much negate LC from most samples

perthresh <- lapply(thresh1,function(t){
  knnreslist <- mclapply(labxf_list,mc.cores=8,FUN=function(v){
    ## reset LC designation based on KNN step 1 and current threshold for KNN1 LC designation
    v[,isLC:=NULL]
    v[,isLC:=ifelse(lcpred.smooth1.k20>=t,1,0)]
    # get nearest neighbors for each cell in the sample
    knnout <- FNN::get.knn(v[,c("xcoord","ycoord")],k = 75) # previously tried with 200, but this designates probably overlarge areas as LC
    
    # get the row indices of those k cells
    neighborslist <- apply(knnout$nn.index,MARGIN=1,FUN=list)
    # goofily, this is returns each entry as a list within a list (e.g., [[1]][[1]], [[2]][[1]], etc.). so fix that:
    neighborslist <- lapply(neighborslist,FUN=unlist)
    # make the list names the cell identifiers as a failsafe against reordering; we will use the list entry names to retrieve the neighbor indices when calculating means below
    names(neighborslist) <- v$rn
    
    # now, get the mean of LC y/n for the k nearest neighbors indexed for each cell
    # (in knnout, the first list item, nn.index, has v's row indices corresponding to those neighbors) 
    # this value is thus a probability 0-1 that, a given cell is a LC type given the LC/nonLC identity of its 20 spatial neighbors
    v[,lcpred.smooth2.k75:=
          apply(.SD, MARGIN = 1, FUN=function(x){
              mean(v[neighborslist[[x]],isLC])}
              )
      ,.SDcols = "rn"]
    return(v)
  })
  return(knnreslist)
})
names(perthresh) <- paste0("k1thresh_",thresh1)


## concatenate back into a single full-dataset table of KNN smoothing predictions per KNN step 1 threshold
perthresh2 <- lapply(perthresh,rbindlist)
perthresh2 <- lapply(perthresh2,as.data.table)
## get rid of some columns no longer needed
perthresh <- lapply(perthresh2,function(x){
  x[,xcoord:=NULL]
  x[,ycoord:=NULL]
  x[,isLC:=NULL]
  
})

saveRDS(perthresh2,"processed-data/00_prelim_procdat_BJM/05_LabelTransfer/03-TestLCknnsmoothing_step2k75.RDS")
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0         parallelly_1.46.1          
 [3] colorout_1.3-3              FNN_1.1.4.1                
 [5] caTools_1.18.3              FastKNN_0.0.1              
 [7] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
 [9] SummarizedExperiment_1.40.0 Biobase_2.70.0             
[11] GenomicRanges_1.62.1        Seqinfo_1.0.0              
[13] IRanges_2.44.0              S4Vectors_0.48.0           
[15] BiocGenerics_0.56.0         generics_0.1.4             
[17] MatrixGenerics_1.22.0       matrixStats_1.5.0          
[19] gridExtra_2.3               ggtext_0.1.2               
[21] ggplot2_4.0.1               data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] gtable_0.3.6        rjson_0.2.23        xfun_0.56           lattice_0.22-7     
 [5] vctrs_0.7.1         tools_4.5.2         bitops_1.0-9        tibble_3.3.1       
 [9] pkgconfig_2.0.3     Matrix_1.7-4        RColorBrewer_1.1-3  S7_0.2.1           
[13] assertthat_0.2.1    lifecycle_1.0.5     compiler_4.5.2      farver_2.1.2       
[17] pdist_1.2.1         codetools_0.2-20    htmltools_0.5.9     yaml_2.3.12        
[21] pillar_1.11.1       rsconnect_1.7.0     DelayedArray_0.36.0 sessioninfo_1.2.3  
[25] magick_2.9.0        abind_1.4-8         tidyselect_1.2.1    digest_0.6.39      
[29] dplyr_1.1.4         rprojroot_2.1.1     fastmap_1.2.0       grid_4.5.2         
[33] here_1.0.2          cli_3.6.5           SparseArray_1.10.8  magrittr_2.0.4     
[37] S4Arrays_1.10.1     dichromat_2.0-0.1   withr_3.0.2         scales_1.4.0       
[41] rmarkdown_2.30      XVector_0.50.0      otel_0.2.0          evaluate_1.0.5     
[45] knitr_1.51          rlang_1.1.7         gridtext_0.1.5      Rcpp_1.1.1         
[49] glue_1.8.0          xml2_1.5.2          rstudioapi_0.18.0   R6_2.6.1           
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-03
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ─────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 assertthat             0.2.1    2019-03-21 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bitops                 1.0-9    2024-10-03 [1] CRAN (R 4.5.0)
 caTools              * 1.18.3   2024-09-04 [1] CRAN (R 4.5.0)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.1.4    2023-11-17 [1] CRAN (R 4.5.0)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 FastKNN              * 0.0.1    2015-02-12 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 FNN                  * 1.1.4.1  2024-09-22 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggplot2              * 4.0.1    2025-11-14 [1] CRAN (R 4.5.2)
 ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] CRAN (R 4.5.2)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pdist                  1.2.1    2022-05-02 [1] CRAN (R 4.5.0)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

────────────────────────────────────────────────────────────────────────────────────