---
title: "05-NMFlabeltransfer_VisToXen"
format: html
date: 2026-01-26
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())


## one time install:github RcppML--has an nmf object class not in CRAN version, but required in singlet 
## spatransfer dependency zdebruine/singlet
# github cindyfang70/spaTransfer (actually called nmfLabelTransfer)

# devtools::install_github('zdebruine/RcppML')
# devtools::install_github('zdebruine/singlet')
# devtools::install_github('cindyfang70/spaTransfer')

library(data.table)
library(RcppML)
library(SingleCellExperiment)
library(SpatialExperiment)
library(Matrix) # otherwise rcppml gives error 'no item called "package:Matrix" on the search list'


## rstudio GUI tweaks
# require(colorout)
# ColorOut()
options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
##

## do not load parallelization libraries here. rcppml is installed from source with OpenMP support (i hope) so lets not overcomplicate the environment.

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```


load Visium data with final labels
```{r}
lcv <- readRDS("../LFF_spatial_LC/processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

vdom <- readRDS("../LFF_spatial_LC/processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
vdom <- vdom$HARMONYlmbna_HDG_SVG_2575
vdom <- vdom[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(vdom,2,"clusid")

vanno <- fread("../LFF_spatial_LC/processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")
vdom <- merge.data.table(vdom,vanno[,.(clusid,anno)],by="clusid")
vdom <- DataFrame(vdom,row.names=vdom$rn)[colnames(lcv),]

colLabels(lcv) <- vdom$anno

rm(vdom,vanno)

unique(lcv$brnum)
lcv$brnum[lcv$brnum=="Br6119(re-dis)"] <- "Br6119"
## drop donors with limited LC and br 1691 (tentatively dropping from Xenium)
lcv <- lcv[,!(lcv$brnum %in% paste0("Br",c(1691,5517,5276,5712)))]


### we need logcounts on here to perform NMF
lcv <- scuttle::logNormCounts(lcv)
```

load xenium data
```{r}
lcx <- readRDS("processed-data/00_prelim_procdat_BJM/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")
unique(lcx$brnum) # donor WITH hemisphere
lcx$donor <- lcx$brnum
lcx$donor[lcx$donor=="Br2305L"] <- "Br2305"
lcx$donor[lcx$donor=="Br1039L"] <- "Br1039"
lcx$donor[lcx$donor=="Br5854L"] <- "Br5854"

stopifnot(all(unique(lcx$donor) %in% unique(lcv$brnum)))
stopifnot(all(unique(lcv$brnum) %in% unique(lcx$donor)))

## isoforms of XBP1 are separate probes for xenium, so we want to add a row that just aggregates the two. we recalculate logcounts during the per-sample loop below anyhow so we can get rid of the normalized counts in the xenium SPE too.
assays(lcx)[[3]] <- NULL
assays(lcx)[[2]] <- NULL
gc(full=T)

```

## add a row for XBP1 gene level to rowData and counts, remove isoforms
```{r}
newg <- as.data.frame(rowData(lcx))
newg[(nrow(newg)+1),"Symbol"] <- "XBP1"
rownames(newg)[nrow(newg)] <- "XBP1"
newg <- as.data.table(newg,keep.rownames=T)
newg[rn=="XBP1",ID:="ENSG00000100219"]
newg[rn=="XBP1",Type:="Gene Expression"]
set(newg,i=nrow(newg),j = c(5:12),value=FALSE)
newg <- newg[!(rn %in% c("XBP1u","XBP1s"))]
newg <- DataFrame(newg,row.names=newg$rn)
newg$rn <- NULL

## remove isoforms from counts matrix
newr <- t(as.matrix(colSums(counts(lcx)[c("XBP1s","XBP1u"),],na.rm=T)))
newct <- as(rbind(as.matrix(counts(lcx)),newr),"dgCMatrix")
rownames(newct)[nrow(newct)] <- "XBP1"
newct <- newct[!(rownames(newct) %in% c("XBP1u","XBP1s")),]

## make sure we have the same genes in both of these matrices, in the same order by row and col
stopifnot(all(rownames(newct)==rownames(newg)))
stopifnot(all(colnames(newct)==colnames(lcx)))

lcx2 <- SpatialExperiment(colData=colData(lcx),rowData=newg,assays=list(counts=newct),spatialCoords = spatialCoords(lcx))

## we need the Symbol column to be gene_name for spaTransfer
colnames(rowData(lcx2))[2] <- "gene_name"

rm(lcx,newct,newg,newr)
gc(full=T)
```

## nmf projection is done using the entire visium dataset per xenium SAMPLE (encoded in brnum, which encodes one tissue section as unique donor-hemisphere in this dataset). using Cindy's implementation of this from https://github.com/LieberInstitute/spatialDLPFC_SCZ_XENIUM/blob/devel/code/analysis/04_label_transfer/01_run_spaTransfer.R

the alpha parameter (elasticnet mixing parameter?) needs to be specified for this to work. based on the example scripts in the draft repo for the spatransfer manuscript, we'll use alpha=0. no clue if that's actually a suggested value or just something there for demonstration purposes though.

```{r}
lcxes <- lapply(unique(lcx2$brnum),function(x){
    lcx2[,lcx2$brnum==x]
})
names(lcxes) <- unique(lcx2$brnum)
lcxes <- lapply(lcxes,scuttle::logNormCounts)

rm(lcx2)
gc(full=T)
### 

# technicalVarName should be the sample identifier in the source(reference) dataset.
# use tol 1e-8 (vs default 1e-5) for higher stringency (passed to nmfLabelTransfer::run_nmf)
# use k=100 for 100 nmf factors
# use 8 threads; prevent data.table parallelization overcommiting threads using setDTthreads
setDTthreads(1,restore_after_fork = FALSE)
options(RcppML.verbose = TRUE)
options(RcppML.threads=8)

# run it
vis_to_xen <- transfer_labels(targets = lcxes,source = lcv,assay = "logcounts",annotationsName = "label",technicalVarName = "sample_id",seed = 42,tol=1e-8,k=100,threads=10,alpha=0) # not sure threads arg is also needed here.

saveRDS(vis_to_xen,"processed-data/00_prelim_procdat_BJM/05-VistoXen_NMFlabelxfer_tol1e8_k100_alph0.RDS")

```

sessioninfo
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] ggrastr_1.0.2               ggtext_0.1.2               
 [3] escheR_1.6.0                ggplot2_4.0.1              
 [5] Matrix_1.7-4                SpatialExperiment_1.16.0   
 [7] data.table_1.17.4           nmfLabelTransfer_0.0.0.9000
 [9] singlet_0.99.8              RcppEigen_0.3.4.0.2        
[11] dplyr_1.1.4                 RcppML_0.5.6               
[13] Seurat_5.4.0                SeuratObject_5.3.0         
[15] sp_2.2-0                    SingleCellExperiment_1.28.1
[17] SummarizedExperiment_1.36.0 Biobase_2.66.0             
[19] GenomicRanges_1.58.0        GenomeInfoDb_1.42.3        
[21] IRanges_2.40.1              S4Vectors_0.44.0           
[23] BiocGenerics_0.52.0         MatrixGenerics_1.18.1      
[25] matrixStats_1.5.0          

loaded via a namespace (and not attached):
  [1] RcppAnnoy_0.0.23        splines_4.4.3           later_1.4.4            
  [4] tibble_3.3.0            polyclip_1.10-7         fastDummies_1.7.5      
  [7] lifecycle_1.0.4         rprojroot_2.1.1         globals_0.18.0         
 [10] lattice_0.22-7          MASS_7.3-65             magrittr_2.0.4         
 [13] limma_3.62.2            plotly_4.11.0           rmarkdown_2.30         
 [16] yaml_2.3.12             httpuv_1.6.16           otel_0.2.0             
 [19] sctransform_0.4.3       spam_2.11-1             spatstat.sparse_3.1-0  
 [22] reticulate_1.44.1       cowplot_1.2.0           pbapply_1.7-4          
 [25] RColorBrewer_1.1-3      abind_1.4-8             zlibbioc_1.52.0        
 [28] Rtsne_0.17              purrr_1.2.0             msigdbr_25.1.1         
 [31] GenomeInfoDbData_1.2.13 ggrepel_0.9.6           irlba_2.3.5.1          
 [34] listenv_0.10.0          spatstat.utils_3.2-1    goftest_1.2-3          
 [37] RSpectra_0.16-2         spatstat.random_3.4-4   fitdistrplus_1.2-6     
 [40] parallelly_1.46.1       commonmark_2.0.0        codetools_0.2-20       
 [43] DelayedArray_0.32.0     xml2_1.5.0              scuttle_1.16.0         
 [46] tidyselect_1.2.1        shape_1.4.6.1           UCSC.utils_1.2.0       
 [49] farver_2.1.2            spatstat.explore_3.7-0  jsonlite_2.0.0         
 [52] progressr_0.18.0        ggridges_0.5.7          survival_3.8-3         
 [55] iterators_1.0.14        foreach_1.5.2           dbscan_1.2.4           
 [58] tools_4.4.3             ica_1.0-3               Rcpp_1.1.1             
 [61] glue_1.8.0              gridExtra_2.3           SparseArray_1.6.2      
 [64] xfun_0.56               here_1.0.2              withr_3.0.2            
 [67] fastmap_1.2.0           litedown_0.8            digest_0.6.39          
 [70] R6_2.6.1                mime_0.13               scattermore_1.2        
 [73] tensor_1.5.1            markdown_2.0            dichromat_2.0-0.1      
 [76] spatstat.data_3.1-9     tidyr_1.3.1             generics_0.1.4         
 [79] httr_1.4.7              htmlwidgets_1.6.4       S4Arrays_1.6.0         
 [82] uwot_0.2.4              pkgconfig_2.0.3         gtable_0.3.6           
 [85] lmtest_0.9-40           S7_0.2.1                XVector_0.46.0         
 [88] htmltools_0.5.8.1       dotCall64_1.2           fgsea_1.32.4           
 [91] scales_1.4.0            png_0.1-8               spatstat.univar_3.1-6  
 [94] knitr_1.51              rstudioapi_0.17.1       reshape2_1.4.5         
 [97] rjson_0.2.23            nlme_3.1-168            curl_7.0.0             
[100] zoo_1.8-14              stringr_1.6.0           KernSmooth_2.23-26     
[103] vipor_0.4.7             parallel_4.4.3          miniUI_0.1.2           
[106] pillar_1.11.1           grid_4.4.3              vctrs_0.7.1            
[109] RANN_2.6.2              promises_1.5.0          beachmat_2.22.0        
[112] xtable_1.8-4            cluster_2.1.8.1         beeswarm_0.4.0         
[115] evaluate_1.0.5          magick_2.9.0            cli_3.6.5              
[118] compiler_4.4.3          rlang_1.1.7             crayon_1.5.3           
[121] future.apply_1.20.1     labeling_0.4.3          ggbeeswarm_0.7.2       
[124] plyr_1.8.9              stringi_1.8.7           viridisLite_0.4.2      
[127] deldir_2.0-4            BiocParallel_1.40.2     assertthat_0.2.1       
[130] babelgene_22.9          lazyeval_0.2.2          spatstat.geom_3.7-0    
[133] glmnet_4.1-10           RcppHNSW_0.6.0          patchwork_1.3.2        
[136] future_1.69.0           statmod_1.5.1           shiny_1.11.1           
[139] ROCR_1.0-12             gridtext_0.1.5          igraph_2.2.1           
[142] fastmatch_1.1-6        
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-01-28
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ─────────────────────────────────────────────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [1] CRAN (R 4.4.1)
 assertthat             0.2.1      2019-03-21 [1] CRAN (R 4.4.0)
 babelgene              22.9       2022-09-29 [1] CRAN (R 4.4.0)
 beachmat               2.22.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 beeswarm               0.4.0      2021-06-01 [1] CRAN (R 4.4.0)
 Biobase              * 2.66.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 BiocGenerics         * 0.52.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 BiocParallel           1.40.2     2025-04-08 [1] Bioconductor
 cli                    3.6.5      2025-04-23 [1] CRAN (R 4.4.1)
 cluster                2.1.8.1    2025-03-12 [1] CRAN (R 4.4.1)
 codetools              0.2-20     2024-03-31 [1] CRAN (R 4.4.3)
 commonmark             2.0.0      2025-07-07 [1] CRAN (R 4.4.1)
 cowplot                1.2.0      2025-07-07 [1] CRAN (R 4.4.1)
 crayon                 1.5.3      2024-06-20 [1] CRAN (R 4.4.0)
 curl                   7.0.0      2025-08-19 [1] CRAN (R 4.4.1)
 data.table           * 1.17.4     2025-05-26 [1] CRAN (R 4.4.3)
 dbscan                 1.2.4      2025-12-19 [1] CRAN (R 4.4.3)
 DelayedArray           0.32.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 deldir                 2.0-4      2024-02-28 [1] CRAN (R 4.4.0)
 dichromat              2.0-0.1    2022-05-02 [1] CRAN (R 4.4.0)
 digest                 0.6.39     2025-11-19 [1] CRAN (R 4.4.3)
 dotCall64              1.2        2024-10-04 [1] CRAN (R 4.4.1)
 dplyr                * 1.1.4      2023-11-17 [1] CRAN (R 4.4.0)
 escheR               * 1.6.0      2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 evaluate               1.0.5      2025-08-27 [1] CRAN (R 4.4.1)
 farver                 2.1.2      2024-05-13 [1] CRAN (R 4.4.0)
 fastDummies            1.7.5      2025-01-20 [1] CRAN (R 4.4.1)
 fastmap                1.2.0      2024-05-15 [1] CRAN (R 4.4.0)
 fastmatch              1.1-6      2024-12-23 [1] CRAN (R 4.4.1)
 fgsea                  1.32.4     2025-03-20 [1] Bioconductor 3.20 (R 4.4.3)
 fitdistrplus           1.2-6      2026-01-24 [1] CRAN (R 4.4.3)
 foreach                1.5.2      2022-02-02 [1] CRAN (R 4.4.0)
 future                 1.69.0     2026-01-16 [1] CRAN (R 4.4.3)
 future.apply           1.20.1     2025-12-09 [1] CRAN (R 4.4.3)
 generics               0.1.4      2025-05-09 [1] CRAN (R 4.4.1)
 GenomeInfoDb         * 1.42.3     2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
 GenomeInfoDbData       1.2.13     2024-12-12 [1] Bioconductor
 GenomicRanges        * 1.58.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 ggbeeswarm             0.7.2      2023-04-29 [1] CRAN (R 4.4.0)
 ggplot2              * 4.0.1      2025-11-14 [1] CRAN (R 4.4.3)
 ggrastr              * 1.0.2      2023-06-01 [1] CRAN (R 4.4.0)
 ggrepel                0.9.6      2024-09-07 [1] CRAN (R 4.4.1)
 ggridges               0.5.7      2025-08-27 [1] CRAN (R 4.4.1)
 ggtext               * 0.1.2      2022-09-16 [1] CRAN (R 4.4.0)
 glmnet                 4.1-10     2025-07-17 [1] CRAN (R 4.4.1)
 globals                0.18.0     2025-05-08 [1] CRAN (R 4.4.1)
 glue                   1.8.0      2024-09-30 [1] CRAN (R 4.4.1)
 goftest                1.2-3      2021-10-07 [1] CRAN (R 4.4.0)
 gridExtra              2.3        2017-09-09 [1] CRAN (R 4.4.0)
 gridtext               0.1.5      2022-09-16 [1] CRAN (R 4.4.0)
 gtable                 0.3.6      2024-10-25 [1] CRAN (R 4.4.1)
 here                   1.0.2      2025-09-15 [1] CRAN (R 4.4.1)
 htmltools              0.5.8.1    2024-04-04 [1] CRAN (R 4.4.0)
 htmlwidgets            1.6.4      2023-12-06 [1] CRAN (R 4.4.0)
 httpuv                 1.6.16     2025-04-16 [1] CRAN (R 4.4.1)
 httr                   1.4.7      2023-08-15 [1] CRAN (R 4.4.0)
 ica                    1.0-3      2022-07-08 [1] CRAN (R 4.4.0)
 igraph                 2.2.1      2025-10-27 [1] CRAN (R 4.4.1)
 IRanges              * 2.40.1     2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
 irlba                  2.3.5.1    2022-10-03 [1] CRAN (R 4.4.0)
 iterators              1.0.14     2022-02-05 [1] CRAN (R 4.4.0)
 jsonlite               2.0.0      2025-03-27 [1] CRAN (R 4.4.1)
 KernSmooth             2.23-26    2025-01-01 [1] CRAN (R 4.4.3)
 knitr                  1.51       2025-12-20 [1] CRAN (R 4.4.3)
 labeling               0.4.3      2023-08-29 [1] CRAN (R 4.4.0)
 later                  1.4.4      2025-08-27 [1] CRAN (R 4.4.1)
 lattice                0.22-7     2025-04-02 [1] CRAN (R 4.4.1)
 lazyeval               0.2.2      2019-03-15 [1] CRAN (R 4.4.0)
 lifecycle              1.0.4      2023-11-07 [1] CRAN (R 4.4.0)
 limma                  3.62.2     2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
 listenv                0.10.0     2025-11-02 [1] CRAN (R 4.4.1)
 litedown               0.8        2025-11-02 [1] CRAN (R 4.4.1)
 lmtest                 0.9-40     2022-03-21 [1] CRAN (R 4.4.0)
 magick                 2.9.0      2025-09-08 [1] CRAN (R 4.4.1)
 magrittr               2.0.4      2025-09-12 [1] CRAN (R 4.4.1)
 markdown               2.0        2025-03-23 [1] CRAN (R 4.4.3)
 MASS                   7.3-65     2025-02-28 [1] CRAN (R 4.4.1)
 Matrix               * 1.7-4      2025-08-28 [1] CRAN (R 4.4.1)
 MatrixGenerics       * 1.18.1     2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
 matrixStats          * 1.5.0      2025-01-07 [1] CRAN (R 4.4.1)
 mime                   0.13       2025-03-17 [1] CRAN (R 4.4.1)
 miniUI                 0.1.2      2025-04-17 [1] CRAN (R 4.4.1)
 msigdbr                25.1.1     2025-07-21 [1] CRAN (R 4.4.1)
 nlme                   3.1-168    2025-03-31 [1] CRAN (R 4.4.1)
 nmfLabelTransfer     * 0.0.0.9000 2026-01-28 [1] Github (cindyfang70/spaTransfer@8b2d090)
 otel                   0.2.0      2025-08-29 [1] CRAN (R 4.4.1)
 parallelly             1.46.1     2026-01-08 [1] CRAN (R 4.4.3)
 patchwork              1.3.2      2025-08-25 [1] CRAN (R 4.4.1)
 pbapply                1.7-4      2025-07-20 [1] CRAN (R 4.4.1)
 pillar                 1.11.1     2025-09-17 [1] CRAN (R 4.4.1)
 pkgconfig              2.0.3      2019-09-22 [1] CRAN (R 4.4.0)
 plotly                 4.11.0     2025-06-19 [1] CRAN (R 4.4.1)
 plyr                   1.8.9      2023-10-02 [1] CRAN (R 4.4.0)
 png                    0.1-8      2022-11-29 [1] CRAN (R 4.4.0)
 polyclip               1.10-7     2024-07-23 [1] CRAN (R 4.4.0)
 progressr              0.18.0     2025-11-06 [1] CRAN (R 4.4.1)
 promises               1.5.0      2025-11-01 [1] CRAN (R 4.4.1)
 purrr                  1.2.0      2025-11-04 [1] CRAN (R 4.4.1)
 R6                     2.6.1      2025-02-15 [1] CRAN (R 4.4.1)
 RANN                   2.6.2      2024-08-25 [1] CRAN (R 4.4.1)
 RColorBrewer           1.1-3      2022-04-03 [1] CRAN (R 4.4.0)
 Rcpp                   1.1.1      2026-01-10 [1] CRAN (R 4.4.3)
 RcppAnnoy              0.0.23     2026-01-12 [1] CRAN (R 4.4.3)
 RcppEigen            * 0.3.4.0.2  2024-08-24 [1] CRAN (R 4.4.1)
 RcppHNSW               0.6.0      2024-02-04 [1] CRAN (R 4.4.1)
 RcppML               * 0.5.6      2026-01-27 [1] Github (zdebruine/RcppML@5449a5b)
 reshape2               1.4.5      2025-11-12 [1] CRAN (R 4.4.1)
 reticulate             1.44.1     2025-11-14 [1] CRAN (R 4.4.3)
 rjson                  0.2.23     2024-09-16 [1] CRAN (R 4.4.1)
 rlang                  1.1.7      2026-01-09 [1] CRAN (R 4.4.3)
 rmarkdown              2.30       2025-09-28 [1] CRAN (R 4.4.1)
 ROCR                   1.0-12     2026-01-23 [1] CRAN (R 4.4.3)
 rprojroot              2.1.1      2025-08-26 [1] CRAN (R 4.4.1)
 RSpectra               0.16-2     2024-07-18 [1] CRAN (R 4.4.0)
 rstudioapi             0.17.1     2024-10-22 [1] CRAN (R 4.4.1)
 Rtsne                  0.17       2023-12-07 [1] CRAN (R 4.4.0)
 S4Arrays               1.6.0      2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 S4Vectors            * 0.44.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 S7                     0.2.1      2025-11-14 [1] CRAN (R 4.4.3)
 scales                 1.4.0      2025-04-24 [1] CRAN (R 4.4.1)
 scattermore            1.2        2023-06-12 [1] CRAN (R 4.4.0)
 sctransform            0.4.3      2026-01-10 [1] CRAN (R 4.4.3)
 scuttle                1.16.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 sessioninfo            1.2.3      2025-02-05 [1] CRAN (R 4.4.1)
 Seurat               * 5.4.0      2025-12-14 [1] CRAN (R 4.4.3)
 SeuratObject         * 5.3.0      2025-12-12 [1] CRAN (R 4.4.3)
 shape                  1.4.6.1    2024-02-23 [1] CRAN (R 4.4.0)
 shiny                  1.11.1     2025-07-03 [1] CRAN (R 4.4.1)
 SingleCellExperiment * 1.28.1     2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
 singlet              * 0.99.8     2026-01-27 [1] Github (zdebruine/singlet@7ce199e)
 sp                   * 2.2-0      2025-02-01 [1] CRAN (R 4.4.1)
 spam                   2.11-1     2025-01-20 [1] CRAN (R 4.4.1)
 SparseArray            1.6.2      2025-02-20 [1] Bioconductor 3.20 (R 4.4.2)
 SpatialExperiment    * 1.16.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 spatstat.data          3.1-9      2025-10-18 [1] CRAN (R 4.4.1)
 spatstat.explore       3.7-0      2026-01-22 [1] CRAN (R 4.4.3)
 spatstat.geom          3.7-0      2026-01-20 [1] CRAN (R 4.4.3)
 spatstat.random        3.4-4      2026-01-21 [1] CRAN (R 4.4.3)
 spatstat.sparse        3.1-0      2024-06-21 [1] CRAN (R 4.4.0)
 spatstat.univar        3.1-6      2026-01-17 [1] CRAN (R 4.4.3)
 spatstat.utils         3.2-1      2026-01-10 [1] CRAN (R 4.4.3)
 statmod                1.5.1      2025-10-09 [1] CRAN (R 4.4.1)
 stringi                1.8.7      2025-03-27 [1] CRAN (R 4.4.1)
 stringr                1.6.0      2025-11-04 [1] CRAN (R 4.4.1)
 SummarizedExperiment * 1.36.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 survival               3.8-3      2024-12-17 [1] CRAN (R 4.4.3)
 tensor                 1.5.1      2025-06-17 [1] CRAN (R 4.4.1)
 tibble                 3.3.0      2025-06-08 [1] CRAN (R 4.4.1)
 tidyr                  1.3.1      2024-01-24 [1] CRAN (R 4.4.0)
 tidyselect             1.2.1      2024-03-11 [1] CRAN (R 4.4.0)
 UCSC.utils             1.2.0      2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 uwot                   0.2.4      2025-11-10 [1] CRAN (R 4.4.1)
 vctrs                  0.7.1      2026-01-23 [1] CRAN (R 4.4.3)
 vipor                  0.4.7      2023-12-18 [1] CRAN (R 4.4.0)
 viridisLite            0.4.2      2023-05-02 [1] CRAN (R 4.4.0)
 withr                  3.0.2      2024-10-28 [1] CRAN (R 4.4.1)
 xfun                   0.56       2026-01-18 [1] CRAN (R 4.4.3)
 xml2                   1.5.0      2025-11-17 [1] CRAN (R 4.4.3)
 xtable                 1.8-4      2019-04-21 [1] CRAN (R 4.4.0)
 XVector                0.46.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 yaml                   2.3.12     2025-12-10 [1] CRAN (R 4.4.3)
 zlibbioc               1.52.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 zoo                    1.8-14     2025-04-10 [1] CRAN (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library
 * ── Packages attached to the search path.

────────────────────────────────────────────────────────────────────────────────────