---
title: "03-WebDiv_labeltransfer_vis"
author: "Bernie Mulvey"
date: "06-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(ggrastr)
library(SpatialExperiment)
library(escheR)
library(Polychrome)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallel)


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

first let's make some plots of the relationship between visium registration scores and manual LC label inclusion
```{r}
xlc <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

wdlab <- fread("processed-data_newseg/04_snSeqLabelTransfer/02-WebDivToXen_xferedlabels.txt")
# wdlab[WebDivLabel %in% grep(WebDivLabel,pattern="ambig",value=T),WebDivLabel:="neurons_ambiguous"]
# wdlab[WebDivLabel %in% grep(WebDivLabel,pattern="oligod",value=T),WebDivLabel:="oligodendrocytes"]
# wdlab <- DataFrame(wdlab,row.names=wdlab$rn)[colnames(xlc),]
wdlab[,WebDivLabel:=ifelse(WebDivLabel %in% c("NE","inhibitory_4"),WebDivLabel,"Other")]
wdlab <- DataFrame(wdlab,row.names=wdlab$rn)[colnames(xlc),]

colLabels(xlc) <- wdlab$WebDivLabel

unique(xlc$label)
# pal <- c("#0F0","#9D0","#7E8","#5B4",
#          "#F00","#D37","#E62","#B40","#A00","#C20","#F90",
#          "#00F",
#          "#F0F",
#          "#B9B9B9","#ECECEC",
#          "#FF5","#CC5",
#          "#989898",
#          "#90F",
#          "#59D")
pal <- c("#00F",
         "#F0F",
         "#B9B9B9")
names(pal) <- c("NE","inhibitory_4","Other")
stopifnot(all(names(pal) %in% unique(xlc$label)))


# names(pal) <- c(paste0("excitatory_",c(26,7,23,30)),
#                 paste0("inhibitory_",c(4,25,14,17,8,20,24)),
#                 "NE",
#                 "5HT_21",
#                 "oligodendrocytes",
#                 "OPCs_27",
#                 "astrocytes_5","astrocytes_29",
#                 "macrophages_microglia_15",
#                 "endothelial_mural_16",
#                 "neurons_ambiguous"
#                 )

xlc$label <- factor(xlc$label,levels=names(pal))
samps <- lapply(unique(xlc$brnum),function(b){xlc[,xlc$brnum==b]})

names(samps) <- unique(xlc$brnum)
rm(xlc)
gc(full=T)

plts <- mapply(s=samps,n=names(samps),SIMPLIFY=FALSE,function(s,n){
  p <- make_escheR(s,spot_size=0.5)
  p <- add_fill(p,"label",point_size=0.5)+    guides(fill=guide_legend(override.aes = list(size=2)))
  p <- p+scale_fill_manual(values=pal)
  p <- p+ggtitle(n)+theme(title=element_text(size=9))
  q <- rasterize(p,layers="Points",dpi=300,dev="cairo_png")
  return(q)
})

png("plots_newseg/04_snSeqLabelTransfer/03a-WebDivLabelxfer_to_xen_inh4LConly.png",height=36,width=18,unit="in",res=150)
do.call("grid.arrange",c(plts,nrow=8))
dev.off()

```


```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] Polychrome_1.5.4            escheR_1.10.0               ggrastr_1.0.2              
 [4] spatialLIBD_1.22.0          FNN_1.1.4.1                 caTools_1.18.3             
 [7] FastKNN_0.0.1               gridExtra_2.3               ggtext_0.1.2               
[10] ggplot2_4.0.2               BiocParallel_1.44.0         parallelly_1.47.0          
[13] colorout_1.3-3              SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
[16] SummarizedExperiment_1.40.0 Biobase_2.70.0              GenomicRanges_1.62.1       
[19] Seqinfo_1.0.0               IRanges_2.44.0              S4Vectors_0.48.1           
[22] BiocGenerics_0.56.0         generics_0.1.4              MatrixGenerics_1.22.0      
[25] matrixStats_1.5.0           data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] later_1.4.5              BiocIO_1.20.0            bitops_1.0-9            
  [4] filelock_1.0.3           tibble_3.3.1             XML_3.99-0.20           
  [7] lifecycle_1.0.5          httr2_1.2.2              edgeR_4.8.2             
 [10] doParallel_1.0.17        rprojroot_2.1.1          lattice_0.22-7          
 [13] magrittr_2.0.4           sass_0.4.10              limma_3.66.0            
 [16] plotly_4.12.0            rmarkdown_2.30           jquerylib_0.1.4         
 [19] yaml_2.3.12              httpuv_1.6.16            otel_0.2.0              
 [22] sessioninfo_1.2.3        cowplot_1.2.0            DBI_1.2.3               
 [25] RColorBrewer_1.1-3       golem_0.5.1              abind_1.4-8             
 [28] purrr_1.2.1              RCurl_1.98-1.17          rappdirs_0.3.4          
 [31] circlize_0.4.17          ggrepel_0.9.6            irlba_2.3.7             
 [34] codetools_0.2-20         DelayedArray_0.36.1      DT_0.34.0               
 [37] scuttle_1.20.0           xml2_1.5.2               tidyselect_1.2.1        
 [40] shape_1.4.6.1            farver_2.1.2             viridis_0.6.5           
 [43] ScaledMatrix_1.18.0      shinyWidgets_0.9.0       BiocFileCache_3.0.0     
 [46] GenomicAlignments_1.46.0 jsonlite_2.0.0           GetoptLong_1.1.0        
 [49] BiocNeighbors_2.4.0      scater_1.38.1            iterators_1.0.14        
 [52] foreach_1.5.2            tools_4.5.2              Rcpp_1.1.1              
 [55] glue_1.8.0               SparseArray_1.10.8       xfun_0.56               
 [58] here_1.0.2               dplyr_1.2.0              withr_3.0.2             
 [61] BiocManager_1.30.27      fastmap_1.2.0            pdist_1.2.1             
 [64] digest_0.6.39            rsvd_1.0.5               R6_2.6.1                
 [67] mime_0.13                colorspace_2.1-2         dichromat_2.0-0.1       
 [70] RSQLite_2.4.5            cigarillo_1.0.0          config_0.3.2            
 [73] tidyr_1.3.2              rtracklayer_1.70.1       httr_1.4.7              
 [76] htmlwidgets_1.6.4        S4Arrays_1.10.1          scatterplot3d_0.3-44    
 [79] pkgconfig_2.0.3          gtable_0.3.6             rsconnect_1.7.0         
 [82] blob_1.3.0               ComplexHeatmap_2.26.1    S7_0.2.1                
 [85] XVector_0.50.0           htmltools_0.5.9          clue_0.3-66             
 [88] scales_1.4.0             png_0.1-8                attempt_0.3.1           
 [91] knitr_1.51               rstudioapi_0.18.0        rjson_0.2.23            
 [94] curl_7.0.0               cachem_1.1.0             GlobalOptions_0.1.3     
 [97] BiocVersion_3.22.0       vipor_0.4.7              AnnotationDbi_1.72.0    
[100] restfulr_0.0.16          pillar_1.11.1            grid_4.5.2              
[103] vctrs_0.7.1              promises_1.5.0           BiocSingular_1.26.0     
[106] dbplyr_2.5.1             beachmat_2.26.0          xtable_1.8-4            
[109] cluster_2.1.8.1          beeswarm_0.4.0           paletteer_1.7.0         
[112] evaluate_1.0.5           magick_2.9.0             cli_3.6.5               
[115] locfit_1.5-9.12          compiler_4.5.2           Rsamtools_2.26.0        
[118] rlang_1.1.7              crayon_1.5.3             labeling_0.4.3          
[121] rematch2_2.1.2           ggbeeswarm_0.7.3         viridisLite_0.4.3       
[124] assertthat_0.2.1         Biostrings_2.78.0        lazyeval_0.2.2          
[127] Matrix_1.7-4             ExperimentHub_3.0.0      benchmarkme_1.0.8       
[130] bit64_4.6.0-1            KEGGREST_1.50.0          statmod_1.5.1           
[133] shiny_1.12.1             AnnotationHub_4.0.0      gridtext_0.1.5          
[136] memoise_2.0.1            bslib_0.10.0             benchmarkmeData_2.0.0   
[139] bit_4.6.0               
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-03
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ─────────────────────────────────────────────────────────────────────────────────
 package              * version   date (UTC) lib source
 abind                  1.4-8     2024-09-12 [1] CRAN (R 4.5.0)
 AnnotationDbi          1.72.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationHub          4.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 assertthat             0.2.1     2019-03-21 [1] CRAN (R 4.5.0)
 attempt                0.3.1     2020-05-03 [1] CRAN (R 4.5.0)
 beachmat               2.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.5.0)
 benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.5.0)
 benchmarkmeData        2.0.0     2026-01-19 [1] CRAN (R 4.5.2)
 Biobase              * 2.70.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocFileCache          3.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocIO                 1.20.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocManager            1.30.27   2025-11-14 [1] CRAN (R 4.5.2)
 BiocNeighbors          2.4.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocVersion            3.22.0    2025-10-07 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0     2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1   2025-01-16 [1] CRAN (R 4.5.0)
 bitops                 1.0-9     2024-10-03 [1] CRAN (R 4.5.0)
 blob                   1.3.0     2026-01-14 [1] CRAN (R 4.5.2)
 bslib                  0.10.0    2026-01-26 [1] CRAN (R 4.5.2)
 cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.5.0)
 caTools              * 1.18.3    2024-09-04 [1] CRAN (R 4.5.0)
 cigarillo              1.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 circlize               0.4.17    2025-12-08 [1] CRAN (R 4.5.2)
 cli                    3.6.5     2025-04-23 [1] CRAN (R 4.5.0)
 clue                   0.3-66    2024-11-13 [1] CRAN (R 4.5.0)
 cluster                2.1.8.1   2025-03-12 [1] CRAN (R 4.5.0)
 codetools              0.2-20    2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3     2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 colorspace             2.1-2     2025-09-22 [1] CRAN (R 4.5.0)
 ComplexHeatmap         2.26.1    2026-01-30 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 config                 0.3.2     2023-08-30 [1] CRAN (R 4.5.0)
 cowplot                1.2.0     2025-07-07 [1] CRAN (R 4.5.0)
 crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0     2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1  2026-01-27 [1] CRAN (R 4.5.2)
 DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.5.0)
 dbplyr                 2.5.1     2025-09-10 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.1    2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1   2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39    2025-11-19 [1] CRAN (R 4.5.2)
 doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.5.0)
 dplyr                  1.2.0     2026-02-03 [1] CRAN (R 4.5.2)
 DT                     0.34.0    2025-09-02 [1] CRAN (R 4.5.0)
 edgeR                  4.8.2     2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 escheR               * 1.10.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 evaluate               1.0.5     2025-08-27 [1] CRAN (R 4.5.0)
 ExperimentHub          3.0.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 farver                 2.1.2     2024-05-13 [1] CRAN (R 4.5.0)
 FastKNN              * 0.0.1     2015-02-12 [1] CRAN (R 4.5.0)
 fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.5.0)
 filelock               1.0.3     2023-12-11 [1] CRAN (R 4.5.0)
 FNN                  * 1.1.4.1   2024-09-22 [1] CRAN (R 4.5.0)
 foreach                1.5.2     2022-02-02 [1] CRAN (R 4.5.0)
 generics             * 0.1.4     2025-05-09 [1] CRAN (R 4.5.0)
 GenomicAlignments      1.46.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 GenomicRanges        * 1.62.1    2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 GetoptLong             1.1.0     2025-11-28 [1] CRAN (R 4.5.2)
 ggbeeswarm             0.7.3     2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2     2026-02-03 [1] CRAN (R 4.5.2)
 ggrastr              * 1.0.2     2023-06-01 [1] CRAN (R 4.5.0)
 ggrepel                0.9.6     2024-09-07 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2     2022-09-16 [1] CRAN (R 4.5.0)
 GlobalOptions          0.1.3     2025-11-28 [1] CRAN (R 4.5.2)
 glue                   1.8.0     2024-09-30 [1] CRAN (R 4.5.0)
 golem                  0.5.1     2024-08-27 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5     2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6     2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2     2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9     2025-12-04 [1] CRAN (R 4.5.2)
 htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.5.0)
 httpuv                 1.6.16    2025-04-16 [1] CRAN (R 4.5.0)
 httr                   1.4.7     2023-08-15 [1] CRAN (R 4.5.0)
 httr2                  1.2.2     2025-12-08 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7     2026-01-30 [1] CRAN (R 4.5.2)
 iterators              1.0.14    2022-02-05 [1] CRAN (R 4.5.0)
 jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.5.0)
 jsonlite               2.0.0     2025-03-27 [1] CRAN (R 4.5.0)
 KEGGREST               1.50.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51      2025-12-20 [1] CRAN (R 4.5.2)
 labeling               0.4.3     2023-08-29 [1] CRAN (R 4.5.0)
 later                  1.4.5     2026-01-08 [1] CRAN (R 4.5.2)
 lattice                0.22-7    2025-04-02 [1] CRAN (R 4.5.2)
 lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.5.0)
 lifecycle              1.0.5     2026-01-08 [1] CRAN (R 4.5.2)
 limma                  3.66.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 locfit                 1.5-9.12  2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0     2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4     2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4     2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0     2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1     2021-11-26 [1] CRAN (R 4.5.0)
 mime                   0.13      2025-03-17 [1] CRAN (R 4.5.0)
 otel                   0.2.0     2025-08-29 [1] CRAN (R 4.5.0)
 paletteer              1.7.0     2026-01-08 [1] CRAN (R 4.5.2)
 parallelly           * 1.47.0    2026-04-17 [1] CRAN (R 4.5.2)
 pdist                  1.2.1     2022-05-02 [1] CRAN (R 4.5.0)
 pillar                 1.11.1    2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.5.0)
 plotly                 4.12.0    2026-01-24 [1] CRAN (R 4.5.2)
 png                    0.1-8     2022-11-29 [1] CRAN (R 4.5.0)
 Polychrome           * 1.5.4     2025-04-06 [1] CRAN (R 4.5.0)
 promises               1.5.0     2025-11-01 [1] CRAN (R 4.5.0)
 purrr                  1.2.1     2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1     2025-02-15 [1] CRAN (R 4.5.0)
 rappdirs               0.3.4     2026-01-17 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1     2026-01-10 [1] CRAN (R 4.5.2)
 RCurl                  1.98-1.17 2025-03-22 [1] CRAN (R 4.5.0)
 rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.5.0)
 restfulr               0.0.16    2025-06-27 [1] CRAN (R 4.5.0)
 rjson                  0.2.23    2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7     2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30      2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1     2025-08-26 [1] CRAN (R 4.5.0)
 Rsamtools              2.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 rsconnect              1.7.0     2025-12-06 [1] CRAN (R 4.5.2)
 RSQLite                2.4.5     2025-11-30 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0    2026-01-16 [1] CRAN (R 4.5.2)
 rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.5.0)
 rtracklayer            1.70.1    2025-12-20 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 S4Arrays               1.10.1    2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1    2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1     2025-11-14 [1] CRAN (R 4.5.2)
 sass                   0.4.10    2025-04-11 [1] CRAN (R 4.5.0)
 ScaledMatrix           1.18.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0     2025-04-24 [1] CRAN (R 4.5.0)
 scater                 1.38.1    2026-03-20 [1] Bioconductor 3.22 (R 4.5.2)
 scatterplot3d          0.3-44    2023-05-05 [1] CRAN (R 4.5.0)
 scuttle                1.20.0    2025-10-30 [1] Bioconductor 3.22 (R 4.5.2)
 Seqinfo              * 1.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3     2025-02-05 [1] CRAN (R 4.5.0)
 shape                  1.4.6.1   2024-02-23 [1] CRAN (R 4.5.0)
 shiny                  1.12.1    2025-12-09 [1] CRAN (R 4.5.2)
 shinyWidgets           0.9.0     2025-02-21 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8    2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 spatialLIBD          * 1.22.0    2025-11-04 [1] Bioconductor 3.22 (R 4.5.2)
 statmod                1.5.1     2025-10-09 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1     2026-01-11 [1] CRAN (R 4.5.2)
 tidyr                  1.3.2     2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1     2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.5.0)
 viridis                0.6.5     2024-01-29 [1] CRAN (R 4.5.0)
 viridisLite            0.4.3     2026-02-04 [1] CRAN (R 4.5.2)
 withr                  3.0.2     2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56      2026-01-18 [1] CRAN (R 4.5.2)
 XML                    3.99-0.20 2025-11-08 [1] CRAN (R 4.5.0)
 xml2                   1.5.2     2026-01-17 [1] CRAN (R 4.5.2)
 xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.5.0)
 XVector                0.50.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12    2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

────────────────────────────────────────────────────────────────────────────────────────────