---
title: "04-webdiv_smoothdLCDefs_regists_and_propsindom"
author: "Bernie Mulvey"
date: "06-03-2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(ggrastr)
library(SpatialExperiment)
library(escheR)


## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallel)


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

first let's make some plots of the relationship between visium registration scores and manual LC label inclusion
```{r}
regs <- fread("processed-data_newseg/04_snSeqLabelTransfer/04c-OneThresh_webdiv_smoothdLC_regist_to_VisLC_ManXenLC_srtNMFpred_wdNMFpred.txt")

regsplt1 <- regs[Ntopdommks %in% c(10,40,70,100,130,160,190,220,250,280,300)]
regsplt1 <- melt.data.table(regsplt1,id.vars = c("domainDef","Ntopdommks","looseInDom","strictInDom","WebDivNMFInDom","k1","k2"))
```

```{r}
pdf("plots_newseg/04_snSeqLabelTransfer/04_KNNsmoothLC_QC_and_spotplots/04a_LFFvis_regist_and_PropWebDivNEinDom_top300mks_k2facet.pdf",height=18,width=5.5)
ggplot(regsplt1[Ntopdommks==300&variable=="LFF_Visium_Domains"],aes(x=WebDivNMFInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=4)+
  ylab("Registration to Visium LC")
dev.off()

pdf("plots_newseg/04_snSeqLabelTransfer/04_KNNsmoothLC_QC_and_spotplots/04a_LFFvis_regist_and_PropManLCStrict_top300mks_k2facet.pdf",height=18,width=5.5)
ggplot(regsplt1[Ntopdommks==300&variable=="LFF_Visium_Domains"],aes(x=strictInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=4)+
  ylab("Registration to Visium LC")
dev.off()

pdf("plots_newseg/04_snSeqLabelTransfer/04_KNNsmoothLC_QC_and_spotplots/04a_LFFvis_regist_and_PropWebDivNEinDom_top40mks_k2facet.pdf",height=18,width=5.5)
ggplot(regsplt1[Ntopdommks==40&variable=="LFF_Visium_Domains"],aes(x=WebDivNMFInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=4)+
  ylab("Registration to Visium LC")
dev.off()

pdf("plots_newseg/04_snSeqLabelTransfer/04_KNNsmoothLC_QC_and_spotplots/04a_LFFvis_regist_and_PropManLCStrict_top40mks_k2facet.pdf",height=18,width=5.5)
ggplot(regsplt1[Ntopdommks==40&variable=="LFF_Visium_Domains"],aes(x=strictInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=4)+
  ylab("Registration to Visium LC")
dev.off()

## proportion of label-xfered LC cells in domain and nmfpred LC cells registration

pdf("plots_newseg/04_snSeqLabelTransfer/04_KNNsmoothLC_QC_and_spotplots/04b_wdNMFLCcells_regist_and_NMFxferLCcellProp_top300mks_k2facet.pdf",height=12,width=5.5)
ggplot(regsplt1[Ntopdommks==300&variable=="nmfpred"],aes(x=WebDivNMFInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=4)+
  ylab("Registration to Cells NMF-Xfered WebDiv NE Label")
dev.off()

pdf("plots_newseg/04_snSeqLabelTransfer/04_KNNsmoothLC_QC_and_spotplots/04b_wdNMFLCcells_regist_and_NMFxferLCcellProp_top40mks_k2facet.pdf",height=12,width=5.5)
ggplot(regsplt1[Ntopdommks==40&variable=="nmfpred"],aes(x=WebDivNMFInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=4)+
  ylab("Registration to Cells NMF-Xfered WebDiv NE Label")
dev.off()

## and how does the domain compare to (Visium LC vs pseudobulking vs nmfpred LC cells)? An ideal domain will be a  stronger match to the Vis domain than to nmfpred LCNE cells alone.
pdf("plots_newseg/04_snSeqLabelTransfer/04_KNNsmoothLC_QC_and_spotplots/04c_VisiumLC-wdNMFprojNEcells_registdiff_top40mks_k1facet.pdf",height=12,width=5.5)
ggplot(regs[Ntopdommks==40],aes(y=(LFF_Visium_Domains-nmfpred),x=k2))+
  geom_point()+
  facet_wrap(~k1)
dev.off()

pdf("plots_newseg/04_snSeqLabelTransfer/04_KNNsmoothLC_QC_and_spotplots/04c_VisiumLC-wdNMFprojNEcells_registdiff_top300mks_k1facet.pdf",height=12,width=5.5)
ggplot(regs[Ntopdommks==300],aes(y=(LFF_Visium_Domains-nmfpred),x=k2))+
  geom_point()+
  facet_wrap(~k1)
dev.off()
```
Ranges that seem promising to zoom in on in code_newseg/04/05:

Vis LC regist by top 40: 0.04/0.361-0.519; 0.08/0.201-0.519; 0.12/0.0401-0.199,0.281-0.439; 0.16/0.361-0.519; 0.2/0.04-0.119

Vis LC regist by top 300: 0.04/0.521-0.679; 0.08/0.441-0.599; 0.12/0.441-0.599; 0.16/0.201-0.359; 0.2/0.121-0.359

NMF wdLC cells by top 40: 0.04/0.441-0.599,0.681-0.76; 0.08/0.361-0.519; 0.12/0.201-0.359; 0.16/0.28-.359,0.441-0.599; 0.2/0.439-0.599

NMF wdLC cells by top 300: 0.04/0.441-0.599; 0.08/0.361-.0519; 0.12/0.279-.439; 0.16/0.281-0.439; 0.2/0.281-0.439

^ these ranges are expanded by 0.79 to each side (to cover everything between the k2 threshold steps of 6/75 in this round of data points).

now, get unique k1-k2 pairs that include the 6/75ths-wise steps we plotted (so we plot them again with fine steps to compare by) --> save to 04d-refine_wdsmoothingthreshes.txt
```{r}
newdt <- as.data.frame(c(rep(0.04,5),rep(0.08,4),rep(0.12,5),rep(0.16,5),rep(0.2,4)))
names(newdt)[1] <- "k1"
newdt$mink2 <- c(0.360,0.520,0.44,0.68,0.44,
                 0.200,0.44,0.36,0.36,
                 0.040,0.280,0.440,0.200,0.28,
                 0.360,0.200,0.280,0.440,0.280,
                 0.120,0.120,0.440,0.280)
newdt$maxk2 <- c(0.520,0.680,0.600,0.760,0.600,
                 0.520,0.600,0.520,0.520,
                 0.200,0.440,0.600,0.360,0.440,
                 0.520,0.360,0.360,0.600,0.440,
                 0.120,0.360,0.600,0.440)


finekpairs <- list()
i <- 1
for (i in c(1:nrow(newdt))){
  l <- newdt$mink2[i]
  h <- newdt$maxk2[i]
  o <- seq(from=l,to=h,by=(1/75))
  o <- as.data.table(cbind(newdt$k1[i],o))
  setnames(o,c("k1","k2thresh"))
  finekpairs[[i]] <- o
  rm(o,l,h)
}
finekpairs <- rbindlist(finekpairs)
## 293 pairs, including duplicate prioritizations and pairs that we plotted here. retain those that are prioritized by at least two metrics OR in seq(0.04,0.76,by=6/75)
finekpairs <- finekpairs[,.N,by=c("k1","k2thresh")]
finekpairs <- unique(finekpairs[N>1|k2thresh %in% round(seq(0.04,0.76,by=6/75),3)]) # 79 pairs, including ones from here that will be replotted as comparators
fwrite(finekpairs,"processed-data_newseg/04_snSeqLabelTransfer/04d-refine_wdsmoothingthreshes.txt",sep='\t',quote=F,row.names=F)
```


R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0         parallelly_1.46.1          
 [3] colorout_1.3-3              escheR_1.10.0              
 [5] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
 [7] SummarizedExperiment_1.40.0 Biobase_2.70.0             
 [9] GenomicRanges_1.62.1        Seqinfo_1.0.0              
[11] IRanges_2.44.0              S4Vectors_0.48.0           
[13] BiocGenerics_0.56.0         generics_0.1.4             
[15] MatrixGenerics_1.22.0       matrixStats_1.5.0          
[17] ggrastr_1.0.2               gridExtra_2.3              
[19] ggtext_0.1.2                ggplot2_4.0.1              
[21] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] gtable_0.3.6        beeswarm_0.4.0      rjson_0.2.23        xfun_0.56          
 [5] lattice_0.22-7      vctrs_0.7.1         tools_4.5.2         tibble_3.3.1       
 [9] pkgconfig_2.0.3     Matrix_1.7-4        RColorBrewer_1.1-3  S7_0.2.1           
[13] lifecycle_1.0.5     compiler_4.5.2      farver_2.1.2        codetools_0.2-20   
[17] vipor_0.4.7         htmltools_0.5.9     yaml_2.3.12         pillar_1.11.1      
[21] rsconnect_1.7.0     DelayedArray_0.36.0 sessioninfo_1.2.3   magick_2.9.0       
[25] abind_1.4-8         tidyselect_1.2.1    digest_0.6.39       dplyr_1.1.4        
[29] rprojroot_2.1.1     fastmap_1.2.0       grid_4.5.2          here_1.0.2         
[33] cli_3.6.5           SparseArray_1.10.8  magrittr_2.0.4      S4Arrays_1.10.1    
[37] dichromat_2.0-0.1   withr_3.0.2         scales_1.4.0        ggbeeswarm_0.7.3   
[41] rmarkdown_2.30      XVector_0.50.0      otel_0.2.0          evaluate_1.0.5     
[45] knitr_1.51          rlang_1.1.7         gridtext_0.1.5      Rcpp_1.1.1         
[49] glue_1.8.0          xml2_1.5.2          rstudioapi_0.18.0   R6_2.6.1           
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-02
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ─────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.1.4    2023-11-17 [1] CRAN (R 4.5.0)
 escheR               * 1.10.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggbeeswarm             0.7.3    2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.1    2025-11-14 [1] CRAN (R 4.5.2)
 ggrastr              * 1.0.2    2023-06-01 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] CRAN (R 4.5.2)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.5.0)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

────────────────────────────────────────────────────────────────────────────────────
