---
title: "06-Take single steps through K2 ranges that look promising for wdxfer and run registrations for them"
format: html
date: "06-03-2026"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
# library(SpatialFeatureExperiment)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}


## rstudio GUI tweaks
require(colorout)
ColorOut()
options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
library(parallel)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = '#000000'), axis.title.x = element_text(size = 9,color = '#000000'), axis.text.y = element_text(size = 8,color = '#000000'), axis.title.y = element_text(size =9,color = '#000000'), plot.title = element_markdown(size = 10,hjust=0.5,color = '#000000'), strip.text = element_text(size=11), legend.text = element_text(size=7,color = '#000000'), legend.title = element_text(size=8,hjust=0.5,color = '#000000'),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position='plot'))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

Candidate ranges identified by plots/04/04 (expanded to 0.08 on each side and condensed to a unique k1-k2 pair table, retaining pairs that were prioritized based on 2+ of 4 registration stats). Most have ~75% nmfpred LC cells in domain, some higher end threshes more like 55%. see full notes on prioritized k1-k2 ranges to zoom in on in plots_newseg/code/04/04.

load candidate ranges, smoothing calcs, xenium, and manual xenium LC assignments
```{r}
refineranges <- fread("processed-data_newseg/04_snSeqLabelTransfer/04d-refine_wdsmoothingthreshes.txt")

## smoothings
dualscores <- readRDS("processed-data_newseg/04_snSeqLabelTransfer/03b-WebDivLCknnsmoothing_step2k75.RDS")
## subset to range needed: 0.04-0.20
dualscores <- dualscores[paste0("k1thresh_",seq(0.04,0.20,by=0.04))]

### lc xenium
lcx <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")
assays(lcx) <- assays(lcx)[1] #dont need normed counts or log counts

### xenium manual LC definition
manlc <- fread("processed-data_newseg/03_ManualLC/01-XprBasedLClabels_and_ManLCcellsWithSegsGainedorAlteredByNM.txt")
manlc <- DataFrame(manlc,row.names=manlc$rn)[colnames(lcx),]
lcx$LCloose <- manlc$LCloose
lcx$LCstrict <- manlc$LCstrict

### webdiv transferred labels themselves
preds <- fread("processed-data_newseg/04_snSeqLabelTransfer/02-WebDivToXen_xferedlabels.txt")

### keep manLC for tabulating % overlap with domain and cells later
manlc <- as.data.table(manlc)
```

### in this domain definition scheme, the cell's K1 score doesn't matter, only the k2 score. there is still a series of k1 thresholds that was used to define "inLC" for k2 score calcs, we just aren't filtering ON individual cell k1 scores.

here, we'll step through the ranges that seemed promising based on two or more registration metrics (somewhat disregarding proportion of NMF-label-transferred NE cells within the domain) in steps of (1/75)

### we already checked these score ranges for having at least 20 cells in at least 16 donors.
```{r}
## avoid floating point bullshit
refineranges[,k1:=round(k1,2)]
refineranges[,k2thresh:=round(k2thresh,3)]

setDTthreads(1,restore_after_fork = FALSE)

regperthresh <- lapply(c(1:nrow(refineranges)),function(i){
  curk1 <- refineranges[i,k1]
  k2thresh <- refineranges[i,k2thresh] - 0.001
  ## subtract each k2 threshold by 0.001 to account for fractional thirds (so that k2s in the result == to e.g. 20/75 (.26666667) are included. they would be <0.267 but > 0.266, while 19/75 would not be.)
  tmpdt <- copy(dualscores[[paste0("k1thresh","_",curk1)]])
  ## cell must meet k2 threshold (thus "1-condition" smoothing)
  tmpdt[,newcol:=ifelse(
      # greaterOrEqual(lcpred.smooth1.k25,k1) &
    greaterOrEqual(lcpred.smooth2.k75,k2thresh),
    "LC","Other")
  ]
    
  if(length(unique(tmpdt$newcol))<2){
    return("NA")
  } else {
    threshname <- paste0("LCsmooth_k1_",curk1,"_k2_",k2thresh)
      setnames(tmpdt,"newcol",threshname)
      tmpdt <- DataFrame(tmpdt[,.(rn,get(threshname))],row.names=tmpdt$rn)[colnames(lcx),]
      colData(lcx)$newcol <- tmpdt[,2] 
      colnames(colData(lcx))[ncol(colData(lcx))] <- threshname

      ## registration stats for this threshold (we then cross correlate those to consistent regist stats from visium and manual xen)

    regstats <- registration_wrapper(sce=lcx,var_registration = threshname,var_sample_id = "brnum",covars = c("Age","Sex","APOE","YRI"),gene_ensembl = "ID",gene_name = "Symbol",min_ncells = 20)
    keepname <- c("ensembl",paste0("t_stat_",c("LC","Other")))
    outtab <- as.data.table(regstats$enrichment)[,..keepname]
    return(outtab)
  }
})
names(regperthresh) <- paste0("LCsmooth_k1_",refineranges$k1,"_k2_",refineranges$k2thresh)

saveRDS(regperthresh,file="processed-data_newseg/04_snSeqLabelTransfer/05-OneThresh_webdiv_fineK2steps_registstats.RDS")
```

## calculate registration stats for xenium manuals, projected weber-divecha labels, and projected visium domains, then can remove those datasets from sesh
```{r}
### weber-divecha NE labels (cell-level, to compare to LC as domain)
preds[WebDivLabel!="NE",WebDivLabel:="No"]
## change NE to LC here for compatibility
preds[WebDivLabel=="NE",WebDivLabel:="LC"]

predstmp <- DataFrame(preds,row.names=preds$rn)[colnames(lcx),]
colData(lcx)$nmfpred <- predstmp$WebDivLabel

# cleanup
rm(predstmp)
gc(full=T)

## run regist for all labels in colData that we want to test
comparegs <- lapply(c("LCloose","LCstrict","nmfpred"),function(l){
  resdat <- registration_wrapper(sce=lcx,var_registration=l,var_sample_id="brnum",covars=c("Age","Sex","APOE","YRI"),gene_ensembl="ID",gene_name="Symbol",min_ncells=20)
  resdat <- as.data.table(resdat$enrichment)
  resdat <- resdat[,.(ensembl,t_stat_LC,t_stat_No)]
  return(resdat)
})
names(comparegs) <- c("LCloose","LCstrict","nmfpred")
```


## load visium data and annots to get regist stats for SRT LC domain
```{r}
lcv <- readRDS("../LFF_spatial_LC/processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

clus <- readRDS("../LFF_spatial_LC/processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
clus <- clus$HARMONYlmbna_HDG_SVG_2575[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(clus,2,"clusid")

anno <- fread("../LFF_spatial_LC/processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")
# this will just be LC/other for registration purposes
anno[,anno:=ifelse(anno=="LC","LC","Other")]

anno <- merge.data.table(clus,anno,by="clusid",all.x=T)
anno <- DataFrame(anno,row.names=anno$rn)[colnames(lcv),]
lcv$vislab <- anno$anno

## subset Visium to samples we want to retain in both assays
unique(lcv$brnum)
lcv$brnum[lcv$brnum=="Br6119(re-dis)"] <- "Br6119"

## drop donors with limited LC
lcv <- lcv[,!(lcv$brnum %in% paste0("Br",c(5517,5276,5712)))]

## check all donors overlap
lcx$donor <- gsub(lcx$brnum,pattern="L",replacement="")
stopifnot(all(unique(lcv$brnum) %in% unique(lcx$donor)))
stopifnot(all(unique(lcx$donor) %in% unique(lcv$brnum)))
### 
## subset LC visium to genes in xenium
tmprd <- as.data.table(rowData(lcv),keep.rownames=T)
tmprd <- tmprd[gene_name %in% rowData(lcx)$Symbol]
keepg <- tmprd$rn

lcv <- lcv[keepg,]

rm(tmprd,lcx,anno,clus)
gc(full=T)
```

ancestry proportion estimates aren't in this visium object, so add them for identical registration modeling
```{r}
flares <- fread("../LFF_spatial_LC/raw-data/ancestryAndHaplotypes/FLARE/global_ancestry_estimates.txt")
flares <- flares[ID %in% lcv$brnum]
stopifnot(all(lcv$brnum %in% flares$ID))

tmpcd <- as.data.table(colData(lcv),keep.rownames=T)
tmpcd <- merge.data.table(tmpcd,flares[,.(ID,YRI)],by.x="brnum",by.y="ID",all.x=T)
stopifnot(all(!is.na(tmpcd$YRI)))

# Also fix APOE genotypes to avoid modeling errors
tmpcd[APOE=="E4/ E4",APOE:="E4hom"]
tmpcd[APOE=="E3/ E4",APOE:="E4het"]
tmpcd[APOE=="E2/ E3",APOE:="E2het"]
tmpcd[APOE=="E2/ E2",APOE:="E2hom"]

## replace coldata
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(lcv),]
tmpcd$rn <- NULL
colData(lcv) <- tmpcd

stopifnot(is.numeric(lcv$Age)&is.numeric(lcv$YRI)&is.character(lcv$APOE)&is.character(lcv$Sex))

rm(tmpcd,flares)
```


### vis lc regist
```{r}
lcvreg <- registration_wrapper(sce=lcv,var_registration = "vislab",var_sample_id = "sample_id",covars = c("Age","Sex","APOE","YRI"),gene_ensembl = "gene_id",gene_name = "gene_name",min_ncells = 10) # one tissue section has a few dozen LC spots at most

lcvreg <- as.data.table(lcvreg$enrichment)
lcvreg <- lcvreg[,.(ensembl,t_stat_LC,t_stat_Other)]
comparegs[[length(comparegs)+1]] <- lcvreg
names(comparegs)[length(comparegs)] <- "LFF_Visium_Domains"

rm(lcv,lcvreg)
gc(full=T)
```

### also get proportion of cells with LC label in each xenium manual annot that are labeled as LC domain
```{r}
propsindoms <- mcmapply(d=dualscores,n=names(dualscores),SIMPLIFY=FALSE,mc.cores=5,FUN=function(d,n){
    curk1 <- as.numeric(gsub(n,pattern="k1thresh_(.*)",replacement="\\1"))
    tmp <- merge.data.table(copy(d),manlc,by="rn")
    tmp <- merge.data.table(tmp,preds,by="rn")
    k2min <- refineranges[k1==curk1,min(k2thresh)]
    k2max <- refineranges[k1==curk1,max(k2thresh)]
    k2steps <- round(seq(k2min,k2max,(1/75)),3)
    k2steps <- k2steps[k2steps %in% refineranges$k2thresh] # some of our target ranges were discontinuous, for made up e.g. 0.04/0.1-0.2 and 0.04/0.4-0.5 
    k2steps <- round(k2steps-0.001,3) # subtract 0.001 to account for fractional thirds as mentioned above
    ## these steps to ssubset to k2 thresholds in the tested ranges doesn't end up working anyhow, so we'll fix it once we concatenate them all outside the loop.
    
    regs <- lapply(k2steps,function(k2){
        setDTthreads(1,restore_after_fork = FALSE)
        
        tmpdt <- copy(tmp)
        tmpdt[,newcol:=ifelse(
          #greaterOrEqual(lcpred.smooth1.k25,k1) &
            greaterOrEqual(lcpred.smooth2.k75,k2),
            "LC","Other")
        ]
        
        looseprop <- ifelse(nrow(tmpdt[LCloose=="LC"])==0,
                            0,
                            nrow(tmpdt[newcol=="LC"&LCloose=="LC"])/
                              nrow(tmpdt[LCloose=="LC"])
        )
        
        strictprop <- ifelse(nrow(tmpdt[LCstrict=="LC"])==0,
                            0,
                            nrow(tmpdt[newcol=="LC"&LCstrict=="LC"])/
                              nrow(tmpdt[LCstrict=="LC"])
        )
        
        nmfprop <- ifelse(nrow(tmpdt[WebDivLabel=="LC"])==0,
                            0,
                            nrow(tmpdt[newcol=="LC"&WebDivLabel=="LC"])/
                              nrow(tmpdt[WebDivLabel=="LC"])
        )

        od <- data.frame(matrix(nrow=1,ncol=4))
        od[1,1] <- curk1
        od[1,2] <- k2
        od[1,3] <- looseprop
        od[1,4] <- strictprop
        od[1,5] <- nmfprop
        od <- as.data.table(od)
        setnames(od,c("k1","k2","looseInDom","strictInDom","WebDivNMFInDom"))

        return(od)
    })
    ordt <- rbindlist(regs)
    return(ordt)
})

propsindoms <- rbindlist(propsindoms)

refineranges.forsubset <- refineranges[,k2thresh:=round(k2thresh-0.001,3)]
propsindoms.fix <- merge.data.table(propsindoms,refineranges.forsubset[,.(k1,k2thresh)],by.x=c("k1","k2"),by.y=c("k1","k2thresh"))
propsindoms.fix <- unique(propsindoms.fix)


fwrite(propsindoms.fix,"processed-data_newseg/04_snSeqLabelTransfer/05b-manualLCcell_wdLCcell_props_in_1thresh_fineK2steps_wdLCdomains.txt",sep='\t',quote=F,row.names=F)
rm(propsindoms,refineranges.forsubset)
```

Run registration for each xenium domain definition against each xenium manual annotation and lc vis label
```{r}
### rejigger the regperthresh thing into one big table
rpt2 <- lapply(names(regperthresh),function(n){
    X <- regperthresh[[n]]
    k1 <- gsub(n,pattern="k1_(.*)_k2.*$",replacement="\\1")
    k2 <- gsub(n,pattern="k1_.*_k2_(.*)$",replacement="\\1")
    newdat <- as.data.table(copy(X))
    newdat <- newdat[,.(ensembl,t_stat_LC)]
    setnames(newdat,2,paste0("wdLCk1_",k1,"_k2_",k2))
    return(newdat)
})

i<-1
for (i in c(1:length(rpt2))){
    if(i==1){
        regperthresh.dt <- copy(rpt2[[i]])
    } else{
        regperthresh.dt <- merge.data.table(regperthresh.dt,rpt2[[i]],by="ensembl")
    }
}

rm(rpt2,i)
```

Now, actually iterate through registrations
```{r}
regres <- mcmapply(m=comparegs,n=names(comparegs),mc.cores=4,SIMPLIFY=FALSE,FUN=function(m,n){
    m <- m[,.(ensembl,t_stat_LC)]
    setnames(m,"t_stat_LC",n)
    
    cortab <- merge.data.table(regperthresh.dt,m,by="ensembl")
    cortab$ensembl <- NULL

    j<-seq(10,300,by=10)
    i<-1
    
    outlist <- list()
    for (i in c(1:(length(j)+1))){
        cormat <- as.data.frame(matrix(nrow=ncol(cortab)-1,ncol=1))
        
        if(i<=length(j)){
            ngene <- j[i]
        } else{
            # no value for top_n --> all genes used
            ngene <- nrow(cortab)
        }
        

        k <- 1
        for (k in c(1:(ncol(cortab)-1))){
            rownames(cormat)[k] <- colnames(cortab)[k]
            colnames(cormat)[1] <- colnames(cortab)[ncol(cortab)]
            
            curthresh <- colnames(cortab)[k]
            curxen <- names(m)[2]
            setorderv(cortab,curthresh,-1)
            
            cormat[k,1] <- cor(cortab[c(1:ngene),..curthresh],cortab[c(1:ngene),..curxen],use="pairwise.complete.obs",method = "spearman")
        }
        outlist[[i]] <- cormat
        names(outlist)[i] <- paste0("DomainTop_",ngene)
    }
    return(outlist)
})

### concat into one big table
regres2 <- list()
i<-1
for (i in c(1:4)){
  x <- regres[[i]]
  j <- 1
  subregs <- list()
  for (j in c(1:length(x))){
    n <- names(x)[j]
    y <- as.data.table(x[[j]],keep.rownames=T)
    y[,Ntopdommks:=as.numeric(gsub(n,pattern="^DomainTop_(.*)$",replacement="\\1"))]
    setnames(y,"rn","domainDef")
    subregs[[j]] <- y
  }
  regres2[[i]] <- subregs
}

regres2.1 <- lapply(regres2,rbindlist)

i<-1
for (i in c(1:length(regres2.1))){
  if(i==1){
    regres.dt <- regres2.1[[i]]
  } else{
    regres.dt <- merge.data.table(regres.dt,regres2.1[[i]],by=c("domainDef","Ntopdommks"))
  }
}
regres.dt <- unique(regres.dt)

### append proportion in domain annots for xeniums. regres.dt has the subtracted 0.001 so these will match
propsindoms.fix[,domainDef:=paste0("wdLC_k1_",k1,"_k2_",k2)]

# something weird happened with the loops above so just fix it here
regres.dt[,domainDef:=gsub(domainDef,pattern="k1_LCsmooth_",replacement="k1_")]
regres.dt[,domainDef:=gsub(domainDef,pattern="wdLCk1",replacement="wdLC_k1")]
regres.dt[,domainDef:=gsub(domainDef,pattern="k2_LCsmooth_",replacement="k2_")]

## these two I cannot figure out the origin of, but they're unnecessary and introduce duplicate rows neither of which will match a domainDef
regres.dt[,domainDef:=gsub(domainDef,pattern="^(.*)\\.x$",replacement="\\1")] 
regres.dt[,domainDef:=gsub(domainDef,pattern="^(.*)\\.y$",replacement="\\1")]
regres.dt <- unique(regres.dt)

regres.dt <- merge.data.table(regres.dt,propsindoms.fix,by="domainDef")

fwrite(regres.dt,"processed-data_newseg/04_snSeqLabelTransfer/05c-1thresh_wdNMF_smoothedLCs_regist_to_manXenLC_visLC_SRTxferLC_finek2steps.txt",quote=F,row.names=F,sep='\t')
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] parallelly_1.47.0           colorout_1.3-3              spatialLIBD_1.22.0         
 [4] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
 [7] Biobase_2.70.0              GenomicRanges_1.62.1        Seqinfo_1.0.0              
[10] IRanges_2.44.0              S4Vectors_0.48.1            BiocGenerics_0.56.0        
[13] generics_0.1.4              MatrixGenerics_1.22.0       matrixStats_1.5.0          
[16] gridExtra_2.3               ggtext_0.1.2                ggplot2_4.0.2              
[19] data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3       rstudioapi_0.18.0        jsonlite_2.0.0          
  [4] shape_1.4.6.1            magrittr_2.0.4           ggbeeswarm_0.7.3        
  [7] magick_2.9.0             farver_2.1.2             rmarkdown_2.30          
 [10] BiocIO_1.20.0            GlobalOptions_0.1.3      vctrs_0.7.1             
 [13] Rsamtools_2.26.0         memoise_2.0.1            config_0.3.2            
 [16] RCurl_1.98-1.17          paletteer_1.7.0          benchmarkme_1.0.8       
 [19] htmltools_0.5.9          S4Arrays_1.10.1          AnnotationHub_4.0.0     
 [22] curl_7.0.0               BiocNeighbors_2.4.0      SparseArray_1.10.8      
 [25] sass_0.4.10              bslib_0.10.0             htmlwidgets_1.6.4       
 [28] httr2_1.2.2              plotly_4.12.0            cachem_1.1.0            
 [31] GenomicAlignments_1.46.0 mime_0.13                lifecycle_1.0.5         
 [34] iterators_1.0.14         pkgconfig_2.0.3          rsvd_1.0.5              
 [37] Matrix_1.7-4             R6_2.6.1                 fastmap_1.2.0           
 [40] shiny_1.12.1             clue_0.3-66              digest_0.6.39           
 [43] colorspace_2.1-2         rematch2_2.1.2           AnnotationDbi_1.72.0    
 [46] scater_1.38.1            rprojroot_2.1.1          irlba_2.3.7             
 [49] ExperimentHub_3.0.0      RSQLite_2.4.5            beachmat_2.26.0         
 [52] filelock_1.0.3           httr_1.4.7               abind_1.4-8             
 [55] compiler_4.5.2           here_1.0.2               bit64_4.6.0-1           
 [58] withr_3.0.2              doParallel_1.0.17        attempt_0.3.1           
 [61] S7_0.2.1                 BiocParallel_1.44.0      viridis_0.6.5           
 [64] DBI_1.2.3                sessioninfo_1.2.3        rappdirs_0.3.4          
 [67] DelayedArray_0.36.1      rjson_0.2.23             tools_4.5.2             
 [70] vipor_0.4.7              otel_0.2.0               beeswarm_0.4.0          
 [73] httpuv_1.6.16            glue_1.8.0               restfulr_0.0.16         
 [76] promises_1.5.0           gridtext_0.1.5           grid_4.5.2              
 [79] rsconnect_1.7.0          cluster_2.1.8.1          gtable_0.3.6            
 [82] tidyr_1.3.2              ScaledMatrix_1.18.0      BiocSingular_1.26.0     
 [85] xml2_1.5.2               XVector_0.50.0           ggrepel_0.9.6           
 [88] BiocVersion_3.22.0       foreach_1.5.2            pillar_1.11.1           
 [91] limma_3.66.0             later_1.4.5              circlize_0.4.17         
 [94] benchmarkmeData_2.0.0    dplyr_1.2.0              BiocFileCache_3.0.0     
 [97] lattice_0.22-7           rtracklayer_1.70.1       bit_4.6.0               
[100] tidyselect_1.2.1         ComplexHeatmap_2.26.1    locfit_1.5-9.12         
[103] scuttle_1.20.0           Biostrings_2.78.0        knitr_1.51              
[106] edgeR_4.8.2              xfun_0.56                statmod_1.5.1           
[109] DT_0.34.0                lazyeval_0.2.2           yaml_2.3.12             
[112] shinyWidgets_0.9.0       cigarillo_1.0.0          evaluate_1.0.5          
[115] codetools_0.2-20         tibble_3.3.1             BiocManager_1.30.27     
[118] cli_3.6.5                xtable_1.8-4             jquerylib_0.1.4         
[121] golem_0.5.1              dichromat_2.0-0.1        Rcpp_1.1.1              
[124] dbplyr_2.5.1             png_0.1-8                XML_3.99-0.20           
[127] blob_1.3.0               bitops_1.0-9             viridisLite_0.4.3       
[130] scales_1.4.0             purrr_1.2.1              crayon_1.5.3            
[133] GetoptLong_1.1.0         rlang_1.1.7              cowplot_1.2.0           
[136] KEGGREST_1.50.0         
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-03
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ─────────────────────────────────────────────────────────────────────────────────
 package              * version   date (UTC) lib source
 abind                  1.4-8     2024-09-12 [1] CRAN (R 4.5.0)
 AnnotationDbi          1.72.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationHub          4.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 attempt                0.3.1     2020-05-03 [1] CRAN (R 4.5.0)
 beachmat               2.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.5.0)
 benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.5.0)
 benchmarkmeData        2.0.0     2026-01-19 [1] CRAN (R 4.5.2)
 Biobase              * 2.70.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocFileCache          3.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocIO                 1.20.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocManager            1.30.27   2025-11-14 [1] CRAN (R 4.5.2)
 BiocNeighbors          2.4.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel           1.44.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocVersion            3.22.0    2025-10-07 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0     2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1   2025-01-16 [1] CRAN (R 4.5.0)
 bitops                 1.0-9     2024-10-03 [1] CRAN (R 4.5.0)
 blob                   1.3.0     2026-01-14 [1] CRAN (R 4.5.2)
 bslib                  0.10.0    2026-01-26 [1] CRAN (R 4.5.2)
 cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.5.0)
 cigarillo              1.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 circlize               0.4.17    2025-12-08 [1] CRAN (R 4.5.2)
 cli                    3.6.5     2025-04-23 [1] CRAN (R 4.5.0)
 clue                   0.3-66    2024-11-13 [1] CRAN (R 4.5.0)
 cluster                2.1.8.1   2025-03-12 [1] CRAN (R 4.5.0)
 codetools              0.2-20    2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3     2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 colorspace             2.1-2     2025-09-22 [1] CRAN (R 4.5.0)
 ComplexHeatmap         2.26.1    2026-01-30 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 config                 0.3.2     2023-08-30 [1] CRAN (R 4.5.0)
 cowplot                1.2.0     2025-07-07 [1] CRAN (R 4.5.0)
 crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0     2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1  2026-01-27 [1] CRAN (R 4.5.2)
 DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.5.0)
 dbplyr                 2.5.1     2025-09-10 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.1    2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1   2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39    2025-11-19 [1] CRAN (R 4.5.2)
 doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.5.0)
 dplyr                  1.2.0     2026-02-03 [1] CRAN (R 4.5.2)
 DT                     0.34.0    2025-09-02 [1] CRAN (R 4.5.0)
 edgeR                  4.8.2     2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate               1.0.5     2025-08-27 [1] CRAN (R 4.5.0)
 ExperimentHub          3.0.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 farver                 2.1.2     2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.5.0)
 filelock               1.0.3     2023-12-11 [1] CRAN (R 4.5.0)
 foreach                1.5.2     2022-02-02 [1] CRAN (R 4.5.0)
 generics             * 0.1.4     2025-05-09 [1] CRAN (R 4.5.0)
 GenomicAlignments      1.46.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 GenomicRanges        * 1.62.1    2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 GetoptLong             1.1.0     2025-11-28 [1] CRAN (R 4.5.2)
 ggbeeswarm             0.7.3     2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2     2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6     2024-09-07 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2     2022-09-16 [1] CRAN (R 4.5.0)
 GlobalOptions          0.1.3     2025-11-28 [1] CRAN (R 4.5.2)
 glue                   1.8.0     2024-09-30 [1] CRAN (R 4.5.0)
 golem                  0.5.1     2024-08-27 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5     2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6     2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2     2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9     2025-12-04 [1] CRAN (R 4.5.2)
 htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.5.0)
 httpuv                 1.6.16    2025-04-16 [1] CRAN (R 4.5.0)
 httr                   1.4.7     2023-08-15 [1] CRAN (R 4.5.0)
 httr2                  1.2.2     2025-12-08 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7     2026-01-30 [1] CRAN (R 4.5.2)
 iterators              1.0.14    2022-02-05 [1] CRAN (R 4.5.0)
 jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.5.0)
 jsonlite               2.0.0     2025-03-27 [1] CRAN (R 4.5.0)
 KEGGREST               1.50.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51      2025-12-20 [1] CRAN (R 4.5.2)
 later                  1.4.5     2026-01-08 [1] CRAN (R 4.5.2)
 lattice                0.22-7    2025-04-02 [1] CRAN (R 4.5.2)
 lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.5.0)
 lifecycle              1.0.5     2026-01-08 [1] CRAN (R 4.5.2)
 limma                  3.66.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 locfit                 1.5-9.12  2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0     2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4     2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4     2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0     2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1     2021-11-26 [1] CRAN (R 4.5.0)
 mime                   0.13      2025-03-17 [1] CRAN (R 4.5.0)
 otel                   0.2.0     2025-08-29 [1] CRAN (R 4.5.0)
 paletteer              1.7.0     2026-01-08 [1] CRAN (R 4.5.2)
 parallelly           * 1.47.0    2026-04-17 [1] CRAN (R 4.5.2)
 pillar                 1.11.1    2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.5.0)
 plotly                 4.12.0    2026-01-24 [1] CRAN (R 4.5.2)
 png                    0.1-8     2022-11-29 [1] CRAN (R 4.5.0)
 promises               1.5.0     2025-11-01 [1] CRAN (R 4.5.0)
 purrr                  1.2.1     2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1     2025-02-15 [1] CRAN (R 4.5.0)
 rappdirs               0.3.4     2026-01-17 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1     2026-01-10 [1] CRAN (R 4.5.2)
 RCurl                  1.98-1.17 2025-03-22 [1] CRAN (R 4.5.0)
 rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.5.0)
 restfulr               0.0.16    2025-06-27 [1] CRAN (R 4.5.0)
 rjson                  0.2.23    2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7     2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30      2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1     2025-08-26 [1] CRAN (R 4.5.0)
 Rsamtools              2.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 rsconnect              1.7.0     2025-12-06 [1] CRAN (R 4.5.2)
 RSQLite                2.4.5     2025-11-30 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0    2026-01-16 [1] CRAN (R 4.5.2)
 rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.5.0)
 rtracklayer            1.70.1    2025-12-20 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 S4Arrays               1.10.1    2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1    2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1     2025-11-14 [1] CRAN (R 4.5.2)
 sass                   0.4.10    2025-04-11 [1] CRAN (R 4.5.0)
 ScaledMatrix           1.18.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0     2025-04-24 [1] CRAN (R 4.5.0)
 scater                 1.38.1    2026-03-20 [1] Bioconductor 3.22 (R 4.5.2)
 scuttle                1.20.0    2025-10-30 [1] Bioconductor 3.22 (R 4.5.2)
 Seqinfo              * 1.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3     2025-02-05 [1] CRAN (R 4.5.0)
 shape                  1.4.6.1   2024-02-23 [1] CRAN (R 4.5.0)
 shiny                  1.12.1    2025-12-09 [1] CRAN (R 4.5.2)
 shinyWidgets           0.9.0     2025-02-21 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8    2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 spatialLIBD          * 1.22.0    2025-11-04 [1] Bioconductor 3.22 (R 4.5.2)
 statmod                1.5.1     2025-10-09 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1     2026-01-11 [1] CRAN (R 4.5.2)
 tidyr                  1.3.2     2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1     2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.5.0)
 viridis                0.6.5     2024-01-29 [1] CRAN (R 4.5.0)
 viridisLite            0.4.3     2026-02-04 [1] CRAN (R 4.5.2)
 withr                  3.0.2     2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56      2026-01-18 [1] CRAN (R 4.5.2)
 XML                    3.99-0.20 2025-11-08 [1] CRAN (R 4.5.0)
 xml2                   1.5.2     2026-01-17 [1] CRAN (R 4.5.2)
 xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.5.0)
 XVector                0.50.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12    2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

────────────────────────────────────────────────────────────────────────────────────────────