---
title: "06-OneThreshSmoothLC_webDivLabelsAndDomain_LUT.Rmd"
format: html
date: "06-03-2026"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
# library(SpatialFeatureExperiment)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}


## rstudio GUI tweaks
require(colorout)
ColorOut()
options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
library(parallel)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)

## ggplot defaults

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

0.08/0.492
0.12/0.412
load smoothings, xenium, and manual xenium LC assignments
```{r}
## smoothings
dualscores <- readRDS("processed-data_newseg/04_snSeqLabelTransfer/03b-WebDivLCknnsmoothing_step2k75.RDS")
dualscores <- dualscores[c("k1thresh_0.08","k1thresh_0.12")]

dualscores8492 <- dualscores$k1thresh_0.08[,domain8492:=ifelse(greaterOrEqual(lcpred.smooth2.k75,round(0.492,3)),"LC","Other")]

dualscores12412 <- dualscores$k1thresh_0.12[,domain12412:=ifelse(greaterOrEqual(lcpred.smooth2.k75,round(0.412,3)),"LC","Other")]

## append nmf predictions
projlabs <- fread("processed-data_newseg/04_snSeqLabelTransfer/02-WebDivToXen_xferedlabels.txt")
dualscores8492 <- merge.data.table(dualscores8492,projlabs,by="rn")
dualscores12412 <- merge.data.table(dualscores12412,projlabs,by="rn")

nrow(dualscores8492[WebDivLabel.x=="NE"]) #11065
nrow(dualscores8492[domain8492=="LC"]) #43683
nrow(dualscores8492[domain8492=="LC"&WebDivLabel.x=="NE"]) # 8873

nrow(dualscores12412[WebDivLabel=="NE"]) # 11065 (total pred cells so same)
nrow(dualscores12412[domain12412=="LC"]) # #37328
nrow(dualscores12412[domain12412=="LC"&WebDivLabel=="NE"]) # 8313

nrow(dualscores8492[domain8492=="LC" & rn %in% dualscores12412[domain12412=="LC",rn]]) 
## ^ 36677; therefore 651 unique cells exclusive to domain12412 and 6355 cells exclusive to domain8492

domlabs <- merge.data.table(dualscores12412[,.(rn,WebDivLabel,domain12412)],dualscores8492[,.(rn,domain8492)],by="rn")
setnames(domlabs,"WebDivLabel","nmfpred")

fwrite(domlabs,"processed-data_newseg/04_snSeqLabelTransfer/06-wdLCsmoothings_LUT.txt",sep='\t',quote=F,row.names=F,col.names=T)
```
seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] parallelly_1.46.1   colorout_1.3-3      data.table_1.18.2.1

loaded via a namespace (and not attached):
 [1] digest_0.6.39     fastmap_1.2.0     xfun_0.56         knitr_1.51       
 [5] htmltools_0.5.9   rmarkdown_2.30    cli_3.6.5         compiler_4.5.2   
 [9] rprojroot_2.1.1   here_1.0.2        rstudioapi_0.18.0 tools_4.5.2      
[13] evaluate_1.0.5    yaml_2.3.12       otel_0.2.0        rlang_1.1.7      