---
title: "02-Examine_MapMyCells_output"
format: html
author: "Bernie Mulvey"
date: "06-16-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
# library(SpatialFeatureExperiment)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
precision <- sqrt(.Machine$double.eps)
(x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
precision <- sqrt(.Machine$double.eps)
(x <= y) | (abs(x-y) <= precision)}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost='localhost')


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = '#000000'), axis.title.x = element_text(size = 9,color = '#000000'), axis.text.y = element_text(size = 8,color = '#000000'), axis.title.y = element_text(size =9,color = '#000000'), plot.title = element_markdown(size = 10,hjust=0.5,color = '#000000'), strip.text = element_text(size=11), legend.text = element_text(size=7,color = '#000000'), legend.title = element_text(size=8,hjust=0.5,color = '#000000'),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position='plot'))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```


read in cell assignments table 
```{r}

mmc <- fread("processed-data_newseg/12_MapMyCells/01-LCX_mapmycellsUploadCtscsv_10xWholeHumanBrain(CCN202210140)_HierarchicalMapping_UTC_1781663294501/01-LCX_mapmycellsUploadCtscsv_10xWholeHumanBrain(CCN202210140)_HierarchicalMapping_UTC_1781663294501.csv",skip = 4)
names(mmc)
# [1] "cell_id"                               
#  [2] "supercluster_label"                    
#  [3] "supercluster_name"                     
#  [4] "supercluster_bootstrapping_probability"
#  [5] "cluster_label"                         
#  [6] "cluster_name"                          
#  [7] "cluster_bootstrapping_probability"     
#  [8] "subcluster_label"                      
#  [9] "subcluster_name"                       
# [10] "subcluster_alias"                      
# [11] "subcluster_bootstrapping_probab

### uncommon clusters--probably cells we would have to throw out because given some other brain region's labels
mmc[,.N,by="supercluster_name"][N<1000]
#  1:               Thalamic excitatory   563
#  2:             Cerebellar inhibitory   954
#  3:                 Lower rhombic lip   174
#  4:        Deep-layer near-projecting   164
#  5:                 Upper rhombic lip   707
#  6:                   MGE interneuron    48
#  7:               Amygdala excitatory    22
#  8: Deep-layer corticothalamic and 6b     9
#  9:                    Choroid plexus    78
# 10:                     Bergmann glia    33
# 11:                   Mammillary body    39
# 12:               Medium spiny neuron     8
# 13:    Upper-layer intratelencephalic    17
# 14:     Eccentric medium spiny neuron     1
# 15:     Deep-layer intratelencephalic    10
# 16:         Hippocampal dentate gyrus     3
# 17:         LAMP5-LHX6 and Chandelier     1

## how many cells not even given a real supercluster ("splatter")?
nrow(mmc[supercluster_name=="Splatter"]) # 70180 or like 15% of cells...cool -____-
```

## ok, cerebellar inhibitory could be legitimate. and the rhombic lips, choroid, interneurons. the others, no way
## ABA clustering doesn't have regions like LC/raphe at the supercluster level of hierarchy--and that dataset had
## virtually no LC-NE cells anyhow. so let's see how the weber-divecha label transferred NE cells interplay here and what kind of cell labels they would be replacing.

```{r}
wdx <- fread("processed-data_newseg/04_snSeqLabelTransfer/06-wdLCsmoothings_LUT.txt")

mmc <- merge.data.table(mmc,wdx,by.x="cell_id",by.y="rn")

ne <- mmc[nmfpred=="NE"]

ggplot(
ne[,.N,by="supercluster_name"],aes(y=supercluster_name,x=N))+geom_col()
## well THAT's interesting. 7k of 11k cells map to splatter of c. but,
#  1400 to astro (!) 
# 880 to oligo
# 619 to vascular
# 609 microglia
```

Well, since these are about as close to "gold standard" labels as we can get, let's see how they look spatially, and behave in terms of DE.
