---
title: "03-Make_MapMyCells_DGElist_withWDxferNEoverride.Rmd"
format: html
author: "Bernie Mulvey"
date: "06-17-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(dreamlet)
library(edgeR)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
# require(colorout)
# ColorOut()
# options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
# ##

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
precision <- sqrt(.Machine$double.eps)
(x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
precision <- sqrt(.Machine$double.eps)
(x <= y) | (abs(x-y) <= precision)}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost='localhost')


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = '#000000'), axis.title.x = element_text(size = 9,color = '#000000'), axis.text.y = element_text(size = 8,color = '#000000'), axis.title.y = element_text(size =9,color = '#000000'), plot.title = element_markdown(size = 10,hjust=0.5,color = '#000000'), strip.text = element_text(size=11), legend.text = element_text(size=7,color = '#000000'), legend.title = element_text(size=8,hjust=0.5,color = '#000000'),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position='plot'))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```


```{r}
xlc <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

## load weber-divecha label transfers to get NE labeled cell ids
wdxfer <- fread("processed-data_newseg/04_snSeqLabelTransfer/06-wdLCsmoothings_LUT.txt")

## load mapmycells output, attach weber divecha label transfers
mmc <- fread("processed-data_newseg/12_MapMyCells/01-LCX_mapmycellsUploadCtscsv_10xWholeHumanBrain(CCN202210140)_HierarchicalMapping_UTC_1781663294501/01-LCX_mapmycellsUploadCtscsv_10xWholeHumanBrain(CCN202210140)_HierarchicalMapping_UTC_1781663294501.csv",skip=4)
mmc <- merge.data.table(mmc,wdxfer[,.(rn,nmfpred)],by.x="cell_id",by.y="rn")

## drop cells assigned to the following clusters IF they aren't assigned the wd-NE label: 
#  1:               Thalamic excitatory   563
#  4:        Deep-layer near-projecting   164
#  6:                   MGE interneuron    48
#  7:               Amygdala excitatory    22
#  8: Deep-layer corticothalamic and 6b     9
# 10:                     Bergmann glia    33
# 11:                   Mammillary body    39
# 12:               Medium spiny neuron     8
# 13:    Upper-layer intratelencephalic    17
# 14:     Eccentric medium spiny neuron     1
# 15:     Deep-layer intratelencephalic    10
# 16:         Hippocampal dentate gyrus     3
# 17:         LAMP5-LHX6 and Chandelier     1

### before dropping mapmycells clusters that are implausible/hyperrare, override with wdNE where applicable
mmc[nmfpred=="NE",supercluster_name:="NE"]
mmc <- mmc[!(supercluster_name %in% c("Thalamic excitatory","Deep-layer near-projecting","Deep-layer corticothalamic and 6b","Deep-layer intratelencephalic","MGE interneuron","Amygdala excitatory","Bergmann glia","Mammillary body","Medium spiny neuron","Eccentric medium spiny neuron","Upper-layer intratelencephalic","Hippocampal dentate gyrus","LAMP5-LHX6 and Chandelier"))]
keepcell <- mmc$cell_id
xlc <- xlc[,keepcell]

## tidy the remaining names
#                       supercluster_name      N
#                                  <char>  <int>
#  1:                           Microglia  35833
#  2:                          Fibroblast  15961
#  3:                            Vascular  75300
#  4:                            Splatter  70180
#  5:                     Oligodendrocyte 121308
#  6:                           Astrocyte 126179
#  7:           Oligodendrocyte precursor  15685
#  8: Committed oligodendrocyte precursor   9225
#  9:                       Miscellaneous   8158
# 10:               Cerebellar inhibitory    954
# 11:         Midbrain-derived inhibitory   3055
# 12:                     CGE interneuron   1972
# 13:                           Ependymal   6095
# 14:                   Lower rhombic lip    174
# 15:                   Upper rhombic lip    707
# 16:                      Choroid plexus     78
mmc[supercluster_name=="Choroid plexus",supercluster_name:="Choroid"]
mmc[supercluster_name=="Upper rhombic lip",supercluster_name:="Upper rhomb. l."]
mmc[supercluster_name=="Lower rhombic lip",supercluster_name:="Lower rhomb. l."]
mmc[supercluster_name=="CGE interneuron",supercluster_name:="CGE iN"]
mmc[supercluster_name=="Midbrain-derived inhibitory",supercluster_name:="Midbrain-der. GABA"]
mmc[supercluster_name=="Cerebellar inhibitory",supercluster_name:="Cerebel. GABA"]
mmc[supercluster_name=="Miscellaneous",supercluster_name:="Misc"]
mmc[supercluster_name=="Committed oligodendrocyte precursor",supercluster_name:="Committed OPC"]
mmc[supercluster_name=="Oligodendrocyte precursor",supercluster_name:="OPC"]
mmc[supercluster_name=="Astrocyte",supercluster_name:="Astro"]
mmc[supercluster_name=="Oligodendrocyte",supercluster_name:="Oligo"]

## for DE runs, we can't have spaces or punctuation in the labels, so change to underscores
mmc[,supercluster_name:=gsub(supercluster_name,pattern=" ",replacement="_")]
mmc[,supercluster_name:=gsub(supercluster_name,pattern="\\.",replacement="_")]
mmc[,supercluster_name:=gsub(supercluster_name,pattern="-",replacement="_")]

mmc <- DataFrame(mmc,row.names=mmc$cell_id)[colnames(xlc),]
xlc$label <- mmc$supercluster_name
```


make DGEList and save
```{r}
ag <- dreamlet::aggregateToPseudoBulk(xlc,assay="counts",sample_id="brnum",cluster_id = "label",BPPARAM = MulticoreParam(8))

## get pbulk metadata
agcd <- as.data.table(colData(ag),keep.rownames=T)
setnames(agcd,"rn","brnum")
## add a simple E4/E2 column
agcd[,APOEsimple:=ifelse(APOE %in% c("E4het","E4hom"),"E4","E2")]

## number of cells per pbulk
agn <- as.data.table(ag@int_colData$n_cells,keep.rownames=T)
setnames(agn,"rn","label")
agn <- melt.data.table(agn,id.vars = "label")
setnames(agn,"variable","brnum")
agn[,brnum:=as.character(brnum)]
agn[,clus_brnum:=paste0(label,"_",brnum)]
agn <- agn[value>=10] ## these are the samples we want to keep to start: 10+ cells for domain type
setnames(agn,"value","ncells")


agm <- as.list(assays(ag))
agc <- unlist(mapply(x=agm,n=names(agm),SIMPLIFY=FALSE,function(x,n){
    x <- x[,colnames(x) %in% agn[label==n,brnum]]
    if(is.vector(x)){return(0)} else {
    sum(greaterOrEqual(colSums(x),2000))}})) ## number of pseudobulks with 10+ cells also having 2000+ counts
## only keep those clusters where we have 10+ such pbulks
kpb <- names(agc[agc>=10])
## filter number of cells table as our pbulk identifier lookup
agn <- agn[label %in% kpb]
## filter pseudobulk count matrices to clusters being retained
agm <- agm[kpb]

## now we want to make one giant sample matrix with the remaining pseudobulks
agm.filt <- mapply(a=agm,n=names(agm),SIMPLIFY=FALSE,function(a,n){
    a <- a[,colnames(a) %in% agn[label==n]$brnum]
    colnames(a) <- paste0(n,"_",colnames(a))
    return(a)
})
agm.filt <- do.call("cbind",agm.filt)
# dim(agm.filt)
# 366 464

# make a per cluster-donor samples table
agn <- merge.data.table(agn,agcd[,.(brnum,sample_id,Sex,APOE,APOEsimple,YRI,Age)],by="brnum")
stopifnot(all(agn$clus_brnum %in% colnames(agm.filt)))
agn <- as.data.frame(agn,row.names=agn$clus_brnum)[colnames(agm.filt),]
gtab <- as.data.frame(rowData(xlc)[,c("Symbol","ID")])
rownames(gtab) <- gtab$Symbol
gtab <- gtab[rownames(agm.filt),]
stopifnot(all(rownames(gtab) %in% rownames(agm.filt)&all(rownames(agm.filt)==rownames(gtab))))
dge <- DGEList(counts=agm.filt, samples=agn,genes=gtab)

## save
saveRDS(dge,"processed-data_newseg/12_MapMyCells/03-MapMyCellsDGEpbulk_wdXferNEoverride.RDS")
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics 
[5] grDevices utils     methods   base     

other attached packages:
 [1] edgeR_4.8.2                
 [2] dreamlet_1.8.0             
 [3] variancePartition_1.40.1   
 [4] BiocParallel_1.44.0        
 [5] limma_3.66.0               
 [6] parallelly_1.47.0          
 [7] colorout_1.3-3             
 [8] Polychrome_1.5.4           
 [9] escheR_1.10.0              
[10] SpatialExperiment_1.20.0   
[11] SingleCellExperiment_1.32.0
[12] SummarizedExperiment_1.40.0
[13] Biobase_2.70.0             
[14] GenomicRanges_1.62.1       
[15] Seqinfo_1.0.0              
[16] IRanges_2.44.0             
[17] S4Vectors_0.48.1           
[18] BiocGenerics_0.56.0        
[19] generics_0.1.4             
[20] MatrixGenerics_1.22.0      
[21] matrixStats_1.5.0          
[22] ggrastr_1.0.2              
[23] gridExtra_2.3              
[24] ggtext_0.1.2               
[25] ggplot2_4.0.2              
[26] data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] splines_4.5.2            
  [2] bitops_1.0-9             
  [3] tibble_3.3.1             
  [4] graph_1.88.1             
  [5] XML_3.99-0.20            
  [6] lifecycle_1.0.5          
  [7] Rdpack_2.6.5             
  [8] mixsqp_0.3-54            
  [9] rprojroot_2.1.1          
 [10] lattice_0.22-7           
 [11] MASS_7.3-65              
 [12] backports_1.5.0          
 [13] magrittr_2.0.4           
 [14] metafor_4.8-0            
 [15] otel_0.2.0               
 [16] sessioninfo_1.2.3        
 [17] DBI_1.2.3                
 [18] minqa_1.2.8              
 [19] RColorBrewer_1.1-3       
 [20] abind_1.4-8              
 [21] EnvStats_3.1.0           
 [22] purrr_1.2.1              
 [23] msigdbr_25.1.1           
 [24] RCurl_1.98-1.17          
 [25] rmeta_3.0                
 [26] ggrepel_0.9.6            
 [27] pbkrtest_0.5.5           
 [28] irlba_2.3.7              
 [29] annotate_1.88.0          
 [30] DelayedMatrixStats_1.32.0
 [31] codetools_0.2-20         
 [32] DelayedArray_0.36.1      
 [33] xml2_1.5.2               
 [34] tidyselect_1.2.1         
 [35] farver_2.1.2             
 [36] lme4_1.1-38              
 [37] mathjaxr_2.0-0           
 [38] iterators_1.0.14         
 [39] progress_1.2.3           
 [40] tools_4.5.2              
 [41] Rcpp_1.1.1               
 [42] zenith_1.12.0            
 [43] glue_1.8.0               
 [44] SparseArray_1.10.8       
 [45] xfun_0.56                
 [46] here_1.0.2               
 [47] dplyr_1.2.0              
 [48] withr_3.0.2              
 [49] numDeriv_2016.8-1.1      
 [50] fastmap_1.2.0            
 [51] boot_1.3-32              
 [52] caTools_1.18.3           
 [53] digest_0.6.39            
 [54] truncnorm_1.0-9          
 [55] R6_2.6.1                 
 [56] colorspace_2.1-2         
 [57] scattermore_1.2          
 [58] gtools_3.9.5             
 [59] dichromat_2.0-0.1        
 [60] RSQLite_2.4.5            
 [61] RhpcBLASctl_0.23-42      
 [62] tidyr_1.3.2              
 [63] corpcor_1.6.10           
 [64] prettyunits_1.2.0        
 [65] httr_1.4.7               
 [66] S4Arrays_1.10.1          
 [67] scatterplot3d_0.3-44     
 [68] pkgconfig_2.0.3          
 [69] gtable_0.3.6             
 [70] blob_1.3.0               
 [71] S7_0.2.1                 
 [72] XVector_0.50.0           
 [73] remaCor_0.0.20           
 [74] zigg_0.0.2               
 [75] GSEABase_1.72.0          
 [76] scales_1.4.0             
 [77] png_0.1-8                
 [78] fANCOVA_0.6-1            
 [79] reformulas_0.4.4         
 [80] ashr_2.2-63              
 [81] knitr_1.51               
 [82] reshape2_1.4.5           
 [83] rjson_0.2.23             
 [84] curl_7.0.0               
 [85] nlme_3.1-168             
 [86] nloptr_2.2.1             
 [87] cachem_1.1.0             
 [88] stringr_1.6.0            
 [89] KernSmooth_2.23-26       
 [90] vipor_0.4.7              
 [91] metadat_1.4-0            
 [92] AnnotationDbi_1.72.0     
 [93] pillar_1.11.1            
 [94] grid_4.5.2               
 [95] vctrs_0.7.1              
 [96] gplots_3.3.0             
 [97] mashr_0.2.79             
 [98] xtable_1.8-4             
 [99] beeswarm_0.4.0           
[100] Rgraphviz_2.54.0         
[101] evaluate_1.0.5           
[102] KEGGgraph_1.70.0         
[103] invgamma_1.2             
[104] magick_2.9.0             
[105] mvtnorm_1.3-3            
[106] cli_3.6.5                
[107] locfit_1.5-9.12          
[108] compiler_4.5.2           
[109] rlang_1.1.7              
[110] crayon_1.5.3             
[111] SQUAREM_2021.1           
[112] labeling_0.4.3           
[113] plyr_1.8.9               
[114] ggbeeswarm_0.7.3         
[115] stringi_1.8.7            
[116] babelgene_22.9           
[117] assertthat_0.2.1         
[118] lmerTest_3.2-0           
[119] Biostrings_2.78.0        
[120] aod_1.3.3                
[121] Matrix_1.7-4             
[122] hms_1.1.4                
[123] sparseMatrixStats_1.22.0 
[124] bit64_4.6.0-1            
[125] KEGGREST_1.50.0          
[126] statmod_1.5.1            
[127] rbibutils_2.4.1          
[128] Rfast_2.1.5.2            
[129] gridtext_0.1.5           
[130] broom_1.0.12             
[131] memoise_2.0.1            
[132] RcppParallel_5.1.11-1    
[133] bit_4.6.0                
[134] EnrichmentBrowser_2.40.0 
─ Session info ──────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       Positron
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-17
 pandoc   NA
 quarto   1.8.27 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [1] CRAN (R 4.5.0)
 annotate               1.88.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationDbi          1.72.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 aod                    1.3.3      2023-12-13 [1] CRAN (R 4.5.0)
 ashr                   2.2-63     2023-08-21 [1] CRAN (R 4.5.0)
 assertthat             0.2.1      2019-03-21 [1] CRAN (R 4.5.0)
 babelgene              22.9       2022-09-29 [1] CRAN (R 4.5.0)
 backports              1.5.0      2024-05-23 [1] CRAN (R 4.5.0)
 beeswarm               0.4.0      2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0      2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1    2025-01-16 [1] CRAN (R 4.5.0)
 bitops                 1.0-9      2024-10-03 [1] CRAN (R 4.5.0)
 blob                   1.3.0      2026-01-14 [1] CRAN (R 4.5.2)
 boot                   1.3-32     2025-08-29 [1] CRAN (R 4.5.2)
 broom                  1.0.12     2026-01-27 [1] CRAN (R 4.5.2)
 cachem                 1.1.0      2024-05-16 [1] CRAN (R 4.5.0)
 caTools                1.18.3     2024-09-04 [1] CRAN (R 4.5.0)
 cli                    3.6.5      2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3      2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 colorspace             2.1-2      2025-09-22 [1] CRAN (R 4.5.0)
 corpcor                1.6.10     2021-09-16 [1] CRAN (R 4.5.0)
 crayon                 1.5.3      2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0      2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1   2026-01-27 [1] CRAN (R 4.5.2)
 DBI                    1.2.3      2024-06-02 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.1     2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 DelayedMatrixStats     1.32.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0      2026-02-03 [1] CRAN (R 4.5.2)
 dreamlet             * 1.8.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 edgeR                * 4.8.2      2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 EnrichmentBrowser      2.40.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 EnvStats               3.1.0      2025-04-24 [1] CRAN (R 4.5.0)
 escheR               * 1.10.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 evaluate               1.0.5      2025-08-27 [1] CRAN (R 4.5.0)
 fANCOVA                0.6-1      2020-11-13 [1] CRAN (R 4.5.0)
 farver                 2.1.2      2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4      2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1     2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 ggbeeswarm             0.7.3      2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2      2026-02-03 [1] CRAN (R 4.5.2)
 ggrastr              * 1.0.2      2023-06-01 [1] CRAN (R 4.5.0)
 ggrepel                0.9.6      2024-09-07 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2      2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0      2024-09-30 [1] CRAN (R 4.5.0)
 gplots                 3.3.0      2025-11-30 [1] CRAN (R 4.5.2)
 graph                  1.88.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 gridExtra            * 2.3        2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5      2022-09-16 [1] CRAN (R 4.5.0)
 GSEABase               1.72.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 gtable                 0.3.6      2024-10-25 [1] CRAN (R 4.5.0)
 gtools                 3.9.5      2023-11-20 [1] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [1] CRAN (R 4.5.0)
 hms                    1.1.4      2025-10-17 [1] CRAN (R 4.5.0)
 httr                   1.4.7      2023-08-15 [1] CRAN (R 4.5.0)
 invgamma               1.2        2025-07-02 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [1] CRAN (R 4.5.2)
 iterators              1.0.14     2022-02-05 [1] CRAN (R 4.5.0)
 KEGGgraph              1.70.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KEGGREST               1.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KernSmooth             2.23-26    2025-01-01 [1] CRAN (R 4.5.2)
 knitr                  1.51       2025-12-20 [1] CRAN (R 4.5.2)
 labeling               0.4.3      2023-08-29 [1] CRAN (R 4.5.0)
 lattice                0.22-7     2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5      2026-01-08 [1] CRAN (R 4.5.2)
 limma                * 3.66.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 lme4                   1.1-38     2025-12-02 [1] CRAN (R 4.5.2)
 lmerTest               3.2-0      2026-01-13 [1] CRAN (R 4.5.2)
 locfit                 1.5-9.12   2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0      2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [1] CRAN (R 4.5.0)
 mashr                  0.2.79     2023-10-18 [1] CRAN (R 4.5.0)
 MASS                   7.3-65     2025-02-28 [1] CRAN (R 4.5.2)
 mathjaxr               2.0-0      2025-12-01 [1] CRAN (R 4.5.2)
 Matrix                 1.7-4      2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1      2021-11-26 [1] CRAN (R 4.5.0)
 metadat                1.4-0      2025-02-04 [1] CRAN (R 4.5.0)
 metafor                4.8-0      2025-01-28 [1] CRAN (R 4.5.0)
 minqa                  1.2.8      2024-08-17 [1] CRAN (R 4.5.0)
 mixsqp                 0.3-54     2023-12-20 [1] CRAN (R 4.5.0)
 msigdbr                25.1.1     2025-07-21 [1] CRAN (R 4.5.2)
 mvtnorm                1.3-3      2025-01-10 [1] CRAN (R 4.5.0)
 nlme                   3.1-168    2025-03-31 [1] CRAN (R 4.5.2)
 nloptr                 2.2.1      2025-03-17 [1] CRAN (R 4.5.0)
 numDeriv               2016.8-1.1 2019-06-06 [1] CRAN (R 4.5.0)
 otel                   0.2.0      2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.47.0     2026-04-17 [1] CRAN (R 4.5.2)
 pbkrtest               0.5.5      2025-07-18 [1] CRAN (R 4.5.0)
 pillar                 1.11.1     2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [1] CRAN (R 4.5.0)
 plyr                   1.8.9      2023-10-02 [1] CRAN (R 4.5.0)
 png                    0.1-8      2022-11-29 [1] CRAN (R 4.5.0)
 Polychrome           * 1.5.4      2025-04-06 [1] CRAN (R 4.5.0)
 prettyunits            1.2.0      2023-09-24 [1] CRAN (R 4.5.0)
 progress               1.2.3      2023-12-06 [1] CRAN (R 4.5.0)
 purrr                  1.2.1      2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1      2025-02-15 [1] CRAN (R 4.5.0)
 rbibutils              2.4.1      2026-01-21 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3      2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [1] CRAN (R 4.5.2)
 RcppParallel           5.1.11-1   2025-08-27 [1] CRAN (R 4.5.0)
 RCurl                  1.98-1.17  2025-03-22 [1] CRAN (R 4.5.0)
 Rdpack                 2.6.5      2026-01-23 [1] CRAN (R 4.5.2)
 reformulas             0.4.4      2026-02-02 [1] CRAN (R 4.5.2)
 remaCor                0.0.20     2025-08-20 [1] CRAN (R 4.5.0)
 reshape2               1.4.5      2025-11-12 [1] CRAN (R 4.5.0)
 Rfast                  2.1.5.2    2025-10-10 [1] CRAN (R 4.5.0)
 Rgraphviz              2.54.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 RhpcBLASctl            0.23-42    2023-02-11 [1] CRAN (R 4.5.0)
 rjson                  0.2.23     2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7      2026-01-09 [1] CRAN (R 4.5.2)
 rmeta                  3.0        2018-03-20 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1      2025-08-26 [1] CRAN (R 4.5.0)
 RSQLite                2.4.5      2025-11-30 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1     2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1      2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0      2025-04-24 [1] CRAN (R 4.5.0)
 scattermore            1.2        2023-06-12 [1] CRAN (R 4.5.0)
 scatterplot3d          0.3-44     2023-05-05 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 sparseMatrixStats      1.22.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 SpatialExperiment    * 1.20.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SQUAREM                2021.1     2021-01-13 [1] CRAN (R 4.5.0)
 statmod                1.5.1      2025-10-09 [1] CRAN (R 4.5.0)
 stringi                1.8.7      2025-03-27 [1] CRAN (R 4.5.0)
 stringr                1.6.0      2025-11-04 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1      2026-01-11 [1] CRAN (R 4.5.2)
 tidyr                  1.3.2      2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [1] CRAN (R 4.5.0)
 truncnorm              1.0-9      2023-03-20 [1] CRAN (R 4.5.0)
 variancePartition    * 1.40.1     2025-12-11 [1] Bioconductor 3.22 (R 4.5.2)
 vctrs                  0.7.1      2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7      2023-12-18 [1] CRAN (R 4.5.0)
 withr                  3.0.2      2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56       2026-01-18 [1] CRAN (R 4.5.2)
 XML                    3.99-0.20  2025-11-08 [1] CRAN (R 4.5.0)
 xml2                   1.5.2      2026-01-17 [1] CRAN (R 4.5.2)
 xtable                 1.8-4      2019-04-21 [1] CRAN (R 4.5.0)
 XVector                0.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 zenith                 1.12.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 zigg                   0.0.2      2025-02-07 [1] CRAN (R 4.5.0)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────