---
title: "04-OneInteractionPerRun_MapMyCellswithWDXferNEoverride"
author: "Bernie Mulvey"
date: "06-17-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(edgeR)
library(limma)


## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```


```{r}
dge <- readRDS("processed-data_newseg/12_MapMyCells/03-MapMyCellsDGEpbulk_wdXferNEoverride.RDS")
```

define functions to run each DE analysis
####################################################################
```{r}
sexfun <- function(dgel,n){
    dgelsex <- dge
    dgelsex$samples$label_sex <- paste0(dgelsex$samples$label, "_", dgelsex$samples$Sex)
    
    # design mat
    des <- model.matrix(~0 + label_sex + APOEsimple + Age + YRI, data = dgelsex$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "label_sex", replacement = "")
    
    
    # filter using design
    dgelsex <- edgeR::normLibSizes(dgelsex)
    keep <- filterByExpr.DGEList(dgelsex, design = des)
    dgelsex <- dgelsex[keep, , keep.lib.sizes = FALSE]
    dgelsex <- edgeR::normLibSizes(dgelsex)
    
    # define contrasts to pass to makeContrasts (we have to do it this way due to environment weirdness with parallelization)    contrasts_list <- list()
    contrasts_list <- list()
    for (current_label in unique(dgelsex$samples$label)) {
        contrast_name <- paste0(current_label, "_Sex")
        contrast_formula <- paste0(current_label, "_M - ", current_label, "_F")
        
        new_contrast <- list(contrast_formula)
        names(new_contrast) <- contrast_name
        contrasts_list <- c(contrasts_list, new_contrast)
    }
    trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))

    ## fit and fit contrasts
    v.swt <- voomLmFit(dgelsex, design = des, block = as.factor(dgelsex$samples$brnum), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    ## get top tables, save an RDS with the DE results and other parts
    sexlabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
    })
    names(sexlabel_detabs) <- colnames(trastmat)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelsex, trasts = trastmat, detabs = sexlabel_detabs)
    
    saveRDS(keeplist, paste0("processed-data_newseg/12_MapMyCells/04bTMP_keeplist_sexlabel_noMixClusts_", n, ".RDS"))
    gc(full = TRUE)
}
```

#####################################################################
##### MODEL 2: E2/E4 carrier status "interaction" edgeR style #######
#####################################################################
```{r}
e2e4fun <- function(dgel,n){dgelE2E4 <- dgel
    dgelE2E4$samples$APOlargeGrp <- ifelse(dgelE2E4$samples$APOE %in% c("E2hom", "E2het"), yes = "E2", no = "E4")
    stopifnot(length(unique(dgelE2E4$samples$APOlargeGrp)) == 2)
    dgelE2E4$samples$label_E2E4 <- paste0(dgelE2E4$samples$label, "_", dgelE2E4$samples$APOlargeGrp)
    
    des <- model.matrix(~0 + label_E2E4 + Sex + Age + YRI, data = dgelE2E4$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "label_E2E4", replacement = "")

    # Filter using the design specified
    dgelE2E4 <- edgeR::normLibSizes(dgelE2E4)
    keep <- filterByExpr.DGEList(dgelE2E4, design = des)
    dgelE2E4 <- dgelE2E4[keep, , keep.lib.sizes = FALSE]
    dgelE2E4 <- edgeR::normLibSizes(dgelE2E4)
    
    # define contrasts to pass to makeContrasts (we have to do it this way due to environment weirdness with parallelization)
    contrasts_list <- list()
    for (current_label in unique(dgelE2E4$samples$label)) {
        contrast_name <- paste0(current_label, "_E4vE2")
        contrast_formula <- paste0(current_label, "_E4 - ", current_label, "_E2")
        
        new_contrast <- list(contrast_formula)
        names(new_contrast) <- contrast_name
        contrasts_list <- c(contrasts_list, new_contrast)
    }
    trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))

    ## fit and fit contrasts
    v.swt <- voomLmFit(dgelE2E4, design = des, block = as.factor(dgelE2E4$samples$brnum), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    ## get top tables, save an RDS with the DE results and other parts
    E2E4label_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
    })
    names(E2E4label_detabs) <- colnames(trastmat)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelE2E4, trasts = trastmat, detabs = E2E4label_detabs)
                     
    saveRDS(keeplist, paste0("processed-data_newseg/12_MapMyCells/04bTMP_keeplist_E2E4label_noMixClusts_", n, ".RDS"))

    gc(full = TRUE)
}
```

##########################################################################
#### MODEL 3: APOE alleles (E2hom/het, 4hom/het) "intxns" edgeR style ####
##########################################################################
```{r}
apogenofun <- function(dgel,n){
    dgelApoAlleles <- dgel
    dgelApoAlleles$samples$label_apo <- paste0(dgelApoAlleles$samples$label, "_", dgelApoAlleles$samples$APOE)
    
    des <- model.matrix(~0 + label_apo + Sex + Age + YRI, data = dgelApoAlleles$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "label_apo", replacement = "")
    
    # filter using the design specified
    dgelApoAlleles <- edgeR::normLibSizes(dgelApoAlleles)
    keep <- filterByExpr.DGEList(dgelApoAlleles, design = des)
    dgelApoAlleles <- dgelApoAlleles[keep, , keep.lib.sizes = FALSE]
    dgelApoAlleles <- edgeR::normLibSizes(dgelApoAlleles)

    contrasts_list <- list()
    for (current_label in unique(dgelApoAlleles$samples$label)) {
        # Create a named vector of contrast formulas for the current label;
        # account for the fact that there are no E2hom pseudobulks for some intra-LC types
        if(!(paste0(current_label,"_E2hom") %in% levels(des))){
          formulas <- c(
              paste0(current_label, "_E2het - (0.5*", current_label, "_E4het + 0.5*", current_label, "_E4hom)"),
              paste0(current_label, "_E4hom - ", current_label, "_E2het"),
              paste0(current_label, "_E4het - ", current_label, "_E2het"),
              paste0(current_label, "_E4hom - ", current_label, "_E4het"),
              paste0(current_label, "_E4hom - ((", current_label, "_E4het + ", current_label, "_E2het)/2)"))
          names(formulas) <- paste0(current_label, c("_E2het_vs_anyE4", "_E4hom_vs_anyE2", "_E4het_vs_anyE2", "_E4hom_vs_E4het", "_E4hom_vs_others"))
        } else {
          formulas <- c(
              paste0(current_label, "_E2hom - (0.5*", current_label, "_E4het + 0.5*", current_label, "_E4hom)"),
              paste0(current_label, "_E2het - (0.5*", current_label, "_E4het + 0.5*", current_label, "_E4hom)"),
              paste0(current_label, "_E4hom - (0.5*", current_label, "_E2het + 0.5*", current_label, "_E2hom)"),
              paste0(current_label, "_E4het - (0.5*", current_label, "_E2het + 0.5*", current_label, "_E2hom)"),
              paste0(current_label, "_E4hom - ", current_label, "_E4het"),
              paste0(current_label, "_E4hom - ((", current_label, "_E4het + ", current_label, "_E2het + ", current_label, "_E2hom)/3)")
          )
          names(formulas) <- paste0(current_label, c("_E2hom_vs_anyE4", "_E2het_vs_anyE4", "_E4hom_vs_anyE2", "_E4het_vs_anyE2", "_E4hom_vs_E4het", "_E4hom_vs_others"))
        }
        contrasts_list <- c(contrasts_list, as.list(formulas))
    }
    trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))

    # fit and fit contrasts
    v.swt <- voomLmFit(dgelApoAlleles, design = des, block = as.factor(dgelApoAlleles$samples$brnum), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    # get toptabs and other components, save
    ApoAllelesLabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
    })
    names(ApoAllelesLabel_detabs) <- colnames(trastmat)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelApoAlleles, trasts = trastmat, detabs = ApoAllelesLabel_detabs)
                     
    saveRDS(keeplist, paste0("processed-data_newseg/12_MapMyCells/04bTMP_keeplist_ApoAllelesLabel_noMixClusts_", n, ".RDS"))
    gc(full = TRUE)
}
```

#############################################################################
#### MODEL 4: label-by-YRI (continuous) "intxns" edgeR style ####
#############################################################################
```{r}
ancesfun <- function(dgel,n){
    dgelances <- dgel
    
    # make design matrices based on YRI being continuous or categorical
    if (is.numeric(dgelances$samples$YRI)) {
        # Continuous YRI setup
        des <- model.matrix(~label:YRI + Sex + APOEsimple + Age, data = dgelances$samples)
        colnames(des) <- gsub(x = colnames(des), pattern = "^label(.*):YRI$", replacement = "\\1_by_YRI")
        trastmat <- paste0(unique(dgelances$samples$label), "_by_YRI") # These are coefficients, not contrasts
        stopifnot(all(trastmat %in% colnames(des)))
    } else {
        # Categorical YRI setup
        dgelances$samples$labAnces <- paste0(dgelances$samples$label, "_", dgelances$samples$YRI)
        des <- model.matrix(~0 + labAnces + Sex + APOEsimple + Age, data = dgelances$samples)
        
        contrasts_list <- list()
        for (current_label in unique(dgelances$samples$label)) {
            contrast_name <- paste0(current_label, "_YRI")
            contrast_formula <- paste0("labAnces", current_label, "_YRI - labAnces", current_label, "_CEU")
            
            new_contrast <- list(contrast_formula)
            names(new_contrast) <- contrast_name
            contrasts_list <- c(contrasts_list, new_contrast)
        }
        trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))
    }
    
    # Filtering using appropriate design from above
    dgelances <- edgeR::normLibSizes(dgelances)
    keep <- filterByExpr.DGEList(dgelances, design = des)
    dgelances <- dgelances[keep, , keep.lib.sizes = FALSE]
    dgelances <- edgeR::normLibSizes(dgelances)

    # Fit
    v.swt <- voomLmFit(dgelances, design = des, block = as.factor(dgelances$samples$brnum), adaptive.span = TRUE, sample.weights = TRUE)
    
    # Fit contrasts or coefficients depending on YRI type
    if (is.numeric(dgelances$samples$YRI)) {
        v.swt.fit <- contrasts.fit(v.swt, coefficients = trastmat)
    } else {
        v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    }
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)
    
    # Get results and other components, save
    yrilabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
    })
    names(yrilabel_detabs) <- colnames(v.swt.fit)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelances, trasts = trastmat, detabs = yrilabel_detabs)

    saveRDS(keeplist, paste0("processed-data_newseg/12_MapMyCells/04bTMP_keeplist_AncesByLabel_noMixClusts_", n, ".RDS"))
    gc(full = TRUE)
}
```

#############################################################################
####### MODEL 5: label*YRI*E4orE2 (cat-cat-cat) "intxns" edgeR style ########
#############################################################################
```{r}
ancesapofun <- function(dgel,n){
    if (is.numeric(dgel$samples$YRI)) {
        message("Subsetting to 75%+ ancestry groups", n)
      newsamp <- as.data.table(dgel$samples,keep.rownames=T)
      newct <- dgel$counts
      
      newsamp <- newsamp[YRI>0.75|YRI<0.25]
      newsamp[,YRIcat:=ifelse(YRI>0.75,"YRI","CEU")]
      newct <- newct[,colnames(newct) %in% newsamp$rn]
      newsamp <- as.data.frame(newsamp,row.names=newsamp$rn)[colnames(newct),]
      newg <- as.data.frame(dgel$genes,row.names=rownames(dgel$genes))[rownames(newct),]
      newdgel <- DGEList(counts=newct,samples=newsamp,genes=newg)
    }
  
    # for the categorical data, set it up and go
    newdgel$samples$APOlargeGrp <- ifelse(newdgel$samples$APOE %in% c("E2hom", "E2het"), yes = "E2", no = "E4")
    newdgel$samples$lab_APOE_PredomAnces <- paste0(newdgel$samples$label, "_", newdgel$samples$APOlargeGrp, "_", newdgel$samples$YRIcat)
    
    # design matrix
    des <- model.matrix(~0 + lab_APOE_PredomAnces + Sex + Age, data = newdgel$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "^lab_APOE_PredomAnces(.*)$", replacement = "\\1")
    
    # filter using the design
    newdgel <- edgeR::normLibSizes(newdgel)
    keep <- filterByExpr.DGEList(newdgel, design = des)
    newdgel <- newdgel[keep, , keep.lib.sizes = FALSE]
    newdgel <- edgeR::normLibSizes(newdgel)

    # define contrasts
    contrasts_list <- list()
    for (current_label in unique(newdgel$samples$label)) {
        formulas <- c(
            paste0(current_label, "_E4_YRI - ", current_label, "_E4_CEU"),
            paste0(current_label, "_E2_YRI - ", current_label, "_E2_CEU"),
            paste0(current_label, "_E4_CEU - ", current_label, "_E2_CEU"),
            paste0(current_label, "_E4_YRI - ", current_label, "_E2_YRI")
        )
        names(formulas) <- paste0(current_label, c("_E4YRI_vs_E4CEU", "_E2YRI_vs_E2CEU", "_E4CEU_vs_E2CEU", "_E4YRI_vs_E2YRI"))
        contrasts_list <- c(contrasts_list, as.list(formulas))
    }
    trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))

    # fit, fit contrasts
    v.swt <- voomLmFit(newdgel, design = des, block = as.factor(newdgel$samples$brnum), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    # get toptabs, other results, and save
    yriapolabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        o <- as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
        #setnames(o,"genes","gene_name") ## force uniformimty with other results
        return(o)
    })
    names(yriapolabel_detabs) <- colnames(v.swt.fit)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = newdgel, trasts = trastmat, detabs = yriapolabel_detabs)
                     
    saveRDS(keeplist, paste0("processed-data_newseg/12_MapMyCells/04bTMP_keeplist_YRIAPOlabel_noMixClusts_", n, ".RDS"))
    gc(full = TRUE)
}
```


## now run all of these simultaneously
```{r}
funs <- list(sexfun,e2e4fun,apogenofun,ancesfun,ancesapofun)

lapply(X=funs,FUN=function(f){
        f(dge,"MapMyCells_wdNEoverride_byLabel")
})
```


## clean up, load the lists and concatenate into one giant result obj per domain definition set
```{r}
rm(list=ls())
gc(full=T)
```


## load the tmp RDSes
```{r}
tmp04 <- list.files("processed-data_newseg/12_MapMyCells",full.names = TRUE)
tmp04 <- grep(tmp04,pattern="04b",value=T)



## combine into one metalist
outs <- lapply(tmp04,readRDS)
names(outs) <- gsub(tmp04,pattern="processed-data_newseg/12_MapMyCells/04bTMP_keeplist_(.*)_noMix.*$",replacement="\\1")

saveRDS(outs,"processed-data_newseg/12_MapMyCells/04-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_MapMyCells-wdNEoverrideByLabel.RDS")
```
## can now delete 04bTMP RDSes.

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices
[5] utils     methods   base     

other attached packages:
[1] BiocParallel_1.44.0 parallelly_1.47.0  
[3] edgeR_4.8.2         limma_3.66.0       
[5] data.table_1.18.2.1

loaded via a namespace (and not attached):
 [1] codetools_0.2-20 xfun_0.56       
 [3] lattice_0.22-7   knitr_1.51      
 [5] cli_3.6.5        grid_4.5.2      
 [7] statmod_1.5.1    compiler_4.5.2  
 [9] rprojroot_2.1.1  here_1.0.2      
[11] tools_4.5.2      evaluate_1.0.5  
[13] locfit_1.5-9.12  otel_0.2.0      
[15] rlang_1.1.7     
> sessioninfo::session_info()
─ Session info ──────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       Positron
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-17
 pandoc   NA
 quarto   1.8.27 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────
 package      * version  date (UTC) lib source
 BiocParallel * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli            3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools      0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 data.table   * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 edgeR        * 4.8.2    2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate       1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 here           1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 knitr          1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice        0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 limma        * 3.66.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 locfit         1.5-9.12 2025-03-05 [1] CRAN (R 4.5.0)
 otel           0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly   * 1.47.0   2026-04-17 [1] CRAN (R 4.5.2)
 rlang          1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rprojroot      2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 sessioninfo    1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 statmod        1.5.1    2025-10-09 [1] CRAN (R 4.5.0)
 xfun           0.56     2026-01-18 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────