---
title: "01a-PreBanksy spatCoord staggering and parameter sets"
author: "Bernie Mulvey"
date: "2026-06-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(scuttle)
library(Banksy)
library(BiocParallel)
library(RcppHNSW) ## faster UMAP in uwot 0.2 (used by banksy UMAP)


## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
```

## load QCed, genes only, NON-log normalized data
```{r}
lcx <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

## re normalization: Banksy uses  NON-log normcounts (per its documentation)
## the loaded SPE already contains the nonlog normaliezd count matrix and is subsetted to gene-targeting probes, so nothing to do on these fronts. lets get rid of the logcounts matrix just to be safe.

assays(lcx)["logcounts"] <- NULL
assayNames(lcx)
# "counts"  "normcounts"
# we just tell banksy in the actual call to use normcounts.
```


## to run Banksy in multisample mode, adjust coordinates so that
## every sample has a unique, non-overlapping set of coords (i.e. to prevent
## cells from a given area on two different slides from colliding)
## code adapted from: https://prabhakarlab.github.io/Banksy/articles/batch-correction.html
```{r}
# get the spatial coords matrix and their rownames to make sure everything matches back up in the end
rownames(spatialCoords(lcx)) <- colnames(lcx)
locs <- cbind(rownames(spatialCoords(lcx)),spatialCoords(lcx)) 
colnames(locs)[1] <- "rn"

# get sample IDs (each brnum entry (brnum+L if left hemi) is a sample id. we have to make into an integer factor here though, which makes the spacing math simpler. we can't do that with a data.table though for some reason
locs <- cbind(locs,factor(lcx$sample_id))
locs <- as.data.table(locs)
setnames(locs,4,"group") # the sample-as-integer-factor column.

locs[,x_centroid:=as.numeric(x_centroid)]
locs[,y_centroid:=as.numeric(y_centroid)]
locs[,group:=as.numeric(group)] # for mathing
setnames(locs,c(2:3),c("sdimx", "sdimy")) # names expected by banksy

# now spread the x coords out so that no two samples overlap in space, even from different slides         
locs[, sdimx := sdimx - min(sdimx), by = group]
global_max <- max(locs$sdimx) * 1.5 # 1.5x the widest sample's width
locs[, sdimx := sdimx + group * global_max] # put that much space between each sample
locs <- as.matrix(as.data.frame(locs[, .(sdimx,sdimy)],row.names=locs$rn))
spatialCoords(lcx) <- locs[colnames(lcx),] # ensuring that ordering of spatialCoords matches everything else

## finally, banksy can pull these from the colData instead of spatialcoords if we're having trouble, so put them in coldata just in case.
colData(lcx)$sdimx <- spatialCoords(lcx)[,1]
colData(lcx)$sdimy <- spatialCoords(lcx)[,2]

# save this SPE for inputting into banksy
saveRDS(lcx,"processed-data_newseg/11_Banksy_clustering/01-staggeredCoord_spe_genetargOnly_NONlogNorm.RDS")
```

## make table of Banksy run parameter sets (weighting of spatial information, number of nearest neighbor cells to be used, leiden clustering resolution on the resulting dimred). each JHPCE job will pull in a row of this table and pass the variables to R/banksy.
we can't parallelize louv/leiden clustering (though we can parallelize banksy steps), so we'll run three resolutions on a multicpu job. so the job params dont need separate resolutions.
```{r}
## set arguments, which we'll try a variety of
lam <- c(0,0.2,0.4,0.6,0.8,1) # 0=spatial info disregarded; 0.2 = suggested "cell typing" setting, 0.8= suggested "domain" setting, 1=fullly spatial weighting, 0.6-0.8 recommended for spatially-informed clusters. So test  (6 values)

gabor <- c(TRUE,FALSE) # some cryptic parameter that has something to do with spatial patterns. still don't have a good answer for like what it does exactly so use both ways. 2 values

# note that kgeom (number of neighboring cells to consider for spatial dimred) is suggested by banksy to range 15-30 for cell-resolution data.
kgeom <- c(15,30) # (2 values)
# 6*2*2 = 24 parameter combos total, each will be clustered at 3 resolutions in one job. get a table with all unique combinations of these 4 things
paramgrid <- as.data.table(expand.grid(lam,gabor,kgeom))
setnames(paramgrid,c("lam","gabor","kgeom"))
fwrite(paramgrid,"processed-data_newseg/11_Banksy_clustering/01-BanksyParamSets.txt",sep='\t',quote=F,row.names=F,col.names = T)
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods  
[8] base     

other attached packages:
 [1] HDF5Array_1.38.0            h5mread_1.2.1              
 [3] rhdf5_2.54.1                DelayedArray_0.36.1        
 [5] SparseArray_1.10.8          S4Arrays_1.10.1            
 [7] abind_1.4-8                 Matrix_1.7-4               
 [9] BiocParallel_1.44.0         parallelly_1.47.0          
[11] colorout_1.3-3              spatialLIBD_1.22.0         
[13] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
[15] SummarizedExperiment_1.40.0 Biobase_2.70.0             
[17] GenomicRanges_1.62.1        Seqinfo_1.0.0              
[19] IRanges_2.44.0              S4Vectors_0.48.1           
[21] BiocGenerics_0.56.0         generics_0.1.4             
[23] MatrixGenerics_1.22.0       matrixStats_1.5.0          
[25] gridExtra_2.3               ggtext_0.1.2               
[27] ggplot2_4.0.2               data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] later_1.4.5              BiocIO_1.20.0           
  [3] bitops_1.0-9             filelock_1.0.3          
  [5] tibble_3.3.1             XML_3.99-0.20           
  [7] lifecycle_1.0.5          httr2_1.2.2             
  [9] edgeR_4.8.2              doParallel_1.0.17       
 [11] rprojroot_2.1.1          lattice_0.22-7          
 [13] magrittr_2.0.4           limma_3.66.0            
 [15] plotly_4.12.0            sass_0.4.10             
 [17] rmarkdown_2.30           jquerylib_0.1.4         
 [19] yaml_2.3.12              httpuv_1.6.16           
 [21] otel_0.2.0               sessioninfo_1.2.3       
 [23] cowplot_1.2.0            DBI_1.2.3               
 [25] RColorBrewer_1.1-3       golem_0.5.1             
 [27] purrr_1.2.1              RCurl_1.98-1.17         
 [29] rappdirs_0.3.4           circlize_0.4.17         
 [31] ggrepel_0.9.6            irlba_2.3.7             
 [33] codetools_0.2-20         DT_0.34.0               
 [35] scuttle_1.20.0           xml2_1.5.2              
 [37] tidyselect_1.2.1         shape_1.4.6.1           
 [39] farver_2.1.2             ScaledMatrix_1.18.0     
 [41] viridis_0.6.5            shinyWidgets_0.9.0      
 [43] BiocFileCache_3.0.0      GenomicAlignments_1.46.0
 [45] jsonlite_2.0.0           GetoptLong_1.1.0        
 [47] BiocNeighbors_2.4.0      scater_1.38.1           
 [49] iterators_1.0.14         foreach_1.5.2           
 [51] tools_4.5.2              Rcpp_1.1.1              
 [53] glue_1.8.0               xfun_0.56               
 [55] here_1.0.2               dplyr_1.2.0             
 [57] withr_3.0.2              BiocManager_1.30.27     
 [59] fastmap_1.2.0            rhdf5filters_1.22.0     
 [61] digest_0.6.39            rsvd_1.0.5              
 [63] R6_2.6.1                 mime_0.13               
 [65] colorspace_2.1-2         dichromat_2.0-0.1       
 [67] RSQLite_2.4.5            cigarillo_1.0.0         
 [69] config_0.3.2             tidyr_1.3.2             
 [71] rtracklayer_1.70.1       httr_1.4.7              
 [73] htmlwidgets_1.6.4        pkgconfig_2.0.3         
 [75] gtable_0.3.6             rsconnect_1.7.0         
 [77] blob_1.3.0               ComplexHeatmap_2.26.1   
 [79] S7_0.2.1                 XVector_0.50.0          
 [81] htmltools_0.5.9          clue_0.3-66             
 [83] scales_1.4.0             png_0.1-8               
 [85] attempt_0.3.1            knitr_1.51              
 [87] rstudioapi_0.18.0        rjson_0.2.23            
 [89] curl_7.0.0               cachem_1.1.0            
 [91] GlobalOptions_0.1.3      BiocVersion_3.22.0      
 [93] vipor_0.4.7              AnnotationDbi_1.72.0    
 [95] restfulr_0.0.16          pillar_1.11.1           
 [97] grid_4.5.2               vctrs_0.7.1             
 [99] promises_1.5.0           BiocSingular_1.26.0     
[101] dbplyr_2.5.1             beachmat_2.26.0         
[103] xtable_1.8-4             cluster_2.1.8.1         
[105] beeswarm_0.4.0           paletteer_1.7.0         
[107] evaluate_1.0.5           magick_2.9.0            
[109] cli_3.6.5                locfit_1.5-9.12         
[111] compiler_4.5.2           Rsamtools_2.26.0        
[113] rlang_1.1.7              crayon_1.5.3            
[115] rematch2_2.1.2           ggbeeswarm_0.7.3        
[117] viridisLite_0.4.3        Biostrings_2.78.0       
[119] lazyeval_0.2.2           ExperimentHub_3.0.0     
[121] benchmarkme_1.0.8        bit64_4.6.0-1           
[123] Rhdf5lib_1.32.0          KEGGREST_1.50.0         
[125] statmod_1.5.1            shiny_1.12.1            
[127] AnnotationHub_4.0.0      gridtext_0.1.5          
[129] memoise_2.0.1            bslib_0.10.0            
[131] benchmarkmeData_2.0.0    bit_4.6.0               
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-12
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────
 package              * version   date (UTC) lib source
 abind                * 1.4-8     2024-09-12 [1] CRAN (R 4.5.0)
 AnnotationDbi          1.72.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationHub          4.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 attempt                0.3.1     2020-05-03 [1] CRAN (R 4.5.0)
 beachmat               2.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.5.0)
 benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.5.0)
 benchmarkmeData        2.0.0     2026-01-19 [1] CRAN (R 4.5.2)
 Biobase              * 2.70.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocFileCache          3.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocIO                 1.20.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocManager            1.30.27   2025-11-14 [1] CRAN (R 4.5.2)
 BiocNeighbors          2.4.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocVersion            3.22.0    2025-10-07 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0     2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1   2025-01-16 [1] CRAN (R 4.5.0)
 bitops                 1.0-9     2024-10-03 [1] CRAN (R 4.5.0)
 blob                   1.3.0     2026-01-14 [1] CRAN (R 4.5.2)
 bslib                  0.10.0    2026-01-26 [1] CRAN (R 4.5.2)
 cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.5.0)
 cigarillo              1.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 circlize               0.4.17    2025-12-08 [1] CRAN (R 4.5.2)
 cli                    3.6.5     2025-04-23 [1] CRAN (R 4.5.0)
 clue                   0.3-66    2024-11-13 [1] CRAN (R 4.5.0)
 cluster                2.1.8.1   2025-03-12 [1] CRAN (R 4.5.0)
 codetools              0.2-20    2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3     2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 colorspace             2.1-2     2025-09-22 [1] CRAN (R 4.5.0)
 ComplexHeatmap         2.26.1    2026-01-30 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 config                 0.3.2     2023-08-30 [1] CRAN (R 4.5.0)
 cowplot                1.2.0     2025-07-07 [1] CRAN (R 4.5.0)
 crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0     2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1  2026-01-27 [1] CRAN (R 4.5.2)
 DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.5.0)
 dbplyr                 2.5.1     2025-09-10 [1] CRAN (R 4.5.0)
 DelayedArray         * 0.36.1    2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1   2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39    2025-11-19 [1] CRAN (R 4.5.2)
 doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.5.0)
 dplyr                  1.2.0     2026-02-03 [1] CRAN (R 4.5.2)
 DT                     0.34.0    2025-09-02 [1] CRAN (R 4.5.0)
 edgeR                  4.8.2     2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate               1.0.5     2025-08-27 [1] CRAN (R 4.5.0)
 ExperimentHub          3.0.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 farver                 2.1.2     2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.5.0)
 filelock               1.0.3     2023-12-11 [1] CRAN (R 4.5.0)
 foreach                1.5.2     2022-02-02 [1] CRAN (R 4.5.0)
 generics             * 0.1.4     2025-05-09 [1] CRAN (R 4.5.0)
 GenomicAlignments      1.46.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 GenomicRanges        * 1.62.1    2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 GetoptLong             1.1.0     2025-11-28 [1] CRAN (R 4.5.2)
 ggbeeswarm             0.7.3     2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2     2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6     2024-09-07 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2     2022-09-16 [1] CRAN (R 4.5.0)
 GlobalOptions          0.1.3     2025-11-28 [1] CRAN (R 4.5.2)
 glue                   1.8.0     2024-09-30 [1] CRAN (R 4.5.0)
 golem                  0.5.1     2024-08-27 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5     2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6     2024-10-25 [1] CRAN (R 4.5.0)
 h5mread              * 1.2.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 HDF5Array            * 1.38.0    2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 here                   1.0.2     2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9     2025-12-04 [1] CRAN (R 4.5.2)
 htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.5.0)
 httpuv                 1.6.16    2025-04-16 [1] CRAN (R 4.5.0)
 httr                   1.4.7     2023-08-15 [1] CRAN (R 4.5.0)
 httr2                  1.2.2     2025-12-08 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7     2026-01-30 [1] CRAN (R 4.5.2)
 iterators              1.0.14    2022-02-05 [1] CRAN (R 4.5.0)
 jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.5.0)
 jsonlite               2.0.0     2025-03-27 [1] CRAN (R 4.5.0)
 KEGGREST               1.50.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51      2025-12-20 [1] CRAN (R 4.5.2)
 later                  1.4.5     2026-01-08 [1] CRAN (R 4.5.2)
 lattice                0.22-7    2025-04-02 [1] CRAN (R 4.5.2)
 lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.5.0)
 lifecycle              1.0.5     2026-01-08 [1] CRAN (R 4.5.2)
 limma                  3.66.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 locfit                 1.5-9.12  2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0     2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4     2025-09-12 [1] CRAN (R 4.5.0)
 Matrix               * 1.7-4     2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0     2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1     2021-11-26 [1] CRAN (R 4.5.0)
 mime                   0.13      2025-03-17 [1] CRAN (R 4.5.0)
 otel                   0.2.0     2025-08-29 [1] CRAN (R 4.5.0)
 paletteer              1.7.0     2026-01-08 [1] CRAN (R 4.5.2)
 parallelly           * 1.47.0    2026-04-17 [1] CRAN (R 4.5.2)
 pillar                 1.11.1    2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.5.0)
 plotly                 4.12.0    2026-01-24 [1] CRAN (R 4.5.2)
 png                    0.1-8     2022-11-29 [1] CRAN (R 4.5.0)
 promises               1.5.0     2025-11-01 [1] CRAN (R 4.5.0)
 purrr                  1.2.1     2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1     2025-02-15 [1] CRAN (R 4.5.0)
 rappdirs               0.3.4     2026-01-17 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1     2026-01-10 [1] CRAN (R 4.5.2)
 RCurl                  1.98-1.17 2025-03-22 [1] CRAN (R 4.5.0)
 rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.5.0)
 restfulr               0.0.16    2025-06-27 [1] CRAN (R 4.5.0)
 rhdf5                * 2.54.1    2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 rhdf5filters           1.22.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Rhdf5lib               1.32.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 rjson                  0.2.23    2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7     2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30      2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1     2025-08-26 [1] CRAN (R 4.5.0)
 Rsamtools              2.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 rsconnect              1.7.0     2025-12-06 [1] CRAN (R 4.5.2)
 RSQLite                2.4.5     2025-11-30 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0    2026-01-16 [1] CRAN (R 4.5.2)
 rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.5.0)
 rtracklayer            1.70.1    2025-12-20 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 S4Arrays             * 1.10.1    2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1    2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1     2025-11-14 [1] CRAN (R 4.5.2)
 sass                   0.4.10    2025-04-11 [1] CRAN (R 4.5.0)
 ScaledMatrix           1.18.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0     2025-04-24 [1] CRAN (R 4.5.0)
 scater                 1.38.1    2026-03-20 [1] Bioconductor 3.22 (R 4.5.2)
 scuttle                1.20.0    2025-10-30 [1] Bioconductor 3.22 (R 4.5.2)
 Seqinfo              * 1.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3     2025-02-05 [1] CRAN (R 4.5.0)
 shape                  1.4.6.1   2024-02-23 [1] CRAN (R 4.5.0)
 shiny                  1.12.1    2025-12-09 [1] CRAN (R 4.5.2)
 shinyWidgets           0.9.0     2025-02-21 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray          * 1.10.8    2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 spatialLIBD          * 1.22.0    2025-11-04 [1] Bioconductor 3.22 (R 4.5.2)
 statmod                1.5.1     2025-10-09 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1     2026-01-11 [1] CRAN (R 4.5.2)
 tidyr                  1.3.2     2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1     2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.5.0)
 viridis                0.6.5     2024-01-29 [1] CRAN (R 4.5.0)
 viridisLite            0.4.3     2026-02-04 [1] CRAN (R 4.5.2)
 withr                  3.0.2     2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56      2026-01-18 [1] CRAN (R 4.5.2)
 XML                    3.99-0.20 2025-11-08 [1] CRAN (R 4.5.0)
 xml2                   1.5.2     2026-01-17 [1] CRAN (R 4.5.2)
 xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.5.0)
 XVector                0.50.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12    2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────