---
title: "02-Register_NonHarmony_Banksys_to_Manual_xLC_and_vLC"
author: "Bernie Mulvey"
date: "12/05/25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load xenium data, banksy clusters, manual xenium LC assignments; add the latter 2 to coldata
```{r}
lcx <- readRDS("processed-data/00_prelim_procdat_BJM/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

manlc <- fread("processed-data/00_prelim_procdat_BJM/03_ManualLC/01-XprBasedLClabels.txt")

banks <- readRDS("processed-data/00_prelim_procdat_BJM/04_Banksy_clusterings/01-banksy_clustruns_outs/01-PrelimBanksys_noHarmony_concat.RDS")

manlc <- merge.data.table(manlc,banks,by="rn")
stopifnot(nrow(manlc)==ncol(lcx))
manlc <- DataFrame(manlc,row.names=manlc$rn)[colnames(lcx),]
manlc$rn  <- NULL
colData(lcx) <- cbind(colData(lcx),manlc)

rm(manlc,banks)
gc(full=T)
```

## get rid of unnecess parts of SPE for RAM purposes
```{r}
## remove logcounts/non-log norm counts/dimreds for memory savings
assays(lcx)[["logcounts"]] <- NULL
assays(lcx)[["normcounts"]] <- NULL
reducedDims(lcx)[1:4] <- NULL # keep the slot itself there but wipe out its contents
gc(full=T)
## remove unnecess columns from colData
tmpcd <- as.data.table(colData(lcx),keep.rownames=T)
keepcols <- c("rn","brnum","Age","Sex","APOE","Ancestry","YRI","CEU","LCloose","LCstrict",grep("clust_M",names(tmpcd),value=T))
tmpcd <- tmpcd[,..keepcols]
setnames(tmpcd,"brnum","sample_id")
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(lcx),]
tmpcd$rn <- NULL
colData(lcx) <- tmpcd
```

## get the LC visium data registration wrapper stats (already run including duplicateCorrelation due to multiple sections per donor in that dataset)
```{r}
visreg <- readRDS("../LFF_spatial_LC/processed-data/13_markers_and_registration_final/01-HDGSVG2575louv1_regWrapOut.RDS")
```

iterate thru spatial registration runs. remember, these just get us the marker stats, and then we have to use some manually written code to actually test the marker correlations (because the built-in command for it has never worked in my hands)
```{r}
mancols <- c("LCloose","LCstrict")

manlc.regs <- mclapply(mancols,mc.cores=2,function(m){
    manout <- registration_wrapper(sce = lcx,var_registration = m,var_sample_id = "sample_id",covars = c("Age","Sex","APOE","YRI"),gene_ensembl = "ID",gene_name = "Symbol",min_ncells = 20)
})
names(manlc.regs) <- mancols

## append the visium registration stats to "manual" LCs list (we want to compare each banksy run to the two manual xenium labelings and the final visium labeling)
manlc.regs[[length(manlc.regs)+1]] <- visreg
names(manlc.regs)[length(manlc.regs)] <- "LFF_Visium_9clust"

###
bcols <- grep("clust_M",colnames(colData(lcx)),value=T)
bankregs <- mclapply(bcols,mc.cores=6,function(b){
    manout <- registration_wrapper(sce = lcx,var_registration = b,var_sample_id = "sample_id",covars = c("Age","Sex","APOE","YRI"),gene_ensembl = "ID",gene_name = "Symbol",min_ncells = 20)
})
names(bankregs) <- bcols

gc(full=T)

## save these so we can load them later without re-running
saveRDS(manlc.regs,"processed-data/00_prelim_procdat_BJM/04_Banksy_clusterings/02-regwrapper_outs_manual_xLCs.RDS")
saveRDS(bankregs,"processed-data/00_prelim_procdat_BJM/04_Banksy_clusterings/02-regwrapper_outs_nonHarmoBanksys.RDS")
```

LEFT OFF HERE -- GET CODE TO TEST THE T STAT CORRELS FOR EACH PAIR OF REGISTRATION STATS

```{r}
### I wrote this code to use inputs already subsetted to just t-statistics from the enrichment test tables. we can put these together in the following way. the xenium's simpler
bankregs <- lapply(bankregs,function(x){
  z <- as.data.table(x$enrichment)
  keepnames <- c("ensembl",grep(names(z),pattern="t_stat",value=T))
  z <- z[,..keepnames]
  # take the name down to just the generic cluster id
  setnames(z,
           c(2:ncol(z)),
           gsub(names(z)[c(2:ncol(z))],pattern="t_stat_",replacement=""))
  return(z)
})

manlc.regs <- lapply(manlc.regs,function(x){
  z <- as.data.table(x$enrichment)
  keepnames <- c("ensembl",grep(names(z),pattern="t_stat",value=T))
  z <- z[,..keepnames]
  # take the name down to just the generic cluster id
  setnames(z,
           c(2:ncol(z)),
           gsub(names(z)[c(2:ncol(z))],pattern="t_stat_",replacement=""))
  return(z)
})

regres <- lapply(manlc.regs,function(m){
  subres <- bpmapply(b=bankregs,n=names(bankregs),BPPARAM=MulticoreParam(4),SIMPLIFY=FALSE,FUN=function(b,n){
    
    j<-seq(10,300,by=10)
    i<-1
    outlist <- list()
    cortab <- merge.data.table(m,b,by="ensembl")
    cortab$ensembl <- NULL
    for (i in c(1:(length(j)+1))){
        if(i<=length(j)){
            ngene <- j[i]
        }
        else{
            # no value for top_n --> all genes used
            ngene <- nrow(cortab)
        }
        # initialize cormat
        nclxen <- ncol(b)
        nclref <- ncol(m)
        cormat <- as.data.frame(matrix(nrow=nclxen-1,ncol=nclref-1))
        # skim the first column, the gene id
        k <- 2
        for (k in c(2:nclxen)){
            rownames(cormat)[k-1] <- colnames(b)[k]
            l <- 2
            for (l in c(2:nclref)){
                colnames(cormat)[l-1] <- colnames(m)[l]
                
                curxen <- colnames(b)[k]
                curref <- colnames(m)[l]
                setorderv(cortab,curxen,-1) # order the rows by current xenium cluster's top marker t stats in descending order, which we get the top N of below
                
                cormat[(k-1),(l-1)] <- cor(cortab[c(1:ngene),..curref],cortab[c(1:ngene),..curxen],use="pairwise.complete.obs",method = "spearman")
            }
        }
        plttab <- as.data.table(cormat,keep.rownames=T)
        setnames(plttab,"rn","Xen_clust")
        outlist[[i]] <- plttab
        names(outlist)[i] <- paste0(n,"_top_",ngene)
    }
    rm(cortab)
    return(outlist)
  })
})

regres2 <- lapply(regres,function(x){
  z <- lapply(x,function(y){
    a <- mapply(b=y,n=names(y),SIMPLIFY=FALSE,FUN=function(b,n){
      b[,Ntopxenmks:=as.numeric(gsub(n,pattern="^clust.*_top_(.*)$",replacement="\\1"))]
      return(b)
    })
    z2 <- rbindlist(a)
    return(z2)
  })
  z3 <- rbindlist(z, idcol="banksy_run")
  return(z3)
})

saveRDS(regres2,"processed-data/00_prelim_procdat_BJM/04_Banksy_clusterings/02-BankyToManual_BanksyToVisium_registration_cors.RDS")
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
─ Session info ─────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-12-07
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ─────────────────────────────────────────────────────────────────────────
 ! package              * version   date (UTC) lib source
   abind                  1.4-8     2024-09-12 [1] CRAN (R 4.4.1)
   AnnotationDbi          1.68.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   AnnotationHub          3.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   attempt                0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat               2.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData        1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase              * 2.66.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocFileCache          2.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocGenerics         * 0.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocIO                 1.16.0    2024-12-12 [1] Github (Bioconductor/BiocIO@8e80624)
   BiocManager            1.30.27   2025-11-14 [1] CRAN (R 4.4.3)
   BiocNeighbors          2.0.1     2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
   BiocParallel         * 1.40.2    2025-04-08 [1] Bioconductor
   BiocSingular           1.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocVersion            3.20.0    2024-05-04 [1] Bioconductor 3.20 (R 4.4.0)
   Biostrings             2.74.1    2024-12-16 [1] Bioconductor 3.20 (R 4.4.2)
   bit                    4.6.0     2025-03-06 [1] CRAN (R 4.4.1)
   bit64                  4.6.0-1   2025-01-16 [1] CRAN (R 4.4.1)
   bitops                 1.0-9     2024-10-03 [1] CRAN (R 4.4.1)
   blob                   1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   bslib                  0.9.0     2025-01-30 [1] CRAN (R 4.4.1)
   cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
 P cli                    3.6.5     2025-04-23 [2] CRAN (R 4.4.1)
   codetools              0.2-20    2024-03-31 [1] CRAN (R 4.4.3)
   colorout             * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   config                 0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                1.2.0     2025-07-07 [1] CRAN (R 4.4.1)
   crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                   7.0.0     2025-08-19 [1] CRAN (R 4.4.1)
   data.table           * 1.17.4    2025-05-26 [1] CRAN (R 4.4.3)
   DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                 2.5.1     2025-09-10 [1] CRAN (R 4.4.1)
   DelayedArray           0.32.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   dichromat              2.0-0.1   2022-05-02 [1] CRAN (R 4.4.0)
   digest                 0.6.39    2025-11-19 [1] CRAN (R 4.4.3)
   doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
   dotCall64              1.2       2024-10-04 [1] CRAN (R 4.4.1)
   dplyr                  1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   DT                     0.34.0    2025-09-02 [1] CRAN (R 4.4.1)
   edgeR                  4.4.2     2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
   evaluate               1.0.5     2025-08-27 [1] CRAN (R 4.4.1)
   ExperimentHub          2.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   farver                 2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fields                 16.3.1    2025-03-08 [1] CRAN (R 4.4.1)
   filelock               1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   foreach                1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics               0.1.4     2025-05-09 [1] CRAN (R 4.4.1)
   GenomeInfoDb         * 1.42.3    2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
   GenomeInfoDbData       1.2.13    2024-12-12 [1] Bioconductor
   GenomicAlignments      1.42.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   GenomicRanges        * 1.58.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   ggbeeswarm             0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2              * 4.0.1     2025-11-14 [1] CRAN (R 4.4.3)
   ggrepel                0.9.6     2024-09-07 [1] CRAN (R 4.4.1)
   ggtext               * 0.1.2     2022-09-16 [1] CRAN (R 4.4.0)
   glue                   1.8.0     2024-09-30 [1] CRAN (R 4.4.1)
   golem                  0.5.1     2024-08-27 [1] CRAN (R 4.4.1)
   gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gridtext               0.1.5     2022-09-16 [1] CRAN (R 4.4.0)
   gtable                 0.3.6     2024-10-25 [1] CRAN (R 4.4.1)
   here                   1.0.2     2025-09-15 [1] CRAN (R 4.4.1)
   htmltools              0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                 1.6.16    2025-04-16 [1] CRAN (R 4.4.1)
   httr                   1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges              * 2.40.1    2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
   irlba                  2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators              1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite               2.0.0     2025-03-27 [1] CRAN (R 4.4.1)
   KEGGREST               1.46.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   knitr                  1.50      2025-03-16 [1] CRAN (R 4.4.2)
   later                  1.4.4     2025-08-27 [1] CRAN (R 4.4.1)
   lattice                0.22-7    2025-04-02 [1] CRAN (R 4.4.1)
   lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle              1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                  3.62.2    2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
   locfit                 1.5-9.12  2025-03-05 [1] CRAN (R 4.4.1)
   magick                 2.9.0     2025-09-08 [1] CRAN (R 4.4.1)
   magrittr               2.0.4     2025-09-12 [1] CRAN (R 4.4.1)
   maps                   3.4.3     2025-05-26 [1] CRAN (R 4.4.1)
   Matrix                 1.7-4     2025-08-28 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.18.1    2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
   matrixStats          * 1.5.0     2025-01-07 [1] CRAN (R 4.4.1)
   memoise                2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   mime                   0.13      2025-03-17 [1] CRAN (R 4.4.1)
   otel                   0.2.0     2025-08-29 [1] CRAN (R 4.4.1)
   paletteer              1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   parallelly           * 1.45.1    2025-07-24 [1] CRAN (R 4.4.1)
   pillar                 1.11.1    2025-09-17 [1] CRAN (R 4.4.1)
   pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                 4.11.0    2025-06-19 [1] CRAN (R 4.4.1)
   png                    0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   promises               1.5.0     2025-11-01 [1] CRAN (R 4.4.1)
   purrr                  1.2.0     2025-11-04 [1] CRAN (R 4.4.1)
   R6                     2.6.1     2025-02-15 [1] CRAN (R 4.4.1)
   rappdirs               0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                   1.1.0     2025-07-02 [1] CRAN (R 4.4.1)
   RCurl                  1.98-1.17 2025-03-22 [1] CRAN (R 4.4.1)
   rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr               0.0.16    2025-06-27 [1] CRAN (R 4.4.1)
   rjson                  0.2.23    2024-09-16 [1] CRAN (R 4.4.1)
 P rlang                * 1.1.6     2025-04-11 [2] CRAN (R 4.4.1)
   rmarkdown              2.30      2025-09-28 [1] CRAN (R 4.4.1)
   rprojroot              2.1.1     2025-08-26 [1] CRAN (R 4.4.1)
   Rsamtools              2.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   RSQLite                2.4.4     2025-11-10 [1] CRAN (R 4.4.1)
   rstudioapi             0.17.1    2024-10-22 [1] CRAN (R 4.4.1)
   rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer            1.66.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   S4Arrays               1.6.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   S4Vectors            * 0.44.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   S7                     0.2.1     2025-11-14 [1] CRAN (R 4.4.3)
   sass                   0.4.10    2025-04-11 [1] CRAN (R 4.4.1)
   ScaledMatrix           1.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   scales                 1.4.0     2025-04-24 [1] CRAN (R 4.4.1)
   scater                 1.34.1    2025-03-03 [1] Bioconductor 3.20 (R 4.4.2)
   scuttle                1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   sessioninfo            1.2.3     2025-02-05 [1] CRAN (R 4.4.1)
   shiny                  1.11.1    2025-07-03 [1] CRAN (R 4.4.1)
   shinyWidgets           0.9.0     2025-02-21 [1] CRAN (R 4.4.1)
   SingleCellExperiment * 1.28.1    2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
   spam                   2.11-1    2025-01-20 [1] CRAN (R 4.4.1)
   SparseArray            1.6.2     2025-02-20 [1] Bioconductor 3.20 (R 4.4.2)
   SpatialExperiment    * 1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   spatialLIBD          * 1.18.0    2024-11-07 [1] Bioconductor 3.20 (R 4.4.3)
   statmod                1.5.1     2025-10-09 [1] CRAN (R 4.4.1)
   SummarizedExperiment * 1.36.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   tibble                 3.3.0     2025-06-08 [1] CRAN (R 4.4.1)
   tidyr                  1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.2.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   vctrs                  0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.2     2024-10-28 [1] CRAN (R 4.4.1)
   xfun                   0.54      2025-10-30 [1] CRAN (R 4.4.1)
   XML                    3.99-0.20 2025-11-08 [1] CRAN (R 4.4.1)
   xml2                   1.5.0     2025-11-17 [1] CRAN (R 4.4.3)
   xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector                0.46.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   yaml                   2.3.10    2024-07-26 [1] CRAN (R 4.4.0)
   zlibbioc               1.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 P ── Loaded and on-disk path mismatch.

────────────────────────────────────────────────────────────────────────────────────
