---
title: "03-Tabulate srtNMFtransfer VisDE Replication"
author: "Bernie Mulvey"
date: "06-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##


## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
precision <- sqrt(.Machine$double.eps)
(x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
precision <- sqrt(.Machine$double.eps)
(x <= y) | (abs(x-y) <= precision)}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

## with LC as a smoothed domain
```{r}
totde <- readRDS("processed-data_newseg/09_VisXferedLabelsDE/02-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_SRTlabelsAsXfered.RDS")
totde <- lapply(totde,function(x){
    z2 <- x$detab
    z2 <- rbindlist(z2,idcol="contrast",fill=T)
    return(z2)
  })
totde <- rbindlist(totde,idcol="modelset")
unfilttot <- copy(totde) # keep a copy of this

visde <- readRDS("../LFF_spatial_LC/processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02b-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_noMixClusts_YRIcontAndCat_JUL2025.RDS")
visdes <- list(visde$cont_ances$E2E4label$detabs,visde$cont_ances$sexlabel$detabs,visde$cat_ances$YRIAPOlabel$detabs)
names(visdes) <- c("E2E4label","sexlabel","YRIAPOlabel")
visdes <- lapply(visdes,rbindlist,idcol="contrast")
visdes <- rbindlist(visdes)


totde <- totde[contrast %in% c(unique(visdes$contrast))]
totde <- totde[,.(contrast,gene_id,gene_name,P.Value,logFC)]
setnames(totde,c("P.Value","logFC"),c("Xen.P.unadj","xlfc"))
## adjPVal and lfc columns will correspond to vis results
totde <- merge.data.table(totde,visdes[adj.P.Val<0.05,.(contrast,gene_id,gene_name,adj.P.Val,logFC)],by=c("contrast","gene_id","gene_name"))


reps <- totde[,.N,by="contrast"]
setnames(reps,"N","VisFDR05_tested")
y <- totde[Xen.P.unadj<0.05,.N,by="contrast"]
setnames(y,"N","XenP05sig")
z <- totde[Xen.P.unadj<0.05&(logFC*xlfc>0),.N,by="contrast"]
setnames(z,"N","XenP05sig_concord")
reps <- merge.data.table(reps,y,by="contrast",all.x=T)
reps <- merge.data.table(reps,z,by="contrast",all.x=T)
reps[,reprate:=XenP05sig_concord/VisFDR05_tested]
reps[,discordrate:=(XenP05sig-XenP05sig_concord)/XenP05sig]

fwrite(reps,"processed-data_newseg/09_VisXferedLabelsDE/03-Tabulate_VisDEreplication_results_percontrast_withXferedSRTlabels.txt",sep='\t',quote=F,row.names=F,col.names=T)

```

```{r}
### which E4-E2 genes replicate for the LC SRT domain in LC cells?
totde[contrast=="LC_E4vE2"&lessOrEqual(Xen.P.unadj,0.05)&xlfc*logFC>0,gene_name]
# ITGB8 NPY

### which E4-E2 DEGs from SRT LC domain replicate in ASTRO cells?
srtLChits <- totde[contrast=="LC_E4vE2"&lessOrEqual(adj.P.Val,0.05),gene_name]
totde[contrast=="Astro_E4vE2"&lessOrEqual(adj.P.Val,0.05)&lessOrEqual(Xen.P.unadj,0.05)&xlfc*logFC>0&gene_name %in% srtLChits,gene_name]
# C3

### are any other E4-E2 DEGs from SRT LC domain nominally DE in ASTRO cells, regardless of SRT astro domain significance? this requires a different visium table where we haven't filtered down to Visium SRT significant hits
setnames(unfilttot,"P.Value","Xen.P.unadj")
unfilttot[contrast=="Astro_E4vE2"&lessOrEqual(Xen.P.unadj,0.05)&gene_name %in% srtLChits,gene_name]
# C3 and also ITGB8
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] BiocParallel_1.44.0 parallelly_1.47.0   colorout_1.3-3     
[4] data.table_1.18.2.1

loaded via a namespace (and not attached):
 [1] digest_0.6.39     codetools_0.2-20  fastmap_1.2.0     xfun_0.56        
 [5] knitr_1.51        htmltools_0.5.9   rmarkdown_2.30    lifecycle_1.0.5  
 [9] cli_3.6.5         rsconnect_1.7.0   compiler_4.5.2    rprojroot_2.1.1  
[13] here_1.0.2        rstudioapi_0.18.0 tools_4.5.2       evaluate_1.0.5   
[17] yaml_2.3.12       otel_0.2.0        rlang_1.1.7      
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-09
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────
 package      * version  date (UTC) lib source
 BiocParallel * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli            3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools      0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout     * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table   * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 digest         0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 evaluate       1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 fastmap        1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 here           1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools      0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 knitr          1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lifecycle      1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 otel           0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly   * 1.47.0   2026-04-17 [1] CRAN (R 4.5.2)
 rlang          1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown      2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot      2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect      1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi     0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 sessioninfo    1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 xfun           0.56     2026-01-18 [1] CRAN (R 4.5.2)
 yaml           2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────