---
title: "02-Calculate_WebDivNmfAstros_forDomainLCcells_nonDomainNEcells"
author: "Bernie Mulvey"
date: "03-12-2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(sf)
library(ggplot2)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)


## ggplot defaults

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load spe, domain labels, and webdiv nmf projection cell type labels
```{r}
lcx <- readRDS("processed-data/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")
tmpcd <- as.data.table(colData(lcx),keep.rownames=T)

nmfNDom <- fread("processed-data/05_LabelTransfer/06-LCsmoothings_LUT.txt")
tmpcd <- merge.data.table(tmpcd,nmfNDom[,.(rn,domain8266,domain12252)],by="rn")

webdiv <- readRDS("processed-data/05_LabelTransfer/07-WebDivToXen_NMFlabelxfer_tol1e8_k66_alph0.RDS")
webdivs <- lapply(webdiv$targets,function(w){
  w$nmf_preds <- as.character(w$nmf_preds)
  y <- as.data.table(colData(w),keep.rownames=T)[,.(rn,nmf_preds)]
  return(y)
})
webdivs <- rbindlist(webdivs)

tmpcd <- merge.data.table(tmpcd,webdivs,by="rn")
setnames(tmpcd,"nmf_preds","webdiv_label")
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(lcx),]
colData(lcx) <- tmpcd

rm(tmpcd,nmfNDom,webdiv,webdivs)
gc(full=T)
```


Now we want to get the number of cells of each type within various distances of each LC cell.
```{r}
celllocs <- cbind(spatialCoords(lcx),as.data.table(colData(lcx))[,.(brnum,APOE,Age,Sex,Ancestry,domain8266,domain12252,webdiv_label,unique_cellid)])

samps <- split(celllocs,by="brnum")

## clean up
rm(lcx,celllocs)
gc(full=T)

## iterate through samples and radii
setDTthreads(1,restore_after_fork = FALSE)
dists <- mclapply(samps,mc.cores=8,FUN=function(s){
    setDTthreads(1,restore_after_fork = FALSE) # prevent resource overcommitment
    
    locs.sf <- st_as_sf(s,coords=c("x_centroid","y_centroid"))
    dists <- c(50,100,200,300,400,500)
    sampres <- lapply(dists,FUN=function(d){
        return(st_is_within_distance(locs.sf,locs.sf,dist=d))
    })
    names(sampres) <- c("d50","d100","d200","d300","d400","d500")
    return(sampres)
})
names(dists) <- names(samps)

## extract the pair IDs. the names of the two columns turn out differently for each sublist entry (radius length), so use .SDcols=1 to get the first column (pair index 1) or =2 for pair index 2 in the calls to lapply(.SD) below


setDTthreads(1,restore_after_fork = FALSE)

# ## this code block is very hard to conceptualize. recommend for demo purposes instead doing
# v=dists[[1]]
# s=samps[[1]]
# n=names(dists)[[1]]
# rad=v[[1]]
## then jumping to line 100 to step through and see what's happening. 

dists2 <- mcmapply(v=dists,s=samps,n=names(dists),
                   SIMPLIFY=FALSE,
                   mc.cores = 4,
                   mc.cleanup = TRUE,
                   FUN=function(v,s,n){
    setDTthreads(1,restore_after_fork = FALSE)

    ret <- lapply(v,function(rad){
        rad2 <- as.data.table(as.data.frame(rad))
        ## these retrieve the cell identifiers based on their row indices (sf returned pairs of row indices to us in the above loop) in the table of coords, cell id, and cell label it was passed
        rad2[,cell1:=lapply(.SD,FUN=function(x){s[x,unique_cellid]}),.SDcols=1]
        rad2[,cell2:=lapply(.SD,FUN=function(x){s[x,unique_cellid]}),.SDcols=2]
        
        # drop self-pairs
        rad2 <- rad2[cell1!=cell2]
        
        # add cell label info now that we have cell ids for each cell in each pair
        z <- merge.data.table(rad2,s[,.(unique_cellid,webdiv_label)],by.x="cell1",by.y="unique_cellid")
        setnames(z,"webdiv_label","webdiv_label1")
        a <- merge.data.table(z,y=s[,.(unique_cellid,webdiv_label)],by.x="cell2",by.y="unique_cellid")
        setnames(a,"webdiv_label","webdiv_label2")
        a[,sample_id:=n]
        
        ## number of NEARBY cells of a given type (label2) for a given cell (cell1). also return the label of the given cell (label1) so we can filter these by target cell type later.
        ## the neighbor table has every pair in both directions (e.g., Cell X near Cell Y and Cell Y near Cell X), so counting from either direction will give the same result. so we can just do this to subset to LC cells (NONunique pairs, just because we love LC):
        a <- a[webdiv_label1=="NE"|webdiv_label2=="NE"]
        
    })
    return(ret)
})

## append domain info for cell 1 (the query cells, NE)
dists2 <- mapply(d=dists2,s=samps,SIMPLIFY=FALSE,function(d,s){
  ret <- lapply(d,function(y){
    y <- merge.data.table(y,s[,.(unique_cellid,domain8266,domain12252)],by.x="cell1",by.y="unique_cellid")
    setnames(y,c("domain8266","domain12252"),c("domain8266_cell1","domain12252_cell1"))
    return(y)
  })
  return(ret)
})

# finally, tabulate the number of neighbors (cell2) of a given type (label2) for for each query cell (cell1). we only need to do this for the reasons above re nonunique pairs:
dists3 <- mclapply(dists2,
                   mc.cores = 4,
                   mc.cleanup = TRUE,
                   FUN=function(v){
  setDTthreads(1,restore_after_fork = FALSE)
  ret <- lapply(v,function(a){
    ## only get info for neighbors of LC cells. keep the LC cells domain info so we can stratify by in or out of domain later if we want.
    ## we simultaneously will get the number of cells of each type (nmfpred2) that are paired to each remaining cell1 (query, i.e. intra-domain LC cell) 
    
    nperclus <- a[webdiv_label1=="NE"] 
    nperclus <- nperclus[,.(.N,domain12252_cell1,domain8266_cell1,webdiv_label1),by=c("cell1","webdiv_label2")]
    nperclus <- unique(nperclus)
    
    return(nperclus)
  })
  names(ret) <- names(v)
  return(ret)
})
```

build...THE MEGA TABLE
```{r}

alldists <- lapply(dists3,function(x){
    y<-rbindlist(x,idcol="rad")
    return(y)
})
alldists <- rbindlist(alldists,idcol="brnum")

rm(dists,dists2,dists3)
gc()

setnames(alldists,"N","N_webdiv_label2")
reord <- c("brnum","rad","domain8266_cell1","domain12252_cell1","webdiv_label1","webdiv_label2","N_webdiv_label2","cell1")
alldist <- alldist[,..reord]

## append metadata for demos matching
tmpcd <- rbindlist(samps)
tmpcd <- unique(tmpcd[,.(brnum,APOE,Age,Sex,Ancestry)])
alldists <- merge.data.table(alldists,tmpcd,by.x="samp",by.y="brnum")
saveRDS(alldists,"processed-data/08_LC_ProximalAstros/02-WebDivNMFpredNeighborCounts_for_NEcells_2DomDefs_ByNECellDomDesig.RDS")
# clean up
rm(tmpcd,samps,reord)
gc(full=T)
```

violin plots of neighbors at rad500
```{r}

ad5 <- melt.data.table(alldists,id.vars = c("brnum","rad","webdiv_label1","webdiv_label2","cell1","APOE","Age","Sex","Ancestry","N_webdiv_label2"))
ad5[,variable:=as.character(variable)]
ad5[,variable:=gsub(variable,pattern="domain(.*)_.*$",replacement=paste0("d","\\1"))]
ad5[,Domain_and_Domainparam:=paste0(variable,"_",value)]

ad5 <- ad5[rad=="d400"&webdiv_label2 %in% grep(webdiv_label2,pattern="ast",value=T)]
## astrocytes_29 is probably a bullshit cell type so drop it
ad5 <- ad5[webdiv_label2!="astrocytes_29"]
ad5[,APOE:=ifelse(APOE %in% c("E2hom","E2het"),"E2","E4")]
ad5[,APOE:=factor(APOE,levels=c("E2","E4"))]

#ad5[,medneigh:=median(Nnmfpred2),by=c("samp","nmfpred2")]
#ad5 <- unique(ad5[,.(samp,APOE,medneigh,nmfpred2,nmfpred1)])
ggplot(ad5,aes(y=N_webdiv_label2,x=Domain_and_Domainparam,fill=APOE))+
  geom_violin()+
  facet_wrap(webdiv_label2~webdiv_label1,scales="free")

ggplot(ad5[variable=="d12252"],aes(x=value,y=N_webdiv_label2,fill=APOE))+
  geom_violin()+
  facet_wrap(webdiv_label2~webdiv_label1,scales="free")+
  xlab("Domain designation of NE cell (Domain '12252')")+
  ylab("Number of Astrocytes within\n400µm of NE cell centroid")
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    parallel  stats     graphics  grDevices utils     methods  
[8] base     

other attached packages:
 [1] ggplot2_4.0.2               sf_1.0-24                  
 [3] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
 [5] SummarizedExperiment_1.40.0 Biobase_2.70.0             
 [7] GenomicRanges_1.62.1        Seqinfo_1.0.0              
 [9] IRanges_2.44.0              S4Vectors_0.48.0           
[11] BiocGenerics_0.56.0         generics_0.1.4             
[13] MatrixGenerics_1.22.0       matrixStats_1.5.0          
[15] BiocParallel_1.44.0         parallelly_1.46.1          
[17] colorout_1.3-3              data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] gtable_0.3.6        rjson_0.2.23        xfun_0.56          
 [4] lattice_0.22-7      vctrs_0.7.1         tools_4.5.2        
 [7] tibble_3.3.1        proxy_0.4-29        pkgconfig_2.0.3    
[10] Matrix_1.7-4        KernSmooth_2.23-26  RColorBrewer_1.1-3 
[13] S7_0.2.1            lifecycle_1.0.5     compiler_4.5.2     
[16] farver_2.1.2        codetools_0.2-20    htmltools_0.5.9    
[19] class_7.3-23        yaml_2.3.12         crayon_1.5.3       
[22] pillar_1.11.1       classInt_0.4-11     DelayedArray_0.36.0
[25] sessioninfo_1.2.3   magick_2.9.0        abind_1.4-8        
[28] tidyselect_1.2.1    digest_0.6.39       dplyr_1.2.0        
[31] labeling_0.4.3      rprojroot_2.1.1     fastmap_1.2.0      
[34] grid_4.5.2          here_1.0.2          cli_3.6.5          
[37] SparseArray_1.10.8  magrittr_2.0.4      S4Arrays_1.10.1    
[40] dichromat_2.0-0.1   e1071_1.7-17        withr_3.0.2        
[43] scales_1.4.0        rmarkdown_2.30      XVector_0.50.0     
[46] otel_0.2.0          datasets_4.5.2      evaluate_1.0.5     
[49] knitr_1.51          rlang_1.1.7         Rcpp_1.1.1         
[52] glue_1.8.0          DBI_1.2.3           rstudioapi_0.18.0  
[55] R6_2.6.1            units_1.0-0        
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-03-11
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ───────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 class                  7.3-23   2025-01-01 [1] CRAN (R 4.5.2)
 classInt               0.4-11   2025-01-08 [1] CRAN (R 4.5.0)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 crayon                 1.5.3    2024-06-20 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DBI                    1.2.3    2024-06-02 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 e1071                  1.7-17   2025-12-18 [1] CRAN (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggplot2              * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KernSmooth             2.23-26  2025-01-01 [1] CRAN (R 4.5.2)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 labeling               0.4.3    2023-08-29 [1] CRAN (R 4.5.0)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 proxy                  0.4-29   2025-12-29 [1] CRAN (R 4.5.2)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 sf                   * 1.0-24   2026-01-13 [1] CRAN (R 4.5.2)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 units                  1.0-0    2025-10-09 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

──────────────────────────────────────────────────────────────────────────────