---
title: "Fig5-VisLabelTransfers_SmoothedLCDomain_ReplicatedE2E4degs"
format: html
date: "03-20-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(ggrastr)
library(escheR)
library(jaffelab)
library(sf)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
##

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
precision <- sqrt(.Machine$double.eps)
(x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
precision <- sqrt(.Machine$double.eps)
(x <= y) | (abs(x-y) <= precision)}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost='localhost')


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = '#000000'), axis.title.x = element_text(size = 9,color = '#000000'), axis.text.y = element_text(size = 8,color = '#000000'), axis.title.y = element_text(size =9,color = '#000000'), plot.title = element_markdown(size = 10,hjust=0.5,color = '#000000'), strip.text = element_text(size=11), legend.text = element_text(size=7,color = '#000000'), legend.title = element_text(size=8,hjust=0.5,color = '#000000'),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position='plot'))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

## Plot from a previously-plotted sample, showing transferred labels themselves, with similar anat to its visium counterpart
## poss brs: Fig 2 (Br5941, Br0946, Br6538, Br1105) or Fig 1 (Br3974)
## try Br5941, Br6538

```{r}
pal <- readRDS("../LFF_spatial_LC/plots/finalpal_somuchfornorainbows.RDS")

xlc <- readRDS("processed-data/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

lxf <- fread("processed-data/05_LabelTransfer/06-LCsmoothings_LUT.txt")
## use the strict domain for plotting
lxf <- lxf[,.(rn,domain12252,nmfpred)]
## convert working names to plotting names for cluster labels
lxf[nmfpred=="Astro_Oligo",nmfpred:="Astro-Oligo Mixed"]
lxf[nmfpred=="Oligo_Astro",nmfpred:="Oligo-Astro Mixed"]
lxf[nmfpred=="ACHE_SERT_Npep",nmfpred:="*ACHE*, *SERT*,<br>Neuropept. Mixed"]
lxf[nmfpred=="ACHE_SERT_LC_WM",nmfpred:="*ACHE*, *SERT*,<br>LC, Oligo Mixed"]
lxf[nmfpred=="Oligo_Complex",nmfpred:="Oligo-Other<br>Types Mixed"]

names(pal)[2] <- "*ACHE*, *SERT*,<br>LC, Oligo Mixed"
names(pal)[3] <- "*ACHE*, *SERT*,<br>Neuropept. Mixed"
names(pal)[9] <- "Oligo-Other<br>Types Mixed"

stopifnot(all(lxf$nmfpred %in% names(pal)))

## add to coldata
tmpcd <- as.data.table(colData(xlc),keep.rownames=T)
tmpcd <- merge.data.table(tmpcd,lxf,by="rn")
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(xlc),]
tmpcd$rn <- NULL
colData(xlc) <- tmpcd

rm(lxf)
```

### make the plots of Br5941 with cell annots, LC domain perimeter drawn on, and an inset that zooms to just the domain part. we'll do whatever rotating is necess. to match visium orientation in illustrator.

### first, determine the domain boundaries.
```{r}
pane <- xlc[,xlc$brnum=="Br5941"]

d <- st_as_sf(as.data.frame(spatialCoords(pane[,pane$domain12252=="LC"])),coords = c("x_centroid","y_centroid"))
    
dhullpt <- st_concave_hull(st_union(d),ratio = 0.2)
## create the data table we will use for overlaying the boundary polygons. geom_path will need the first point to also be popped into the last slot to close the shape
dompoly <- rbind(dhullpt[[1]][[1]],
                  dhullpt[[1]][[1]][1,])
dompoly <- cbind(dompoly,rep("LC",nrow(dompoly)))

dompoly <- as.data.table(dompoly)

setnames(dompoly,c("xpol","ypol","dompol"))

dompoly[,xpol:=as.numeric(xpol)]
dompoly[,ypol:=as.numeric(ypol)]

rm(d,dhullpt)
```

### then, make full-section escher plot with the domain perimeter overlaid...
```{r}
# sizepal <- c(0.5,rep(0.2,8)) <- we need to pass a vector of point sizes to have the polygon-over-spot plotting scheme work here:

sizesman <- ifelse(pane$nmfpred=="LC",0.5,0.2)

plt <- make_escheR(pane,y_reverse=FALSE)
plt <- plt |> add_fill("nmfpred",size=0.2,point_size = 0.2)


## assign spot colors based on our existing palette
p <- plt +
  labs(fill="Transferred\nVisium Label")+
  scale_fill_manual(values=pal)

## override the escheR size settings with the spot sizes determined above:
p@layers[[1]]$aes_params$size <- sizesman


## draw LC domain boundary
p <- p + 
  geom_path(data=dompoly,aes(x=xpol,y=ypol,group=dompol,col=dompol))+
  scale_color_manual(values=pal)

## legend specifications
p <- p +
  guides(fill=guide_legend(override.aes=list(size=1.5)),size="none",color="none")

## title
p <- p + ggtitle("Br5941 (E4/E4, AA, M)")

## for consistency/illustrator editing simplifying
p <- p+ theme(panel.grid = element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      axis.ticks = element_blank(),
      legend.text = element_markdown(size=6,margin=margin(0,0,0,0,"in")),
      legend.key.size = ggplot2::unit(0.075,"in"),
      legend.title=element_text(size=7,hjust=0.5),
      legend.background = element_blank(),
      legend.box = element_blank(),
      legend.box.spacing = ggplot2::unit(0,"in"),
      legend.key = element_blank(),
      legend.key.height=ggplot2::unit(0.0125,"in"),
      legend.key.width=ggplot2::unit(0.0125,"in"),
      legend.key.spacing.y = ggplot2::unit(0.04,"in"),
      title=element_text(size=10,hjust=0.5))
      # legend.position.inside = c(0.19,0.165),
      # legend.position="inside")

p <- rasterize(p,layers="Points",dpi=600,dev="cairo_png")

pdf(paste0("plots/manuscript/Fig5_astrogenes_in_LC/Fig5W-Predlabels_wDomainBounds_Br5941.pdf"),height=2,width=3.25)
print(p)
dev.off()
```

#### and an inset that just spans the LC area
```{r}
### get the min and max x and y coords of the LC polygon, add a little buffer space, and get cell IDs for all cells that correspond
zoomcoord <- as.data.table(spatialCoords(pane),keep.rownames=T)
range(dompoly$xpol) # 140-1277
range(dompoly$ypol) # 2307-4030

zoomcoord <- zoomcoord[greaterOrEqual(x_centroid,50)&lessOrEqual(x_centroid,1350)]
zoomcoord <- zoomcoord[greaterOrEqual(y_centroid,2250)&lessOrEqual(y_centroid,4100)]

pane2 <- pane[,zoomcoord$rn]


sizesman <- ifelse(pane2$nmfpred=="LC",0.75,0.5)

plt <- make_escheR(pane2,y_reverse=FALSE)
plt <- plt |> add_fill("nmfpred",size=0.2,point_size = 0.2)


## assign spot colors based on our existing palette
p <- plt + 
  scale_fill_manual(values=pal)

## override the escheR size settings with the spot sizes determined above:
p@layers[[1]]$aes_params$size <- sizesman


## draw LC domain boundary
p <- p + 
  geom_path(data=dompoly,aes(x=xpol,y=ypol,group=dompol,col=dompol))+
  scale_color_manual(values=pal)

## no legends needed
p <- p +
  guides(fill="none",size="none",color="none")

## no title

## for consistency/illustrator editing simplifying
p <- p+ theme(panel.grid = element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      axis.ticks = element_blank(),
      legend.text = element_markdown(size=6,margin=margin(0,0,0,0,"in")),
      legend.key.size = ggplot2::unit(0.075,"in"),
      legend.title=element_text(size=7,hjust=0.5),
      legend.background = element_blank(),
      legend.box = element_blank(),
      legend.box.spacing = ggplot2::unit(0,"in"),
      legend.key = element_blank(),
      legend.key.height=ggplot2::unit(0.0125,"in"),
      legend.key.width=ggplot2::unit(0.0125,"in"),
      legend.key.spacing.y = ggplot2::unit(0.04,"in"),
      title=element_text(size=10,hjust=0.5))
      # legend.position.inside = c(0.19,0.165),
      # legend.position="inside")

p <- rasterize(p,layers="Points",dpi=600,dev="cairo_png")

pdf(paste0("plots/manuscript/Fig5_astrogenes_in_LC/Fig5Winset-Predlabels_wDomainBounds_Br5941.pdf"),height=2.25,width=1.5)
print(p)
dev.off()
```
We looked at whether the Astro cells or the LC cells expressed astrocyte markers identified as Visium LC APOE-DEGs, including PLPP3, FGFR3, NTSR2. (Pseudobulk expression plots of these and DBH, TH in these two Xenium clusters, ostensibly showing that the expression is in the Astro cluster.)
```{r}
xactxfer <- readRDS("processed-data/06_DomainDE/07-visLabsAsXfered_voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_withinLCVisLabs.RDS")

xact.12252 <- xactxfer$E2E4label$voomLFt$EList$E[c("DBH","PLPP3","FGFR3","NTSR2"),]
# reconfigure data for jaffelab::cleaningy
# get sample metadata in the same arrangement for LC and astro pbulks
keepsamps <- grep(colnames(xact.12252),pattern="^LC_Br|^Astro_Br",value=T)
xcd <- xactxfer$E2E4label$dgel$samples[keepsamps,]
xact.12252 <- xact.12252[,keepsamps]

# plot: domain 8266
# ensure factors are set for model matrix
xcd$APOEsimple <- factor(xcd$APOEsimple,levels = c("E2","E4"))
xcd$Sex <- factor(xcd$Sex)
xcd$Age <- as.numeric(xcd$Age)
xcd$YRI <- as.numeric(xcd$YRI)
# model matrix

modmat <- model.matrix(~APOEsimple+Sex+Age+YRI,data = xcd)
## correcting for all covars (i.e., retaining no variables in the model matrix)
testres <- jaffelab::cleaningY(xact.12252,modmat,1)


pltg <- as.data.table(cbind(t(testres),xcd$label,xcd$brnum))
setnames(pltg,c((ncol(pltg)-1):ncol(pltg)),c("label","brnum"))
pltg <- as.data.table(cbind(pltg[,lapply(.SD,as.numeric),.SDcols=c(1:(ncol(pltg)-2))],pltg[,.(label,brnum)]))

pltg <- melt.data.table(pltg,id.vars=c("label","brnum"))
pltg[,titleg:=paste0("*",variable,"*")]
pltg[,label:=as.character(label)]

pdf("plots/manuscript/Fig5_astrogenes_in_LC/Fig5X-bulk_Xpr_DBH_and_astroDEGs.pdf",height=3,width=2.5)
ggplot(pltg,aes(x=label,y=value,col=label))+
    geom_boxplot(outliers = F,linewidth=0.375)+
    geom_point(alpha=0.7,size=1,stroke = 0.1,position = position_jitter(width=0.2))+
    facet_wrap(.~titleg,nrow=2)+
    xlab("Cell Label Transferred from Visium")+
    ylab("Pseudobulk Expression (Corrected for Age,\nSex, APOE Status, Genomic Ancestry)")+
    scale_color_manual(values=pal)+
    guides(color="none")+
    theme(panel.grid.major = element_line(linewidth=0.25),
         panel.grid.minor=element_line(linewidth=0.125),
         axis.text.x=element_text(size=8),
         axis.text.y=element_text(size=8),
         axis.title.x=element_text(size=9),
         axis.title.y=element_text(size=9),
         strip.text = element_markdown(size=9,margin=margin(0,0,-0.025,0,"in")),
         axis.ticks = element_line(linewidth=0.25),
         strip.background = element_blank())
dev.off()

rm(pltg,modmat,testres,xact.12252,xcd,xactxfer)
```

show an LC smoothing on Br5941 or Br6538. we'll do whatever rotation necess to match visium panels' orientation in illustrator
```{r}
smoovepans <- list(Br5941=xlc[,xlc$brnum=="Br5941"],Br6538=xlc[,xlc$brnum=="Br6538"])

dompal <- c(pal[1],"#C3C3C3")
names(dompal) <- c("LC","Other")
sizepal <- c(0.4,0.25)
names(sizepal) <- c("LC","Other")

lapply(names(smoovepans),function(n){
    smoove <- smoovepans[[n]]
    plt <- make_escheR(smoove,y_reverse=FALSE)
    plt <- plt |> add_fill("domain12252",size=0.2,point_size = 0.2)

    ## override the escheR size settings to pass our size pal
    plt@layers[[1]]$aes_params$size <- NULL
    plt <- plt+aes(size=domain12252)+
        scale_size_manual(values=sizepal)

    
    if(n=="Br6538"){ # Br6538 (E2/E3, EA, M)
        p <- plt + ggtitle(paste0(n," (E2/E3\nEA, M)"))
    } else if (n=="Br5941"){ # Br5941 (E4/E4, AA, M)
        p <- plt + ggtitle(paste0(n," (E4/E4\nAA, M)"))
    }
    p <- p+ 
    labs(fill="Domain")+
    guides(fill=guide_legend(override.aes=list(size=1.5)),size="none")+
    scale_fill_manual(values=dompal)+
    theme(panel.grid = element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_markdown(size=6,margin=margin(0,0,0,0,"in")),
        legend.key.size = ggplot2::unit(0.075,"in"),
        legend.title=element_text(size=7,hjust=0.5),
        legend.background = element_blank(),
        legend.margin = margin(0,-0.3,0,0,"in"),
        legend.box = element_blank(),
        legend.box.spacing = ggplot2::unit(0,"in"),
        legend.key = element_blank(),
        legend.key.height=ggplot2::unit(0.0125,"in"),
        legend.key.width=ggplot2::unit(0.0125,"in"),
        legend.key.spacing.y = ggplot2::unit(0.04,"in"),
        title=element_text(size=10,hjust=0.5))
        # legend.position.inside = c(0.19,0.165),
        # legend.position="inside")
    p <- rasterize(p,layers="Points",dpi=600,dev="cairo_png")
    pdf(paste0("plots/manuscript/Fig5_astrogenes_in_LC/Fig5Y-SmoothedLC_",n,".pdf"),height=2.25,width=1.5)
    print(p)
    dev.off()
})
```

pseudobulk cleaning y plots for Xenium-replicated LC domain astrocyte DEGs
```{r}
domde <- readRDS("processed-data/06_DomainDE/02-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_XferedVisLabels.RDS")

dom.12252 <- domde$dom12252$E2E4label$voomLFt$EList$E[c("PLPP3","FGFR3","NTSR2"),]
# reconfigure data for jaffelab::cleaningy
# get sample metadata in the same arrangement for LC and astro pbulks
keepsamps <- grep(colnames(dom.12252),pattern="LC_Br",value=T)
xcd <- domde$dom12252$E2E4label$dgel$samples[keepsamps,]
dom.12252 <- dom.12252[,keepsamps]

# plot: domain 8266
# ensure factors are set for model matrix
xcd$APOEsimple <- factor(xcd$APOEsimple,levels = c("E2","E4"))
xcd$Sex <- factor(xcd$Sex)
xcd$Age <- as.numeric(xcd$Age)
xcd$YRI <- as.numeric(xcd$YRI)

modmat <-  model.matrix(~0+APOEsimple+Age+YRI+Sex,data = xcd)
degs <- jaffelab::cleaningY(dom.12252,modmat,2) # correct for terms that arent the two levels of APOEsimple

pltg <- as.data.table(cbind(t(degs),xcd$APOEsimple,xcd$brnum))
setnames(pltg,c((ncol(pltg)-1):ncol(pltg)),c("APOE","brnum"))
pltg <- as.data.table(cbind(pltg[,lapply(.SD,as.numeric),.SDcols=c(1:(ncol(pltg)-2))],pltg[,.(APOE,brnum)]))

pltg <- melt.data.table(pltg,id.vars=c("APOE","brnum"))
pltg[,titleg:=paste0("*",variable,"*")]
pltg[,APOE:=as.character(APOE)]
pltg[,APOE:=ifelse(APOE=="1","E2","E4")]

pdf("plots/manuscript/Fig5_astrogenes_in_LC/Fig5Z-Pbulk_astrogenes_byAPOE_inLCdomain12252.pdf",height=1.5,width=3)
ggplot(pltg,aes(x=APOE,y=value,col=APOE))+
    geom_boxplot(outliers = F,linewidth=0.375)+
    geom_point(alpha=0.7,size=1,stroke = 0.1,position = position_jitter(width=0.2))+
    facet_wrap(.~titleg,nrow=1,scales = "free_y")+
    xlab("APOE Status")+
    ylab("LC Domain Expression")+
    scale_color_manual(values=c("E2"="#009E73","E4"="#D55E00"))+
    guides(color="none")+
    theme(panel.grid.major = element_line(linewidth=0.25),
         panel.grid.minor=element_line(linewidth=0.125),
         axis.text.x=element_text(size=8),
         axis.text.y=element_text(size=8),
         axis.title.x=element_text(size=9),
         axis.title.y=element_text(size=9),
         strip.text = element_markdown(size=9,margin=margin(0,0,-0.025,0,"in")),
         axis.ticks = element_line(linewidth=0.25),
         strip.background = element_blank())
dev.off()
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices
[5] utils     methods   base     

other attached packages:
 [1] jaffelab_0.99.34           
 [2] rafalib_1.0.4              
 [3] BiocParallel_1.44.0        
 [4] parallelly_1.46.1          
 [5] escheR_1.10.0              
 [6] ggrastr_1.0.2              
 [7] SpatialExperiment_1.20.0   
 [8] SingleCellExperiment_1.32.0
 [9] SummarizedExperiment_1.40.0
[10] Biobase_2.70.0             
[11] GenomicRanges_1.62.1       
[12] Seqinfo_1.0.0              
[13] IRanges_2.44.0             
[14] S4Vectors_0.48.0           
[15] BiocGenerics_0.56.0        
[16] generics_0.1.4             
[17] MatrixGenerics_1.22.0      
[18] matrixStats_1.5.0          
[19] gridExtra_2.3              
[20] ggtext_0.1.2               
[21] ggplot2_4.0.2              
[22] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] tidyselect_1.2.1    dplyr_1.2.0        
 [3] vipor_0.4.7         farver_2.1.2       
 [5] S7_0.2.1            lifecycle_1.0.5    
 [7] statmod_1.5.1       magrittr_2.0.4     
 [9] compiler_4.5.2      rlang_1.1.7        
[11] tools_4.5.2         knitr_1.51         
[13] S4Arrays_1.10.1     labeling_0.4.3     
[15] here_1.0.2          DelayedArray_0.36.0
[17] xml2_1.5.2          RColorBrewer_1.1-3 
[19] abind_1.4-8         purrr_1.2.1        
[21] withr_3.0.2         grid_4.5.2         
[23] googledrive_2.1.2   edgeR_4.8.2        
[25] paletteer_1.7.0     scales_1.4.0       
[27] MASS_7.3-65         dichromat_2.0-0.1  
[29] cli_3.6.5           ragg_1.5.0         
[31] otel_0.2.0          rjson_0.2.23       
[33] sessioninfo_1.2.3   commonmark_2.0.0   
[35] ggbeeswarm_0.7.3    stringr_1.6.0      
[37] splines_4.5.2       parallel_4.5.2     
[39] XVector_0.50.0      vctrs_0.7.1        
[41] Matrix_1.7-4        litedown_0.9       
[43] beeswarm_0.4.0      systemfonts_1.3.1  
[45] magick_2.9.0        locfit_1.5-9.12    
[47] limma_3.66.0        glue_1.8.0         
[49] rematch2_2.1.2      codetools_0.2-20   
[51] stringi_1.8.7       gtable_0.3.6       
[53] tibble_3.3.1        pillar_1.11.1      
[55] R6_2.6.1            textshaping_1.0.4  
[57] rprojroot_2.1.1     evaluate_1.0.5     
[59] lattice_0.22-7      markdown_2.0       
[61] gridtext_0.1.5      segmented_2.2-1    
[63] gargle_1.6.1        Rcpp_1.1.1         
[65] SparseArray_1.10.8  nlme_3.1-168       
[67] xfun_0.56           fs_1.6.6           
[69] pkgconfig_2.0.3    
> sessioninfo::session_info()
─ Session info ──────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       Positron
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-03-19
 pandoc   NA
 quarto   1.8.27 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 commonmark             2.0.0    2025-07-07 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 edgeR                  4.8.2    2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 escheR               * 1.10.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fs                     1.6.6    2025-04-12 [1] CRAN (R 4.5.0)
 gargle                 1.6.1    2026-01-29 [1] CRAN (R 4.5.2)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggbeeswarm             0.7.3    2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggrastr              * 1.0.2    2023-06-01 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 googledrive            2.1.2    2025-09-10 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 jaffelab             * 0.99.34  2026-02-24 [1] Github (LieberInstitute/jaffelab@e7a26ca)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 labeling               0.4.3    2023-08-29 [1] CRAN (R 4.5.0)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 limma                  3.66.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 litedown               0.9      2025-12-18 [1] CRAN (R 4.5.2)
 locfit                 1.5-9.12 2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 markdown               2.0      2025-03-23 [1] CRAN (R 4.5.0)
 MASS                   7.3-65   2025-02-28 [1] CRAN (R 4.5.2)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 nlme                   3.1-168  2025-03-31 [1] CRAN (R 4.5.2)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 paletteer              1.7.0    2026-01-08 [1] CRAN (R 4.5.2)
 parallelly           * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 purrr                  1.2.1    2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 rafalib              * 1.0.4    2025-04-08 [1] CRAN (R 4.5.0)
 ragg                   1.5.0    2025-09-02 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rematch2               2.1.2    2020-05-01 [1] CRAN (R 4.5.0)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 segmented              2.2-1    2026-01-29 [1] CRAN (R 4.5.2)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 statmod                1.5.1    2025-10-09 [1] CRAN (R 4.5.0)
 stringi                1.8.7    2025-03-27 [1] CRAN (R 4.5.0)
 stringr                1.6.0    2025-11-04 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 systemfonts            1.3.1    2025-10-01 [1] CRAN (R 4.5.0)
 textshaping            1.0.4    2025-10-10 [1] CRAN (R 4.5.0)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.5.0)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────