---
title: "05-NMFlabeltransfer_PreandPost_smoother"
author: "Bernie Mulvey"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(ggrastr)
library(SpatialExperiment)
library(escheR)
library(nmfLabelTransfer)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load xenium sample list with projected unsmoothed domain assigns 
```{r}
xenprj <- readRDS("processed-data/00_prelim_procdat_BJM/05-VistoXen_NMFlabelxfer_tol1e8_k100_alph0.RDS")

xenprj <- xenprj$targets
## fix weirdness with how the projected labels get stored
xenprj <- lapply(xenprj,function(x){
  x$nmfpred_unsmooth <- as.character(x$nmf_preds)
  return(x)
})
```

fix a bug in smoother function
```{r}
smootherfix <- function (labels_curr, locs, k = 10L, props = NULL, max_iter = 10, 
  verbose = TRUE) 
{
  if (is.null(props)) {
    props <- rep(0.5, length(unique(labels_curr)))
    names(props) <- unique(labels_curr)
  }
  knn <- kNN(locs, k = k)$id
  N <- nrow(knn)
  labels_raw <- labels_curr
  labels_update <- labels_curr
  if (max_iter == -1) 
    max_iter <- Inf
  iter <- 1
  while (iter < max_iter) {
    if (verbose) {
      message("Iteration ", iter)
    }
    for (i in seq(N)) {
      neighbor_labels <- labels_curr[c(i, knn[i, ])]
      neighbor_props <- table(neighbor_labels)/length(neighbor_labels)
      cell_type_i <- labels_curr[[i]]
      prop_thres <- props[cell_type_i]
      if (any(neighbor_props > prop_thres)) {
        labels_update[i] <- names(which.max(neighbor_props))
      }
      else {
        labels_update[i] <- labels_curr[i]
      }
    }
    change <- sum(labels_update != labels_curr)
    if (verbose) {
      message("Change: ", change)
    }
    if (change == 0) {
      break
    }
    labels_curr <- labels_update
    iter <- iter + 1
  }
  labels_update
}
```

## note that, as the function is written above, custom proportion thresholds per cell type will behave differently than I would assume they aim to:

lets take a hypothetical scenario where I have three labels: oligo, vascular, and LC. I set custom prop thresholds:
Oligo 0.65
LC 0.15
Vascular 0.3

Now, I reach an LC cell with the following neighbor proportions:
Oligo: 0.6
LC: 0.2
Vascular: 0.2

The Oligo 0.6 fulfills the >prop_thres condition, because for an LC cell prop_thres is 0.25. We would then be updating the label to oligo not because it passed the oligo threshold of 0.65, but because it passed the LC threshold of 0.15 (in which case we would want to give it the label of LC. Or that’s what I would interpret this parameter to be trying to do, anyhow)

So we can really only use the defaults for now.

#### smoove with default proportions (0.5 for all labels), k=25, 100 iter ####

adapting from (ttps://github.com/LieberInstitute/spatialDLPFC_SCZ_XENIUM/blob/devel/code/analysis/04_label_transfer/01_run_spaTransfer.R 
```{r}
xenprj2 <- mclapply(xenprj,mc.cores=6,function(x){
  custprops <- table(x$nmfpred_unsmooth)/length(x$nmfpred_unsmooth)
  custprops <- custprops*1.5
  x$nmfpred_smooth <- smootherfix(
    labels_curr = colData(x)[,"nmfpred_unsmooth"],
    locs=spatialCoords(x),
    k=25,
    max_iter = -1)
  return(x)})
```

## load palette
```{r}
pal <- readRDS("../LFF_spatial_LC/plots/finalpal_somuchfornorainbows.RDS")

names(pal)
unique(xenprj2[[1]]$nmfpred_smooth)

names(pal)[2] <- "ACHE_SERT_LC_WM"
names(pal)[3] <- "ACHE_SERT_Npep"
names(pal)[5] <- "Astro_Oligo"
names(pal)[8] <- "Oligo_Astro"
names(pal)[9] <- "Oligo_Complex"

stopifnot(all(names(pal) %in% xenprj2$Br0942$nmfpred_smooth))
```

plot of unsmoothed and defaultish smoothed
```{r}
plts <- lapply(xenprj2,function(spe){
  p <- make_escheR(spe, y_reverse=FALSE)
  p <- p |> add_fill("nmfpred_unsmooth",size=0.7,point_size=0.7)
  p <- p+scale_fill_manual(
    name = "Unsmoothed Domain",
    values = pal,
    guide = guide_legend(override.aes = list(size = 3)))
  p <- rasterize(p,layers="Points",dpi=300,dev="cairo_png")
  return(p)
})


pdf("plots/00_prelimplotsBJM/05_Visdomaintransfer/01a-VisiumLabelsTransferred_nosmoothing.pdf",height=36,width=22)
do.call("grid.arrange",c(plts,ncol=4))
dev.off()

#############################

plts <- lapply(xenprj2,function(spe){
  p <- make_escheR(spe, y_reverse=FALSE)
  p <- p |> add_fill("nmfpred_smooth",size=0.7,point_size=0.7)
  p <- p+scale_fill_manual(
    name = "Smoothed Domain",
    values = pal,
    guide = guide_legend(override.aes = list(size = 3)))
  p <- rasterize(p,layers="Points",dpi=300,dev="cairo_png")
  return(p)
})


pdf("plots/00_prelimplotsBJM/05_Visdomaintransfer/01b-VisiumLabelsTransferred_smoothed_k25_100iter_defaultProps.pdf",height=36,width=22)
do.call("grid.arrange",c(plts,ncol=4))
dev.off()
```

So, some tuning will be needed.
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] ggrastr_1.0.2               dbscan_1.2.4               
 [3] nmfLabelTransfer_0.0.0.9000 singlet_0.99.8             
 [5] RcppEigen_0.3.4.0.2         dplyr_1.1.4                
 [7] RcppML_0.5.6                Seurat_5.4.0               
 [9] SeuratObject_5.3.0          sp_2.2-0                   
[11] BiocParallel_1.40.2         parallelly_1.46.1          
[13] colorout_1.3-0.2            escheR_1.6.0               
[15] SpatialExperiment_1.16.0    SingleCellExperiment_1.28.1
[17] SummarizedExperiment_1.36.0 Biobase_2.66.0             
[19] GenomicRanges_1.58.0        GenomeInfoDb_1.42.3        
[21] IRanges_2.40.1              S4Vectors_0.44.0           
[23] BiocGenerics_0.52.0         MatrixGenerics_1.18.1      
[25] matrixStats_1.5.0           gridExtra_2.3              
[27] ggtext_0.1.2                ggplot2_4.0.1              
[29] data.table_1.17.4          

loaded via a namespace (and not attached):
  [1] RcppAnnoy_0.0.23        splines_4.4.3           later_1.4.4            
  [4] tibble_3.3.0            polyclip_1.10-7         fastDummies_1.7.5      
  [7] lifecycle_1.0.4         rprojroot_2.1.1         globals_0.18.0         
 [10] lattice_0.22-7          MASS_7.3-65             magrittr_2.0.4         
 [13] limma_3.62.2            plotly_4.11.0           rmarkdown_2.30         
 [16] yaml_2.3.12             httpuv_1.6.16           otel_0.2.0             
 [19] sctransform_0.4.3       spam_2.11-1             spatstat.sparse_3.1-0  
 [22] reticulate_1.44.1       cowplot_1.2.0           pbapply_1.7-4          
 [25] RColorBrewer_1.1-3      abind_1.4-8             zlibbioc_1.52.0        
 [28] Rtsne_0.17              purrr_1.2.0             msigdbr_25.1.1         
 [31] GenomeInfoDbData_1.2.13 ggrepel_0.9.6           irlba_2.3.5.1          
 [34] listenv_0.10.0          spatstat.utils_3.2-1    goftest_1.2-3          
 [37] RSpectra_0.16-2         spatstat.random_3.4-4   fitdistrplus_1.2-6     
 [40] codetools_0.2-20        DelayedArray_0.32.0     xml2_1.5.0             
 [43] tidyselect_1.2.1        shape_1.4.6.1           UCSC.utils_1.2.0       
 [46] farver_2.1.2            spatstat.explore_3.7-0  jsonlite_2.0.0         
 [49] progressr_0.18.0        ggridges_0.5.7          survival_3.8-3         
 [52] iterators_1.0.14        foreach_1.5.2           tools_4.4.3            
 [55] ica_1.0-3               Rcpp_1.1.1              glue_1.8.0             
 [58] SparseArray_1.6.2       xfun_0.56               here_1.0.2             
 [61] withr_3.0.2             fastmap_1.2.0           digest_0.6.39          
 [64] R6_2.6.1                mime_0.13               scattermore_1.2        
 [67] tensor_1.5.1            dichromat_2.0-0.1       spatstat.data_3.1-9    
 [70] tidyr_1.3.1             generics_0.1.4          httr_1.4.7             
 [73] htmlwidgets_1.6.4       S4Arrays_1.6.0          uwot_0.2.4             
 [76] pkgconfig_2.0.3         gtable_0.3.6            lmtest_0.9-40          
 [79] S7_0.2.1                XVector_0.46.0          htmltools_0.5.8.1      
 [82] dotCall64_1.2           fgsea_1.32.4            scales_1.4.0           
 [85] png_0.1-8               knn.covertree_1.0       spatstat.univar_3.1-6  
 [88] knitr_1.51              rstudioapi_0.17.1       reshape2_1.4.5         
 [91] rjson_0.2.23            nlme_3.1-168            curl_7.0.0             
 [94] zoo_1.8-14              stringr_1.6.0           KernSmooth_2.23-26     
 [97] vipor_0.4.7             miniUI_0.1.2            pillar_1.11.1          
[100] grid_4.4.3              vctrs_0.7.1             RANN_2.6.2             
[103] promises_1.5.0          xtable_1.8-4            cluster_2.1.8.1        
[106] beeswarm_0.4.0          evaluate_1.0.5          magick_2.9.0           
[109] cli_3.6.5               compiler_4.4.3          rlang_1.1.7            
[112] crayon_1.5.3            future.apply_1.20.1     labeling_0.4.3         
[115] ggbeeswarm_0.7.2        plyr_1.8.9              stringi_1.8.7          
[118] viridisLite_0.4.2       deldir_2.0-4            assertthat_0.2.1       
[121] babelgene_22.9          lazyeval_0.2.2          spatstat.geom_3.7-0    
[124] glmnet_4.1-10           Matrix_1.7-4            RcppHNSW_0.6.0         
[127] patchwork_1.3.2         future_1.69.0           statmod_1.5.1          
[130] shiny_1.11.1            ROCR_1.0-12             gridtext_0.1.5         
[133] igraph_2.2.1            fastmatch_1.1-6        
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-01-29
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ─────────────────────────────────────────────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [1] CRAN (R 4.4.1)
 assertthat             0.2.1      2019-03-21 [1] CRAN (R 4.4.0)
 babelgene              22.9       2022-09-29 [1] CRAN (R 4.4.0)
 beeswarm               0.4.0      2021-06-01 [1] CRAN (R 4.4.0)
 Biobase              * 2.66.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 BiocGenerics         * 0.52.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 BiocParallel         * 1.40.2     2025-04-08 [1] Bioconductor
 cli                    3.6.5      2025-04-23 [1] CRAN (R 4.4.1)
 cluster                2.1.8.1    2025-03-12 [1] CRAN (R 4.4.1)
 codetools              0.2-20     2024-03-31 [1] CRAN (R 4.4.3)
 colorout             * 1.3-0.2    2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
 cowplot                1.2.0      2025-07-07 [1] CRAN (R 4.4.1)
 crayon                 1.5.3      2024-06-20 [1] CRAN (R 4.4.0)
 curl                   7.0.0      2025-08-19 [1] CRAN (R 4.4.1)
 data.table           * 1.17.4     2025-05-26 [1] CRAN (R 4.4.3)
 dbscan               * 1.2.4      2025-12-19 [1] CRAN (R 4.4.3)
 DelayedArray           0.32.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 deldir                 2.0-4      2024-02-28 [1] CRAN (R 4.4.0)
 dichromat              2.0-0.1    2022-05-02 [1] CRAN (R 4.4.0)
 digest                 0.6.39     2025-11-19 [1] CRAN (R 4.4.3)
 dotCall64              1.2        2024-10-04 [1] CRAN (R 4.4.1)
 dplyr                * 1.1.4      2023-11-17 [1] CRAN (R 4.4.0)
 escheR               * 1.6.0      2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 evaluate               1.0.5      2025-08-27 [1] CRAN (R 4.4.1)
 farver                 2.1.2      2024-05-13 [1] CRAN (R 4.4.0)
 fastDummies            1.7.5      2025-01-20 [1] CRAN (R 4.4.1)
 fastmap                1.2.0      2024-05-15 [1] CRAN (R 4.4.0)
 fastmatch              1.1-6      2024-12-23 [1] CRAN (R 4.4.1)
 fgsea                  1.32.4     2025-03-20 [1] Bioconductor 3.20 (R 4.4.3)
 fitdistrplus           1.2-6      2026-01-24 [1] CRAN (R 4.4.3)
 foreach                1.5.2      2022-02-02 [1] CRAN (R 4.4.0)
 future                 1.69.0     2026-01-16 [1] CRAN (R 4.4.3)
 future.apply           1.20.1     2025-12-09 [1] CRAN (R 4.4.3)
 generics               0.1.4      2025-05-09 [1] CRAN (R 4.4.1)
 GenomeInfoDb         * 1.42.3     2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
 GenomeInfoDbData       1.2.13     2024-12-12 [1] Bioconductor
 GenomicRanges        * 1.58.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 ggbeeswarm             0.7.2      2023-04-29 [1] CRAN (R 4.4.0)
 ggplot2              * 4.0.1      2025-11-14 [1] CRAN (R 4.4.3)
 ggrastr              * 1.0.2      2023-06-01 [1] CRAN (R 4.4.0)
 ggrepel                0.9.6      2024-09-07 [1] CRAN (R 4.4.1)
 ggridges               0.5.7      2025-08-27 [1] CRAN (R 4.4.1)
 ggtext               * 0.1.2      2022-09-16 [1] CRAN (R 4.4.0)
 glmnet                 4.1-10     2025-07-17 [1] CRAN (R 4.4.1)
 globals                0.18.0     2025-05-08 [1] CRAN (R 4.4.1)
 glue                   1.8.0      2024-09-30 [1] CRAN (R 4.4.1)
 goftest                1.2-3      2021-10-07 [1] CRAN (R 4.4.0)
 gridExtra            * 2.3        2017-09-09 [1] CRAN (R 4.4.0)
 gridtext               0.1.5      2022-09-16 [1] CRAN (R 4.4.0)
 gtable                 0.3.6      2024-10-25 [1] CRAN (R 4.4.1)
 here                   1.0.2      2025-09-15 [1] CRAN (R 4.4.1)
 htmltools              0.5.8.1    2024-04-04 [1] CRAN (R 4.4.0)
 htmlwidgets            1.6.4      2023-12-06 [1] CRAN (R 4.4.0)
 httpuv                 1.6.16     2025-04-16 [1] CRAN (R 4.4.1)
 httr                   1.4.7      2023-08-15 [1] CRAN (R 4.4.0)
 ica                    1.0-3      2022-07-08 [1] CRAN (R 4.4.0)
 igraph                 2.2.1      2025-10-27 [1] CRAN (R 4.4.1)
 IRanges              * 2.40.1     2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
 irlba                  2.3.5.1    2022-10-03 [1] CRAN (R 4.4.0)
 iterators              1.0.14     2022-02-05 [1] CRAN (R 4.4.0)
 jsonlite               2.0.0      2025-03-27 [1] CRAN (R 4.4.1)
 KernSmooth             2.23-26    2025-01-01 [1] CRAN (R 4.4.3)
 knitr                  1.51       2025-12-20 [1] CRAN (R 4.4.3)
 knn.covertree          1.0        2019-10-28 [1] CRAN (R 4.4.0)
 labeling               0.4.3      2023-08-29 [1] CRAN (R 4.4.0)
 later                  1.4.4      2025-08-27 [1] CRAN (R 4.4.1)
 lattice                0.22-7     2025-04-02 [1] CRAN (R 4.4.1)
 lazyeval               0.2.2      2019-03-15 [1] CRAN (R 4.4.0)
 lifecycle              1.0.4      2023-11-07 [1] CRAN (R 4.4.0)
 limma                  3.62.2     2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
 listenv                0.10.0     2025-11-02 [1] CRAN (R 4.4.1)
 lmtest                 0.9-40     2022-03-21 [1] CRAN (R 4.4.0)
 magick                 2.9.0      2025-09-08 [1] CRAN (R 4.4.1)
 magrittr               2.0.4      2025-09-12 [1] CRAN (R 4.4.1)
 MASS                   7.3-65     2025-02-28 [1] CRAN (R 4.4.1)
 Matrix                 1.7-4      2025-08-28 [1] CRAN (R 4.4.1)
 MatrixGenerics       * 1.18.1     2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
 matrixStats          * 1.5.0      2025-01-07 [1] CRAN (R 4.4.1)
 mime                   0.13       2025-03-17 [1] CRAN (R 4.4.1)
 miniUI                 0.1.2      2025-04-17 [1] CRAN (R 4.4.1)
 msigdbr                25.1.1     2025-07-21 [1] CRAN (R 4.4.1)
 nlme                   3.1-168    2025-03-31 [1] CRAN (R 4.4.1)
 nmfLabelTransfer     * 0.0.0.9000 2026-01-28 [1] Github (cindyfang70/spaTransfer@8b2d090)
 otel                   0.2.0      2025-08-29 [1] CRAN (R 4.4.1)
 parallelly           * 1.46.1     2026-01-08 [1] CRAN (R 4.4.3)
 patchwork              1.3.2      2025-08-25 [1] CRAN (R 4.4.1)
 pbapply                1.7-4      2025-07-20 [1] CRAN (R 4.4.1)
 pillar                 1.11.1     2025-09-17 [1] CRAN (R 4.4.1)
 pkgconfig              2.0.3      2019-09-22 [1] CRAN (R 4.4.0)
 plotly                 4.11.0     2025-06-19 [1] CRAN (R 4.4.1)
 plyr                   1.8.9      2023-10-02 [1] CRAN (R 4.4.0)
 png                    0.1-8      2022-11-29 [1] CRAN (R 4.4.0)
 polyclip               1.10-7     2024-07-23 [1] CRAN (R 4.4.0)
 progressr              0.18.0     2025-11-06 [1] CRAN (R 4.4.1)
 promises               1.5.0      2025-11-01 [1] CRAN (R 4.4.1)
 purrr                  1.2.0      2025-11-04 [1] CRAN (R 4.4.1)
 R6                     2.6.1      2025-02-15 [1] CRAN (R 4.4.1)
 RANN                   2.6.2      2024-08-25 [1] CRAN (R 4.4.1)
 RColorBrewer           1.1-3      2022-04-03 [1] CRAN (R 4.4.0)
 Rcpp                   1.1.1      2026-01-10 [1] CRAN (R 4.4.3)
 RcppAnnoy              0.0.23     2026-01-12 [1] CRAN (R 4.4.3)
 RcppEigen            * 0.3.4.0.2  2024-08-24 [1] CRAN (R 4.4.1)
 RcppHNSW               0.6.0      2024-02-04 [1] CRAN (R 4.4.1)
 RcppML               * 0.5.6      2026-01-27 [1] Github (zdebruine/RcppML@5449a5b)
 reshape2               1.4.5      2025-11-12 [1] CRAN (R 4.4.1)
 reticulate             1.44.1     2025-11-14 [1] CRAN (R 4.4.3)
 rjson                  0.2.23     2024-09-16 [1] CRAN (R 4.4.1)
 rlang                  1.1.7      2026-01-09 [1] CRAN (R 4.4.3)
 rmarkdown              2.30       2025-09-28 [1] CRAN (R 4.4.1)
 ROCR                   1.0-12     2026-01-23 [1] CRAN (R 4.4.3)
 rprojroot              2.1.1      2025-08-26 [1] CRAN (R 4.4.1)
 RSpectra               0.16-2     2024-07-18 [1] CRAN (R 4.4.0)
 rstudioapi             0.17.1     2024-10-22 [1] CRAN (R 4.4.1)
 Rtsne                  0.17       2023-12-07 [1] CRAN (R 4.4.0)
 S4Arrays               1.6.0      2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 S4Vectors            * 0.44.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 S7                     0.2.1      2025-11-14 [1] CRAN (R 4.4.3)
 scales                 1.4.0      2025-04-24 [1] CRAN (R 4.4.1)
 scattermore            1.2        2023-06-12 [1] CRAN (R 4.4.0)
 sctransform            0.4.3      2026-01-10 [1] CRAN (R 4.4.3)
 sessioninfo            1.2.3      2025-02-05 [1] CRAN (R 4.4.1)
 Seurat               * 5.4.0      2025-12-14 [1] CRAN (R 4.4.3)
 SeuratObject         * 5.3.0      2025-12-12 [1] CRAN (R 4.4.3)
 shape                  1.4.6.1    2024-02-23 [1] CRAN (R 4.4.0)
 shiny                  1.11.1     2025-07-03 [1] CRAN (R 4.4.1)
 SingleCellExperiment * 1.28.1     2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
 singlet              * 0.99.8     2026-01-27 [1] Github (zdebruine/singlet@7ce199e)
 sp                   * 2.2-0      2025-02-01 [1] CRAN (R 4.4.1)
 spam                   2.11-1     2025-01-20 [1] CRAN (R 4.4.1)
 SparseArray            1.6.2      2025-02-20 [1] Bioconductor 3.20 (R 4.4.2)
 SpatialExperiment    * 1.16.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 spatstat.data          3.1-9      2025-10-18 [1] CRAN (R 4.4.1)
 spatstat.explore       3.7-0      2026-01-22 [1] CRAN (R 4.4.3)
 spatstat.geom          3.7-0      2026-01-20 [1] CRAN (R 4.4.3)
 spatstat.random        3.4-4      2026-01-21 [1] CRAN (R 4.4.3)
 spatstat.sparse        3.1-0      2024-06-21 [1] CRAN (R 4.4.0)
 spatstat.univar        3.1-6      2026-01-17 [1] CRAN (R 4.4.3)
 spatstat.utils         3.2-1      2026-01-10 [1] CRAN (R 4.4.3)
 statmod                1.5.1      2025-10-09 [1] CRAN (R 4.4.1)
 stringi                1.8.7      2025-03-27 [1] CRAN (R 4.4.1)
 stringr                1.6.0      2025-11-04 [1] CRAN (R 4.4.1)
 SummarizedExperiment * 1.36.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 survival               3.8-3      2024-12-17 [1] CRAN (R 4.4.3)
 tensor                 1.5.1      2025-06-17 [1] CRAN (R 4.4.1)
 tibble                 3.3.0      2025-06-08 [1] CRAN (R 4.4.1)
 tidyr                  1.3.1      2024-01-24 [1] CRAN (R 4.4.0)
 tidyselect             1.2.1      2024-03-11 [1] CRAN (R 4.4.0)
 UCSC.utils             1.2.0      2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 uwot                   0.2.4      2025-11-10 [1] CRAN (R 4.4.1)
 vctrs                  0.7.1      2026-01-23 [1] CRAN (R 4.4.3)
 vipor                  0.4.7      2023-12-18 [1] CRAN (R 4.4.0)
 viridisLite            0.4.2      2023-05-02 [1] CRAN (R 4.4.0)
 withr                  3.0.2      2024-10-28 [1] CRAN (R 4.4.1)
 xfun                   0.56       2026-01-18 [1] CRAN (R 4.4.3)
 xml2                   1.5.0      2025-11-17 [1] CRAN (R 4.4.3)
 xtable                 1.8-4      2019-04-21 [1] CRAN (R 4.4.0)
 XVector                0.46.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 yaml                   2.3.12     2025-12-10 [1] CRAN (R 4.4.3)
 zlibbioc               1.52.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 zoo                    1.8-14     2025-04-10 [1] CRAN (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library
 * ── Packages attached to the search path.

────────────────────────────────────────────────────────────────────────────────────