---
title: "04-LCsmooth_knnstep1_k20_step2_k75_thresholdsViz"
author: "Bernie Mulvey"
date: "03-05-2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(ggrastr)
library(SpatialExperiment)
library(escheR)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallel)


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

first let's make some plots of the relationship between visium registration scores and manual LC label inclusion
```{r}
regs <- fread("processed-data/05_LabelTransfer/04c-NMF_to_KNNksmoothedLCs_regist_to_manXenLC_visLC.txt")

regsplt1 <- regs[Ntopdommks %in% c(10,40,70,100,130,160,190,220,250,280,300)]
regsplt1 <- melt.data.table(regsplt1,id.vars = c("domainDef","Ntopdommks","looseInDom","strictInDom","nmfpredInDom","k1","k2"))

pdf("plots/05_Visdomaintransfer/04a_LFFvis_regist_andStrictManCellprop_top300mks_k2facet.pdf",height=18,width=5.5)
ggplot(regsplt1[Ntopdommks==300&variable=="LFF_Visium"],aes(x=strictInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=4)+
  ylab("Registration to Visium LC")
dev.off()

## proportion of label-xfered LC cells in domain and visium LC registration

pdf("plots/05_Visdomaintransfer/04a_VisLC_regist_and_NMFxferLCcellProp_top300mks_k2facet.pdf",height=12,width=5.5)
ggplot(regsplt1[Ntopdommks==300&variable=="LFF_Visium"],aes(x=nmfpredInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=4)+
  ylab("Registration to Visium LC")
dev.off()

###### registrations with diff number markers for each cellprop #######

pdf("plots/05_Visdomaintransfer/04b_XenManStrict_regist_andStrictManCellprop_top300mks_k2facet.pdf",height=18,width=5.5)
ggplot(regsplt1[Ntopdommks==300&variable=="LFF_Xenium_strict"],aes(x=strictInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=4)+
  ylab("Registration to Strict-Definition Xenium Manual LC Cells")
dev.off()


pdf("plots/05_Visdomaintransfer/04b_XenLoose_regist_andLooseManCellprop_top300mks_k2facet.pdf",height=18,width=5.5)
ggplot(regsplt1[Ntopdommks==300&variable=="LFF_Xenium_loose"],aes(x=looseInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=4)+
  ylab("Registration to Loose-Definition Xenium Manual LC Cells")
dev.off()

# 

pdf("plots/05_Visdomaintransfer/04b_VisLC_regist_and_NMFxferLCcellProp_top300mks_k2facet.pdf",height=18,width=5.5)
ggplot(regsplt1[Ntopdommks==300&variable=="LFF_Visium"],aes(x=nmfpredInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=4)+
  ylab("Registration to Visium LC")
dev.off()


## how much better does each  domain definition register to Visium compared to strictly defined LC cells alone?
pdf("plots/05_Visdomaintransfer/04c_LFFvis-LFFXenstrict_registdiff_top300mks_k1facet.pdf",height=12,width=5.5)
ggplot(regs[Ntopdommks==300],aes(y=(LFF_Visium-LFF_Xenium_strict),x=k2))+
  geom_point()+
  facet_wrap(~k1)
dev.off()

pdf("plots/05_Visdomaintransfer/04c_LFFvis-LFFXenstrict_registdiff_top50mks_k1facet.pdf",height=12,width=5.5)
ggplot(regs[Ntopdommks==50],aes(y=(LFF_Visium-LFF_Xenium_strict),x=k2))+
  geom_point()+
  facet_wrap(~k1)
dev.off()

pdf("plots/05_Visdomaintransfer/04c_LFFvis-LFFXenstrict_registdiff_top20mks_k1facet.pdf",height=12,width=5.5)
ggplot(regs[Ntopdommks==20],aes(y=(LFF_Visium-LFF_Xenium_strict),x=k2))+
  geom_point()+
  facet_wrap(~k1)
dev.off()

### and loose?
pdf("plots/05_Visdomaintransfer/04d_LFFvis-LFFXenloose_registdiff_top300mks_k1facet.pdf",height=12,width=5.5)
ggplot(regs[Ntopdommks==300],aes(y=(LFF_Visium-LFF_Xenium_loose),x=k2))+
  geom_point()+
  facet_wrap(~k1)
dev.off()

pdf("plots/05_Visdomaintransfer/04d_LFFvis-LFFXenloose_registdiff_top50mks_k1facet.pdf",height=12,width=5.5)
ggplot(regs[Ntopdommks==50],aes(y=(LFF_Visium-LFF_Xenium_loose),x=k2))+
  geom_point()+
  facet_wrap(~k1)
dev.off()

pdf("plots/05_Visdomaintransfer/04d_LFFvis-LFFXenloose_registdiff_top20mks_k1facet.pdf",height=12,width=5.5)
ggplot(regs[Ntopdommks==20],aes(y=(LFF_Visium-LFF_Xenium_loose),x=k2))+
  geom_point()+
  facet_wrap(~k1)
dev.off()
```

## get KNN smoothing and transferred labels
```{r}
dualscores <- readRDS("processed-data/05_LabelTransfer/03-LCknnsmoothing_step2k75.RDS")

## we want to label individual "LC" cells that fall outside the domain to assess their boundaries, so retrieve the cellwise visium-nmf-predicted-labels. 
xferlab <- readRDS("processed-data/05_LabelTransfer/02-VistoXen_NMFlabelxfer_tol1e8_k63_alph0.RDS")
visxfer <- lapply(xferlab$targets,function(x){
    y <- as.data.table(cbind(rownames(colData(x)),as.character(x$nmf_preds)))
    setnames(y,c("rn","nmfpred"))
    return(y)
})
visxfer <- rbindlist(visxfer)
rm(xferlab)
gc(full=T)

## now attach
dualscores <- lapply(dualscores,function(x){
  y <- merge.data.table(x,visxfer,by="rn")
  return(y)
})


spe <- readRDS("processed-data/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

logcounts(spe) <- NULL #dont need it
```

```{r}
pal <- readRDS("../LFF_spatial_LC/plots/finalpal_somuchfornorainbows.RDS")

names(pal)
pal <- pal[c(1,6,4)]
names(pal)[2] <- "nonDom_NE"
names(pal)[3] <- "Other"
```

## run through the combos that seem promising on both visium registration and LC cell retention
# plots with one donor's xenium sample and all visium samples in a row for each threshold
```{r}
lcv <- readRDS("../LFF_spatial_LC/processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

vdom <- readRDS("../LFF_spatial_LC/processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
vdom <- vdom$HARMONYlmbna_HDG_SVG_2575
vdom <- vdom[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(vdom,2,"clusid")

vanno <- fread("../LFF_spatial_LC/processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")
vdom <- merge.data.table(vdom,vanno[,.(clusid,anno)],by="clusid")
vdom[,anno:=ifelse(anno=="LC","LC","Other")]
vdom <- DataFrame(vdom,row.names=vdom$rn)[colnames(lcv),]

colLabels(lcv) <- vdom$anno

rm(vdom,vanno)

lcv$brnum[lcv$brnum=="Br6119(re-dis)"] <- "Br6119"
## drop donors with limited LC
lcv <- lcv[,!(lcv$brnum %in% paste0("Br",c(5517,5276,5712)))]

## we also want to match left/right sections here so change brnum to reflect the hemi of origin
lcv$brnum <- paste0(lcv$brnum,lcv$hemisphere)
lcv$brnum <- gsub(lcv$brnum,pattern="(Br.*)R",replacement="\\1")
sum(unique(lcv$brnum) %in% unique(spe$brnum))==length(unique(spe$brnum)) # k we good
```

### one plot file per threshold set and incorporating visium samples

## for selected k1/k2 pairs based on plots 04a-f. initial sweep leads me to these (k1,k2):
.05 / 0.04 thru 0.6 for completeness
.1 / 0.04-0.44
.15 / 0.04-0.28

```{r}
k1thresh <- round(seq(0.05,0.15,0.05),2)
setDTthreads(1,restore_after_fork = FALSE)
lapply(k1thresh,function(k1){
  if (k1==0.05){
    k2s <- round(seq(0.04,0.6,0.08),2)
  } else if (k1==0.1){
    k2s <- round(seq(0.04,0.44,0.08),2)
  } else if (k1==0.15){
    k2s <- round(seq(0.04,0.28,0.08),2)
  } 

  n <- paste0("k1thresh_",k1)
  d <- copy(dualscores[[n]]) # single brackets don't return a d.t. i guess, but double brackets do

  lapply(k2s,function(k2){
    d <- as.data.table(d)
    d[,isLC:=ifelse(lcpred.smooth1.k20>=k1 & lcpred.smooth2.k75>=k2,"LC","Other")]
    ## mark cells outside the defined domain that nmf labeled LC.
    d[nmfpred=="nonDom_NE",isLC:=ifelse(isLC=="LC","LC","nonDom_NE")]
    d2 <- DataFrame(d,row.names=d$rn)[colnames(spe),]
    spe$pltclus <- d2$isLC
    
    plts <- list()
    j<-1
    for (j in c(1:length(unique(spe$brnum)))){
      s <- unique(spe$brnum)[j]
      sublist <- list()
      pltclus.size <- spe[,spe$brnum==s]$pltclus
      px <- make_escheR(spe[,spe$brnum==s], y_reverse=FALSE)
      px <- rasterize(add_fill(p = px,var="pltclus",point_size=0.75),layers="Points",dpi=300,dev="cairo_png")
      px <- px+scale_fill_manual(values = pal)
      
      # since escheR is making the size aesthetic inaccessible, override it:
      px@layers[[1]]$aes_params$size <- NULL
      ## if we specify a conditional to aes, we can then just set a true/false value for the scale values (e.g. below, size is the result of evaluating the logical)
      px <- px+aes(size=pltclus=="nonDom_NE")+
      scale_size_manual(values = c("TRUE" = 2.25, "FALSE" = 0.75))
      ## carry on
      
      px <- px+guides(fill="none",color="none",size="none")
      px <- px+ggtitle(s)

      sublist[[1]] <- px

      i<-1
      for (i in c(1:length(unique(lcv[,lcv$brnum==s]$sample_id)))){
        curs <- unique(lcv[,lcv$brnum==s]$sample_id)[i]
        curd <- lcv[,lcv$brnum==s & lcv$sample_id==curs]
        px <- make_escheR(curd, y_reverse=FALSE)
        px <- rasterize(add_fill(p=px,var="label",size=1.25,point_size=1.25),layers="Points",dpi=300,dev="cairo_png")
        px <- px+scale_fill_manual(
          values = pal)+guides(color="none",fill="none")
        px <- px+ggtitle(curs)
        sublist[[(i+1)]] <- px
      }
      retobj <- gridExtra::arrangeGrob(grobs = sublist, ncol = length(sublist))
      plts[[j]] <- retobj
    }
    # flat_plts <- do.call(c, plts)
    # args_list <- c(flat_plts, list(nrow = length(plts)))
    png(paste0("plots/05_Visdomaintransfer/04e-NMFtransferLC_smoothedDom_knn20_k1_",k1,"_knn75_k2_",k2,"_VisCompare.png"),height=4*length(plts),width=14,units = "in",res=150)
    grid.arrange(grobs=plts,ncol=1)
    dev.off()
  })
})

```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0         parallelly_1.46.1          
 [3] colorout_1.3-3              escheR_1.10.0              
 [5] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
 [7] SummarizedExperiment_1.40.0 Biobase_2.70.0             
 [9] GenomicRanges_1.62.1        Seqinfo_1.0.0              
[11] IRanges_2.44.0              S4Vectors_0.48.0           
[13] BiocGenerics_0.56.0         generics_0.1.4             
[15] MatrixGenerics_1.22.0       matrixStats_1.5.0          
[17] ggrastr_1.0.2               gridExtra_2.3              
[19] ggtext_0.1.2                ggplot2_4.0.1              
[21] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] gtable_0.3.6        beeswarm_0.4.0      rjson_0.2.23        xfun_0.56          
 [5] lattice_0.22-7      vctrs_0.7.1         tools_4.5.2         tibble_3.3.1       
 [9] pkgconfig_2.0.3     Matrix_1.7-4        RColorBrewer_1.1-3  S7_0.2.1           
[13] lifecycle_1.0.5     compiler_4.5.2      farver_2.1.2        codetools_0.2-20   
[17] vipor_0.4.7         htmltools_0.5.9     yaml_2.3.12         pillar_1.11.1      
[21] rsconnect_1.7.0     DelayedArray_0.36.0 sessioninfo_1.2.3   magick_2.9.0       
[25] abind_1.4-8         tidyselect_1.2.1    digest_0.6.39       dplyr_1.1.4        
[29] rprojroot_2.1.1     fastmap_1.2.0       grid_4.5.2          here_1.0.2         
[33] cli_3.6.5           SparseArray_1.10.8  magrittr_2.0.4      S4Arrays_1.10.1    
[37] dichromat_2.0-0.1   withr_3.0.2         scales_1.4.0        ggbeeswarm_0.7.3   
[41] rmarkdown_2.30      XVector_0.50.0      otel_0.2.0          evaluate_1.0.5     
[45] knitr_1.51          rlang_1.1.7         gridtext_0.1.5      Rcpp_1.1.1         
[49] glue_1.8.0          xml2_1.5.2          rstudioapi_0.18.0   R6_2.6.1           
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-02
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ─────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.1.4    2023-11-17 [1] CRAN (R 4.5.0)
 escheR               * 1.10.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggbeeswarm             0.7.3    2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.1    2025-11-14 [1] CRAN (R 4.5.2)
 ggrastr              * 1.0.2    2023-06-01 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] CRAN (R 4.5.2)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.5.0)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

────────────────────────────────────────────────────────────────────────────────────
