---
title: "04-ONEcond_KNNsmooth_finestep_regist.Rmd"
format: html
date: "06-05-2026"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
# library(SpatialFeatureExperiment)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}


## rstudio GUI tweaks
require(colorout)
ColorOut()
options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
library(parallel)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = '#000000'), axis.title.x = element_text(size = 9,color = '#000000'), axis.text.y = element_text(size = 8,color = '#000000'), axis.title.y = element_text(size =9,color = '#000000'), plot.title = element_markdown(size = 10,hjust=0.5,color = '#000000'), strip.text = element_text(size=11), legend.text = element_text(size=7,color = '#000000'), legend.title = element_text(size=8,hjust=0.5,color = '#000000'),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position='plot'))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

load smoothings, xenium, and manual xenium LC assignments
```{r}
## smoothings
dualscores <- readRDS("processed-data_newseg/07_visLabelTransfer/02b-srtXferLCknnsmoothing_step2k75.RDS")
## only testing up to k1 0.2
dualscores <- dualscores[1:5]

### lc xenium
lcx <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")
assays(lcx) <- assays(lcx)[1] #dont need normed counts or log counts

### xenium manual LC definition
manlc <- fread("processed-data_newseg/03_ManualLC/01-XprBasedLClabels_and_ManLCcellsWithSegsGainedorAlteredByNM.txt")
manlc <- DataFrame(manlc,row.names=manlc$rn)[colnames(lcx),]
lcx$LCloose <- manlc$LCloose
lcx$LCstrict <- manlc$LCstrict

### nmf-transferred labels
xferlab <- readRDS("processed-data_newseg/07_visLabelTransfer/01-VistoXen_NMFlabelxfer_tol1e8_k63_alph0.RDS")

preds <- lapply(xferlab$targets,function(x){
    y <- as.data.table(cbind(rownames(colData(x)),as.character(x$nmf_preds)))
    setnames(y,c("rn","nmfpred"))
    return(y)
})
preds <- rbindlist(preds)
rm(xferlab)

### keep manLC for tabulating % overlap with domain and cells later
manlc <- as.data.table(manlc)
```

### here, the cell's K1 score doesn't matter, only the k2 score. there is still a series of k1 thresholds that was used to define "inLC" for k2 score calcs, we just aren't filtering on what a cell's own k1 score came out to.

now we have some k1 smoothings / k2 thresholds that we like:
0.08-0.36-0.52 (expanded range for fine stepping: 0.281-0.52)
0.12-0.28 (expanded range for fine stepping: 0.12-0.28)
0.16-originals were too meh visually, try fine stepping 0.04-0.28
0.2-" " " " try fine stepping 0.04-0.2

## make a table of these; steps are 1/75 (size of k2 neighborhood) - 0.001 to ensure inclusion/exclusion of previously examined k1s of interest/noninterest ##
```{r}
refineranges <- list(
  "0.08"=round(signif(seq(0.28,0.52,by=(1/75))),3),
  "0.12"=round(signif(seq(0.2,0.28,by=(1/75))),3),
  "0.16"=round(signif(seq(0.04,0.281,by=(1/75))),3),
  "0.2"=round(signif(seq(0.04,0.200,by=(1/75))),3))
refineranges<-lapply(refineranges,as.data.table)
refineranges <- rbindlist(refineranges,idcol="k1")
setnames(refineranges,2,"k2thresh")

# then lop of 0.001 to ensure fractional thirds aren't problematic (.267 > 20/75, but we want to test k2thresh of ≥20/75:
refineranges[,k2thresh:=k2thresh-0.001]
```

### we already checked these score ranges for having at least 20 cells in at least 16 donors.
```{r}
## avoid floating point bullshit
refineranges[,k1:=round(as.numeric(k1),2)]

setDTthreads(1,restore_after_fork = FALSE)

regperthresh <- lapply(c(1:nrow(refineranges)),function(i){
  curk1 <- refineranges[i,k1]
  k2thresh <- refineranges[i,k2thresh] - 0.001
  ## subtract each k2 threshold by 0.001 to account for fractional thirds (so that k2s in the result == to e.g. 20/75 (.26666667) are included. they would be <0.267 but > 0.266, while 19/75 would not be.)
  tmpdt <- copy(dualscores[[paste0("k1thresh","_",curk1)]])
  ## cell must meet k2 threshold (thus "1-condition" smoothing)
  tmpdt[,newcol:=ifelse(
      # greaterOrEqual(lcpred.smooth1.k25,k1) &
    greaterOrEqual(lcpred.smooth2.k75,k2thresh),
    "LC","Other")
  ]
    
  if(length(unique(tmpdt$newcol))<2){
    return("NA")
  } else {
    threshname <- paste0("LCsmooth_k1_",curk1,"_k2_",k2thresh)
      setnames(tmpdt,"newcol",threshname)
      tmpdt <- DataFrame(tmpdt[,.(rn,get(threshname))],row.names=tmpdt$rn)[colnames(lcx),]
      colData(lcx)$newcol <- tmpdt[,2] 
      colnames(colData(lcx))[ncol(colData(lcx))] <- threshname

      ## registration stats for this threshold (we then cross correlate those to consistent regist stats from visium and manual xen)

    regstats <- registration_wrapper(sce=lcx,var_registration = threshname,var_sample_id = "brnum",covars = c("Age","Sex","APOE","YRI"),gene_ensembl = "ID",gene_name = "Symbol",min_ncells = 20)
    keepname <- c("ensembl",paste0("t_stat_",c("LC","Other")))
    outtab <- as.data.table(regstats$enrichment)[,..keepname]
    return(outtab)
  }
})
names(regperthresh) <- paste0("LCsmooth_k1_",refineranges$k1,"_k2_",refineranges$k2thresh)

saveRDS(regperthresh,file="processed-data_newseg/07_visLabelTransfer/04a-ONEcond_srtLCsmooth_knnstep1_k25_step2_k75_registPerThresh_finesteps.RDS")
```

## calculate registration stats for xenium manuals and visium domains, then can remove those datasets from sesh
```{r}
lcxLoreg <- registration_wrapper(sce=lcx,var_registration="LCloose",var_sample_id="brnum",covars=c("Age","Sex","APOE","YRI"),gene_ensembl="ID",gene_name="Symbol",min_ncells=20)

lcxLoreg <- as.data.table(lcxLoreg$enrichment)
lcxLoreg <- lcxLoreg[,.(ensembl,t_stat_LC,t_stat_No)]
comparegs <- list()
comparegs[[length(comparegs)+1]] <- lcxLoreg
names(comparegs)[length(comparegs)] <- "LFF_Xenium_loose"

###
lcxHireg <- registration_wrapper(sce=lcx,var_registration="LCstrict",var_sample_id="brnum",covars=c("Age","Sex","APOE","YRI"),gene_ensembl="ID",gene_name="Symbol",min_ncells=20)

lcxHireg <- as.data.table(lcxHireg$enrichment)
lcxHireg <- lcxHireg[,.(ensembl,t_stat_LC,t_stat_No)]
comparegs[[length(comparegs)+1]] <- lcxHireg
names(comparegs)[length(comparegs)] <- "LFF_Xenium_strict"

###
predstmp <- copy(preds)
predstmp[nmfpred!="LC",nmfpred:="No"]
predstmp <- DataFrame(predstmp,row.names=predstmp$rn)[colnames(lcx),]
colData(lcx)$nmfpred <- predstmp$nmfpred
rm(predstmp)
gc()

lcxnmfp <- registration_wrapper(sce=lcx,var_registration="nmfpred",var_sample_id="brnum",covars=c("Age","Sex","APOE","YRI"),gene_ensembl="ID",gene_name="Symbol",min_ncells=20)

lcxnmfp <- as.data.table(lcxnmfp$enrichment)
lcxnmfp <- lcxnmfp[,.(ensembl,t_stat_LC,t_stat_No)]
comparegs[[length(comparegs)+1]] <- lcxnmfp
names(comparegs)[length(comparegs)] <- "Xenium_All_NMFpred_LCcells"

### 
rm(lcxHireg,lcxLoreg,lcxnmfp)
gc(full=T)
```

## visium data and annots
```{r}
lcv <- readRDS("../LFF_spatial_LC/processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

clus <- readRDS("../LFF_spatial_LC/processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
clus <- clus$HARMONYlmbna_HDG_SVG_2575[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(clus,2,"clusid")

anno <- fread("../LFF_spatial_LC/processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")
# this will just be LC/other for registration purposes
anno[,anno:=ifelse(anno=="LC","LC","Other")]

anno <- merge.data.table(clus,anno,by="clusid",all.x=T)
anno <- DataFrame(anno,row.names=anno$rn)[colnames(lcv),]
lcv$vislab <- anno$anno

## subset Visium to samples we want to retain in both assays
unique(lcv$brnum)
lcv$brnum[lcv$brnum=="Br6119(re-dis)"] <- "Br6119"

## drop donors with limited LC
lcv <- lcv[,!(lcv$brnum %in% paste0("Br",c(5517,5276,5712)))]

## check all donors overlap
lcx$donor <- gsub(lcx$brnum,pattern="L",replacement="")
stopifnot(all(unique(lcv$brnum) %in% unique(lcx$donor)))
stopifnot(all(unique(lcx$donor) %in% unique(lcv$brnum)))
### 
## subset LC visium to genes in xenium
tmprd <- as.data.table(rowData(lcv),keep.rownames=T)
tmprd <- tmprd[gene_id %in% rowData(lcx)$ID]
keepg <- tmprd$rn

lcv <- lcv[keepg,]

rm(tmprd,lcx,anno,clus)
gc(full=T)
```

ancestry proportion estimates aren't in this visium object, so add them for identical registration modeling
```{r}
flares <- fread("../LFF_spatial_LC/raw-data/ancestryAndHaplotypes/FLARE/global_ancestry_estimates.txt")
flares <- flares[ID %in% lcv$brnum]
stopifnot(all(lcv$brnum %in% flares$ID))

tmpcd <- as.data.table(colData(lcv),keep.rownames=T)
tmpcd <- merge.data.table(tmpcd,flares[,.(ID,YRI)],by.x="brnum",by.y="ID",all.x=T)
stopifnot(all(!is.na(tmpcd$YRI)))

# Also fix APOE genotypes to avoid modeling errors
tmpcd[APOE=="E4/ E4",APOE:="E4hom"]
tmpcd[APOE=="E3/ E4",APOE:="E4het"]
tmpcd[APOE=="E2/ E3",APOE:="E2het"]
tmpcd[APOE=="E2/ E2",APOE:="E2hom"]

## replace coldata
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(lcv),]
tmpcd$rn <- NULL
colData(lcv) <- tmpcd

stopifnot(is.numeric(lcv$Age)&is.numeric(lcv$YRI)&is.character(lcv$APOE)&is.character(lcv$Sex))

rm(tmpcd,flares)
```


### vis lc regist
```{r}
lcvreg <- registration_wrapper(sce=lcv,var_registration = "vislab",var_sample_id = "sample_id",covars = c("Age","Sex","APOE","YRI"),gene_ensembl = "gene_id",gene_name = "gene_name",min_ncells = 10) # one tissue section has a few dozen LC spots at most

lcvreg <- as.data.table(lcvreg$enrichment)
lcvreg <- lcvreg[,.(ensembl,t_stat_LC,t_stat_Other)]
comparegs[[length(comparegs)+1]] <- lcvreg
names(comparegs)[length(comparegs)] <- "LFF_Visium"

rm(lcv,lcvreg)
gc(full=T)
```

### also get proportion of cells with LC label in each xenium manual annot that are labeled as LC domain
```{r}
propsindoms <- mcmapply(d=dualscores,n=names(dualscores),SIMPLIFY=FALSE,mc.cores=5,FUN=function(d,n){
    curk1 <- as.numeric(gsub(n,pattern="k1thresh_(.*)",replacement="\\1"))
    tmp <- merge.data.table(copy(d),manlc,by="rn")
    tmp <- merge.data.table(tmp,preds,by="rn")
    k2s <- refineranges[k1==curk1,k2thresh] ## == works because we rounded
    regs <- lapply(k2s,function(k2){
        setDTthreads(1,restore_after_fork = FALSE)
        
        tmpdt <- copy(tmp)
        tmpdt[,newcol:=ifelse(
          #greaterOrEqual(lcpred.smooth1.k25,k1) &
            greaterOrEqual(lcpred.smooth2.k75,k2),
            "LC","Other")
        ]
        
        looseprop <- ifelse(nrow(tmpdt[LCloose=="LC"])==0,
                            0,
                            nrow(tmpdt[newcol=="LC"&LCloose=="LC"])/
                              nrow(tmpdt[LCloose=="LC"])
        )
        
        strictprop <- ifelse(nrow(tmpdt[LCstrict=="LC"])==0,
                            0,
                            nrow(tmpdt[newcol=="LC"&LCstrict=="LC"])/
                              nrow(tmpdt[LCstrict=="LC"])
        )
        
        nmfprop <- ifelse(nrow(tmpdt[nmfpred=="LC"])==0,
                            0,
                            nrow(tmpdt[newcol=="LC"&nmfpred=="LC"])/
                              nrow(tmpdt[nmfpred=="LC"])
        )

        od <- data.frame(matrix(nrow=1,ncol=5))
        od[1,1] <- curk1
        od[1,2] <- k2
        od[1,3] <- looseprop
        od[1,4] <- strictprop
        od[1,5] <- nmfprop
        od <- as.data.table(od)
        setnames(od,c("k1","k2","looseInDom","strictInDom","nmfpredInDom"))

        return(od)
    })
    ordt <- rbindlist(regs)
    return(ordt)
})
propsindoms <- rbindlist(propsindoms)


fwrite(propsindoms,"processed-data_newseg/07_visLabelTransfer/04b-ONEcond_prop_manualLCcells_inEach_SRTpredSmoothDom_finek2steps.txt",sep='\t',quote=F,row.names=F)

```

Run registration for each xenium domain definition against each xenium manual annotation and lc vis label
```{r}
### rejigger the regperthresh thing into one big table
rpt <- mapply(X=regperthresh,Y=names(regperthresh),SIMPLIFY=FALSE,function(X,Y){
    k1 <- gsub(Y,pattern="LCsmooth_k1_(.*)_k2.*$",replacement="\\1")
    k2 <- gsub(Y,pattern="LCsmooth_k1_.*_k2_(.*)$",replacement="\\1")
    newdat <- copy(X[,.(ensembl,t_stat_LC)])
    setnames(newdat,2,paste0("LCk1_",k1,"_k2_",k2))
    return(newdat)
})

for (i in c(1:length(rpt))){
        if(i==1){
            regperthresh.dt <- copy(rpt[[i]])
        } else{
            regperthresh.dt <- merge.data.table(regperthresh.dt,rpt[[i]],by="ensembl",all.x=T)
        }
}
rm(rpt,i)
```

Now, actually iterate through registrations
```{r}
regres <- mcmapply(m=comparegs,n=names(comparegs),mc.cores=4,SIMPLIFY=FALSE,FUN=function(m,n){
    m <- m[,.(ensembl,t_stat_LC)]
    setnames(m,"t_stat_LC",n)
    
    cortab <- merge.data.table(regperthresh.dt,m,by="ensembl")
    cortab$ensembl <- NULL

    j<-seq(10,300,by=10)
    i<-1
    
    outlist <- list()
    for (i in c(1:(length(j)+1))){
        cormat <- as.data.frame(matrix(nrow=ncol(cortab)-1,ncol=1))
        
        if(i<=length(j)){
            ngene <- j[i]
        } else{
            # no value for top_n --> all genes used
            ngene <- nrow(cortab)
        }
        

        k <- 1
        for (k in c(1:(ncol(cortab)-1))){
            rownames(cormat)[k] <- colnames(cortab)[k]
            colnames(cormat)[1] <- colnames(cortab)[ncol(cortab)]
            
            curthresh <- colnames(cortab)[k]
            curxen <- names(m)[2]
            setorderv(cortab,curthresh,-1)
            
            cormat[k,1] <- cor(cortab[c(1:ngene),..curthresh],cortab[c(1:ngene),..curxen],use="pairwise.complete.obs",method = "spearman")
        }
        outlist[[i]] <- cormat
        names(outlist)[i] <- paste0("DomainTop_",ngene)
    }
    keepol <- unique(
      names(outlist)[
        names(outlist) %in%
          paste0("DomainTop_",
                 unique(c(seq(10,nrow(cortab),10),nrow(cortab)))
                )
      ]
    )
    outlist <- outlist[keepol] # account for repeat correlations for ngene > nrow(cortab)
    return(outlist)
})

### concat into one big table
regres2 <- lapply(regres,function(x){
    z <- mapply(y=x,n=names(x),SIMPLIFY=FALSE,FUN=function(y,n){
        y <- as.data.table(y,keep.rownames=T)
        y[,Ntopdommks:=as.numeric(gsub(n,pattern="^DomainTop_(.*)$",replacement="\\1"))]
        setnames(y,"rn","domainDef")
        return(y)
    })
    rettab <- rbindlist(z)
    return(rettab)
})
regres3 <- merge.data.table(regres2[[1]],regres2[[2]],by=c("domainDef","Ntopdommks"))
regres4 <- merge.data.table(regres3,regres2[[3]],by=c("domainDef","Ntopdommks"))
regres5 <- merge.data.table(regres4,regres2[[4]],by=c("domainDef","Ntopdommks"),all.x=T)
### append proportion in domain annots for xeniums. add back 0.001 for matching
propsindoms[,domainDef:=paste0("LC_k1_",k1,"_k2_",k2)]

# something weird happened with the loops above so just fix it here
regres5[,domainDef:=gsub(domainDef,pattern="LCk1",replacement="LC_k1")]

regres5 <- merge.data.table(regres5,propsindoms,by="domainDef")

fwrite(regres5,"processed-data_newseg/07_visLabelTransfer/04c-ONEcond_NMF_to_KNNsmoothed_NMFsrtLCs_regist_to_manXenLC_visLC_finek2steps.txt",quote=F,row.names=F,sep='\t')
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods  
[8] base     

other attached packages:
 [1] parallelly_1.47.0           colorout_1.3-3             
 [3] spatialLIBD_1.22.0          SpatialExperiment_1.20.0   
 [5] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
 [7] Biobase_2.70.0              GenomicRanges_1.62.1       
 [9] Seqinfo_1.0.0               IRanges_2.44.0             
[11] S4Vectors_0.48.1            BiocGenerics_0.56.0        
[13] generics_0.1.4              MatrixGenerics_1.22.0      
[15] matrixStats_1.5.0           gridExtra_2.3              
[17] ggtext_0.1.2                ggplot2_4.0.2              
[19] data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3       rstudioapi_0.18.0       
  [3] jsonlite_2.0.0           shape_1.4.6.1           
  [5] magrittr_2.0.4           ggbeeswarm_0.7.3        
  [7] magick_2.9.0             farver_2.1.2            
  [9] rmarkdown_2.30           BiocIO_1.20.0           
 [11] GlobalOptions_0.1.3      vctrs_0.7.1             
 [13] Rsamtools_2.26.0         memoise_2.0.1           
 [15] config_0.3.2             RCurl_1.98-1.17         
 [17] paletteer_1.7.0          benchmarkme_1.0.8       
 [19] htmltools_0.5.9          S4Arrays_1.10.1         
 [21] AnnotationHub_4.0.0      curl_7.0.0              
 [23] BiocNeighbors_2.4.0      SparseArray_1.10.8      
 [25] sass_0.4.10              bslib_0.10.0            
 [27] htmlwidgets_1.6.4        httr2_1.2.2             
 [29] plotly_4.12.0            cachem_1.1.0            
 [31] GenomicAlignments_1.46.0 mime_0.13               
 [33] lifecycle_1.0.5          iterators_1.0.14        
 [35] pkgconfig_2.0.3          rsvd_1.0.5              
 [37] Matrix_1.7-4             R6_2.6.1                
 [39] fastmap_1.2.0            shiny_1.12.1            
 [41] clue_0.3-66              digest_0.6.39           
 [43] colorspace_2.1-2         rematch2_2.1.2          
 [45] AnnotationDbi_1.72.0     scater_1.38.1           
 [47] rprojroot_2.1.1          irlba_2.3.7             
 [49] ExperimentHub_3.0.0      RSQLite_2.4.5           
 [51] beachmat_2.26.0          filelock_1.0.3          
 [53] httr_1.4.7               abind_1.4-8             
 [55] compiler_4.5.2           here_1.0.2              
 [57] bit64_4.6.0-1            withr_3.0.2             
 [59] doParallel_1.0.17        attempt_0.3.1           
 [61] S7_0.2.1                 BiocParallel_1.44.0     
 [63] viridis_0.6.5            DBI_1.2.3               
 [65] sessioninfo_1.2.3        rappdirs_0.3.4          
 [67] DelayedArray_0.36.1      rjson_0.2.23            
 [69] tools_4.5.2              vipor_0.4.7             
 [71] otel_0.2.0               beeswarm_0.4.0          
 [73] httpuv_1.6.16            glue_1.8.0              
 [75] restfulr_0.0.16          promises_1.5.0          
 [77] gridtext_0.1.5           grid_4.5.2              
 [79] rsconnect_1.7.0          cluster_2.1.8.1         
 [81] gtable_0.3.6             tidyr_1.3.2             
 [83] ScaledMatrix_1.18.0      BiocSingular_1.26.0     
 [85] xml2_1.5.2               XVector_0.50.0          
 [87] ggrepel_0.9.6            BiocVersion_3.22.0      
 [89] foreach_1.5.2            pillar_1.11.1           
 [91] limma_3.66.0             later_1.4.5             
 [93] circlize_0.4.17          benchmarkmeData_2.0.0   
 [95] dplyr_1.2.0              BiocFileCache_3.0.0     
 [97] lattice_0.22-7           rtracklayer_1.70.1      
 [99] bit_4.6.0                tidyselect_1.2.1        
[101] ComplexHeatmap_2.26.1    locfit_1.5-9.12         
[103] scuttle_1.20.0           Biostrings_2.78.0       
[105] knitr_1.51               edgeR_4.8.2             
[107] xfun_0.56                statmod_1.5.1           
[109] DT_0.34.0                lazyeval_0.2.2          
[111] yaml_2.3.12              shinyWidgets_0.9.0      
[113] cigarillo_1.0.0          evaluate_1.0.5          
[115] codetools_0.2-20         tibble_3.3.1            
[117] BiocManager_1.30.27      cli_3.6.5               
[119] xtable_1.8-4             jquerylib_0.1.4         
[121] golem_0.5.1              dichromat_2.0-0.1       
[123] Rcpp_1.1.1               dbplyr_2.5.1            
[125] png_0.1-8                XML_3.99-0.20           
[127] blob_1.3.0               bitops_1.0-9            
[129] viridisLite_0.4.3        scales_1.4.0            
[131] purrr_1.2.1              crayon_1.5.3            
[133] GetoptLong_1.1.0         rlang_1.1.7             
[135] cowplot_1.2.0            KEGGREST_1.50.0         
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-05
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────
 package              * version   date (UTC) lib source
 abind                  1.4-8     2024-09-12 [1] CRAN (R 4.5.0)
 AnnotationDbi          1.72.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationHub          4.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 attempt                0.3.1     2020-05-03 [1] CRAN (R 4.5.0)
 beachmat               2.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.5.0)
 benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.5.0)
 benchmarkmeData        2.0.0     2026-01-19 [1] CRAN (R 4.5.2)
 Biobase              * 2.70.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocFileCache          3.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocIO                 1.20.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocManager            1.30.27   2025-11-14 [1] CRAN (R 4.5.2)
 BiocNeighbors          2.4.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel           1.44.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocVersion            3.22.0    2025-10-07 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0     2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1   2025-01-16 [1] CRAN (R 4.5.0)
 bitops                 1.0-9     2024-10-03 [1] CRAN (R 4.5.0)
 blob                   1.3.0     2026-01-14 [1] CRAN (R 4.5.2)
 bslib                  0.10.0    2026-01-26 [1] CRAN (R 4.5.2)
 cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.5.0)
 cigarillo              1.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 circlize               0.4.17    2025-12-08 [1] CRAN (R 4.5.2)
 cli                    3.6.5     2025-04-23 [1] CRAN (R 4.5.0)
 clue                   0.3-66    2024-11-13 [1] CRAN (R 4.5.0)
 cluster                2.1.8.1   2025-03-12 [1] CRAN (R 4.5.0)
 codetools              0.2-20    2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3     2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 colorspace             2.1-2     2025-09-22 [1] CRAN (R 4.5.0)
 ComplexHeatmap         2.26.1    2026-01-30 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 config                 0.3.2     2023-08-30 [1] CRAN (R 4.5.0)
 cowplot                1.2.0     2025-07-07 [1] CRAN (R 4.5.0)
 crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0     2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1  2026-01-27 [1] CRAN (R 4.5.2)
 DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.5.0)
 dbplyr                 2.5.1     2025-09-10 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.1    2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1   2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39    2025-11-19 [1] CRAN (R 4.5.2)
 doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.5.0)
 dplyr                  1.2.0     2026-02-03 [1] CRAN (R 4.5.2)
 DT                     0.34.0    2025-09-02 [1] CRAN (R 4.5.0)
 edgeR                  4.8.2     2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate               1.0.5     2025-08-27 [1] CRAN (R 4.5.0)
 ExperimentHub          3.0.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 farver                 2.1.2     2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.5.0)
 filelock               1.0.3     2023-12-11 [1] CRAN (R 4.5.0)
 foreach                1.5.2     2022-02-02 [1] CRAN (R 4.5.0)
 generics             * 0.1.4     2025-05-09 [1] CRAN (R 4.5.0)
 GenomicAlignments      1.46.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 GenomicRanges        * 1.62.1    2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 GetoptLong             1.1.0     2025-11-28 [1] CRAN (R 4.5.2)
 ggbeeswarm             0.7.3     2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2     2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6     2024-09-07 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2     2022-09-16 [1] CRAN (R 4.5.0)
 GlobalOptions          0.1.3     2025-11-28 [1] CRAN (R 4.5.2)
 glue                   1.8.0     2024-09-30 [1] CRAN (R 4.5.0)
 golem                  0.5.1     2024-08-27 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5     2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6     2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2     2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9     2025-12-04 [1] CRAN (R 4.5.2)
 htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.5.0)
 httpuv                 1.6.16    2025-04-16 [1] CRAN (R 4.5.0)
 httr                   1.4.7     2023-08-15 [1] CRAN (R 4.5.0)
 httr2                  1.2.2     2025-12-08 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7     2026-01-30 [1] CRAN (R 4.5.2)
 iterators              1.0.14    2022-02-05 [1] CRAN (R 4.5.0)
 jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.5.0)
 jsonlite               2.0.0     2025-03-27 [1] CRAN (R 4.5.0)
 KEGGREST               1.50.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51      2025-12-20 [1] CRAN (R 4.5.2)
 later                  1.4.5     2026-01-08 [1] CRAN (R 4.5.2)
 lattice                0.22-7    2025-04-02 [1] CRAN (R 4.5.2)
 lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.5.0)
 lifecycle              1.0.5     2026-01-08 [1] CRAN (R 4.5.2)
 limma                  3.66.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 locfit                 1.5-9.12  2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0     2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4     2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4     2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0     2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1     2021-11-26 [1] CRAN (R 4.5.0)
 mime                   0.13      2025-03-17 [1] CRAN (R 4.5.0)
 otel                   0.2.0     2025-08-29 [1] CRAN (R 4.5.0)
 paletteer              1.7.0     2026-01-08 [1] CRAN (R 4.5.2)
 parallelly           * 1.47.0    2026-04-17 [1] CRAN (R 4.5.2)
 pillar                 1.11.1    2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.5.0)
 plotly                 4.12.0    2026-01-24 [1] CRAN (R 4.5.2)
 png                    0.1-8     2022-11-29 [1] CRAN (R 4.5.0)
 promises               1.5.0     2025-11-01 [1] CRAN (R 4.5.0)
 purrr                  1.2.1     2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1     2025-02-15 [1] CRAN (R 4.5.0)
 rappdirs               0.3.4     2026-01-17 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1     2026-01-10 [1] CRAN (R 4.5.2)
 RCurl                  1.98-1.17 2025-03-22 [1] CRAN (R 4.5.0)
 rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.5.0)
 restfulr               0.0.16    2025-06-27 [1] CRAN (R 4.5.0)
 rjson                  0.2.23    2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7     2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30      2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1     2025-08-26 [1] CRAN (R 4.5.0)
 Rsamtools              2.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 rsconnect              1.7.0     2025-12-06 [1] CRAN (R 4.5.2)
 RSQLite                2.4.5     2025-11-30 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0    2026-01-16 [1] CRAN (R 4.5.2)
 rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.5.0)
 rtracklayer            1.70.1    2025-12-20 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 S4Arrays               1.10.1    2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1    2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1     2025-11-14 [1] CRAN (R 4.5.2)
 sass                   0.4.10    2025-04-11 [1] CRAN (R 4.5.0)
 ScaledMatrix           1.18.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0     2025-04-24 [1] CRAN (R 4.5.0)
 scater                 1.38.1    2026-03-20 [1] Bioconductor 3.22 (R 4.5.2)
 scuttle                1.20.0    2025-10-30 [1] Bioconductor 3.22 (R 4.5.2)
 Seqinfo              * 1.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3     2025-02-05 [1] CRAN (R 4.5.0)
 shape                  1.4.6.1   2024-02-23 [1] CRAN (R 4.5.0)
 shiny                  1.12.1    2025-12-09 [1] CRAN (R 4.5.2)
 shinyWidgets           0.9.0     2025-02-21 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8    2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 spatialLIBD          * 1.22.0    2025-11-04 [1] Bioconductor 3.22 (R 4.5.2)
 statmod                1.5.1     2025-10-09 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1     2026-01-11 [1] CRAN (R 4.5.2)
 tidyr                  1.3.2     2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1     2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.5.0)
 viridis                0.6.5     2024-01-29 [1] CRAN (R 4.5.0)
 viridisLite            0.4.3     2026-02-04 [1] CRAN (R 4.5.2)
 withr                  3.0.2     2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56      2026-01-18 [1] CRAN (R 4.5.2)
 XML                    3.99-0.20 2025-11-08 [1] CRAN (R 4.5.0)
 xml2                   1.5.2     2026-01-17 [1] CRAN (R 4.5.2)
 xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.5.0)
 XVector                0.50.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12    2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────