---
title: "05-ONEcond_srtXfer_LabelsAndSmoothLCLUT"
format: html
date: "06-06-2026"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
# library(SpatialFeatureExperiment)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}


## rstudio GUI tweaks
require(colorout)
ColorOut()
options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
library(parallel)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)

## ggplot defaults

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

.08/.426
0.12/199
16/119
20 079
load smoothings, xenium, and manual xenium LC assignments
```{r}
## smoothings
dualscores <- readRDS("processed-data_newseg/07_visLabelTransfer/02b-srtXferLCknnsmoothing_step2k75.RDS")
dualscores <- dualscores[c("k1thresh_0.08","k1thresh_0.12","k1thresh_0.16","k1thresh_0.2")]

dualscores8426 <- dualscores$k1thresh_0.08[,domain8426:=ifelse(greaterOrEqual(lcpred.smooth2.k75,round(0.4260,3)),"LC","Other")]

dualscores12199 <- dualscores$k1thresh_0.12[,domain12199:=ifelse(greaterOrEqual(lcpred.smooth2.k75,round(0.1990,3)),"LC","Other")]

dualscores16119 <- dualscores$k1thresh_0.16[,domain16119:=ifelse(greaterOrEqual(lcpred.smooth2.k75,round(0.1190,3)),"LC","Other")]

dualscores20079 <- dualscores$k1thresh_0.2[,domain20079:=ifelse(greaterOrEqual(lcpred.smooth2.k75,round(0.0790,3)),"LC","Other")]

srtdoms <- merge.data.table(dualscores8426[,.(rn,domain8426)],dualscores12199[,.(rn,domain12199)],by="rn")
srtdoms <- merge.data.table(srtdoms,dualscores16119[,.(rn,domain16119)],by="rn")
srtdoms <- merge.data.table(srtdoms,dualscores20079[,.(rn,domain20079)],by="rn")

## append nmf predictions
projlabs <- readRDS("processed-data_newseg/07_visLabelTransfer/01-VistoXen_NMFlabelxfer_tol1e8_k63_alph0.RDS")
projlabs <- projlabs$targets
projlabs <- lapply(projlabs,function(x){
    x$nmf_preds <- as.character(x$nmf_preds)
    y <- as.data.table(colData(x),keep.rownames=T)[,.(rn,nmf_preds)]
    setnames(y,2,"SRTnmfpred")
    return(y)
})
projlabs <- rbindlist(projlabs)

## sanity check, merge
stopifnot(length(unique(projlabs$rn))==length(unique(srtdoms$rn)))
srtdoms <- merge.data.table(srtdoms,projlabs,by="rn")
```

Descriptive stats per domain definition:
```{r}
##
nrow(srtdoms[SRTnmfpred=="LC"]) #8026
nrow(srtdoms[domain8426=="LC"]) #38185
nrow(srtdoms[domain8426=="LC"&SRTnmfpred=="LC"]) # 5826

nrow(srtdoms[SRTnmfpred=="LC"]) # 8026 (total pred cells so same)
nrow(srtdoms[domain12199=="LC"]) # #44189
nrow(srtdoms[domain12199=="LC"&SRTnmfpred=="LC"]) # 5995

nrow(srtdoms[domain16119=="LC"]) #38685
nrow(srtdoms[domain16119=="LC"&SRTnmfpred=="LC"]) # 5539

nrow(srtdoms[domain20079=="LC"]) # #32865
nrow(srtdoms[domain20079=="LC"&SRTnmfpred=="LC"]) # 5080

tabols <- melt.data.table(srtdoms,id.vars = c("rn","SRTnmfpred"))
tabols <- tabols[value=="LC"]
tabols[,timesInDom:=.N,by="rn"]

### number of cells in any domain definition: 47798 (9.72%) 

## number of cells in all domain definitions: 28562 (59.8% of cells in any LC domain)
# length(tabols[timesInDom==4,unique(rn)])# /length(unique(tabols$rn))

## number of cells unique to each definition
tabols[timesInDom==1][,.N,by="variable"]
#       variable     N
#         <fctr> <int>
# 1:  domain8426  1906
# 2: domain12199  2933
# 3: domain16119   564
# 4: domain20079   335
```

save
```{r}
fwrite(srtdoms,"processed-data_newseg/07_visLabelTransfer/05-srtLCsmoothDoms_and_NMFpreds_lut.txt",sep='\t',quote=F,row.names=F,col.names=T)
```
seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] parallelly_1.46.1   colorout_1.3-3      data.table_1.18.2.1

loaded via a namespace (and not attached):
 [1] digest_0.6.39     fastmap_1.2.0     xfun_0.56         knitr_1.51       
 [5] htmltools_0.5.9   rmarkdown_2.30    cli_3.6.5         compiler_4.5.2   
 [9] rprojroot_2.1.1   here_1.0.2        rstudioapi_0.18.0 tools_4.5.2      
[13] evaluate_1.0.5    yaml_2.3.12       otel_0.2.0        rlang_1.1.7      