---
title: "05-ONEcond_KNNsmooth_finestep_LUT.Rmd"
format: html
date: "03-10-2026"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
# library(SpatialFeatureExperiment)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}


## rstudio GUI tweaks
require(colorout)
ColorOut()
options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
library(parallel)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)

## ggplot defaults

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

two smoothings to use (let's see how different they are or maybe itll just be one)
load smoothings, xenium, and manual xenium LC assignments
```{r}
## smoothings
dualscores <- readRDS("processed-data/05_LabelTransfer/03-LCknnsmoothing_step2k75.RDS")
dualscores <- dualscores[c("k1thresh_0.08","k1thresh_0.12")]

dualscores8266 <- dualscores$k1thresh_0.08[,domain8266:=ifelse(greaterOrEqual(lcpred.smooth2.k75,0.266),"LC","Other")]

dualscores12252 <- dualscores$k1thresh_0.12[,domain12252:=ifelse(greaterOrEqual(lcpred.smooth2.k75,0.252),"LC","Other")]

nrow(dualscores12252[domain=="LC"]) #36830
nrow(dualscores8266[domain=="LC"]) #55761
nrow(dualscores8266[domain=="LC" & rn %in% dualscores12252[domain=="LC",rn]]) # 36822, so all but 8 of the cells identified by in the .12/.252 smoothing are in the more permissive 0.08/.266 smoothing.

# attach nmf-predicted labels to this
nmfpredlabs <- readRDS("processed-data/05_LabelTransfer/02-VistoXen_NMFlabelxfer_tol1e8_k63_alph0.RDS")
preds <- lapply(nmfpredlabs$targets,function(x){
  x$nmf_preds <- as.character(x$nmf_preds)
  y <- as.data.table(colData(x),keep.rownames=T)[,.(rn,nmf_preds)]
  setnames(y,2,"nmfpred")
  return(y)
})
preds <- rbindlist(preds)

domlabs <- merge.data.table(dualscores8266,dualscores12252[,.(rn,domain12252)],by="rn")
domlabs <- merge.data.table(domlabs,preds,by="rn")

domlabs[,lcpred.smooth1.k25:=NULL]
domlabs[,lcpred.smooth2.k75:=NULL]

fwrite(domlabs,"processed-data/05_LabelTransfer/06-LCsmoothings_LUT.txt",sep='\t',quote=F,row.names=F,col.names=T)
```
seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] parallelly_1.46.1   colorout_1.3-3      data.table_1.18.2.1

loaded via a namespace (and not attached):
 [1] digest_0.6.39     fastmap_1.2.0     xfun_0.56         knitr_1.51       
 [5] htmltools_0.5.9   rmarkdown_2.30    cli_3.6.5         compiler_4.5.2   
 [9] rprojroot_2.1.1   here_1.0.2        rstudioapi_0.18.0 tools_4.5.2      
[13] evaluate_1.0.5    yaml_2.3.12       otel_0.2.0        rlang_1.1.7      
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-03-11
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ───────────────────────────────────────────────────────────────────
 package     * version  date (UTC) lib source
 cli           3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 colorout    * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table  * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 digest        0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 evaluate      1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 fastmap       1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 here          1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools     0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 knitr         1.51     2025-12-20 [1] CRAN (R 4.5.2)
 otel          0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly  * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 rlang         1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown     2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot     2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rstudioapi    0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 sessioninfo   1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 xfun          0.56     2026-01-18 [1] CRAN (R 4.5.2)
 yaml          2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

──────────────────────────────────────────────────────────────────────────────