---
title: "05b-MolExperiments_to_DAPIonly_SPEs"
author: "Bernie Mulvey"
date: "04-16-26"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(MoleculeExperiment)
library(SpatialExperiment)
library(SpatialFeatureExperiment)
library(spatialLIBD)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}


## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

see preceding script for the temporary copies of NMDAPI that were pulled and converted for MolExpt compatibility

```{r}
brnums <- list.files("code/01_spe/05_DAPI_nuclearCount_spe/xendaptmp",full.names=T)

nucspes <- lapply(brnums,function(b){
  
  xendat <- MoleculeExperiment::readXenium(b,keepCols="all",addBoundaries="nucleus")
  
  xendat@molecules$detected[[1]] <- lapply(xendat@molecules$detected[[1]], FUN=function(g){return(g[which(g$qv>20),])})
  
   ## tabulate to nuclear counts using countMolecules, which will return a spatialExperiment object
    ## note that inexplicably the countMolecules function uses two more cores than whatever the user specifies.
    xendat <- MoleculeExperiment::countMolecules(xendat,moleculesAssay="detected",boundariesAssay="nucleus",nCores = 6)
    ## the cell identifiers come back different from our existing SPEs, so for compatability, substitute the "." for a "_". (Every cell's nucleus has the same ID as its cell, so we can check this against the cell-level DAPI-only SPE at the end, filter nuclei based on filtered cells, and replace the colData with that from the cell-level). 
    colnames(xendat) <- gsub(colnames(xendat),pattern="\\.",replacement="_")
    ## also for compatability, drop the dash from left hemisphere samples
    colnames(xendat) <- gsub(colnames(xendat),pattern="-L",replacement="L")
    
    ## add a colData column with the (undashed) brain # just in case
    # colData(xendat)$brnum <- gsub(basename(b),pattern="-L",replacement="L")
    
    return(xendat)
})
names(nucspes) <- basename(brnums)
# copy in case something goes wrong below..
holdnucs <- nucspes

### now concatenate these all into one giant SPE...
allnucs <- nucspes[[1]]
i<-2
for (i in c(2:length(nucspes))){
  allnucs <- cbind(allnucs,nucspes[[i]])
  nucspes[[i]] <- "done"
}

dim(allnucs)
# 541 490851
# ok we can get rid of the copy
rm(holdnucs)
gc(full=T)

## now fix colnames (cell/nuc IDs) to match existing SPEs
colnames(allnucs) <- gsub(colnames(allnucs),pattern="\\.",replacement="_")
colnames(allnucs) <- gsub(colnames(allnucs),pattern="-L",replacement="L")

### cleanup
rm(nucspes,holdnucs,i,brnums,brs)
gc()
```

## save an unfiltered DAPI-only nuclei SPE that includes control probes, depending what we might decide we want to do downstream
```{r}
saveRDS(allnucs,"processed-data/01_spe/05a_DAPIonly_nuclearCounts_SPE_unfiltered.RDS")
```

load filtered, DAPI-only cell level SPE to make sure all nuclei corresponding to retained DAPI-only cells are here
```{r}
dapcell <- readRDS("processed-data/02_QC/07b-DAPIonly_spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

sum(colnames(allnucs) %in% colnames(dapcell))==ncol(dapcell)
# TRUE
# YUSSSSS
keepcell <- colnames(dapcell)
## filter to nuclei of the retained cells
allnucs2 <- allnucs[,keepcell]

## clone colData
cellcd <- as.data.table(colData(dapcell),keep.rownames=T)
keepn <- grep(names(cellcd),pattern="outlier|subsets|total|sizeFactor|sum|detected|counts|cell_area|uniqueg",value=T,invert=T)
cellcd <- cellcd[,..keepn]
setnames(cellcd,"cell_id","nucleus_id")
setnames(cellcd,"unique_cellid","unique_nuc_id")
stopifnot(all(cellcd$rn==cellcd$unique_nuc_id))
stopifnot(all(cellcd$rn %in% colnames(allnucs2)))

cellcd <- DataFrame(cellcd,row.names=cellcd$rn)[colnames(allnucs2),]
cellcd$rn <- NULL

colData(allnucs2) <- cellcd

## for the filtered-to-retained cells one, we no longer need negative control probes either
keepg <- grep(rownames(allnucs2),pattern="Unassigned|NegControl|Deprecated",value=T,invert=T)
## save unfiltered though, depending what we might want to do with it.
saveRDS(allnucs2,"processed-data/01_spe/05b_DAPIonly_nuclearCounts_SPE_retainedDAPIcellsOnly.RDS")
```
seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0             parallelly_1.46.1              
 [3] colorout_1.3-3                  spatialLIBD_1.22.0             
 [5] SpatialFeatureExperiment_1.12.1 MoleculeExperiment_1.10.0      
 [7] SpatialExperiment_1.20.0        SingleCellExperiment_1.32.0    
 [9] SummarizedExperiment_1.40.0     Biobase_2.70.0                 
[11] GenomicRanges_1.62.1            Seqinfo_1.0.0                  
[13] IRanges_2.44.0                  S4Vectors_0.48.0               
[15] BiocGenerics_0.56.0             generics_0.1.4                 
[17] MatrixGenerics_1.22.0           matrixStats_1.5.0              
[19] data.table_1.18.2.1            

loaded via a namespace (and not attached):
  [1] spatialreg_1.4-2          bitops_1.0-9              sf_1.0-24                
  [4] EBImage_4.52.0            httr_1.4.7                RColorBrewer_1.1-3       
  [7] doParallel_1.0.17         tools_4.5.2               R6_2.6.1                 
 [10] DT_0.34.0                 HDF5Array_1.38.0          lazyeval_0.2.2           
 [13] rhdf5filters_1.22.0       GetoptLong_1.1.0          sp_2.2-0                 
 [16] gridExtra_2.3             cli_3.6.5                 sandwich_3.1-1           
 [19] sass_0.4.10               mvtnorm_1.3-3             S7_0.2.1                 
 [22] proxy_0.4-29              Rsamtools_2.26.0          R.utils_2.13.0           
 [25] dichromat_2.0-0.1         scater_1.38.0             sessioninfo_1.2.3        
 [28] attempt_0.3.1             limma_3.66.0              rstudioapi_0.18.0        
 [31] RSQLite_2.4.5             shape_1.4.6.1             BiocIO_1.20.0            
 [34] spdep_1.4-1               dplyr_1.2.0               Matrix_1.7-4             
 [37] ggbeeswarm_0.7.3          abind_1.4-8               R.methodsS3_1.8.2        
 [40] terra_1.8-93              lifecycle_1.0.5           multcomp_1.4-29          
 [43] yaml_2.3.12               edgeR_4.8.2               rhdf5_2.54.1             
 [46] SparseArray_1.10.8        BiocFileCache_3.0.0       paletteer_1.7.0          
 [49] grid_4.5.2                blob_1.3.0                promises_1.5.0           
 [52] dqrng_0.4.1               ExperimentHub_3.0.0       crayon_1.5.3             
 [55] lattice_0.22-7            beachmat_2.26.0           cowplot_1.2.0            
 [58] cigarillo_1.0.0           KEGGREST_1.50.0           magick_2.9.0             
 [61] zeallot_0.2.0             pillar_1.11.1             knitr_1.51               
 [64] ComplexHeatmap_2.26.1     rjson_0.2.23              boot_1.3-32              
 [67] codetools_0.2-20          wk_0.9.5                  glue_1.8.0               
 [70] vctrs_0.7.1               png_0.1-8                 gtable_0.3.6             
 [73] rematch2_2.1.2            cachem_1.1.0              xfun_0.56                
 [76] S4Arrays_1.10.1           mime_0.13                 DropletUtils_1.30.0      
 [79] coda_0.19-4.1             survival_3.8-6            sfheaders_0.4.5          
 [82] iterators_1.0.14          units_1.0-0               statmod_1.5.1            
 [85] TH.data_1.1-5             nlme_3.1-168              bit64_4.6.0-1            
 [88] filelock_1.0.3            rprojroot_2.1.1           bslib_0.10.0             
 [91] irlba_2.3.7               vipor_0.4.7               KernSmooth_2.23-26       
 [94] otel_0.2.0                colorspace_2.1-2          spData_2.3.4             
 [97] DBI_1.2.3                 tidyselect_1.2.1          bit_4.6.0                
[100] compiler_4.5.2            curl_7.0.0                httr2_1.2.2              
[103] BiocNeighbors_2.4.0       h5mread_1.2.1             DelayedArray_0.36.0      
[106] plotly_4.12.0             rtracklayer_1.70.1        scales_1.4.0             
[109] classInt_0.4-11           rappdirs_0.3.4            tiff_0.1-12              
[112] digest_0.6.39             fftwtools_0.9-11          rmarkdown_2.30           
[115] benchmarkmeData_2.0.0     XVector_0.50.0            htmltools_0.5.9          
[118] pkgconfig_2.0.3           jpeg_0.1-11               sparseMatrixStats_1.22.0 
[121] dbplyr_2.5.1              fastmap_1.2.0             rlang_1.1.7.9000         
[124] GlobalOptions_0.1.3       htmlwidgets_1.6.4         shiny_1.12.1             
[127] DelayedMatrixStats_1.32.0 farver_2.1.2              jquerylib_0.1.4          
[130] zoo_1.8-15                jsonlite_2.0.0            config_0.3.2             
[133] R.oo_1.27.1               BiocSingular_1.26.0       RCurl_1.98-1.17          
[136] magrittr_2.0.4            scuttle_1.20.0            s2_1.1.9                 
[139] Rhdf5lib_1.32.0           Rcpp_1.1.1                viridis_0.6.5            
[142] MASS_7.3-65               AnnotationHub_4.0.0       parallel_4.5.2           
[145] ggrepel_0.9.6             deldir_2.0-4              Biostrings_2.78.0        
[148] splines_4.5.2             circlize_0.4.17           locfit_1.5-9.12          
[151] ScaledMatrix_1.18.0       LearnBayes_2.15.2         BiocVersion_3.22.0       
[154] XML_3.99-0.20             evaluate_1.0.5            golem_0.5.1              
[157] BiocManager_1.30.27       foreach_1.5.2             httpuv_1.6.16            
[160] tidyr_1.3.2               purrr_1.2.1               benchmarkme_1.0.8        
[163] clue_0.3-66               ggplot2_4.0.2             rsvd_1.0.5               
[166] xtable_1.8-4              restfulr_0.0.16           e1071_1.7-17             
[169] later_1.4.5               viridisLite_0.4.3         class_7.3-23             
[172] tibble_3.3.1              memoise_2.0.1             beeswarm_0.4.0           
[175] AnnotationDbi_1.72.0      GenomicAlignments_1.46.0  cluster_2.1.8.1          
[178] shinyWidgets_0.9.0        here_1.0.2               
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-04-17
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────────────
 package                  * version    date (UTC) lib source
 abind                      1.4-8      2024-09-12 [2] CRAN (R 4.5.0)
 AnnotationDbi              1.72.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 AnnotationHub              4.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 attempt                    0.3.1      2020-05-03 [2] CRAN (R 4.5.0)
 beachmat                   2.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 beeswarm                   0.4.0      2021-06-01 [2] CRAN (R 4.5.0)
 benchmarkme                1.0.8      2022-06-12 [2] CRAN (R 4.5.0)
 benchmarkmeData            2.0.0      2026-01-19 [2] CRAN (R 4.5.2)
 Biobase                  * 2.70.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocFileCache              3.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics             * 0.56.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocIO                     1.20.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocManager                1.30.27    2025-11-14 [2] CRAN (R 4.5.2)
 BiocNeighbors              2.4.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocParallel             * 1.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocSingular               1.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocVersion                3.22.0     2025-10-07 [2] Bioconductor 3.22 (R 4.5.1)
 Biostrings                 2.78.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 bit                        4.6.0      2025-03-06 [2] CRAN (R 4.5.0)
 bit64                      4.6.0-1    2025-01-16 [2] CRAN (R 4.5.0)
 bitops                     1.0-9      2024-10-03 [2] CRAN (R 4.5.0)
 blob                       1.3.0      2026-01-14 [2] CRAN (R 4.5.2)
 boot                       1.3-32     2025-08-29 [2] CRAN (R 4.5.2)
 bslib                      0.10.0     2026-01-26 [2] CRAN (R 4.5.2)
 cachem                     1.1.0      2024-05-16 [2] CRAN (R 4.5.0)
 cigarillo                  1.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 circlize                   0.4.17     2025-12-08 [2] CRAN (R 4.5.2)
 class                      7.3-23     2025-01-01 [2] CRAN (R 4.5.2)
 classInt                   0.4-11     2025-01-08 [2] CRAN (R 4.5.0)
 cli                        3.6.5      2025-04-23 [2] CRAN (R 4.5.0)
 clue                       0.3-66     2024-11-13 [2] CRAN (R 4.5.0)
 cluster                    2.1.8.1    2025-03-12 [2] CRAN (R 4.5.0)
 coda                       0.19-4.1   2024-01-31 [2] CRAN (R 4.5.0)
 codetools                  0.2-20     2024-03-31 [2] CRAN (R 4.5.2)
 colorout                 * 1.3-3      2026-01-31 [2] Github (jalvesaq/colorout@64863bb)
 colorspace                 2.1-2      2025-09-22 [2] CRAN (R 4.5.0)
 ComplexHeatmap             2.26.1     2026-01-30 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 config                     0.3.2      2023-08-30 [2] CRAN (R 4.5.0)
 cowplot                    1.2.0      2025-07-07 [2] CRAN (R 4.5.0)
 crayon                     1.5.3      2024-06-20 [2] CRAN (R 4.5.0)
 curl                       7.0.0      2025-08-19 [2] CRAN (R 4.5.0)
 data.table               * 1.18.2.1   2026-01-27 [2] CRAN (R 4.5.2)
 DBI                        1.2.3      2024-06-02 [2] CRAN (R 4.5.0)
 dbplyr                     2.5.1      2025-09-10 [2] CRAN (R 4.5.0)
 DelayedArray               0.36.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 DelayedMatrixStats         1.32.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 deldir                     2.0-4      2024-02-28 [2] CRAN (R 4.5.0)
 dichromat                  2.0-0.1    2022-05-02 [2] CRAN (R 4.5.0)
 digest                     0.6.39     2025-11-19 [2] CRAN (R 4.5.2)
 doParallel                 1.0.17     2022-02-07 [2] CRAN (R 4.5.0)
 dplyr                      1.2.0      2026-02-03 [2] CRAN (R 4.5.2)
 dqrng                      0.4.1      2024-05-28 [2] CRAN (R 4.5.0)
 DropletUtils               1.30.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 DT                         0.34.0     2025-09-02 [2] CRAN (R 4.5.0)
 e1071                      1.7-17     2025-12-18 [2] CRAN (R 4.5.2)
 EBImage                    4.52.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 edgeR                      4.8.2      2025-12-23 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate                   1.0.5      2025-08-27 [2] CRAN (R 4.5.0)
 ExperimentHub              3.0.0      2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 farver                     2.1.2      2024-05-13 [2] CRAN (R 4.5.0)
 fastmap                    1.2.0      2024-05-15 [2] CRAN (R 4.5.0)
 fftwtools                  0.9-11     2021-03-01 [2] CRAN (R 4.5.0)
 filelock                   1.0.3      2023-12-11 [2] CRAN (R 4.5.0)
 foreach                    1.5.2      2022-02-02 [2] CRAN (R 4.5.0)
 generics                 * 0.1.4      2025-05-09 [2] CRAN (R 4.5.0)
 GenomicAlignments          1.46.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 GenomicRanges            * 1.62.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 GetoptLong                 1.1.0      2025-11-28 [2] CRAN (R 4.5.2)
 ggbeeswarm                 0.7.3      2025-11-29 [2] CRAN (R 4.5.2)
 ggplot2                    4.0.2      2026-02-03 [2] CRAN (R 4.5.2)
 ggrepel                    0.9.6      2024-09-07 [2] CRAN (R 4.5.0)
 GlobalOptions              0.1.3      2025-11-28 [2] CRAN (R 4.5.2)
 glue                       1.8.0      2024-09-30 [2] CRAN (R 4.5.0)
 golem                      0.5.1      2024-08-27 [2] CRAN (R 4.5.0)
 gridExtra                  2.3        2017-09-09 [2] CRAN (R 4.5.0)
 gtable                     0.3.6      2024-10-25 [2] CRAN (R 4.5.0)
 h5mread                    1.2.1      2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 HDF5Array                  1.38.0     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 here                       1.0.2      2025-09-15 [2] CRAN (R 4.5.0)
 htmltools                  0.5.9      2025-12-04 [2] CRAN (R 4.5.2)
 htmlwidgets                1.6.4      2023-12-06 [2] CRAN (R 4.5.0)
 httpuv                     1.6.16     2025-04-16 [2] CRAN (R 4.5.0)
 httr                       1.4.7      2023-08-15 [2] CRAN (R 4.5.0)
 httr2                      1.2.2      2025-12-08 [2] CRAN (R 4.5.2)
 IRanges                  * 2.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 irlba                      2.3.7      2026-01-30 [2] CRAN (R 4.5.2)
 iterators                  1.0.14     2022-02-05 [2] CRAN (R 4.5.0)
 jpeg                       0.1-11     2025-03-21 [2] CRAN (R 4.5.0)
 jquerylib                  0.1.4      2021-04-26 [2] CRAN (R 4.5.0)
 jsonlite                   2.0.0      2025-03-27 [2] CRAN (R 4.5.0)
 KEGGREST                   1.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 KernSmooth                 2.23-26    2025-01-01 [2] CRAN (R 4.5.2)
 knitr                      1.51       2025-12-20 [2] CRAN (R 4.5.2)
 later                      1.4.5      2026-01-08 [2] CRAN (R 4.5.2)
 lattice                    0.22-7     2025-04-02 [2] CRAN (R 4.5.2)
 lazyeval                   0.2.2      2019-03-15 [2] CRAN (R 4.5.0)
 LearnBayes                 2.15.2     2026-01-31 [2] CRAN (R 4.5.2)
 lifecycle                  1.0.5      2026-01-08 [2] CRAN (R 4.5.2)
 limma                      3.66.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 locfit                     1.5-9.12   2025-03-05 [2] CRAN (R 4.5.0)
 magick                     2.9.0      2025-09-08 [2] CRAN (R 4.5.0)
 magrittr                   2.0.4      2025-09-12 [2] CRAN (R 4.5.0)
 MASS                       7.3-65     2025-02-28 [2] CRAN (R 4.5.2)
 Matrix                     1.7-4      2025-08-28 [2] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics           * 1.22.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 matrixStats              * 1.5.0      2025-01-07 [2] CRAN (R 4.5.0)
 memoise                    2.0.1      2021-11-26 [2] CRAN (R 4.5.0)
 mime                       0.13       2025-03-17 [2] CRAN (R 4.5.0)
 MoleculeExperiment       * 1.10.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 multcomp                   1.4-29     2025-10-20 [2] CRAN (R 4.5.0)
 mvtnorm                    1.3-3      2025-01-10 [2] CRAN (R 4.5.0)
 nlme                       3.1-168    2025-03-31 [2] CRAN (R 4.5.2)
 otel                       0.2.0      2025-08-29 [2] CRAN (R 4.5.0)
 paletteer                  1.7.0      2026-01-08 [2] CRAN (R 4.5.2)
 parallelly               * 1.46.1     2026-01-08 [2] CRAN (R 4.5.2)
 pillar                     1.11.1     2025-09-17 [2] CRAN (R 4.5.0)
 pkgconfig                  2.0.3      2019-09-22 [2] CRAN (R 4.5.0)
 plotly                     4.12.0     2026-01-24 [2] CRAN (R 4.5.2)
 png                        0.1-8      2022-11-29 [2] CRAN (R 4.5.0)
 promises                   1.5.0      2025-11-01 [2] CRAN (R 4.5.0)
 proxy                      0.4-29     2025-12-29 [2] CRAN (R 4.5.2)
 purrr                      1.2.1      2026-01-09 [2] CRAN (R 4.5.2)
 R.methodsS3                1.8.2      2022-06-13 [2] CRAN (R 4.5.0)
 R.oo                       1.27.1     2025-05-02 [2] CRAN (R 4.5.0)
 R.utils                    2.13.0     2025-02-24 [2] CRAN (R 4.5.0)
 R6                         2.6.1      2025-02-15 [2] CRAN (R 4.5.0)
 rappdirs                   0.3.4      2026-01-17 [2] CRAN (R 4.5.2)
 RColorBrewer               1.1-3      2022-04-03 [2] CRAN (R 4.5.0)
 Rcpp                       1.1.1      2026-01-10 [2] CRAN (R 4.5.2)
 RCurl                      1.98-1.17  2025-03-22 [2] CRAN (R 4.5.0)
 rematch2                   2.1.2      2020-05-01 [2] CRAN (R 4.5.0)
 restfulr                   0.0.16     2025-06-27 [2] CRAN (R 4.5.0)
 rhdf5                      2.54.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 rhdf5filters               1.22.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 Rhdf5lib                   1.32.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 rjson                      0.2.23     2024-09-16 [2] CRAN (R 4.5.0)
 rlang                      1.1.7.9000 2026-02-04 [1] Github (r-lib/rlang@bae1fe1)
 rmarkdown                  2.30       2025-09-28 [2] CRAN (R 4.5.0)
 rprojroot                  2.1.1      2025-08-26 [2] CRAN (R 4.5.0)
 Rsamtools                  2.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 RSQLite                    2.4.5      2025-11-30 [2] CRAN (R 4.5.2)
 rstudioapi                 0.18.0     2026-01-16 [2] CRAN (R 4.5.2)
 rsvd                       1.0.5      2021-04-16 [2] CRAN (R 4.5.0)
 rtracklayer                1.70.1     2025-12-20 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 s2                         1.1.9      2025-05-23 [2] CRAN (R 4.5.0)
 S4Arrays                   1.10.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 S4Vectors                * 0.48.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 S7                         0.2.1      2025-11-14 [2] CRAN (R 4.5.2)
 sandwich                   3.1-1      2024-09-15 [2] CRAN (R 4.5.0)
 sass                       0.4.10     2025-04-11 [2] CRAN (R 4.5.0)
 ScaledMatrix               1.18.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 scales                     1.4.0      2025-04-24 [2] CRAN (R 4.5.0)
 scater                     1.38.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 scuttle                    1.20.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 Seqinfo                  * 1.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 sessioninfo                1.2.3      2025-02-05 [2] CRAN (R 4.5.0)
 sf                         1.0-24     2026-01-13 [2] CRAN (R 4.5.2)
 sfheaders                  0.4.5      2025-11-25 [2] CRAN (R 4.5.2)
 shape                      1.4.6.1    2024-02-23 [2] CRAN (R 4.5.0)
 shiny                      1.12.1     2025-12-09 [2] CRAN (R 4.5.2)
 shinyWidgets               0.9.0      2025-02-21 [2] CRAN (R 4.5.0)
 SingleCellExperiment     * 1.32.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 sp                         2.2-0      2025-02-01 [2] CRAN (R 4.5.0)
 SparseArray                1.10.8     2025-12-18 [2] Bioconductor 3.22 (R 4.5.2)
 sparseMatrixStats          1.22.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 SpatialExperiment        * 1.20.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 SpatialFeatureExperiment * 1.12.1     2025-11-04 [2] Bioconductor 3.22 (R 4.5.2)
 spatialLIBD              * 1.22.0     2025-11-04 [2] Bioconductor 3.22 (R 4.5.2)
 spatialreg                 1.4-2      2025-09-06 [2] CRAN (R 4.5.0)
 spData                     2.3.4      2025-01-08 [2] CRAN (R 4.5.0)
 spdep                      1.4-1      2025-08-31 [2] CRAN (R 4.5.0)
 statmod                    1.5.1      2025-10-09 [2] CRAN (R 4.5.0)
 SummarizedExperiment     * 1.40.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 survival                   3.8-6      2026-01-16 [2] CRAN (R 4.5.2)
 terra                      1.8-93     2026-01-12 [2] CRAN (R 4.5.2)
 TH.data                    1.1-5      2025-11-17 [2] CRAN (R 4.5.2)
 tibble                     3.3.1      2026-01-11 [2] CRAN (R 4.5.2)
 tidyr                      1.3.2      2025-12-19 [2] CRAN (R 4.5.2)
 tidyselect                 1.2.1      2024-03-11 [2] CRAN (R 4.5.0)
 tiff                       0.1-12     2023-11-28 [2] CRAN (R 4.5.0)
 units                      1.0-0      2025-10-09 [2] CRAN (R 4.5.0)
 vctrs                      0.7.1      2026-01-23 [2] CRAN (R 4.5.2)
 vipor                      0.4.7      2023-12-18 [2] CRAN (R 4.5.0)
 viridis                    0.6.5      2024-01-29 [2] CRAN (R 4.5.0)
 viridisLite                0.4.3      2026-02-04 [2] CRAN (R 4.5.2)
 wk                         0.9.5      2025-12-18 [2] CRAN (R 4.5.2)
 xfun                       0.56       2026-01-18 [2] CRAN (R 4.5.2)
 XML                        3.99-0.20  2025-11-08 [2] CRAN (R 4.5.0)
 xtable                     1.8-4      2019-04-21 [2] CRAN (R 4.5.0)
 XVector                    0.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 yaml                       2.3.12     2025-12-10 [2] CRAN (R 4.5.2)
 zeallot                    0.2.0      2025-05-27 [2] CRAN (R 4.5.0)
 zoo                        1.8-15     2025-12-15 [2] CRAN (R 4.5.2)

 [1] /Users/bmulvey/Library/R/arm64/4.5/library
 [2] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────────────