---
title: "04-Compare_DomainLabXfer_toVis"
author: "Bernie Mulvey"
date: "03-01-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(ggrastr)
library(gridExtra)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 7,color = "#000000"), axis.title.x = element_text(size = 8,color = "#000000"), axis.text.y = element_text(size = 7,color = "#000000"), axis.title.y = element_text(size =8,color = "#000000"), plot.title = element_markdown(size = 8,hjust=0.5,color = "#000000"), strip.text = element_text(size=8), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=7,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.25),panel.grid.major = element_line(linewidth=0.25),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
fullde <- readRDS("processed-data/06_DomainDE/02-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_XferedVisLabels.RDS")

fulldes <- list(
  dom0528=list(fullde$dom0528$ApoAllelesLabel$detabs,fullde$dom0528$E2E4label$detabs,fullde$dom0528$sexlabel$detabs,fullde$dom0528$YRIAPOlabel$detabs),
  dom1020=list(fullde$dom1020$ApoAllelesLabel$detabs,fullde$dom1020$E2E4label$detabs,fullde$dom1020$sexlabel$detabs,fullde$dom1020$YRIAPOlabel$detabs),
  dom1512=list(fullde$dom1512$ApoAllelesLabel$detabs,fullde$dom1512$E2E4label$detabs,fullde$dom1512$sexlabel$detabs,fullde$dom1512$YRIAPOlabel$detabs))

names(fulldes[[1]]) <- c("ApoAllelesLabel","E2E4label","sexlabel","YRIAPOlabel")
names(fulldes[[2]]) <- c("ApoAllelesLabel","E2E4label","sexlabel","YRIAPOlabel")
names(fulldes[[3]]) <- c("ApoAllelesLabel","E2E4label","sexlabel","YRIAPOlabel")

subde <- readRDS("processed-data/06_DomainDE/04-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_withinLCVisLabs.RDS")

subdes <- list(
  dom0528=list(subde$dom0528$ApoAllelesLabel$detabs,subde$dom0528$E2E4label$detabs,subde$dom0528$sexlabel$detabs,subde$dom0528$YRIAPOlabel$detabs),
  dom1020=list(subde$dom1020$ApoAllelesLabel$detabs,subde$dom1020$E2E4label$detabs,subde$dom1020$sexlabel$detabs,subde$dom1020$YRIAPOlabel$detabs),
  dom1512=list(subde$dom1512$ApoAllelesLabel$detabs,subde$dom1512$E2E4label$detabs,subde$dom1512$sexlabel$detabs,subde$dom1512$YRIAPOlabel$detabs))


names(subdes[[1]]) <- c("ApoAllelesLabel","E2E4label","sexlabel","YRIAPOlabel")
names(subdes[[2]]) <- c("ApoAllelesLabel","E2E4label","sexlabel","YRIAPOlabel")
names(subdes[[3]]) <- c("ApoAllelesLabel","E2E4label","sexlabel","YRIAPOlabel")


rm(fullde,subde)
```

```{r}
fulldes <- lapply(fulldes,function(x){
  z <- lapply(x,function(a){
    y <- rbindlist(a,idcol="contrast",fill=T)
    return(y)
  })
  y <- rbindlist(z,fill=T)
  return(y)
})

fulldes <- lapply(fulldes,function(x){
  x[contrast %in% grep(contrast,pattern="LC_outOfDomain",value=T),LCsub:="outDom"]
  x[contrast %in% grep(contrast,pattern="LCdomain",value=T),LCsub:="inDom"]
    
  x[contrast %in% grep(contrast,pattern="LC_outOfDomain_",value=T),contrast:=gsub(contrast,pattern="LC_outOfDomain_",replacement="LC_")]
  x[contrast %in% grep(contrast,pattern="LCdomain_",value=T),contrast:=gsub(contrast,pattern="LCdomain_",replacement="LC_")]
  
  return(x)
})

subdes <- lapply(subdes,function(x){
  z <- lapply(x,function(y){
    a <- rbindlist(y,idcol="contrast")
    return(a)
  })
  z <- rbindlist(z,fill=T)
  ## some tables don't have a gene_name column, but all have an RN column with symbol so propagate for uniformity
  z[,gene_name:=rn]
  return(z)
})

fulldes <- lapply(fulldes,function(x){
  setnames(x,"t","tstatfull")
  return(x)
})


subdes <- lapply(subdes,function(x){
  setnames(x,"t","tstatinLCDom")
  return(x)
})

fulldes <- lapply(fulldes,function(x){
  y <- x[,.(gene_name,contrast,tstatfull,LCsub)]
  return(y)
})

subdes <- lapply(subdes,function(x){
  y <- x[,.(gene_name,contrast,tstatinLCDom)]
  return(y)
})
```

```{r}
jtdes <- mapply(fulldes,subdes,SIMPLIFY=FALSE,FUN=function(x,y){
  z <- merge.data.table(x,y,by=c("contrast","gene_name"))
  z[!is.na(LCsub),contrast:=paste0(contrast,"_",LCsub)]
  return(z)
})

keepcont <- unique(c(jtdes$dom0528$contrast,jtdes$dom1020$contrast,jtdes$dom1512$contrast))
## except LC contrasts for which the plots are going to need to be a little different
lccont <- grep(keepcont,pattern="LC",value=T)
keepcont <- grep(keepcont,pattern="LC",value=T,invert=T)

pltset <- function(con){
  i<-1
  plts<-list()
  for (i in c(1:length(keepcont))){
    d <- keepcont[i]
    j<-1
    subplt <- list()
    for (j in c(1:length(jtdes))){
      x <- jtdes[[j]]
      n <- names(jtdes)[j]
      pltdat <- x[contrast==d,]
      minx <- min(c(pltdat$tstatfull,pltdat$tstatinLCDom),na.rm=T)
      maxx <- max(c(pltdat$tstatfull,pltdat$tstatinLCDom),na.rm=T)
      mm <- max(abs(c(minx,maxx)))
      if(is.infinite(mm)){
        mm <- 5
      }
      p <- ggplot(pltdat,aes(x=tstatfull,y=tstatinLCDom))+
        geom_point(size=0.5,alpha=0.5)+
        labs(title=paste0(d),x="Xenwide DE t",y=paste0("t-stat: Cell Type DE within LC\nDomain (smoothing",n,")"))+
        scale_x_continuous(limits=c(-mm,mm),expand=c(0,0))+
        scale_y_continuous(limits=c(-mm,mm),expand=c(0,0))
        geom_abline(slope=1,intercept=0,color="red",linetype="dashed")
      subplt[[j]] <- p
    }
    retobj <- gridExtra::arrangeGrob(grobs = subplt, ncol = length(subplt))
    plts[[i]] <- retobj
  }
  dev.off()
  return(plts)
}


### all tests
allplt <- pltset(keepcont)
png("plots/06_DomainDE/04a-Systematic_FullXen_LCXenDom_labelwise_tstat_comparisons.png",height=length(keepcont)*3,width=9,unit="in",res=150)
grid.arrange(grobs=allplt,ncol=1)
dev.off()

### non single-geno tests
keepcont <- grep(keepcont,pattern="hom|het",value=T,invert=T)
mainsplt <- pltset(keepcont)
png("plots/06_DomainDE/04b-Systematic_FullXen_LCXenDom_labelwise_tstat_comparisons_mainDEtestsonly.png",height=length(keepcont)*3,width=9,unit="in",res=150)
grid.arrange(grobs=mainsplt,ncol=1)
dev.off()

rm(keepcont,allplt,mainsplt,pltset)
```

## for LC contrasts: we want a plot of domain DE ("indom") vs LC domain LC cells, and vs non-LC domain LC cells, as well as those two compared to each other.
```{r}
lccont <- unique(gsub(lccont,pattern="_outDom|_inDom",replacement=""))

pltset <- function(con){
  plts <- lapply(con,function(d){
    subplt <- mapply(x=jtdes,n=names(jtdes),SIMPLIFY=FALSE,function(x,n){
      pltdat <- copy(x[contrast %in% grep(contrast,pattern=d,value=T)]) # both indomain and out of domain contrasts
      domainde <- pltdat[LCsub=="inDom"]
      setnames(domainde,"tstatfull","tstat_LCdomain")
      
      nondomcells <- pltdat[LCsub=="outDom"]
      setnames(nondomcells,"tstatfull","tstat_nonDomainLCcells")
      
      newpltdat <- merge.data.table(domainde[,.(gene_name,tstat_LCdomain)],nondomcells[,.(gene_name,tstat_nonDomainLCcells,tstatinLCDom)],by="gene_name")
      setnames(newpltdat,"tstatinLCDom","tstat_LCdomainLCcells")
      
      minx <- min(c(newpltdat$tstat_LCdomain,newpltdat$tstat_nonDomainLCcells,newpltdat$tstat_LCdomainLCcells),na.rm=T)
      maxx <- max(c(newpltdat$tstat_LCdomain,newpltdat$tstat_nonDomainLCcells,newpltdat$tstat_LCdomainLCcells),na.rm=T)
      mm <- max(abs(c(minx,maxx)))
      if(is.infinite(mm)){
        mm <- 5
      }
      
      p1 <- ggplot(newpltdat,aes(x=tstat_LCdomain,y=tstat_LCdomainLCcells))+
        geom_point(size=0.5,alpha=0.5)+
        labs(title=paste0(d),x="LC Domain DE tstat",y=paste0("LC Cells of LC Domain \n(smoothing",n,") tstat"))+
        scale_x_continuous(limits=c(-mm,mm),expand=c(0,0))+
        scale_y_continuous(limits=c(-mm,mm),expand=c(0,0))+
        geom_abline(slope=1,intercept=0,color="red",linetype="dashed")
        
        p2 <- ggplot(newpltdat,aes(x=tstat_LCdomain,y=tstat_nonDomainLCcells))+
        geom_point(size=0.5,alpha=0.5)+
        labs(title=paste0(d),x="LC Domain DE tstat",y=paste0("LC Cells OUTSIDE Domain \n(smoothing",n,") tstat"))+
        scale_x_continuous(limits=c(-mm,mm),expand=c(0,0))+
        scale_y_continuous(limits=c(-mm,mm),expand=c(0,0))+
        geom_abline(slope=1,intercept=0,color="red",linetype="dashed")
        
        p3 <- ggplot(newpltdat,aes(x=tstat_LCdomainLCcells,y=tstat_nonDomainLCcells))+
        geom_point(size=0.5,alpha=0.5)+
        labs(title=paste0(d),x=paste0("LC Cells of LC Domain \n(smoothing",n,") tstat"),y=paste0("LC Cells OUTSIDE Domain","\n(smoothing",n,") tstat"))+
        scale_x_continuous(limits=c(-mm,mm),expand=c(0,0))+
        scale_y_continuous(limits=c(-mm,mm),expand=c(0,0))+
        geom_abline(slope=1,intercept=0,color="red",linetype="dashed")
    
        retobj <- gridExtra::arrangeGrob(grobs = list(p1,p2,p3), ncol = 3)
        return(retobj)
    })
    subret <- gridExtra::arrangeGrob(grobs=subplt,ncol=1)
    return(subret)
  })
  dev.off()
  return(plts)
}

plts <- pltset(lccont)
png("plots/06_DomainDE/04c-Systematic_LCdomCells_nonDomLCcells_LCwholedom_DEtstat_comparisons.png",height=length(lccont)*9,width=9,unit="in",res=150)
grid.arrange(grobs=plts,ncol=1)
dev.off()

## make a separate one thats cleaner without all the subgeno analyses
keepcont <- grep(lccont,pattern="E4hom|E4het|E2hom|E2het",value=T,invert=T)
plts <- pltset(keepcont)
png("plots/06_DomainDE/04d-Systematic_LCdomCells_nonDomLCcells_LCwholedom_DEtstat_comparisons_mainDEtestsonly.png",height=length(keepcont)*9,width=9,unit="in",res=150)
grid.arrange(grobs=plts,ncol=1)
dev.off()

```


```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] gridExtra_2.3       ggrastr_1.0.2       ggtext_0.1.2        ggplot2_4.0.2      
 [5] edgeR_4.8.2         limma_3.66.0        BiocParallel_1.44.0 parallelly_1.46.1  
 [9] colorout_1.3-3      data.table_1.18.2.1

loaded via a namespace (and not attached):
 [1] generics_0.1.4     xml2_1.5.2         stringi_1.8.7      lattice_0.22-7    
 [5] digest_0.6.39      magrittr_2.0.4     evaluate_1.0.5     grid_4.5.2        
 [9] RColorBrewer_1.1-3 fastmap_1.2.0      rprojroot_2.1.1    scales_1.4.0      
[13] codetools_0.2-20   cli_3.6.5          rlang_1.1.7        litedown_0.9      
[17] commonmark_2.0.0   withr_3.0.2        yaml_2.3.12        otel_0.2.0        
[21] ggbeeswarm_0.7.3   tools_4.5.2        dplyr_1.2.0        locfit_1.5-9.12   
[25] here_1.0.2         vctrs_0.7.1        R6_2.6.1           lifecycle_1.0.5   
[29] stringr_1.6.0      vipor_0.4.7        pkgconfig_2.0.3    beeswarm_0.4.0    
[33] pillar_1.11.1      gtable_0.3.6       glue_1.8.0         Rcpp_1.1.1        
[37] statmod_1.5.1      xfun_0.56          tibble_3.3.1       tidyselect_1.2.1  
[41] rstudioapi_0.18.0  knitr_1.51         dichromat_2.0-0.1  farver_2.1.2      
[45] htmltools_0.5.9    rmarkdown_2.30     labeling_0.4.3     compiler_4.5.2    
[49] S7_0.2.1           markdown_2.0       gridtext_0.1.5    
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-03-02
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package      * version  date (UTC) lib source
 beeswarm       0.4.0    2021-06-01 [1] CRAN (R 4.5.0)
 BiocParallel * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli            3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools      0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout     * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 commonmark     2.0.0    2025-07-07 [1] CRAN (R 4.5.0)
 data.table   * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 dichromat      2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest         0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr          1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 edgeR        * 4.8.2    2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate       1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver         2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap        1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics       0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 ggbeeswarm     0.7.3    2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2      * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggrastr      * 1.0.2    2023-06-01 [1] CRAN (R 4.5.0)
 ggtext       * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue           1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra    * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext       0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable         0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here           1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools      0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 knitr          1.51     2025-12-20 [1] CRAN (R 4.5.2)
 labeling       0.4.3    2023-08-29 [1] CRAN (R 4.5.0)
 lattice        0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle      1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 limma        * 3.66.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 litedown       0.9      2025-12-18 [1] CRAN (R 4.5.2)
 locfit         1.5-9.12 2025-03-05 [1] CRAN (R 4.5.0)
 magrittr       2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 markdown       2.0      2025-03-23 [1] CRAN (R 4.5.0)
 otel           0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly   * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pillar         1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig      2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 R6             2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer   1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp           1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rlang          1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown      2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot      2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rstudioapi     0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S7             0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales         1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 sessioninfo    1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 statmod        1.5.1    2025-10-09 [1] CRAN (R 4.5.0)
 stringi        1.8.7    2025-03-27 [1] CRAN (R 4.5.0)
 stringr        1.6.0    2025-11-04 [1] CRAN (R 4.5.0)
 tibble         3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect     1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs          0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 vipor          0.4.7    2023-12-18 [1] CRAN (R 4.5.0)
 withr          3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun           0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2           1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 yaml           2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────