---
title: "05-PlotPbulkXpr_VisLCdomain_XenLCdomainReplicated_E2E4DEGs"
format: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(jaffelab)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
##

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
precision <- sqrt(.Machine$double.eps)
(x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
precision <- sqrt(.Machine$double.eps)
(x <= y) | (abs(x-y) <= precision)}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost='localhost')


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = '#000000'), axis.title.x = element_text(size = 9,color = '#000000'), axis.text.y = element_text(size = 8,color = '#000000'), axis.title.y = element_text(size =9,color = '#000000'), plot.title = element_markdown(size = 10,hjust=0.5,color = '#000000'), strip.text = element_text(size=11), legend.text = element_text(size=7,color = '#000000'), legend.title = element_text(size=8,hjust=0.5,color = '#000000'),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position='plot'))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

plot each visium LC domain E4-E2 DEG that replicates in both xenium LC domains:

AIF1
AQP4
C3
EGFR
FGFR3
GJA1
IDH2
ITGB8
NPY
NTSR2
PLPP3
SLC7A11
SPARC
SPON1
TRIL

```{r}
xde <- readRDS("processed-data/06_DomainDE/02-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_XferedVisLabels.RDS")

xde.8266 <- xde$dom8266$E2E4label$voomLFt$EList$E[c("AIF1", "AQP4", "C3", "EGFR", "FGFR3", "GJA1", "IDH2", "ITGB8", "NPY", "NTSR2", "PLPP3", "SLC7A11", "SPARC", "SPON1", "TRIL"),]
xde.12252 <- xde$dom12252$E2E4label$voomLFt$EList$E[c("AIF1", "AQP4", "C3", "EGFR", "FGFR3", "GJA1", "IDH2", "ITGB8", "NPY", "NTSR2", "PLPP3", "SLC7A11", "SPARC", "SPON1", "TRIL"),]
```

reconfigure data for jaffelab::cleaningy
```{r}
# get sample metadata in the same arrangement for LC pbulks
keepsamps <- grep(colnames(xde.8266),pattern="LC_Br",value=T)
cd8266 <- xde$dom8266$E2E4label$dgel$samples[keepsamps,]
xde.8266 <- xde.8266[,keepsamps]

keepsamps <- grep(colnames(xde.12252),pattern="LC_Br",value=T)
cd12252 <- xde$dom12252$E2E4label$dgel$samples[keepsamps, ]
xde.12252 <- xde.12252[, keepsamps]
```
plot: domain 8266
```{r}
# ensure factors are set for model matrix
cd8266$APOEsimple <- factor(cd8266$APOEsimple,levels = c("E2","E4"))
cd8266$Sex <- factor(cd8266$Sex)
# model matrix
modmat <- model.matrix(~0+APOEsimple+Sex+Age+YRI,data = cd8266)
## not correcting for E2/E4 (i.e., first 2 columns of the model matrix)
testres <- jaffelab::cleaningY(xde.8266,modmat,2)


testres2.8266 <- as.data.table(cbind(t(testres),cd8266$APOEsimple,cd8266$brnum,cd8266$Sex))
setnames(testres2.8266,c((ncol(testres2.8266)-2):ncol(testres2.8266)),c("APOE","brnum","sex"))
testres2.8266 <- as.data.table(cbind(testres2.8266[,lapply(.SD,as.numeric),.SDcols=c(1:(ncol(testres2.8266)-3))],testres2.8266[,.(APOE,brnum,sex)]))

testres2.8266 <- melt.data.table(testres2.8266,id.vars=c("APOE","brnum","sex"))
testres2.8266[,titleg:=paste0("*",variable,"*")]
testres2.8266[,APOE:=as.character(APOE)]
testres2.8266[,APOE:=ifelse(APOE=="1","E2","E4")]

pdf("plots/06_DomainDE/05-LCdomain_VisDE_replicatedE2E4gene_pbulkboxplots.pdf",height=9,width=6,onefile=TRUE)
ggplot(testres2.8266,aes(x=APOE,y=value))+
    geom_boxplot(outliers = F,linewidth = 0.25)+
    geom_point(alpha=0.7,size=1,stroke = 0,position = position_jitter(width=0.2),aes(stroke="black",color=APOE))+
    facet_wrap(.~titleg,nrow=3,scales="free_y")+
    labs(y="Pseudobulk Xenium LC Domain (Permissive Smoothing)\n Expression corrected for Sex, Age, and Genomic Ancestry",
         x="APOE Carrier Type")+
    guides(color="none",fill="none")+
    theme(strip.background = element_blank(),
          strip.text = element_markdown(size=9,hjust=0.5),
          axis.text.x=element_text(size=8,angle=30,hjust=1,vjust=1))
dev.off()
```

plot: domain 12252
```{r}
# ensure factors are set for model matrix
cd12252$APOEsimple <- factor(cd12252$APOEsimple, levels = c("E2", "E4"))
cd12252$Sex <- factor(cd12252$Sex)
# model matrix
modmat <- model.matrix(~ 0 + APOEsimple + Sex + Age + YRI, data = cd12252)
## not correcting for E2/E4 (i.e., first 2 columns of the model matrix)
testres <- jaffelab::cleaningY(xde.12252, modmat, 2)


testres2.12252 <- as.data.table(cbind(t(testres), cd12252$APOEsimple, cd12252$brnum, cd12252$Sex))
setnames(testres2.12252, c((ncol(testres2.12252) - 2):ncol(testres2.12252)), c("APOE", "brnum", "sex"))
testres2.12252 <- as.data.table(cbind(testres2.12252[, lapply(.SD, as.numeric), .SDcols = c(1:(ncol(testres2.12252) - 3))], testres2.12252[, .(APOE, brnum, sex)]))

testres2.12252 <- melt.data.table(testres2.12252, id.vars = c("APOE", "brnum", "sex"))
testres2.12252[, titleg := paste0("*", variable, "*")]
testres2.12252[, APOE := as.character(APOE)]
testres2.12252[, APOE := ifelse(APOE == "1", "E2", "E4")]

pdf("plots/06_DomainDE/05-LCdomain_VisDE_replicatedE2E4gene_pbulkboxplots_StrictSmoothing.pdf", height = 9, width = 6)
ggplot(testres2.12252, aes(x = APOE, y = value)) +
    geom_boxplot(outliers = FALSE, linewidth = 0.25) +
    geom_point(alpha = 0.7, size = 1, stroke = 0, position = position_jitter(width = 0.2), aes(color = APOE)) +
    facet_wrap(. ~ titleg, nrow = 3, scales = "free_y") +
    labs(
        y = "Pseudobulk Xenium LC Domain (Strict Smoothing)\n Expression corrected for Sex, Age, and Genomic Ancestry",
        x = "APOE Carrier Type"
    ) +
    guides(color = "none", fill = "none") +
    theme(
        strip.background = element_blank(),
        strip.text = element_markdown(size = 9, hjust = 0.5),
        axis.text.x = element_text(size = 8, angle = 30, hjust = 1, vjust = 1)
    )
dev.off()
```

session inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    parallel  stats     graphics 
[5] grDevices utils     methods   base     

other attached packages:
 [1] jaffelab_0.99.34           
 [2] rafalib_1.0.4              
 [3] dreamlet_1.8.0             
 [4] variancePartition_1.40.1   
 [5] limma_3.66.0               
 [6] SpatialExperiment_1.20.0   
 [7] SingleCellExperiment_1.32.0
 [8] SummarizedExperiment_1.40.0
 [9] Biobase_2.70.0             
[10] GenomicRanges_1.62.1       
[11] Seqinfo_1.0.0              
[12] IRanges_2.44.0             
[13] S4Vectors_0.48.0           
[14] BiocGenerics_0.56.0        
[15] generics_0.1.4             
[16] MatrixGenerics_1.22.0      
[17] matrixStats_1.5.0          
[18] gridExtra_2.3              
[19] ggtext_0.1.2               
[20] ggplot2_4.0.2              
[21] BiocParallel_1.44.0        
[22] parallelly_1.46.1          
[23] colorout_1.3-3             
[24] data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] splines_4.5.2            
  [2] bitops_1.0-9             
  [3] tibble_3.3.1             
  [4] graph_1.88.1             
  [5] XML_3.99-0.20            
  [6] lifecycle_1.0.5          
  [7] Rdpack_2.6.5             
  [8] mixsqp_0.3-54            
  [9] edgeR_4.8.2              
 [10] rprojroot_2.1.1          
 [11] lattice_0.22-7           
 [12] MASS_7.3-65              
 [13] backports_1.5.0          
 [14] magrittr_2.0.4           
 [15] metafor_4.8-0            
 [16] otel_0.2.0               
 [17] DBI_1.2.3                
 [18] minqa_1.2.8              
 [19] RColorBrewer_1.1-3       
 [20] abind_1.4-8              
 [21] EnvStats_3.1.0           
 [22] purrr_1.2.1              
 [23] msigdbr_25.1.1           
 [24] rmeta_3.0                
 [25] RCurl_1.98-1.17          
 [26] ggrepel_0.9.6            
 [27] pbkrtest_0.5.5           
 [28] irlba_2.3.7              
 [29] annotate_1.88.0          
 [30] commonmark_2.0.0         
 [31] DelayedMatrixStats_1.32.0
 [32] codetools_0.2-20         
 [33] DelayedArray_0.36.0      
 [34] xml2_1.5.2               
 [35] tidyselect_1.2.1         
 [36] farver_2.1.2             
 [37] lme4_1.1-38              
 [38] googledrive_2.1.2        
 [39] mathjaxr_2.0-0           
 [40] iterators_1.0.14         
 [41] systemfonts_1.3.1        
 [42] segmented_2.2-1          
 [43] tools_4.5.2              
 [44] progress_1.2.3           
 [45] ragg_1.5.0               
 [46] Rcpp_1.1.1               
 [47] zenith_1.12.0            
 [48] glue_1.8.0               
 [49] SparseArray_1.10.8       
 [50] xfun_0.56                
 [51] here_1.0.2               
 [52] dplyr_1.2.0              
 [53] withr_3.0.2              
 [54] numDeriv_2016.8-1.1      
 [55] fastmap_1.2.0            
 [56] boot_1.3-32              
 [57] litedown_0.9             
 [58] caTools_1.18.3           
 [59] digest_0.6.39            
 [60] truncnorm_1.0-9          
 [61] R6_2.6.1                 
 [62] textshaping_1.0.4        
 [63] scattermore_1.2          
 [64] gtools_3.9.5             
 [65] markdown_2.0             
 [66] dichromat_2.0-0.1        
 [67] RSQLite_2.4.5            
 [68] RhpcBLASctl_0.23-42      
 [69] tidyr_1.3.2              
 [70] corpcor_1.6.10           
 [71] prettyunits_1.2.0        
 [72] httr_1.4.7               
 [73] S4Arrays_1.10.1          
 [74] pkgconfig_2.0.3          
 [75] gtable_0.3.6             
 [76] blob_1.3.0               
 [77] S7_0.2.1                 
 [78] XVector_0.50.0           
 [79] remaCor_0.0.20           
 [80] zigg_0.0.2               
 [81] GSEABase_1.72.0          
 [82] scales_1.4.0             
 [83] png_0.1-8                
 [84] fANCOVA_0.6-1            
 [85] reformulas_0.4.4         
 [86] ashr_2.2-63              
 [87] knitr_1.51               
 [88] reshape2_1.4.5           
 [89] rjson_0.2.23             
 [90] curl_7.0.0               
 [91] nlme_3.1-168             
 [92] nloptr_2.2.1             
 [93] cachem_1.1.0             
 [94] stringr_1.6.0            
 [95] KernSmooth_2.23-26       
 [96] vipor_0.4.7              
 [97] metadat_1.4-0            
 [98] AnnotationDbi_1.72.0     
 [99] pillar_1.11.1            
[100] grid_4.5.2               
[101] vctrs_0.7.1              
[102] gplots_3.3.0             
[103] mashr_0.2.79             
[104] xtable_1.8-4             
[105] beeswarm_0.4.0           
[106] Rgraphviz_2.54.0         
[107] evaluate_1.0.5           
[108] KEGGgraph_1.70.0         
[109] invgamma_1.2             
[110] magick_2.9.0             
[111] mvtnorm_1.3-3            
[112] cli_3.6.5                
[113] locfit_1.5-9.12          
[114] compiler_4.5.2           
[115] rlang_1.1.7              
[116] crayon_1.5.3             
[117] SQUAREM_2021.1           
[118] labeling_0.4.3           
[119] fs_1.6.6                 
[120] plyr_1.8.9               
[121] ggbeeswarm_0.7.3         
[122] stringi_1.8.7            
[123] babelgene_22.9           
[124] assertthat_0.2.1         
[125] lmerTest_3.2-0           
[126] Biostrings_2.78.0        
[127] aod_1.3.3                
[128] Matrix_1.7-4             
[129] hms_1.1.4                
[130] sparseMatrixStats_1.22.0 
[131] bit64_4.6.0-1            
[132] KEGGREST_1.50.0          
[133] statmod_1.5.1            
[134] rbibutils_2.4.1          
[135] Rfast_2.1.5.2            
[136] gridtext_0.1.5           
[137] gargle_1.6.1             
[138] broom_1.0.12             
[139] memoise_2.0.1            
[140] RcppParallel_5.1.11-1    
[141] bit_4.6.0                
[142] EnrichmentBrowser_2.40.0 
─ Session info ──────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       Positron
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-03-18
 pandoc   NA
 quarto   1.8.27 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [1] CRAN (R 4.5.0)
 annotate               1.88.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationDbi          1.72.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 aod                    1.3.3      2023-12-13 [1] CRAN (R 4.5.0)
 ashr                   2.2-63     2023-08-21 [1] CRAN (R 4.5.0)
 assertthat             0.2.1      2019-03-21 [1] CRAN (R 4.5.0)
 babelgene              22.9       2022-09-29 [1] CRAN (R 4.5.0)
 backports              1.5.0      2024-05-23 [1] CRAN (R 4.5.0)
 beeswarm               0.4.0      2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0      2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1    2025-01-16 [1] CRAN (R 4.5.0)
 bitops                 1.0-9      2024-10-03 [1] CRAN (R 4.5.0)
 blob                   1.3.0      2026-01-14 [1] CRAN (R 4.5.2)
 boot                   1.3-32     2025-08-29 [1] CRAN (R 4.5.2)
 broom                  1.0.12     2026-01-27 [1] CRAN (R 4.5.2)
 cachem                 1.1.0      2024-05-16 [1] CRAN (R 4.5.0)
 caTools                1.18.3     2024-09-04 [1] CRAN (R 4.5.0)
 cli                    3.6.5      2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3      2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 commonmark             2.0.0      2025-07-07 [1] CRAN (R 4.5.0)
 corpcor                1.6.10     2021-09-16 [1] CRAN (R 4.5.0)
 crayon                 1.5.3      2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0      2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1   2026-01-27 [1] CRAN (R 4.5.2)
 DBI                    1.2.3      2024-06-02 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 DelayedMatrixStats     1.32.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0      2026-02-03 [1] CRAN (R 4.5.2)
 dreamlet             * 1.8.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 edgeR                  4.8.2      2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 EnrichmentBrowser      2.40.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 EnvStats               3.1.0      2025-04-24 [1] CRAN (R 4.5.0)
 evaluate               1.0.5      2025-08-27 [1] CRAN (R 4.5.0)
 fANCOVA                0.6-1      2020-11-13 [1] CRAN (R 4.5.0)
 farver                 2.1.2      2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [1] CRAN (R 4.5.0)
 fs                     1.6.6      2025-04-12 [1] CRAN (R 4.5.0)
 gargle                 1.6.1      2026-01-29 [1] CRAN (R 4.5.2)
 generics             * 0.1.4      2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggbeeswarm             0.7.3      2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2      2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6      2024-09-07 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2      2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0      2024-09-30 [1] CRAN (R 4.5.0)
 googledrive            2.1.2      2025-09-10 [1] CRAN (R 4.5.0)
 gplots                 3.3.0      2025-11-30 [1] CRAN (R 4.5.2)
 graph                  1.88.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 gridExtra            * 2.3        2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5      2022-09-16 [1] CRAN (R 4.5.0)
 GSEABase               1.72.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 gtable                 0.3.6      2024-10-25 [1] CRAN (R 4.5.0)
 gtools                 3.9.5      2023-11-20 [1] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [1] CRAN (R 4.5.0)
 hms                    1.1.4      2025-10-17 [1] CRAN (R 4.5.0)
 httr                   1.4.7      2023-08-15 [1] CRAN (R 4.5.0)
 invgamma               1.2        2025-07-02 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [1] CRAN (R 4.5.2)
 iterators              1.0.14     2022-02-05 [1] CRAN (R 4.5.0)
 jaffelab             * 0.99.34    2026-02-24 [1] Github (LieberInstitute/jaffelab@e7a26ca)
 KEGGgraph              1.70.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KEGGREST               1.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KernSmooth             2.23-26    2025-01-01 [1] CRAN (R 4.5.2)
 knitr                  1.51       2025-12-20 [1] CRAN (R 4.5.2)
 labeling               0.4.3      2023-08-29 [1] CRAN (R 4.5.0)
 lattice                0.22-7     2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5      2026-01-08 [1] CRAN (R 4.5.2)
 limma                * 3.66.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 litedown               0.9        2025-12-18 [1] CRAN (R 4.5.2)
 lme4                   1.1-38     2025-12-02 [1] CRAN (R 4.5.2)
 lmerTest               3.2-0      2026-01-13 [1] CRAN (R 4.5.2)
 locfit                 1.5-9.12   2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0      2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [1] CRAN (R 4.5.0)
 markdown               2.0        2025-03-23 [1] CRAN (R 4.5.0)
 mashr                  0.2.79     2023-10-18 [1] CRAN (R 4.5.0)
 MASS                   7.3-65     2025-02-28 [1] CRAN (R 4.5.2)
 mathjaxr               2.0-0      2025-12-01 [1] CRAN (R 4.5.2)
 Matrix                 1.7-4      2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1      2021-11-26 [1] CRAN (R 4.5.0)
 metadat                1.4-0      2025-02-04 [1] CRAN (R 4.5.0)
 metafor                4.8-0      2025-01-28 [1] CRAN (R 4.5.0)
 minqa                  1.2.8      2024-08-17 [1] CRAN (R 4.5.0)
 mixsqp                 0.3-54     2023-12-20 [1] CRAN (R 4.5.0)
 msigdbr                25.1.1     2025-07-21 [1] CRAN (R 4.5.2)
 mvtnorm                1.3-3      2025-01-10 [1] CRAN (R 4.5.0)
 nlme                   3.1-168    2025-03-31 [1] CRAN (R 4.5.2)
 nloptr                 2.2.1      2025-03-17 [1] CRAN (R 4.5.0)
 numDeriv               2016.8-1.1 2019-06-06 [1] CRAN (R 4.5.0)
 otel                   0.2.0      2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1     2026-01-08 [1] CRAN (R 4.5.2)
 pbkrtest               0.5.5      2025-07-18 [1] CRAN (R 4.5.0)
 pillar                 1.11.1     2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [1] CRAN (R 4.5.0)
 plyr                   1.8.9      2023-10-02 [1] CRAN (R 4.5.0)
 png                    0.1-8      2022-11-29 [1] CRAN (R 4.5.0)
 prettyunits            1.2.0      2023-09-24 [1] CRAN (R 4.5.0)
 progress               1.2.3      2023-12-06 [1] CRAN (R 4.5.0)
 purrr                  1.2.1      2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1      2025-02-15 [1] CRAN (R 4.5.0)
 rafalib              * 1.0.4      2025-04-08 [1] CRAN (R 4.5.0)
 ragg                   1.5.0      2025-09-02 [1] CRAN (R 4.5.0)
 rbibutils              2.4.1      2026-01-21 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3      2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [1] CRAN (R 4.5.2)
 RcppParallel           5.1.11-1   2025-08-27 [1] CRAN (R 4.5.0)
 RCurl                  1.98-1.17  2025-03-22 [1] CRAN (R 4.5.0)
 Rdpack                 2.6.5      2026-01-23 [1] CRAN (R 4.5.2)
 reformulas             0.4.4      2026-02-02 [1] CRAN (R 4.5.2)
 remaCor                0.0.20     2025-08-20 [1] CRAN (R 4.5.0)
 reshape2               1.4.5      2025-11-12 [1] CRAN (R 4.5.0)
 Rfast                  2.1.5.2    2025-10-10 [1] CRAN (R 4.5.0)
 Rgraphviz              2.54.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 RhpcBLASctl            0.23-42    2023-02-11 [1] CRAN (R 4.5.0)
 rjson                  0.2.23     2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7      2026-01-09 [1] CRAN (R 4.5.2)
 rmeta                  3.0        2018-03-20 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1      2025-08-26 [1] CRAN (R 4.5.0)
 RSQLite                2.4.5      2025-11-30 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1      2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0      2025-04-24 [1] CRAN (R 4.5.0)
 scattermore            1.2        2023-06-12 [1] CRAN (R 4.5.0)
 segmented              2.2-1      2026-01-29 [1] CRAN (R 4.5.2)
 Seqinfo              * 1.0.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 sparseMatrixStats      1.22.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 SpatialExperiment    * 1.20.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SQUAREM                2021.1     2021-01-13 [1] CRAN (R 4.5.0)
 statmod                1.5.1      2025-10-09 [1] CRAN (R 4.5.0)
 stringi                1.8.7      2025-03-27 [1] CRAN (R 4.5.0)
 stringr                1.6.0      2025-11-04 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 systemfonts            1.3.1      2025-10-01 [1] CRAN (R 4.5.0)
 textshaping            1.0.4      2025-10-10 [1] CRAN (R 4.5.0)
 tibble                 3.3.1      2026-01-11 [1] CRAN (R 4.5.2)
 tidyr                  1.3.2      2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [1] CRAN (R 4.5.0)
 truncnorm              1.0-9      2023-03-20 [1] CRAN (R 4.5.0)
 variancePartition    * 1.40.1     2025-12-11 [1] Bioconductor 3.22 (R 4.5.2)
 vctrs                  0.7.1      2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7      2023-12-18 [1] CRAN (R 4.5.0)
 withr                  3.0.2      2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56       2026-01-18 [1] CRAN (R 4.5.2)
 XML                    3.99-0.20  2025-11-08 [1] CRAN (R 4.5.0)
 xml2                   1.5.2      2026-01-17 [1] CRAN (R 4.5.2)
 xtable                 1.8-4      2019-04-21 [1] CRAN (R 4.5.0)
 XVector                0.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 zenith                 1.12.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 zigg                   0.0.2      2025-02-07 [1] CRAN (R 4.5.0)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────