---
title: "02-Tabulate_CellNeighborBanksyLabels"
author: "Bernie Mulvey"
date: "06-22-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(sf)
library(sfheaders)
## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##


## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallel)


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```


load table of all cells' coordinates
load banksy label assignments + cluster annotations
```{r}
cellcoord <- fread("processed-data_newseg/14_CellTypeProximity/retainedCellBoundaries.txt.gz")

cluslabels <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_assignmentsMergeClusts.txt")
anno <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_mergeClusts_Annots.txt")
anno[,anno:=gsub(anno,pattern=" ",replacement="_")]
anno[,anno:=gsub(anno,pattern="\\.",replacement="")]
cluslabels <- merge.data.table(cluslabels,anno,by="clusid")

## drop cells in the discarded doublet-ish clluster including from the coordinate table
cluslabels <- cluslabels[anno!="DISCARD"]
cellcoord <- cellcoord[cell_id %in% cluslabels$rn]
cellcoord <- merge.data.table(cellcoord,cluslabels,by.x="cell_id",by.y="rn")
cellcoord[,clusid:=NULL]

rm(anno,cluslabels)
gc()
```

## split into individual sample coordinate tables; use SF to encode the polygons (cells) by their vertices
```{r}
cellcoord[,brnum:=gsub(cell_id,pattern="^(.*)_.*$",replacement="\\1")]
cellcoordlist <- split(cellcoord,cellcoord$brnum)
rm(cellcoord)

## make sf objects of query cells (LCNE) and all others
cellcoord_sfs <- lapply(cellcoordlist,function(dt){
  lcsf <- sf_polygon(
    obj=dt[anno=="LCNE"],
    x="vertex_x",
    y="vertex_y",
    polygon_id="cell_id",
    keep=TRUE
  )
  
  nonlcsf <- sf_polygon(
  obj = dt[anno!="LCNE"],
  x = "vertex_x",
  y = "vertex_y",
  polygon_id = "cell_id",
  keep = TRUE # retain the annotation column as well
  )
  
  retlist <- list(lcnes=lcsf,others=nonlcsf)
  return(retlist)
})

rm(cellcoordlist)
gc(full=T)

## calculate how many cells within 10, 20, 30, 40, 50, 100, 200, 300 units of each LCNE cell
dists <- c(10,20,30,40,50,100,150,200,300)

setDTthreads(1,restore_after_fork = FALSE)

## this is surprisingly memory efficient. much faster if the parallelization is across the distance thresholds than if parallelizing across samples.
anses <- lapply(cellcoord_sfs,function(x){
  distans <- mclapply(dists,mc.cores=9,function(t){
    setDTthreads(1,restore_after_fork=FALSE)
    ## get the row indices of each LC cell and of non-LC neighbors within distance t
    ans <- st_is_within_distance(x=x$lcnes,y=x$others,dist = t)
    ## result is a list of other cell row indices, each list corresp to one lcne row idx
    ## make each list entry into a one column data.table as an intermediate step to combining them all into one table of neighbors with index information retained 
    ans2 <- lapply(ans,as.data.table)
    ## concatenate all the other-cell-index one-column data tables, with a second column indicating the row index (list entry #) of the corresponding LC cell
    ans2 <- rbindlist(ans2,idcol="lc_cell_idx")
    setnames(ans2,2,"other_cell_idx")
    
    ## for ease of use, fetch the SPE cell_id for each LC cell from the sf object
    ans2[,lc_cell_id:=x$lcnes$cell_id[lc_cell_idx]]
    ## and likewise for the found neighboring cells
    ans2[,other_cell_id:=x$others$cell_id[other_cell_idx]]
    ## then get the annotation assigned to each of the non-LC cells found within current distance thresh
    ans2[,other_cell_label:=x$others$anno[other_cell_idx]]
    ## tabulate the number of each kind of annotation found as a neighbor to that LC neuron at this thresh--this is our result table for this distance threshold
    ans3 <- ans2[,.N,by=c("lc_cell_id","other_cell_label")]
    return(ans3)
  })
  names(distans) <- paste0(dists,"u") # make character to safely propagate into a column
  # concatenate tabulated cells-of-annotation-type-within-distance results for all LC cells at all distance thresholds, retaining info re: the distance threshold applied
  alldistans <- rbindlist(distans,idcol="dist_thresh")
  alldistans[,dist_thresh:=as.numeric(gsub(dist_thresh,pattern="u",replacement=""))]
  ## return our object: one table for a sample, with each LC-NE cell and the number of nearby cells of each label type at each of several distance thresholds.
  return(alldistans)
})
alldistans2 <- rbindlist(anses,idcol="brnum")
length(unique(alldistans2$lc_cell_id)) # 21334, consistent with number of cells getting LCNE label from banksy, so we good.
```

### save after quick peep to make sure it looks ok
```{r}
fwrite(alldistans2,"processed-data_newseg/14_CellTypeProximity/02-CellTypes_within_9distances__ofEachLCNEcell_by_brnum.txt.gz",sep='\t',quote=F,row.names=F,col.names=T)
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] parallelly_1.47.0           colorout_1.3-3              sfheaders_0.4.5            
 [4] sf_1.0-24                   SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
 [7] SummarizedExperiment_1.40.0 Biobase_2.70.0              GenomicRanges_1.62.1       
[10] Seqinfo_1.0.0               IRanges_2.44.0              S4Vectors_0.48.1           
[13] BiocGenerics_0.56.0         generics_0.1.4              MatrixGenerics_1.22.0      
[16] matrixStats_1.5.0           gridExtra_2.3               ggtext_0.1.2               
[19] ggplot2_4.0.2               data.table_1.18.4          

loaded via a namespace (and not attached):
 [1] gtable_0.3.6        rjson_0.2.23        xfun_0.56           lattice_0.22-7     
 [5] vctrs_0.7.1         tools_4.5.2         tibble_3.3.1        proxy_0.4-29       
 [9] R.oo_1.27.1         pkgconfig_2.0.3     Matrix_1.7-4        KernSmooth_2.23-26 
[13] RColorBrewer_1.1-3  S7_0.2.1            lifecycle_1.0.5     compiler_4.5.2     
[17] farver_2.1.2        htmltools_0.5.9     class_7.3-23        yaml_2.3.12        
[21] pillar_1.11.1       R.utils_2.13.0      rsconnect_1.7.0     classInt_0.4-11    
[25] DelayedArray_0.36.1 sessioninfo_1.2.3   magick_2.9.0        abind_1.4-8        
[29] tidyselect_1.2.1    digest_0.6.39       dplyr_1.2.0         rprojroot_2.1.1    
[33] fastmap_1.2.0       grid_4.5.2          here_1.0.2          cli_3.6.5          
[37] SparseArray_1.10.8  magrittr_2.0.4      S4Arrays_1.10.1     dichromat_2.0-0.1  
[41] e1071_1.7-17        withr_3.0.2         scales_1.4.0        rmarkdown_2.30     
[45] XVector_0.50.0      otel_0.2.0          R.methodsS3_1.8.2   evaluate_1.0.5     
[49] knitr_1.51          rlang_1.1.7         gridtext_0.1.5      Rcpp_1.1.1         
[53] glue_1.8.0          DBI_1.2.3           xml2_1.5.2          rstudioapi_0.18.0  
[57] R6_2.6.1            units_1.0-0        
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-22
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────────────────
 package              * version date (UTC) lib source
 abind                  1.4-8   2024-09-12 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 class                  7.3-23  2025-01-01 [1] CRAN (R 4.5.2)
 classInt               0.4-11  2025-01-08 [1] CRAN (R 4.5.0)
 cli                    3.6.5   2025-04-23 [1] CRAN (R 4.5.0)
 colorout             * 1.3-3   2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.4  2026-05-06 [1] CRAN (R 4.5.2)
 DBI                    1.2.3   2024-06-02 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.1  2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1 2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39  2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0   2026-02-03 [1] CRAN (R 4.5.2)
 e1071                  1.7-17  2025-12-18 [1] CRAN (R 4.5.2)
 evaluate               1.0.5   2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2   2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4   2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1  2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 ggplot2              * 4.0.2   2026-02-03 [1] CRAN (R 4.5.2)
 ggtext               * 0.1.2   2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0   2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3     2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5   2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6   2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2   2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9   2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KernSmooth             2.23-26 2025-01-01 [1] CRAN (R 4.5.2)
 knitr                  1.51    2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7  2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5   2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0   2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4   2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4   2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0   2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0   2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.47.0  2026-04-17 [1] CRAN (R 4.5.2)
 pillar                 1.11.1  2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.5.0)
 proxy                  0.4-29  2025-12-29 [1] CRAN (R 4.5.2)
 R.methodsS3            1.8.2   2022-06-13 [1] CRAN (R 4.5.0)
 R.oo                   1.27.1  2025-05-02 [1] CRAN (R 4.5.0)
 R.utils                2.13.0  2025-02-24 [1] CRAN (R 4.5.0)
 R6                     2.6.1   2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3   2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1   2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23  2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7   2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30    2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1   2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0   2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0  2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1  2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1  2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1   2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0   2025-04-24 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3   2025-02-05 [1] CRAN (R 4.5.0)
 sf                   * 1.0-24  2026-01-13 [1] CRAN (R 4.5.2)
 sfheaders            * 0.4.5   2025-11-25 [1] CRAN (R 4.5.2)
 SingleCellExperiment * 1.32.0  2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8  2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0  2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0  2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1   2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.5.0)
 units                  1.0-0   2025-10-09 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1   2026-01-23 [1] CRAN (R 4.5.2)
 withr                  3.0.2   2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56    2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2   2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12  2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────────────────