---
title: "03-CellsOfType_as_pctage_or_per_area_of_wdLCdomains"
author: "Bernie Mulvey"
date: "06-22-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(sf)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##


## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

since LCNE cells were identified using NM segmentation and kept overlapping NM and DAPI regions as distinct "nuclei" (with overlap zone going to the DAPI boundary), discontinuous NM segs or an NM seg and a DAPI seg within one biological LC neuron could result in 2 or more unique xenium cells. thus it is possible that we are double-counting neighbor cells for such neurons (e.g., an astrocyte equidistant from an NM body and a nucleus that are parts of the same LC neuron would be flagged as a neighbor for each of two Xenium LCNE "cells"). this is a problem specific to our approach to capturing LC-NE cells in the Xenium samples, whereas other cell types should not have this problem.

a robust second approach will therefore be to determine how many of each non-LCNE cell type fall within the Xenium LC domain. These could be either normalized to the percentage of total cells in the domain (which might vary depending on tissue quality/area of the LC, which we don't have info on), or area of the domain.

## load cell-domain assignments, banksy labels, and tabulate proportion of each label within  domain. include the DISCARD cluster here, more for the purposes of calculating an accurate denominator than because we're interested in knowing how many of these artifactual doublet xenium cells are within LC. 
```{r}
bsys <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_assignmentsMergeClusts.txt")
anno <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_mergeClusts_Annots.txt")
bsys <- merge.data.table(bsys,anno,by="clusid")

wdlcs <- fread("processed-data_newseg/04_snSeqLabelTransfer/06-wdLCsmoothings_LUT.txt")
wdlcs <- wdlcs[,.(rn,domain12412,domain8492)] # indicates whether a given cell is in LC domain
setnames(wdlcs,c(2,3),c("wd_domain12412","wd_domain8492"))

srtlcs <- fread("processed-data_newseg/07_visLabelTransfer/05-srtLCsmoothDoms_and_NMFpreds_lut.txt")
srtlcs <- srtlcs[,.(rn,domain8426,domain12199,domain16119,domain20079)]
setnames(srtlcs,c(2:5),c("srt_domain8426","srt_domain12199","srt_domain16119","srt_domain20079"))

bsys <- merge.data.table(bsys,wdlcs,by="rn")
bsys <- merge.data.table(bsys,srtlcs,by="rn")
## extract brnum (sample) from cell barcode in rn
bsys[,brnum:=gsub(rn,pattern="^(.*)_.*$",replacement="\\1")]
```

```{r}
domdefs <- c("wd_domain12412","wd_domain8492","srt_domain8426","srt_domain12199","srt_domain16119","srt_domain20079")
anses <- lapply(domdefs,function(x){
  tmp <- copy(bsys)[get(x)=="LC"]
  ans <- tmp[,.N,by=c("brnum","anno")]
  totcells <- tmp[,.N,by="brnum"]
  setnames(totcells,"N","LCdomainCellNum")
  ans <- merge.data.table(ans,totcells,by="brnum")
  ans[,prop_of_LCdomain_cells:=N/LCdomainCellNum]
  return(ans)
})
names(anses) <- domdefs

ans <- rbindlist(anses,idcol="LCDomainDef")

fwrite(ans,"processed-data_newseg/14_CellTypeProximity/03-BsyLabelProportions_in_wdDomains_and_srtDomains.txt",sep='\t',quote=F,row.names=F,col.names=T)
```

the slightly trickier one will be defining the domain areas. we'll need the cell polygon coordinates again:
```{r}
allpoly <- fread("processed-data_newseg/14_CellTypeProximity/retainedCellBoundaries.txt.gz")

domdefs <- c("wd_domain12412","wd_domain8492","srt_domain8426","srt_domain12199","srt_domain16119","srt_domain20079")

area_anses <- lapply(domdefs,function(x){
  tmp <- copy(bsys)[get(x)=="LC"]
  tmp <- merge.data.table(tmp,allpoly,by.x="rn",by.y="cell_id")
  lcdomlist <- split(tmp,tmp$brnum)
  
  ## get area of LC domain for each sample
  lcareas <- lapply(lcdomlist,function(l){
    # subset your data to the cells that make up your zone (already done)
    # Define the gap distance-this should be large enough to bridge the space between dots in the same domain zone, but small enough not to merge distinct, separate ones.
    # test values using br5415 -- should get 11 separate LC domain areas. testing this on several addnl brains with multiple discontig LC domain pieces returned the correct number of domains in each case.
    gap_distance <- 30 ## 
    # Then, apply st_union() to merge touching geometries, and st_cast() to break apart the unconnected bodies.
    lccellpolygons_sf <- sf_polygon(
      obj = l,
      x = "vertex_x",
      y = "vertex_y",
      polygon_id = "cell_id",
      keep = TRUE # Retains the label and other attributes
    )
    buffered_sf <- st_buffer(lccellpolygons_sf, dist = gap_distance)
    merged_sf <- st_union(buffered_sf)
    lczonepolys <- st_cast(merged_sf, "POLYGON")
    # retract the the buffer area we applied in Step 1 to make cells touch
    # This pulls the boundary back to meet the original cell polygons
    tight_zones_sfc <- st_buffer(lczonepolys, dist = -gap_distance)
    zones_sf <- st_sf(geometry = tight_zones_sfc)
    zones_sf$area <- st_area(zones_sf) # get the area of each lc domain zone
    lcdomtotarea <- sum(zones_sf$area) # get the total area of the lc domain zones
    return(lcdomtotarea)
  })
  lcareas <- lapply(lcareas,as.data.table)
  lcareas <- rbindlist(lcareas,idcol="brnum")
  setnames(lcareas,"V1","LCdomarea")
  
  ## get total intra-LC cells per annotation type as before
  ans <- tmp[,.N,by=c("brnum","anno")]
  
  ## append LC domain total area
  ans <- merge.data.table(ans,lcareas,by="brnum")
  
  ## calculate number of cells of type anno per area
  ans[,annoPerLCdomArea:=N/LCdomarea]
  return(ans)
})
names(area_anses) <- domdefs

saveRDS(area_anses,"processed-data_newseg/14_CellTypeProximity/03_BsyLabelCounts_per_LCDomainArea.RDS")
```


seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] parallelly_1.47.0           colorout_1.3-3              sfheaders_0.4.5            
 [4] sf_1.0-24                   SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
 [7] SummarizedExperiment_1.40.0 Biobase_2.70.0              GenomicRanges_1.62.1       
[10] Seqinfo_1.0.0               IRanges_2.44.0              S4Vectors_0.48.1           
[13] BiocGenerics_0.56.0         generics_0.1.4              MatrixGenerics_1.22.0      
[16] matrixStats_1.5.0           gridExtra_2.3               ggtext_0.1.2               
[19] ggplot2_4.0.2               data.table_1.18.4          

loaded via a namespace (and not attached):
 [1] gtable_0.3.6        rjson_0.2.23        xfun_0.56           lattice_0.22-7     
 [5] vctrs_0.7.1         tools_4.5.2         tibble_3.3.1        proxy_0.4-29       
 [9] R.oo_1.27.1         pkgconfig_2.0.3     Matrix_1.7-4        KernSmooth_2.23-26 
[13] RColorBrewer_1.1-3  S7_0.2.1            lifecycle_1.0.5     compiler_4.5.2     
[17] farver_2.1.2        htmltools_0.5.9     class_7.3-23        yaml_2.3.12        
[21] pillar_1.11.1       R.utils_2.13.0      rsconnect_1.7.0     classInt_0.4-11    
[25] DelayedArray_0.36.1 sessioninfo_1.2.3   magick_2.9.0        abind_1.4-8        
[29] tidyselect_1.2.1    digest_0.6.39       dplyr_1.2.0         rprojroot_2.1.1    
[33] fastmap_1.2.0       grid_4.5.2          here_1.0.2          cli_3.6.5          
[37] SparseArray_1.10.8  magrittr_2.0.4      S4Arrays_1.10.1     dichromat_2.0-0.1  
[41] e1071_1.7-17        withr_3.0.2         scales_1.4.0        rmarkdown_2.30     
[45] XVector_0.50.0      otel_0.2.0          R.methodsS3_1.8.2   evaluate_1.0.5     
[49] knitr_1.51          rlang_1.1.7         gridtext_0.1.5      Rcpp_1.1.1         
[53] glue_1.8.0          DBI_1.2.3           xml2_1.5.2          rstudioapi_0.18.0  
[57] R6_2.6.1            units_1.0-0        
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-22
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────────────────
 package              * version date (UTC) lib source
 abind                  1.4-8   2024-09-12 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 class                  7.3-23  2025-01-01 [1] CRAN (R 4.5.2)
 classInt               0.4-11  2025-01-08 [1] CRAN (R 4.5.0)
 cli                    3.6.5   2025-04-23 [1] CRAN (R 4.5.0)
 colorout             * 1.3-3   2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.4  2026-05-06 [1] CRAN (R 4.5.2)
 DBI                    1.2.3   2024-06-02 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.1  2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1 2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39  2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0   2026-02-03 [1] CRAN (R 4.5.2)
 e1071                  1.7-17  2025-12-18 [1] CRAN (R 4.5.2)
 evaluate               1.0.5   2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2   2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4   2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1  2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 ggplot2              * 4.0.2   2026-02-03 [1] CRAN (R 4.5.2)
 ggtext               * 0.1.2   2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0   2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3     2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5   2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6   2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2   2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9   2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KernSmooth             2.23-26 2025-01-01 [1] CRAN (R 4.5.2)
 knitr                  1.51    2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7  2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5   2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0   2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4   2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4   2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0   2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0   2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.47.0  2026-04-17 [1] CRAN (R 4.5.2)
 pillar                 1.11.1  2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.5.0)
 proxy                  0.4-29  2025-12-29 [1] CRAN (R 4.5.2)
 R.methodsS3            1.8.2   2022-06-13 [1] CRAN (R 4.5.0)
 R.oo                   1.27.1  2025-05-02 [1] CRAN (R 4.5.0)
 R.utils                2.13.0  2025-02-24 [1] CRAN (R 4.5.0)
 R6                     2.6.1   2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3   2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1   2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23  2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7   2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30    2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1   2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0   2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0  2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1  2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1  2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1   2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0   2025-04-24 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3   2025-02-05 [1] CRAN (R 4.5.0)
 sf                   * 1.0-24  2026-01-13 [1] CRAN (R 4.5.2)
 sfheaders            * 0.4.5   2025-11-25 [1] CRAN (R 4.5.2)
 SingleCellExperiment * 1.32.0  2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8  2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0  2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0  2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1   2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.5.0)
 units                  1.0-0   2025-10-09 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1   2026-01-23 [1] CRAN (R 4.5.2)
 withr                  3.0.2   2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56    2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2   2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12  2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────────────────