---
title: "02_MapMyCells_visualize"
author: "Bernie Mulvey"
date: "06-17-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(ggrastr)
library(SpatialExperiment)
library(escheR)
library(Polychrome)

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallel)


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
xlc <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

mmc <- fread("processed-data_newseg/12_MapMyCells/01-LCX_mapmycellsUploadCtscsv_10xWholeHumanBrain(CCN202210140)_HierarchicalMapping_UTC_1781663294501/01-LCX_mapmycellsUploadCtscsv_10xWholeHumanBrain(CCN202210140)_HierarchicalMapping_UTC_1781663294501.csv",skip=4)


## drop the following clusters: 
#  1:               Thalamic excitatory   563
#  4:        Deep-layer near-projecting   164
#  6:                   MGE interneuron    48
#  7:               Amygdala excitatory    22
#  8: Deep-layer corticothalamic and 6b     9
# 10:                     Bergmann glia    33
# 11:                   Mammillary body    39
# 12:               Medium spiny neuron     8
# 13:    Upper-layer intratelencephalic    17
# 14:     Eccentric medium spiny neuron     1
# 15:     Deep-layer intratelencephalic    10
# 16:         Hippocampal dentate gyrus     3
# 17:         LAMP5-LHX6 and Chandelier     1

mmc <- mmc[!(supercluster_name %in% c("Thalamic excitatory","Deep-layer near-projecting","Deep-layer corticothalamic and 6b","Deep-layer intratelencephalic","MGE interneuron","Amygdala excitatory","Bergmann glia","Mammillary body","Medium spiny neuron","Eccentric medium spiny neuron","Upper-layer intratelencephalic","Hippocampal dentate gyrus","LAMP5-LHX6 and Chandelier"))]
keepcell <- mmc$cell_id
xlc <- xlc[,keepcell]

## tidy the remaining names
#                       supercluster_name      N
#                                  <char>  <int>
#  1:                           Microglia  35833
#  2:                          Fibroblast  15961
#  3:                            Vascular  75300
#  4:                            Splatter  70180
#  5:                     Oligodendrocyte 121308
#  6:                           Astrocyte 126179
#  7:           Oligodendrocyte precursor  15685
#  8: Committed oligodendrocyte precursor   9225
#  9:                       Miscellaneous   8158
# 10:               Cerebellar inhibitory    954
# 11:         Midbrain-derived inhibitory   3055
# 12:                     CGE interneuron   1972
# 13:                           Ependymal   6095
# 14:                   Lower rhombic lip    174
# 15:                   Upper rhombic lip    707
# 16:                      Choroid plexus     78
mmc[supercluster_name=="Choroid plexus",supercluster_name:="Choroid"]
mmc[supercluster_name=="Upper rhombic lip",supercluster_name:="Upper rhomb. l."]
mmc[supercluster_name=="Lower rhombic lip",supercluster_name:="Lower rhomb. l."]
mmc[supercluster_name=="CGE interneuron",supercluster_name:="CGE iN"]
mmc[supercluster_name=="Midbrain-derived inhibitory",supercluster_name:="Midbrain-der. GABA"]
mmc[supercluster_name=="Cerebellar inhibitory",supercluster_name:="Cerebel. GABA"]
mmc[supercluster_name=="Miscellaneous",supercluster_name:="Misc"]
mmc[supercluster_name=="Committed oligodendrocyte precursor",supercluster_name:="Committed OPC"]
mmc[supercluster_name=="Oligodendrocyte precursor",supercluster_name:="OPC"]
mmc[supercluster_name=="Astrocyte",supercluster_name:="Astro"]
mmc[supercluster_name=="Oligodendrocyte",supercluster_name:="Oligo"]


mmc <- DataFrame(mmc,row.names=mmc$cell_id)[colnames(xlc),]
xlc$label <- mmc$supercluster_name

```

```{r}
unique(xlc$label)
pal <- c("#0F0","#7C2","#5A4",
         "#B40","#F00","#F90", # "#D37", "#E62","#A00","#C20",,
         "#00F",
         "#F0F","#59D","#90F",
         # "#D37","#D37",
         "#979696","#CAC9CA", #,"#ECECEC",
         "#DC3","#FF0",
         "#000",
         "#7b8418"
)

names(pal) <- c("OPC","Committed OPC","Oligo",
"Fibroblast","Vascular","Microglia",
"Misc",
"Midbrain-der. GABA","CGE iN","Cerebel. GABA",
"Lower rhomb. l.","Upper rhomb. l.",
"Ependymal","Choroid",
"Astro",
"Splatter")
stopifnot(all(names(pal) %in% unique(xlc$label)))


# names(pal) <- c(paste0("excitatory_",c(26,7,23,30)),
#                 paste0("inhibitory_",c(4,25,14,17,8,20,24)),
#                 "NE",
#                 "5HT_21",
#                 "oligodendrocytes",
#                 "OPCs_27",
#                 "astrocytes_5","astrocytes_29",
#                 "macrophages_microglia_15",
#                 "endothelial_mural_16",
#                 "neurons_ambiguous"
#                 )

xlc$label <- factor(xlc$label,levels=names(pal))
samps <- lapply(unique(xlc$brnum),function(b){xlc[,xlc$brnum==b]})

names(samps) <- unique(xlc$brnum)
rm(xlc)
gc(full=T)

plts <- mapply(s=samps,n=names(samps),SIMPLIFY=FALSE,function(s,n){
  p <- make_escheR(s,spot_size=0.5)
  p <- add_fill(p,"label",point_size=0.5)+    guides(fill=guide_legend(override.aes = list(size=2)))
  p <- p+scale_fill_manual(values=pal)
  p <- p+ggtitle(n)+theme(title=element_text(size=9))
  q <- rasterize(p,layers="Points",dpi=300,dev="cairo_png")
  return(q)
})

png("plots_newseg/12_MapMyCells/02-MapMyCells_superclusts.png",height=36,width=18,unit="in",res=150)
do.call("grid.arrange",c(plts,nrow=8))
dev.off()

```

Well, the only spatial feature we can really go on among these labels is the ependyma/subependyma (astrocyte-rich), which does look like it mapped correctly. But boy is the rest ugly. and not just because of an ad hoc palette. Nonetheless, for exploratory purposes, let's try DE retaining these labels (but overriding them if weber-divecha labeled a cell as a LC-NE.)
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics 
[5] grDevices utils     datasets  methods  
[9] base     

other attached packages:
 [1] parallelly_1.47.0          
 [2] colorout_1.3-3             
 [3] Polychrome_1.5.4           
 [4] escheR_1.10.0              
 [5] SpatialExperiment_1.20.0   
 [6] SingleCellExperiment_1.32.0
 [7] SummarizedExperiment_1.40.0
 [8] Biobase_2.70.0             
 [9] GenomicRanges_1.62.1       
[10] Seqinfo_1.0.0              
[11] IRanges_2.44.0             
[12] S4Vectors_0.48.1           
[13] BiocGenerics_0.56.0        
[14] generics_0.1.4             
[15] MatrixGenerics_1.22.0      
[16] matrixStats_1.5.0          
[17] ggrastr_1.0.2              
[18] gridExtra_2.3              
[19] ggtext_0.1.2               
[20] ggplot2_4.0.2              
[21] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] SparseArray_1.10.8   xml2_1.5.2          
 [3] lattice_0.22-7       magrittr_2.0.4      
 [5] evaluate_1.0.5       grid_4.5.2          
 [7] RColorBrewer_1.1-3   Matrix_1.7-4        
 [9] rprojroot_2.1.1      scales_1.4.0        
[11] abind_1.4-8          cli_3.6.5           
[13] rlang_1.1.7          scatterplot3d_0.3-44
[15] XVector_0.50.0       DelayedArray_0.36.1 
[17] withr_3.0.2          otel_0.2.0          
[19] S4Arrays_1.10.1      ggbeeswarm_0.7.3    
[21] tools_4.5.2          dplyr_1.2.0         
[23] colorspace_2.1-2     here_1.0.2          
[25] vctrs_0.7.1          R6_2.6.1            
[27] magick_2.9.0         lifecycle_1.0.5     
[29] vipor_0.4.7          pkgconfig_2.0.3     
[31] beeswarm_0.4.0       pillar_1.11.1       
[33] gtable_0.3.6         glue_1.8.0          
[35] Rcpp_1.1.1           xfun_0.56           
[37] tibble_3.3.1         tidyselect_1.2.1    
[39] knitr_1.51           dichromat_2.0-0.1   
[41] rjson_0.2.23         farver_2.1.2        
[43] labeling_0.4.3       compiler_4.5.2      
[45] S7_0.2.1             gridtext_0.1.5      
> sessioninfo::session_info()
─ Session info ──────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       Positron
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-17
 pandoc   NA
 quarto   1.8.27 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 colorspace             2.1-2    2025-09-22 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.1   2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 escheR               * 1.10.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 ggbeeswarm             0.7.3    2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggrastr              * 1.0.2    2023-06-01 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 labeling               0.4.3    2023-08-29 [1] CRAN (R 4.5.0)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.47.0   2026-04-17 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 Polychrome           * 1.5.4    2025-04-06 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1   2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 scatterplot3d          0.3-44   2023-05-05 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.5.0)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.