---
title: "04b-MakeLabelXfer_PlusSmoothLC_tab"
format: html
date: "2026-03-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost='localhost')


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = '#000000'), axis.title.x = element_text(size = 9,color = '#000000'), axis.text.y = element_text(size = 8,color = '#000000'), axis.title.y = element_text(size =9,color = '#000000'), plot.title = element_markdown(size = 10,hjust=0.5,color = '#000000'), strip.text = element_text(size=11), legend.text = element_text(size=7,color = '#000000'), legend.title = element_text(size=8,hjust=0.5,color = '#000000'),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position='plot'))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

0.15 0.12
0.1 0.36
0.05 0.36
```{r}
smooves <- readRDS("processed-data/05_LabelTransfer/03-LCknnsmoothing_step2k75.RDS")
smoove0536 <- smooves$k1thresh_0.05
smoove1036 <- smooves$k1thresh_0.1
smoove1512 <- smooves$k1thresh_0.15

smoove0536[,LCsmooth0536:=ifelse(lcpred.smooth1.k20>=0.05 & lcpred.smooth2.k75>=0.36,"LCdomain","Other")]
smoove1036[,LCsmooth1036:=ifelse(lcpred.smooth1.k20>=0.1 & lcpred.smooth2.k75>=0.36,"LCdomain","Other")]
smoove1512[,LCsmooth1512:=ifelse(lcpred.smooth1.k20>=0.15 & lcpred.smooth2.k75>=0.12,"LCdomain","Other")]

labelxfers <- readRDS("processed-data/05_LabelTransfer/02-VistoXen_NMFlabelxfer_tol1e8_k63_alph0.RDS")
preds <- lapply(labelxfers$targets,function(x){
    y <- as.data.table(cbind(rownames(colData(x)),as.character(x$nmf_preds)))
    setnames(y,c("rn","nmfpred"))
    return(y)
})
preds <- rbindlist(preds)

preds <- merge.data.table(preds,smoove0536[,.(rn,LCsmooth0536)],by="rn")
preds <- merge.data.table(preds,smoove1036[,.(rn,LCsmooth1036)],by="rn")
preds <- merge.data.table(preds,smoove1512[,.(rn,LCsmooth1512)],by="rn")

nrow(preds[LCsmooth0536=="LCdomain"&nmfpred=="LC"])/nrow(preds[nmfpred=="LC"]) # 64.4% of LC cells in domain
nrow(preds[LCsmooth1036=="LCdomain"&nmfpred=="LC"])/nrow(preds[nmfpred=="LC"]) # 55.8% of LC cells in domain
nrow(preds[LCsmooth1512=="LCdomain"&nmfpred=="LC"])/nrow(preds[nmfpred=="LC"]) # 48.3% of LC cells in domain

### create a set of overriding transferred label columns, where a cell keeps its transferred label unless its in the LC domain
### for handling out-of-smoothed-domain LC cells, change those labels to be clear about what they are
preds[,LCsmooth0536_orxfer:=ifelse(LCsmooth0536=="LCdomain","LCdomain",nmfpred)]
preds[LCsmooth0536_orxfer=="LC",LCsmooth0536_orxfer:="LC_outOfDomain"]
preds[,LCsmooth1036_orxfer:=ifelse(LCsmooth1036=="LCdomain","LCdomain",nmfpred)]
preds[LCsmooth1036_orxfer=="LC",LCsmooth1036_orxfer:="LC_outOfDomain"]
preds[,LCsmooth1512_orxfer:=ifelse(LCsmooth1512=="LCdomain","LCdomain",nmfpred)]
preds[LCsmooth1512_orxfer=="LC",LCsmooth1512_orxfer:="LC_outOfDomain"]

fwrite(preds,"processed-data/05_LabelTransfer/04d-TransferredLabels_and_SmoothLCdomain_assigns.txt",sep='\t',quote=F,row.names=F)
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices
[5] utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0        
 [2] parallelly_1.46.1          
 [3] SpatialExperiment_1.20.0   
 [4] SingleCellExperiment_1.32.0
 [5] SummarizedExperiment_1.40.0
 [6] Biobase_2.70.0             
 [7] GenomicRanges_1.62.1       
 [8] Seqinfo_1.0.0              
 [9] IRanges_2.44.0             
[10] S4Vectors_0.48.0           
[11] BiocGenerics_0.56.0        
[12] generics_0.1.4             
[13] MatrixGenerics_1.22.0      
[14] matrixStats_1.5.0          
[15] gridExtra_2.3              
[16] ggtext_0.1.2               
[17] ggplot2_4.0.2              
[18] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] SparseArray_1.10.8  xml2_1.5.2         
 [3] lattice_0.22-7      magrittr_2.0.4     
 [5] evaluate_1.0.5      grid_4.5.2         
 [7] RColorBrewer_1.1-3  R.oo_1.27.1        
 [9] rprojroot_2.1.1     Matrix_1.7-4       
[11] R.utils_2.13.0      scales_1.4.0       
[13] codetools_0.2-20    abind_1.4-8        
[15] cli_3.6.5           rlang_1.1.7        
[17] XVector_0.50.0      R.methodsS3_1.8.2  
[19] withr_3.0.2         DelayedArray_0.36.0
[21] otel_0.2.0          S4Arrays_1.10.1    
[23] parallel_4.5.2      tools_4.5.2        
[25] dplyr_1.2.0         here_1.0.2         
[27] vctrs_0.7.1         R6_2.6.1           
[29] magick_2.9.0        lifecycle_1.0.5    
[31] pkgconfig_2.0.3     pillar_1.11.1      
[33] gtable_0.3.6        glue_1.8.0         
[35] Rcpp_1.1.1          xfun_0.56          
[37] tibble_3.3.1        tidyselect_1.2.1   
[39] knitr_1.51          dichromat_2.0-0.1  
[41] rjson_0.2.23        farver_2.1.2       
[43] compiler_4.5.2      S7_0.2.1           
[45] gridtext_0.1.5     
> sessioninfo::session_info()
─ Session info ──────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       Positron
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-12
 pandoc   NA
 quarto   1.8.27 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggplot2              * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 R.methodsS3            1.8.2    2022-06-13 [1] CRAN (R 4.5.0)
 R.oo                   1.27.1   2025-05-02 [1] CRAN (R 4.5.0)
 R.utils                2.13.0   2025-02-24 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────