---
title: "Samui_SPE"
author: "Bernie Mulvey"
date: "07-09-2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##


## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
spe <- readRDS("processed-data_newseg/02_QC/05a-spe_filtered_genetargonly_log-and-nonlog-norms_withLowresHEs.RDS")

colData(spe) <- colData(spe)[,c(17,18,1:3,13:14,19:25)]
tmpcd <- as.data.table(colData(spe),keep.rownames=T)

### domain and label transfer assignments
srtdomlut <- fread("processed-data_newseg/07_visLabelTransfer/05-srtLCsmoothDoms_and_NMFpreds_lut.txt")
names(srtdomlut)
setnames(srtdomlut,c(2:6),c("SmoothedLC8_SRTxfer","SmoothedLC12_SRTxfer","SmoothedLC16_SRTxfer","SmoothedLC20_SRTxfer","Xfered_SRT_Label"))

wddomlut <- fread("processed-data_newseg/04_snSeqLabelTransfer/06-wdLCsmoothings_LUT.txt")
setnames(wddomlut,c(2:4),c("Xfered_snRNA_Label","SmoothedLC12_snRNAxfer","SmoothedLC8_snRNAxfer"))

domlut <- merge.data.table(srtdomlut,wddomlut,by="rn")


bsy <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_assignmentsMergeClusts.txt")
an <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_mergeClusts_Annots.txt")
bsy <- merge.data.table(bsy,an,by="clusid")
bsy[anno=="DISCARD",anno:="DROPPED_CLUSTER"]
bsy[,clusid:=NULL]
luts <- merge.data.table(domlut,bsy,by="rn")

setnames(luts,"anno","BanksyCluster_Label")
tmpcd <- merge.data.table(tmpcd,luts,by="rn")
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(spe),]
tmpcd$rn <- NULL
colData(spe) <- tmpcd

spe <- scater::computeLibraryFactors(spe)
spe <- scater::logNormCounts(spe)

saveRDS(spe,"processed-data_newseg/16_app-prep/01-SamuiSPE.RDS")
```

```{r}
sessionInfo()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0         parallelly_1.47.0           colorout_1.3-3             
 [4] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
 [7] Biobase_2.70.0              GenomicRanges_1.62.1        Seqinfo_1.0.0              
[10] IRanges_2.44.0              S4Vectors_0.48.1            BiocGenerics_0.56.0        
[13] generics_0.1.4              MatrixGenerics_1.22.0       matrixStats_1.5.0          
[16] gridExtra_2.3               ggtext_0.1.2                ggplot2_4.0.2              
[19] data.table_1.18.4          

loaded via a namespace (and not attached):
 [1] beeswarm_0.4.0      gtable_0.3.6        rjson_0.2.23        xfun_0.56          
 [5] ggrepel_0.9.6       lattice_0.22-7      vctrs_0.7.1         tools_4.5.2        
 [9] parallel_4.5.2      tibble_3.3.1        BiocNeighbors_2.4.0 pkgconfig_2.0.3    
[13] Matrix_1.7-4        RColorBrewer_1.1-3  S7_0.2.1            lifecycle_1.0.5    
[17] compiler_4.5.2      farver_2.1.2        codetools_0.2-20    vipor_0.4.7        
[21] htmltools_0.5.9     yaml_2.3.12         pillar_1.11.1       rsconnect_1.7.0    
[25] DelayedArray_0.36.1 viridis_0.6.5       magick_2.9.0        abind_1.4-8        
[29] rsvd_1.0.5          tidyselect_1.2.1    digest_0.6.39       BiocSingular_1.26.0
[33] dplyr_1.2.0         rprojroot_2.1.1     fastmap_1.2.0       grid_4.5.2         
[37] here_1.0.2          cli_3.6.5           SparseArray_1.10.8  scater_1.38.1      
[41] magrittr_2.0.4      S4Arrays_1.10.1     dichromat_2.0-0.1   withr_3.0.2        
[45] scales_1.4.0        ggbeeswarm_0.7.3    rmarkdown_2.30      XVector_0.50.0     
[49] otel_0.2.0          ScaledMatrix_1.18.0 beachmat_2.26.0     evaluate_1.0.5     
[53] knitr_1.51          viridisLite_0.4.3   irlba_2.3.7         rlang_1.1.7        
[57] gridtext_0.1.5      Rcpp_1.1.1          glue_1.8.0          scuttle_1.20.0     
[61] xml2_1.5.2          rstudioapi_0.18.0   R6_2.6.1   