---
title: "01-MakeDGEs_wdLCdomain_Banksyclusts"
author: "Bernie Mulvey"
date: "06-21-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(dreamlet)
library(edgeR)
library(limma)
library(SpatialExperiment)
library(WeberDivechaLCdata)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##


greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

DE using the nmfpred labels directly--disregarding the multi-cell-type LC domain here.

```{r}
xlc <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")
tmpcd <- as.data.table(colData(xlc),keep.rownames=T)

## we need the annotated banksy labels and weber-divecha LC domains (for intra-domain cell type DE)
bsy <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_assignmentsMergeClusts.txt")
anno <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_mergeClusts_Annots.txt")
bsy <- merge.data.table(bsy,anno,by="clusid")
bsy[,anno:=gsub(anno,pattern=" ",replacement="_")]
bsy[,anno:=gsub(anno,pattern="\\.",replacement="")]
tmpcd <- merge.data.table(bsy[,.(rn,anno)],tmpcd,by="rn")

wddom <- fread("processed-data_newseg/04_snSeqLabelTransfer/06-wdLCsmoothings_LUT.txt")
tmpcd <- merge.data.table(tmpcd,wddom[,.(rn,domain12412,domain8492)],by="rn")

## remove discard cluster
tmpcd <- tmpcd[anno!="DISCARD"]
xlc <- xlc[,tmpcd$rn]

tmpcd[,APOEsimple:=ifelse(APOE %in% c("E4hom","E4het"),"E4","E2")]
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(xlc),]
tmpcd$rn <- NULL

colData(xlc) <- tmpcd

## three SPEs: xlc12, xlc8, xlc (all)
xlc12 <- xlc[,xlc$domain12412=="LC"]
xlc8 <- xlc[,xlc$domain8492=="LC"]
spes <- list("domain12412"=xlc12,"domain8492"=xlc8,"wholetissue"=xlc)
rm(xlc,xlc12,xlc8,bsy,anno,wddom)
gc(full=T)
```

```{r}
dgels <- lapply(spes,function(l){
  ag <- dreamlet::aggregateToPseudoBulk(l,assay="counts",sample_id="brnum",cluster_id = "anno",BPPARAM = MulticoreParam(8))
    
  agcd <- as.data.table(colData(ag),keep.rownames=T)
  setnames(agcd,"rn","brnum")
  
  agn <- as.data.table(ag@int_colData$n_cells,keep.rownames=T)
  setnames(agn,"rn","clus")
  agn <- melt.data.table(agn,id.vars = "clus")
  setnames(agn,"variable","brnum")
  agn[,brnum:=as.character(brnum)]
  agn[,clus_brnum:=paste0(clus,"_",brnum)]
  agn <- agn[value>=10] ## these are the samples we want to keep to start: 10+ cells for domain type
  setnames(agn,"value","ncells")
  
    
  agm <- as.list(assays(ag))
  agc <- unlist(mapply(x=agm,n=names(agm),SIMPLIFY=FALSE,function(x,n){
    x <- x[,colnames(x) %in% agn[clus==n,brnum]]
    if(is.vector(x)){return(0)} else {
    sum(greaterOrEqual(colSums(x),2000))}})) ## number of pseudobulks with 10+ cells also having 2000+ counts ## number of pseudobulks with 2000+ counts in at least 10 samples
  ## only keep those where at least 10 samples meet that
  kpb <- names(agc[agc>=10])
  agn <- agn[clus %in% kpb]
  agm <- agm[kpb]
    
    ## now we want to make one giant sample matrix with the remaining samples
  agm.filt <- mapply(a=agm,n=names(agm),SIMPLIFY=FALSE,function(a,n){
    a <- a[,colnames(a) %in% agn[clus==n]$brnum]
    colnames(a) <- paste0(n,"_",colnames(a))
    return(a)
  })
  agm.filt <- do.call("cbind",agm.filt)
    
  # make a per cluster-donor samples table
  agn <- merge.data.table(agn,agcd[,.(brnum,sample_id,Sex,APOE,APOEsimple,YRI,Age)],by="brnum")
  stopifnot(all(agn$clus_brnum %in% colnames(agm.filt)))
  agn <- as.data.frame(agn,row.names=agn$clus_brnum)[colnames(agm.filt),]
  dge <- DGEList(counts=agm.filt, samples=agn)
  return(dge)
})
names(dgels) <- c("domain12412","domain8492","wholetissue")

saveRDS(dgels,"processed-data_newseg/13_BanksyDE/01-BanksyLabels_wdLCdomains_DGELists.RDS")
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] parallelly_1.47.0           colorout_1.3-3             
 [3] WeberDivechaLCdata_1.12.0   ExperimentHub_3.0.0        
 [5] AnnotationHub_4.0.0         BiocFileCache_3.0.0        
 [7] dbplyr_2.5.1                SpatialExperiment_1.20.0   
 [9] edgeR_4.8.2                 dreamlet_1.8.0             
[11] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
[13] Biobase_2.70.0              GenomicRanges_1.62.1       
[15] Seqinfo_1.0.0               IRanges_2.44.0             
[17] S4Vectors_0.48.1            BiocGenerics_0.56.0        
[19] generics_0.1.4              MatrixGenerics_1.22.0      
[21] matrixStats_1.5.0           variancePartition_1.40.1   
[23] BiocParallel_1.44.0         limma_3.66.0               
[25] ggplot2_4.0.2               data.table_1.18.4          

loaded via a namespace (and not attached):
  [1] splines_4.5.2             filelock_1.0.3            bitops_1.0-9             
  [4] tibble_3.3.1              graph_1.88.1              XML_3.99-0.20            
  [7] httr2_1.2.2               lifecycle_1.0.5           Rdpack_2.6.5             
 [10] mixsqp_0.3-54             rprojroot_2.1.1           lattice_0.22-7           
 [13] MASS_7.3-65               backports_1.5.0           magrittr_2.0.4           
 [16] metafor_4.8-0             rmarkdown_2.30            yaml_2.3.12              
 [19] otel_0.2.0                DBI_1.2.3                 minqa_1.2.8              
 [22] RColorBrewer_1.1-3        abind_1.4-8               EnvStats_3.1.0           
 [25] purrr_1.2.1               rmeta_3.0                 msigdbr_25.1.1           
 [28] RCurl_1.98-1.17           rappdirs_0.3.4            ggrepel_0.9.6            
 [31] pbkrtest_0.5.5            irlba_2.3.7               annotate_1.88.0          
 [34] DelayedMatrixStats_1.32.0 codetools_0.2-20          DelayedArray_0.36.1      
 [37] tidyselect_1.2.1          farver_2.1.2              lme4_1.1-38              
 [40] mathjaxr_2.0-0            iterators_1.0.14          tools_4.5.2              
 [43] progress_1.2.3            Rcpp_1.1.1                zenith_1.12.0            
 [46] glue_1.8.0                SparseArray_1.10.8        xfun_0.56                
 [49] here_1.0.2                dplyr_1.2.0               withr_3.0.2              
 [52] numDeriv_2016.8-1.1       BiocManager_1.30.27       fastmap_1.2.0            
 [55] boot_1.3-32               caTools_1.18.3            digest_0.6.39            
 [58] truncnorm_1.0-9           R6_2.6.1                  scattermore_1.2          
 [61] gtools_3.9.5              dichromat_2.0-0.1         RSQLite_2.4.5            
 [64] RhpcBLASctl_0.23-42       tidyr_1.3.2               corpcor_1.6.10           
 [67] prettyunits_1.2.0         httr_1.4.7                S4Arrays_1.10.1          
 [70] pkgconfig_2.0.3           gtable_0.3.6              rsconnect_1.7.0          
 [73] blob_1.3.0                S7_0.2.1                  XVector_0.50.0           
 [76] remaCor_0.0.20            htmltools_0.5.9           zigg_0.0.2               
 [79] GSEABase_1.72.0           scales_1.4.0              png_0.1-8                
 [82] fANCOVA_0.6-1             reformulas_0.4.4          ashr_2.2-63              
 [85] knitr_1.51                rstudioapi_0.18.0         rjson_0.2.23             
 [88] reshape2_1.4.5            curl_7.0.0                nlme_3.1-168             
 [91] nloptr_2.2.1              cachem_1.1.0              stringr_1.6.0            
 [94] BiocVersion_3.22.0        KernSmooth_2.23-26        parallel_4.5.2           
 [97] vipor_0.4.7               metadat_1.4-0             AnnotationDbi_1.72.0     
[100] pillar_1.11.1             grid_4.5.2                vctrs_0.7.1              
[103] gplots_3.3.0              mashr_0.2.79              xtable_1.8-4             
[106] beeswarm_0.4.0            Rgraphviz_2.54.0          evaluate_1.0.5           
[109] KEGGgraph_1.70.0          magick_2.9.0              invgamma_1.2             
[112] mvtnorm_1.3-3             cli_3.6.5                 locfit_1.5-9.12          
[115] compiler_4.5.2            rlang_1.1.7               crayon_1.5.3             
[118] SQUAREM_2021.1            plyr_1.8.9                ggbeeswarm_0.7.3         
[121] stringi_1.8.7             assertthat_0.2.1          babelgene_22.9           
[124] lmerTest_3.2-0            Biostrings_2.78.0         aod_1.3.3                
[127] Matrix_1.7-4              hms_1.1.4                 sparseMatrixStats_1.22.0 
[130] bit64_4.6.0-1             KEGGREST_1.50.0           statmod_1.5.1            
[133] rbibutils_2.4.1           Rfast_2.1.5.2             broom_1.0.12             
[136] memoise_2.0.1             RcppParallel_5.1.11-1     bit_4.6.0                
[139] EnrichmentBrowser_2.40.0 
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-21
 rstudio  2026.05.1+225 Golden Wattle (desktop)
 pandoc   3.8.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.9.37 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [1] CRAN (R 4.5.0)
 annotate               1.88.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationDbi          1.72.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationHub        * 4.0.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 aod                    1.3.3      2023-12-13 [1] CRAN (R 4.5.0)
 ashr                   2.2-63     2023-08-21 [1] CRAN (R 4.5.0)
 assertthat             0.2.1      2019-03-21 [1] CRAN (R 4.5.0)
 babelgene              22.9       2022-09-29 [1] CRAN (R 4.5.0)
 backports              1.5.0      2024-05-23 [1] CRAN (R 4.5.0)
 beeswarm               0.4.0      2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocFileCache        * 3.0.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocManager            1.30.27    2025-11-14 [1] CRAN (R 4.5.2)
 BiocParallel         * 1.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocVersion            3.22.0     2025-10-07 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0      2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1    2025-01-16 [1] CRAN (R 4.5.0)
 bitops                 1.0-9      2024-10-03 [1] CRAN (R 4.5.0)
 blob                   1.3.0      2026-01-14 [1] CRAN (R 4.5.2)
 boot                   1.3-32     2025-08-29 [1] CRAN (R 4.5.2)
 broom                  1.0.12     2026-01-27 [1] CRAN (R 4.5.2)
 cachem                 1.1.0      2024-05-16 [1] CRAN (R 4.5.0)
 caTools                1.18.3     2024-09-04 [1] CRAN (R 4.5.0)
 cli                    3.6.5      2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3      2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 corpcor                1.6.10     2021-09-16 [1] CRAN (R 4.5.0)
 crayon                 1.5.3      2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0      2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.4     2026-05-06 [1] CRAN (R 4.5.2)
 DBI                    1.2.3      2024-06-02 [1] CRAN (R 4.5.0)
 dbplyr               * 2.5.1      2025-09-10 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.1     2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 DelayedMatrixStats     1.32.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0      2026-02-03 [1] CRAN (R 4.5.2)
 dreamlet             * 1.8.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 edgeR                * 4.8.2      2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 EnrichmentBrowser      2.40.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 EnvStats               3.1.0      2025-04-24 [1] CRAN (R 4.5.0)
 evaluate               1.0.5      2025-08-27 [1] CRAN (R 4.5.0)
 ExperimentHub        * 3.0.0      2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 fANCOVA                0.6-1      2020-11-13 [1] CRAN (R 4.5.0)
 farver                 2.1.2      2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [1] CRAN (R 4.5.0)
 filelock               1.0.3      2023-12-11 [1] CRAN (R 4.5.0)
 generics             * 0.1.4      2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1     2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 ggbeeswarm             0.7.3      2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2      2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6      2024-09-07 [1] CRAN (R 4.5.0)
 glue                   1.8.0      2024-09-30 [1] CRAN (R 4.5.0)
 gplots                 3.3.0      2025-11-30 [1] CRAN (R 4.5.2)
 graph                  1.88.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 GSEABase               1.72.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 gtable                 0.3.6      2024-10-25 [1] CRAN (R 4.5.0)
 gtools                 3.9.5      2023-11-20 [1] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [1] CRAN (R 4.5.0)
 hms                    1.1.4      2025-10-17 [1] CRAN (R 4.5.0)
 htmltools              0.5.9      2025-12-04 [1] CRAN (R 4.5.2)
 httr                   1.4.7      2023-08-15 [1] CRAN (R 4.5.0)
 httr2                  1.2.2      2025-12-08 [1] CRAN (R 4.5.2)
 invgamma               1.2        2025-07-02 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [1] CRAN (R 4.5.2)
 iterators              1.0.14     2022-02-05 [1] CRAN (R 4.5.0)
 KEGGgraph              1.70.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KEGGREST               1.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KernSmooth             2.23-26    2025-01-01 [1] CRAN (R 4.5.2)
 knitr                  1.51       2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7     2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5      2026-01-08 [1] CRAN (R 4.5.2)
 limma                * 3.66.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 lme4                   1.1-38     2025-12-02 [1] CRAN (R 4.5.2)
 lmerTest               3.2-0      2026-01-13 [1] CRAN (R 4.5.2)
 locfit                 1.5-9.12   2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0      2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [1] CRAN (R 4.5.0)
 mashr                  0.2.79     2023-10-18 [1] CRAN (R 4.5.0)
 MASS                   7.3-65     2025-02-28 [1] CRAN (R 4.5.2)
 mathjaxr               2.0-0      2025-12-01 [1] CRAN (R 4.5.2)
 Matrix                 1.7-4      2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1      2021-11-26 [1] CRAN (R 4.5.0)
 metadat                1.4-0      2025-02-04 [1] CRAN (R 4.5.0)
 metafor                4.8-0      2025-01-28 [1] CRAN (R 4.5.0)
 minqa                  1.2.8      2024-08-17 [1] CRAN (R 4.5.0)
 mixsqp                 0.3-54     2023-12-20 [1] CRAN (R 4.5.0)
 msigdbr                25.1.1     2025-07-21 [1] CRAN (R 4.5.2)
 mvtnorm                1.3-3      2025-01-10 [1] CRAN (R 4.5.0)
 nlme                   3.1-168    2025-03-31 [1] CRAN (R 4.5.2)
 nloptr                 2.2.1      2025-03-17 [1] CRAN (R 4.5.0)
 numDeriv               2016.8-1.1 2019-06-06 [1] CRAN (R 4.5.0)
 otel                   0.2.0      2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.47.0     2026-04-17 [1] CRAN (R 4.5.2)
 pbkrtest               0.5.5      2025-07-18 [1] CRAN (R 4.5.0)
 pillar                 1.11.1     2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [1] CRAN (R 4.5.0)
 plyr                   1.8.9      2023-10-02 [1] CRAN (R 4.5.0)
 png                    0.1-8      2022-11-29 [1] CRAN (R 4.5.0)
 prettyunits            1.2.0      2023-09-24 [1] CRAN (R 4.5.0)
 progress               1.2.3      2023-12-06 [1] CRAN (R 4.5.0)
 purrr                  1.2.1      2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1      2025-02-15 [1] CRAN (R 4.5.0)
 rappdirs               0.3.4      2026-01-17 [1] CRAN (R 4.5.2)
 rbibutils              2.4.1      2026-01-21 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3      2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [1] CRAN (R 4.5.2)
 RcppParallel           5.1.11-1   2025-08-27 [1] CRAN (R 4.5.0)
 RCurl                  1.98-1.17  2025-03-22 [1] CRAN (R 4.5.0)
 Rdpack                 2.6.5      2026-01-23 [1] CRAN (R 4.5.2)
 reformulas             0.4.4      2026-02-02 [1] CRAN (R 4.5.2)
 remaCor                0.0.20     2025-08-20 [1] CRAN (R 4.5.0)
 reshape2               1.4.5      2025-11-12 [1] CRAN (R 4.5.0)
 Rfast                  2.1.5.2    2025-10-10 [1] CRAN (R 4.5.0)
 Rgraphviz              2.54.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 RhpcBLASctl            0.23-42    2023-02-11 [1] CRAN (R 4.5.0)
 rjson                  0.2.23     2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7      2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30       2025-09-28 [1] CRAN (R 4.5.0)
 rmeta                  3.0        2018-03-20 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1      2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0      2025-12-06 [1] CRAN (R 4.5.2)
 RSQLite                2.4.5      2025-11-30 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0     2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1     2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1      2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0      2025-04-24 [1] CRAN (R 4.5.0)
 scattermore            1.2        2023-06-12 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 sparseMatrixStats      1.22.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 SpatialExperiment    * 1.20.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SQUAREM                2021.1     2021-01-13 [1] CRAN (R 4.5.0)
 statmod                1.5.1      2025-10-09 [1] CRAN (R 4.5.0)
 stringi                1.8.7      2025-03-27 [1] CRAN (R 4.5.0)
 stringr                1.6.0      2025-11-04 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1      2026-01-11 [1] CRAN (R 4.5.2)
 tidyr                  1.3.2      2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [1] CRAN (R 4.5.0)
 truncnorm              1.0-9      2023-03-20 [1] CRAN (R 4.5.0)
 variancePartition    * 1.40.1     2025-12-11 [1] Bioconductor 3.22 (R 4.5.2)
 vctrs                  0.7.1      2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7      2023-12-18 [1] CRAN (R 4.5.0)
 WeberDivechaLCdata   * 1.12.0     2025-11-04 [1] Bioconductor 3.22 (R 4.5.2)
 withr                  3.0.2      2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56       2026-01-18 [1] CRAN (R 4.5.2)
 XML                    3.99-0.20  2025-11-08 [1] CRAN (R 4.5.0)
 xtable                 1.8-4      2019-04-21 [1] CRAN (R 4.5.0)
 XVector                0.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12     2025-12-10 [1] CRAN (R 4.5.2)
 zenith                 1.12.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 zigg                   0.0.2      2025-02-07 [1] CRAN (R 4.5.0)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────