---
title: "07-MakeDGEs_srtLCdomain_Banksyclusts_CatYRIDonsOnly"
author: "Bernie Mulvey"
date: "07-11-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(dreamlet)
library(edgeR)
library(limma)
library(SpatialExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##


greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

DE using the banksy labels directly and banksy labels within each SRT-transfer dfeined lc domain.
these dgelists are for a sensitivity analysis where we only consider the subset of donors used in the ancestry-stratified E2-E4 tests. 
```{r}
xlc <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")
tmpcd <- as.data.table(colData(xlc),keep.rownames=T)

## we need the annotated banksy labels and weber-divecha LC domains (for intra-domain cell type DE)
bsy <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_assignmentsMergeClusts.txt")
anno <- fread("processed-data_newseg/11_Banksy_clustering/04-M0lam02kg30res05_mergeClusts_Annots.txt")
bsy <- merge.data.table(bsy,anno,by="clusid")
bsy[,anno:=gsub(anno,pattern=" ",replacement="_")]
bsy[,anno:=gsub(anno,pattern="\\.",replacement="")]
tmpcd <- merge.data.table(bsy[,.(rn,anno)],tmpcd,by="rn")

srtdom <- fread("processed-data_newseg/07_visLabelTransfer/05-srtLCsmoothDoms_and_NMFpreds_lut.txt")
tmpcd <- merge.data.table(tmpcd,srtdom[,.(rn,domain8426,domain12199,domain16119,domain20079)],by="rn")

## remove discard cluster
tmpcd <- tmpcd[anno!="DISCARD"]
xlc <- xlc[,tmpcd$rn]

### !!! ancestry filtering
tmpcd <- tmpcd[greaterOrEqual(YRI,0.75)|lessOrEqual(YRI,0.25)]
xlc <- xlc[,tmpcd$rn]
## length(unique(tmpcd$brnum)) # 27
## length(unique(tmpcd$donor)) # 26 -- ends up 25 for domain 20079 only.

## 2-condition APOE genotype:
tmpcd[,APOEsimple:=ifelse(APOE %in% c("E4hom","E4het"),"E4","E2")]

tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(xlc),]
tmpcd$rn <- NULL

colData(xlc) <- tmpcd

## five SPEs: xlc8/12/16/20 and xlc (all)
xlc8 <- xlc[,xlc$domain8426=="LC"]
xlc12 <- xlc[,xlc$domain12199=="LC"]
xlc16 <- xlc[,xlc$domain16119=="LC"]
xlc20 <- xlc[,xlc$domain20079=="LC"]
spes <- list("domain8426"=xlc8,"domain12199"=xlc12,"domain16119"=xlc16,"domain20079"=xlc20,"wholetissue"=xlc)
rm(xlc,xlc12,xlc8,xlc16,xlc20,bsy,anno,srtdom)
gc(full=T)
```

```{r}
dgels <- lapply(spes,function(l){
  ag <- dreamlet::aggregateToPseudoBulk(l,assay="counts",sample_id="brnum",cluster_id = "anno",BPPARAM = MulticoreParam(8))
    
  agcd <- as.data.table(colData(ag),keep.rownames=T)
  setnames(agcd,"rn","brnum")
  
  agn <- as.data.table(ag@int_colData$n_cells,keep.rownames=T)
  setnames(agn,"rn","clus")
  agn <- melt.data.table(agn,id.vars = "clus")
  setnames(agn,"variable","brnum")
  agn[,brnum:=as.character(brnum)]
  agn[,clus_brnum:=paste0(clus,"_",brnum)]
  agn <- agn[value>=10] ## these are the samples we want to keep to start: 10+ cells for domain type
  setnames(agn,"value","ncells")
  
    
  agm <- as.list(assays(ag))
  agc <- unlist(mapply(x=agm,n=names(agm),SIMPLIFY=FALSE,function(x,n){
    x <- x[,colnames(x) %in% agn[clus==n,brnum]]
    if(is.vector(x)){return(0)} else {
    sum(greaterOrEqual(colSums(x),2000))}})) ## number of pseudobulks with 10+ cells also having 2000+ counts ## number of pseudobulks with 2000+ counts in at least 10 samples
  ## only keep those where at least 10 samples meet that
  kpb <- names(agc[agc>=10])
  agn <- agn[clus %in% kpb]
  agm <- agm[kpb]
    
    ## now we want to make one giant sample matrix with the remaining samples
  agm.filt <- mapply(a=agm,n=names(agm),SIMPLIFY=FALSE,function(a,n){
    a <- a[,colnames(a) %in% agn[clus==n]$brnum]
    colnames(a) <- paste0(n,"_",colnames(a))
    return(a)
  })
  agm.filt <- do.call("cbind",agm.filt)
    
  # make a per cluster-donor samples table
  agn <- merge.data.table(agn,agcd[,.(brnum,sample_id,Sex,APOE,APOEsimple,YRI,Age)],by="brnum")
  stopifnot(all(agn$clus_brnum %in% colnames(agm.filt)))
  agn <- as.data.frame(agn,row.names=agn$clus_brnum)[colnames(agm.filt),]
  dge <- DGEList(counts=agm.filt, samples=agn)
  return(dge)
})
names(dgels) <- names(spes)

saveRDS(dgels,"processed-data_newseg/13_BanksyDE/07-BanksyLabels_srtLCdomains_DGELists_catYRIdonorsOnly.RDS")
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices
[5] utils     methods   base     

other attached packages:
 [1] parallelly_1.47.0          
 [2] SpatialExperiment_1.20.0   
 [3] edgeR_4.8.2                
 [4] dreamlet_1.8.0             
 [5] SingleCellExperiment_1.32.0
 [6] SummarizedExperiment_1.40.0
 [7] Biobase_2.70.0             
 [8] GenomicRanges_1.62.1       
 [9] Seqinfo_1.0.0              
[10] IRanges_2.44.0             
[11] S4Vectors_0.48.1           
[12] BiocGenerics_0.56.0        
[13] generics_0.1.4             
[14] MatrixGenerics_1.22.0      
[15] matrixStats_1.5.0          
[16] variancePartition_1.40.1   
[17] BiocParallel_1.44.0        
[18] limma_3.66.0               
[19] ggplot2_4.0.2              
[20] data.table_1.18.4          

loaded via a namespace (and not attached):
  [1] mathjaxr_2.0-0           
  [2] RColorBrewer_1.1-3       
  [3] magrittr_2.0.4           
  [4] magick_2.9.0             
  [5] ggbeeswarm_0.7.3         
  [6] farver_2.1.2             
  [7] nloptr_2.2.1             
  [8] vctrs_0.7.1              
  [9] memoise_2.0.1            
 [10] minqa_1.2.8              
 [11] DelayedMatrixStats_1.32.0
 [12] RCurl_1.98-1.17          
 [13] SQUAREM_2021.1           
 [14] mixsqp_0.3-54            
 [15] progress_1.2.3           
 [16] S4Arrays_1.10.1          
 [17] curl_7.0.0               
 [18] truncnorm_1.0-9          
 [19] broom_1.0.12             
 [20] SparseArray_1.10.8       
 [21] mashr_0.2.79             
 [22] KernSmooth_2.23-26       
 [23] pbkrtest_0.5.5           
 [24] plyr_1.8.9               
 [25] cachem_1.1.0             
 [26] lifecycle_1.0.5          
 [27] iterators_1.0.14         
 [28] pkgconfig_2.0.3          
 [29] Matrix_1.7-4             
 [30] R6_2.6.1                 
 [31] fastmap_1.2.0            
 [32] rbibutils_2.4.1          
 [33] digest_0.6.39            
 [34] numDeriv_2016.8-1.1      
 [35] AnnotationDbi_1.72.0     
 [36] rprojroot_2.1.1          
 [37] irlba_2.3.7              
 [38] RSQLite_2.4.5            
 [39] invgamma_1.2             
 [40] metadat_1.4-0            
 [41] httr_1.4.7               
 [42] abind_1.4-8              
 [43] compiler_4.5.2           
 [44] here_1.0.2               
 [45] bit64_4.6.0-1            
 [46] aod_1.3.3                
 [47] withr_3.0.2              
 [48] S7_0.2.1                 
 [49] backports_1.5.0          
 [50] metafor_4.8-0            
 [51] DBI_1.2.3                
 [52] gplots_3.3.0             
 [53] MASS_7.3-65              
 [54] DelayedArray_0.36.1      
 [55] rjson_0.2.23             
 [56] corpcor_1.6.10           
 [57] gtools_3.9.5             
 [58] caTools_1.18.3           
 [59] tools_4.5.2              
 [60] vipor_0.4.7              
 [61] otel_0.2.0               
 [62] beeswarm_0.4.0           
 [63] msigdbr_25.1.1           
 [64] remaCor_0.0.20           
 [65] glue_1.8.0               
 [66] nlme_3.1-168             
 [67] grid_4.5.2               
 [68] reshape2_1.4.5           
 [69] gtable_0.3.6             
 [70] tidyr_1.3.2              
 [71] hms_1.1.4                
 [72] XVector_0.50.0           
 [73] ggrepel_0.9.6            
 [74] pillar_1.11.1            
 [75] stringr_1.6.0            
 [76] babelgene_22.9           
 [77] splines_4.5.2            
 [78] dplyr_1.2.0              
 [79] lattice_0.22-7           
 [80] bit_4.6.0                
 [81] annotate_1.88.0          
 [82] tidyselect_1.2.1         
 [83] locfit_1.5-9.12          
 [84] Biostrings_2.78.0        
 [85] knitr_1.51               
 [86] reformulas_0.4.4         
 [87] scattermore_1.2          
 [88] RhpcBLASctl_0.23-42      
 [89] xfun_0.56                
 [90] statmod_1.5.1            
 [91] KEGGgraph_1.70.0         
 [92] stringi_1.8.7            
 [93] boot_1.3-32              
 [94] evaluate_1.0.5           
 [95] codetools_0.2-20         
 [96] tibble_3.3.1             
 [97] Rgraphviz_2.54.0         
 [98] graph_1.88.1             
 [99] cli_3.6.5                
[100] RcppParallel_5.1.11-1    
[101] zenith_1.12.0            
[102] xtable_1.8-4             
[103] Rdpack_2.6.5             
[104] dichromat_2.0-0.1        
[105] Rcpp_1.1.1               
[106] zigg_0.0.2               
[107] rmeta_3.0                
[108] EnvStats_3.1.0           
[109] png_0.1-8                
[110] Rfast_2.1.5.2            
[111] XML_3.99-0.20            
[112] parallel_4.5.2           
[113] assertthat_0.2.1         
[114] blob_1.3.0               
[115] prettyunits_1.2.0        
[116] sparseMatrixStats_1.22.0 
[117] bitops_1.0-9             
[118] lme4_1.1-38              
[119] mvtnorm_1.3-3            
[120] GSEABase_1.72.0          
[121] lmerTest_3.2-0           
[122] scales_1.4.0             
[123] purrr_1.2.1              
[124] crayon_1.5.3             
[125] fANCOVA_0.6-1            
[126] rlang_1.1.7              
[127] EnrichmentBrowser_2.40.0 
[128] ashr_2.2-63              
[129] KEGGREST_1.50.0          
> sessioninfo::session_info()
─ Session info ──────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       Positron
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-07-11
 pandoc   NA
 quarto   1.9.38 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [1] CRAN (R 4.5.0)
 annotate               1.88.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationDbi          1.72.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 aod                    1.3.3      2023-12-13 [1] CRAN (R 4.5.0)
 ashr                   2.2-63     2023-08-21 [1] CRAN (R 4.5.0)
 assertthat             0.2.1      2019-03-21 [1] CRAN (R 4.5.0)
 babelgene              22.9       2022-09-29 [1] CRAN (R 4.5.0)
 backports              1.5.0      2024-05-23 [1] CRAN (R 4.5.0)
 beeswarm               0.4.0      2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0      2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1    2025-01-16 [1] CRAN (R 4.5.0)
 bitops                 1.0-9      2024-10-03 [1] CRAN (R 4.5.0)
 blob                   1.3.0      2026-01-14 [1] CRAN (R 4.5.2)
 boot                   1.3-32     2025-08-29 [1] CRAN (R 4.5.2)
 broom                  1.0.12     2026-01-27 [1] CRAN (R 4.5.2)
 cachem                 1.1.0      2024-05-16 [1] CRAN (R 4.5.0)
 caTools                1.18.3     2024-09-04 [1] CRAN (R 4.5.0)
 cli                    3.6.5      2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [1] CRAN (R 4.5.2)
 corpcor                1.6.10     2021-09-16 [1] CRAN (R 4.5.0)
 crayon                 1.5.3      2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0      2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.4     2026-05-06 [1] CRAN (R 4.5.2)
 DBI                    1.2.3      2024-06-02 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.1     2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 DelayedMatrixStats     1.32.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0      2026-02-03 [1] CRAN (R 4.5.2)
 dreamlet             * 1.8.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 edgeR                * 4.8.2      2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 EnrichmentBrowser      2.40.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 EnvStats               3.1.0      2025-04-24 [1] CRAN (R 4.5.0)
 evaluate               1.0.5      2025-08-27 [1] CRAN (R 4.5.0)
 fANCOVA                0.6-1      2020-11-13 [1] CRAN (R 4.5.0)
 farver                 2.1.2      2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4      2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1     2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 ggbeeswarm             0.7.3      2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2      2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6      2024-09-07 [1] CRAN (R 4.5.0)
 glue                   1.8.0      2024-09-30 [1] CRAN (R 4.5.0)
 gplots                 3.3.0      2025-11-30 [1] CRAN (R 4.5.2)
 graph                  1.88.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 GSEABase               1.72.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 gtable                 0.3.6      2024-10-25 [1] CRAN (R 4.5.0)
 gtools                 3.9.5      2023-11-20 [1] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [1] CRAN (R 4.5.0)
 hms                    1.1.4      2025-10-17 [1] CRAN (R 4.5.0)
 httr                   1.4.7      2023-08-15 [1] CRAN (R 4.5.0)
 invgamma               1.2        2025-07-02 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [1] CRAN (R 4.5.2)
 iterators              1.0.14     2022-02-05 [1] CRAN (R 4.5.0)
 KEGGgraph              1.70.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KEGGREST               1.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KernSmooth             2.23-26    2025-01-01 [1] CRAN (R 4.5.2)
 knitr                  1.51       2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7     2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5      2026-01-08 [1] CRAN (R 4.5.2)
 limma                * 3.66.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 lme4                   1.1-38     2025-12-02 [1] CRAN (R 4.5.2)
 lmerTest               3.2-0      2026-01-13 [1] CRAN (R 4.5.2)
 locfit                 1.5-9.12   2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0      2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [1] CRAN (R 4.5.0)
 mashr                  0.2.79     2023-10-18 [1] CRAN (R 4.5.0)
 MASS                   7.3-65     2025-02-28 [1] CRAN (R 4.5.2)
 mathjaxr               2.0-0      2025-12-01 [1] CRAN (R 4.5.2)
 Matrix                 1.7-4      2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1      2021-11-26 [1] CRAN (R 4.5.0)
 metadat                1.4-0      2025-02-04 [1] CRAN (R 4.5.0)
 metafor                4.8-0      2025-01-28 [1] CRAN (R 4.5.0)
 minqa                  1.2.8      2024-08-17 [1] CRAN (R 4.5.0)
 mixsqp                 0.3-54     2023-12-20 [1] CRAN (R 4.5.0)
 msigdbr                25.1.1     2025-07-21 [1] CRAN (R 4.5.2)
 mvtnorm                1.3-3      2025-01-10 [1] CRAN (R 4.5.0)
 nlme                   3.1-168    2025-03-31 [1] CRAN (R 4.5.2)
 nloptr                 2.2.1      2025-03-17 [1] CRAN (R 4.5.0)
 numDeriv               2016.8-1.1 2019-06-06 [1] CRAN (R 4.5.0)
 otel                   0.2.0      2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.47.0     2026-04-17 [1] CRAN (R 4.5.2)
 pbkrtest               0.5.5      2025-07-18 [1] CRAN (R 4.5.0)
 pillar                 1.11.1     2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [1] CRAN (R 4.5.0)
 plyr                   1.8.9      2023-10-02 [1] CRAN (R 4.5.0)
 png                    0.1-8      2022-11-29 [1] CRAN (R 4.5.0)
 prettyunits            1.2.0      2023-09-24 [1] CRAN (R 4.5.0)
 progress               1.2.3      2023-12-06 [1] CRAN (R 4.5.0)
 purrr                  1.2.1      2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1      2025-02-15 [1] CRAN (R 4.5.0)
 rbibutils              2.4.1      2026-01-21 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3      2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [1] CRAN (R 4.5.2)
 RcppParallel           5.1.11-1   2025-08-27 [1] CRAN (R 4.5.0)
 RCurl                  1.98-1.17  2025-03-22 [1] CRAN (R 4.5.0)
 Rdpack                 2.6.5      2026-01-23 [1] CRAN (R 4.5.2)
 reformulas             0.4.4      2026-02-02 [1] CRAN (R 4.5.2)
 remaCor                0.0.20     2025-08-20 [1] CRAN (R 4.5.0)
 reshape2               1.4.5      2025-11-12 [1] CRAN (R 4.5.0)
 Rfast                  2.1.5.2    2025-10-10 [1] CRAN (R 4.5.0)
 Rgraphviz              2.54.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 RhpcBLASctl            0.23-42    2023-02-11 [1] CRAN (R 4.5.0)
 rjson                  0.2.23     2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7      2026-01-09 [1] CRAN (R 4.5.2)
 rmeta                  3.0        2018-03-20 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1      2025-08-26 [1] CRAN (R 4.5.0)
 RSQLite                2.4.5      2025-11-30 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1     2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1      2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0      2025-04-24 [1] CRAN (R 4.5.0)
 scattermore            1.2        2023-06-12 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 sparseMatrixStats      1.22.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 SpatialExperiment    * 1.20.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SQUAREM                2021.1     2021-01-13 [1] CRAN (R 4.5.0)
 statmod                1.5.1      2025-10-09 [1] CRAN (R 4.5.0)
 stringi                1.8.7      2025-03-27 [1] CRAN (R 4.5.0)
 stringr                1.6.0      2025-11-04 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1      2026-01-11 [1] CRAN (R 4.5.2)
 tidyr                  1.3.2      2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [1] CRAN (R 4.5.0)
 truncnorm              1.0-9      2023-03-20 [1] CRAN (R 4.5.0)
 variancePartition    * 1.40.1     2025-12-11 [1] Bioconductor 3.22 (R 4.5.2)
 vctrs                  0.7.1      2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7      2023-12-18 [1] CRAN (R 4.5.0)
 withr                  3.0.2      2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56       2026-01-18 [1] CRAN (R 4.5.2)
 XML                    3.99-0.20  2025-11-08 [1] CRAN (R 4.5.0)
 xtable                 1.8-4      2019-04-21 [1] CRAN (R 4.5.0)
 XVector                0.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 zenith                 1.12.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 zigg                   0.0.2      2025-02-07 [1] CRAN (R 4.5.0)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────