---
title: "03-AstroProportion_and_NEAstro_Ratios"
format: html
date: "03-16-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)

## rstudio GUI tweaks
options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
##

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
precision <- sqrt(.Machine$double.eps)
(x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
precision <- sqrt(.Machine$double.eps)
(x <= y) | (abs(x-y) <= precision)}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost='localhost')


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = '#000000'), axis.title.x = element_text(size = 9,color = '#000000'), axis.text.y = element_text(size = 8,color = '#000000'), axis.title.y = element_text(size =9,color = '#000000'), plot.title = element_markdown(size = 10,hjust=0.5,color = '#000000'), strip.text = element_text(size=11), legend.text = element_text(size=7,color = '#000000'), legend.title = element_text(size=8,hjust=0.5,color = '#000000'),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position='plot'))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

Absolute astrocyte %age in domain:
```{r}
restabs <- fread("processed-data/08_LC_ProximalAstros/03-AstroCells_as_Proportion_of_LCdomaincells_restabs.txt")


png("plots/08_LC_ProximalAstros/03-CellTypeProps_inLCdomain_by_domainAndLabelXferMethod.png",height=600,width=1800)
ggplot(restabs,aes(x=APOEcarriertype,y=propOfLCdom))+
    geom_boxplot()+
    facet_grid(domain~labeling+cellLabel)+
    ylab("Proportion of Domain cells with label")+
    xlab("APOE carrier type")
dev.off()
```

Astrocyte to NE ratios:
```{r}
restabs2 <- restabs[cellLabel %in% grep(cellLabel,pattern="NE|LC|^Astro",value=T)]
restabs2 <- restabs2[cellLabel %in% grep(cellLabel,pattern="Oligo|SERT",value=T,invert=T)]
restab2 <- dcast(restabs2,brnum+domain+labeling+APOEcarriertype~cellLabel,value.var="N")
restab2[,wd_AstNEratio:=Astrocytes/NE]
restab2[,vis_astLCratio:=Astro/LC]

restab3 <- melt.data.table(restab2,id.vars=c("brnum","domain","labeling","APOEcarriertype"),measure.vars = c("wd_AstNEratio", "vis_astLCratio"), 
    variable.name = "ratio_type", 
    value.name = "ratio_val", 
    na.rm = TRUE
)

png("plots/08_LC_ProximalAstros/03-AstroNE_ratio_by_APOE_domain_and_labelsXfered.png",height=600,width=600)
ggplot(restab3,aes(x=APOEcarriertype,y=ratio_val))+
    geom_boxplot()+
    facet_grid(domain~labeling)+
    ylab("Intra-domain Astro:NE ratio")+
    xlab("APOE carrier type")
dev.off()


t.test(restab3[APOEcarriertype=="E2"&ratio_type=="wd_AstNEratio"&domain=="domain8266",ratio_val],restab3[APOEcarriertype=="E4"&ratio_type=="wd_AstNEratio"&domain=="domain8266",ratio_val],var.equal = T)
t.test(restab3[APOEcarriertype=="E2"&ratio_type=="vis_astLCratio"&domain=="domain8266",ratio_val],restab3[APOEcarriertype=="E4"&ratio_type=="vis_astLCratio"&domain=="domain8266",ratio_val],var.equal = T)

t.test(restab3[APOEcarriertype=="E2"&ratio_type=="wd_AstNEratio"&domain=="domain12252",ratio_val],restab3[APOEcarriertype=="E4"&ratio_type=="wd_AstNEratio"&domain=="domain12252",ratio_val],var.equal = T)
t.test(restab3[APOEcarriertype=="E2"&ratio_type=="vis_astLCratio"&domain=="domain12252",ratio_val],restab3[APOEcarriertype=="E4"&ratio_type=="vis_astLCratio"&domain=="domain12252",ratio_val],var.equal = T)

```

### seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils    
[5] methods   base     

other attached packages:
[1] bettermc_1.2.2.9000 BiocParallel_1.44.0
[3] parallelly_1.46.1   gridExtra_2.3      
[5] ggtext_0.1.2        ggplot2_4.0.2      
[7] data.table_1.18.2.1

loaded via a namespace (and not attached):
 [1] gtable_0.3.6       dplyr_1.2.0       
 [3] compiler_4.5.2     tidyselect_1.2.1  
 [5] Rcpp_1.1.1         xml2_1.5.2        
 [7] parallel_4.5.2     dichromat_2.0-0.1 
 [9] textshaping_1.0.4  systemfonts_1.3.1 
[11] scales_1.4.0       here_1.0.2        
[13] R6_2.6.1           labeling_0.4.3    
[15] generics_0.1.4     knitr_1.51        
[17] tibble_3.3.1       rprojroot_2.1.1   
[19] pillar_1.11.1      RColorBrewer_1.1-3
[21] rlang_1.1.7        xfun_0.56         
[23] S7_0.2.1           otel_0.2.0        
[25] cli_3.6.5          withr_3.0.2       
[27] magrittr_2.0.4     grid_4.5.2        
[29] gridtext_0.1.5     lifecycle_1.0.5   
[31] vctrs_0.7.1        evaluate_1.0.5    
[33] glue_1.8.0         farver_2.1.2      
[35] sessioninfo_1.2.3  ragg_1.5.0        
[37] codetools_0.2-20   tools_4.5.2       
[39] pkgconfig_2.0.3   
> sessioninfo::session_info()
─ Session info ──────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       Positron
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-03-16
 pandoc   NA
 quarto   1.8.27 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────
 package      * version    date (UTC) lib source
 bettermc     * 1.2.2.9000 2026-03-16 [2] Github (akersting/bettermc@d6ec1aa)
 BiocParallel * 1.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli            3.6.5      2025-04-23 [1] CRAN (R 4.5.0)
 codetools      0.2-20     2024-03-31 [1] CRAN (R 4.5.2)
 data.table   * 1.18.2.1   2026-01-27 [1] CRAN (R 4.5.2)
 dichromat      2.0-0.1    2022-05-02 [1] CRAN (R 4.5.0)
 dplyr          1.2.0      2026-02-03 [1] CRAN (R 4.5.2)
 evaluate       1.0.5      2025-08-27 [1] CRAN (R 4.5.0)
 farver         2.1.2      2024-05-13 [1] CRAN (R 4.5.0)
 generics       0.1.4      2025-05-09 [1] CRAN (R 4.5.0)
 ggplot2      * 4.0.2      2026-02-03 [1] CRAN (R 4.5.2)
 ggtext       * 0.1.2      2022-09-16 [1] CRAN (R 4.5.0)
 glue           1.8.0      2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra    * 2.3        2017-09-09 [1] CRAN (R 4.5.0)
 gridtext       0.1.5      2022-09-16 [1] CRAN (R 4.5.0)
 gtable         0.3.6      2024-10-25 [1] CRAN (R 4.5.0)
 here           1.0.2      2025-09-15 [1] CRAN (R 4.5.0)
 knitr          1.51       2025-12-20 [1] CRAN (R 4.5.2)
 labeling       0.4.3      2023-08-29 [1] CRAN (R 4.5.0)
 lifecycle      1.0.5      2026-01-08 [1] CRAN (R 4.5.2)
 magrittr       2.0.4      2025-09-12 [1] CRAN (R 4.5.0)
 otel           0.2.0      2025-08-29 [1] CRAN (R 4.5.0)
 parallelly   * 1.46.1     2026-01-08 [1] CRAN (R 4.5.2)
 pillar         1.11.1     2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig      2.0.3      2019-09-22 [1] CRAN (R 4.5.0)
 R6             2.6.1      2025-02-15 [1] CRAN (R 4.5.0)
 ragg           1.5.0      2025-09-02 [1] CRAN (R 4.5.0)
 RColorBrewer   1.1-3      2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp           1.1.1      2026-01-10 [1] CRAN (R 4.5.2)
 rlang          1.1.7      2026-01-09 [1] CRAN (R 4.5.2)
 rprojroot      2.1.1      2025-08-26 [1] CRAN (R 4.5.0)
 S7             0.2.1      2025-11-14 [1] CRAN (R 4.5.2)
 scales         1.4.0      2025-04-24 [1] CRAN (R 4.5.0)
 sessioninfo    1.2.3      2025-02-05 [1] CRAN (R 4.5.0)
 systemfonts    1.3.1      2025-10-01 [1] CRAN (R 4.5.0)
 textshaping    1.0.4      2025-10-10 [1] CRAN (R 4.5.0)
 tibble         3.3.1      2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect     1.2.1      2024-03-11 [1] CRAN (R 4.5.0)
 vctrs          0.7.1      2026-01-23 [1] CRAN (R 4.5.2)
 withr          3.0.2      2024-10-28 [1] CRAN (R 4.5.0)
 xfun           0.56       2026-01-18 [1] CRAN (R 4.5.2)
 xml2           1.5.2      2026-01-17 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────