---
title: "05-Tabulate VisDE Replication"
author: "Bernie Mulvey"
date: "03-18-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##


## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
precision <- sqrt(.Machine$double.eps)
(x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
precision <- sqrt(.Machine$double.eps)
(x <= y) | (abs(x-y) <= precision)}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

## with LC as a smoothed domain
```{r}
totde <- readRDS("processed-data/06_DomainDE/02-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_XferedVisLabels.RDS")
totde <- lapply(totde,function(x){
  y <- lapply(x,function(z){
    z2 <- z$detab
    z2 <- rbindlist(z2,idcol="contrast",fill=T)
    return(z2)
  })
  rbindlist(y,idcol="modelset")
})


visde <- readRDS("../LFF_spatial_LC/processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02b-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_noMixClusts_YRIcontAndCat_JUL2025.RDS")
visdes <- list(visde$cont_ances$E2E4label$detabs,visde$cont_ances$sexlabel$detabs,visde$cat_ances$YRIAPOlabel$detabs)
names(visdes) <- c("E2E4label","sexlabel","YRIAPOlabel")
visdes <- lapply(visdes,rbindlist,idcol="contrast")
visdes <- rbindlist(visdes)

totde <- lapply(totde,function(d){
  z <- d
  z[,contrast:=gsub(contrast,pattern="LCdomain",replacement="LC")]
  z <- z[contrast %in% c(unique(visdes$contrast))]
  z <- z[,.(contrast,gene_name,P.Value,logFC)]
  setnames(z,c("P.Value","logFC"),c("Xen.P.unadj","xlfc"))
  z <- merge.data.table(z,visdes[adj.P.Val<0.05,.(contrast,gene_name,adj.P.Val,logFC)],by=c("contrast","gene_name"))
  return(z)
})

reps <- lapply(totde,function(d){
  x <- d[,.N,by="contrast"]
  setnames(x,"N","VisFDR05_tested")
  y <- d[Xen.P.unadj<0.05,.N,by="contrast"]
  setnames(y,"N","XenP05sig")
  z <- d[Xen.P.unadj<0.05&(logFC*xlfc>0),.N,by="contrast"]
  setnames(z,"N","XenP05sig_concord")
  x <- merge.data.table(x,y,by="contrast",all.x=T)
  x <- merge.data.table(x,z,by="contrast",all.x=T)
  x[,reprate:=XenP05sig_concord/VisFDR05_tested]
  x[,discordrate:=(XenP05sig-XenP05sig_concord)/XenP05sig]
  return(x)
})


mapply(x=reps,n=names(reps),SIMPLIFY = FALSE,FUN = function(x,n){
  fwrite(x,paste0("processed-data/06_DomainDE/05-Tabulate_VisDEreplication_results_percontrast_withSmooth",n,".txt"),sep='\t',quote=F,row.names=F,col.names=T)
})

rep8266 <- reps$dom8266[,.(contrast,VisFDR05_tested,reprate)]
setnames(rep8266,"reprate","reprate8266")
rep12252 <- reps$dom12252[,.(contrast,reprate)]
setnames(rep12252,"reprate","reprate12252")

reps2 <- merge.data.table(rep8266,rep12252,by="contrast")


fwrite(reps2,"processed-data/06_DomainDE/05a-Replication_Rates_by_DomainDefinition_withLCsmoothing.txt",sep='\t',quote=F,row.names=T)
```

### which E4-E2 genes replicate for the LC domain?
```{r}
lce4.8266 <- totde$dom8266[contrast=="LC_E4vE2"&lessOrEqual(Xen.P.unadj,0.05)&xlfc*logFC>0]
lce4.12252 <- totde$dom12252[contrast=="LC_E4vE2"&lessOrEqual(Xen.P.unadj,0.05)&xlfc*logFC>0]
sharedg <- lce4.8266[gene_name %in% lce4.12252$gene_name,.(xlfc,gene_name)]
```
AIF1
AQP4
C3
EGFR
FGFR3
GJA1
IDH2
ITGB8
NPY
NTSR2
PLPP3
SLC7A11
SPARC
SPON1
TRIL


### with intra-LC only, direct label transfers
```{r}
totde <- readRDS("processed-data/06_DomainDE/04-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_withinLCVisLabs.RDS")
totde <- lapply(totde,function(x){
  y <- lapply(x,function(z){
    z2 <- z$detab
    z2 <- rbindlist(z2,idcol="contrast",fill=T)
    return(z2)
  })
  rbindlist(y,idcol="modelset")
})


visde <- readRDS("../LFF_spatial_LC/processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02b-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_noMixClusts_YRIcontAndCat_JUL2025.RDS")
visdes <- list(visde$cont_ances$E2E4label$detabs,visde$cont_ances$sexlabel$detabs,visde$cat_ances$YRIAPOlabel$detabs)
names(visdes) <- c("E2E4label","sexlabel","YRIAPOlabel")
visdes <- lapply(visdes,rbindlist,idcol="contrast")
visdes <- rbindlist(visdes)

totde <- lapply(totde,function(d){
  z <- d
  z[,contrast:=gsub(contrast,pattern="LCdomain",replacement="LC")]
  z <- z[contrast %in% c(unique(visdes$contrast))]
  z <- z[,.(contrast,gene_name,P.Value,logFC)]
  setnames(z,c("P.Value","logFC"),c("Xen.P.unadj","xlfc"))
  z <- merge.data.table(z,visdes[adj.P.Val<0.05,.(contrast,gene_name,adj.P.Val,logFC)],by=c("contrast","gene_name"))
  return(z)
})

reps <- lapply(totde,function(d){
  x <- d[,.N,by="contrast"]
  setnames(x,"N","VisFDR05_tested")
  y <- d[Xen.P.unadj<0.05,.N,by="contrast"]
  setnames(y,"N","XenP05sig")
  z <- d[Xen.P.unadj<0.05&(logFC*xlfc>0),.N,by="contrast"]
  setnames(z,"N","XenP05sig_concord")
  x <- merge.data.table(x,y,by="contrast",all.x=T)
  x <- merge.data.table(x,z,by="contrast",all.x=T)
  x[,reprate:=XenP05sig_concord/VisFDR05_tested]
  x[,discordrate:=(XenP05sig-XenP05sig_concord)/XenP05sig]
  return(x)
})


mapply(x=reps,n=names(reps),SIMPLIFY = FALSE,FUN = function(x,n){
  fwrite(x,paste0("processed-data/06_DomainDE/05-Tabulate_VisDEreplication_results_withinLCdomainONLY_percontrast",n,".txt"),sep='\t',quote=F,row.names=F,col.names=T)
})

rep8266 <- reps$dom8266[,.(contrast,VisFDR05_tested,reprate)]
setnames(rep8266,"reprate","reprate8266")
rep12252 <- reps$dom12252[,.(contrast,reprate)]
setnames(rep12252,"reprate","reprate12252")

reps2 <- merge.data.table(rep8266,rep12252,by="contrast")

fwrite(reps2,"processed-data/06_DomainDE/05b-Vis domain DE replication per label only within smoothed LC.txt",sep='\t',quote=F,row.names=T) 
```



```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] ggplot2_4.0.2       BiocParallel_1.44.0 parallelly_1.46.1   colorout_1.3-3     
[5] edgeR_4.8.2         limma_3.66.0        data.table_1.18.2.1

loaded via a namespace (and not attached):
 [1] gtable_0.3.6       dplyr_1.2.0        compiler_4.5.2     tidyselect_1.2.1  
 [5] dichromat_2.0-0.1  scales_1.4.0       yaml_2.3.12        fastmap_1.2.0     
 [9] statmod_1.5.1      lattice_0.22-7     here_1.0.2         R6_2.6.1          
[13] labeling_0.4.3     generics_0.1.4     knitr_1.51         tibble_3.3.1      
[17] rprojroot_2.1.1    pillar_1.11.1      RColorBrewer_1.1-3 rlang_1.1.7       
[21] xfun_0.56          S7_0.2.1           otel_0.2.0         cli_3.6.5         
[25] withr_3.0.2        magrittr_2.0.4     digest_0.6.39      grid_4.5.2        
[29] locfit_1.5-9.12    rstudioapi_0.18.0  lifecycle_1.0.5    vctrs_0.7.1       
[33] evaluate_1.0.5     glue_1.8.0         farver_2.1.2       sessioninfo_1.2.3 
[37] codetools_0.2-20   rmarkdown_2.30     pkgconfig_2.0.3    tools_4.5.2       
[41] htmltools_0.5.9   
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-28
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package      * version  date (UTC) lib source
 BiocParallel * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli            3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools      0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout     * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table   * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 dichromat      2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest         0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr          1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 edgeR        * 4.8.2    2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate       1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver         2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap        1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics       0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 ggplot2      * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 glue           1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gtable         0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here           1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools      0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 knitr          1.51     2025-12-20 [1] CRAN (R 4.5.2)
 labeling       0.4.3    2023-08-29 [1] CRAN (R 4.5.0)
 lattice        0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle      1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 limma        * 3.66.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 locfit         1.5-9.12 2025-03-05 [1] CRAN (R 4.5.0)
 magrittr       2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 otel           0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly   * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pillar         1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig      2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 R6             2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer   1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 rlang          1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown      2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot      2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rstudioapi     0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S7             0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales         1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 sessioninfo    1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 statmod        1.5.1    2025-10-09 [1] CRAN (R 4.5.0)
 tibble         3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect     1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs          0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 withr          3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun           0.56     2026-01-18 [1] CRAN (R 4.5.2)
 yaml           2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────