---
title: "06-Make_VisLabelsAsXfered_DGELists"
author: "Bernie Mulvey"
date: "03-18-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(dreamlet)
library(edgeR)
library(limma)
library(SpatialExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

DE using the nmfpred labels directly--disregarding the multi-cell-type LC domain here.

```{r}
xlc <- readRDS("processed-data/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")
tmpcd <- as.data.table(colData(xlc),keep.rownames=T)

lcd <- fread("processed-data/05_LabelTransfer/06-LCsmoothings_LUT.txt")

tmpcd <- merge.data.table(tmpcd,lcd[,.(rn,nmfpred)],by="rn")

tmpcd[,APOEsimple:=ifelse(APOE %in% c("E4hom","E4het"),"E4","E2")]
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(xlc),]
tmpcd$rn <- NULL

colData(xlc) <- tmpcd
```

```{r}
ag <- dreamlet::aggregateToPseudoBulk(xlc,assay="counts",sample_id="brnum",cluster_id = "nmfpred",BPPARAM = MulticoreParam(8))
  
agcd <- as.data.table(colData(ag),keep.rownames=T)
setnames(agcd,"rn","brnum")

agn <- as.data.table(ag@int_colData$n_cells,keep.rownames=T)
setnames(agn,"rn","clus")
agn <- melt.data.table(agn,id.vars = "clus")
setnames(agn,"variable","brnum")
agn[,brnum:=as.character(brnum)]
agn[,clus_brnum:=paste0(clus,"_",brnum)]
agn <- agn[value>=10] ## these are the samples we want to keep to start: 10+ cells for domain type
setnames(agn,"value","ncells")

  
agm <- as.list(assays(ag))
agc <- unlist(lapply(agm,function(x){sum(colSums(x)>2000)})) ## number of pseudobulks with 2000+ counts
## only keep those where at least 10 samples meet that
kpb <- names(agc[agc>=10])
agn <- agn[clus %in% kpb]
agm <- agm[kpb]
  
  ## now we want to make one giant sample matrix with the remaining samples
agm.filt <- mapply(a=agm,n=names(agm),SIMPLIFY=FALSE,function(a,n){
  a <- a[,colnames(a) %in% agn[clus==n]$brnum]
  colnames(a) <- paste0(n,"_",colnames(a))
  return(a)
})
agm.filt <- do.call("cbind",agm.filt)
  
# make a per cluster-donor samples table
agn <- merge.data.table(agn,agcd[,.(brnum,sample_id,Sex,APOE,APOEsimple,YRI,Age)],by="brnum")
stopifnot(all(agn$clus_brnum %in% colnames(agm.filt)))
agn <- as.data.frame(agn,row.names=agn$clus_brnum)[colnames(agm.filt),]
dge <- DGEList(counts=agm.filt, samples=agn)

saveRDS(dge,"processed-data/06_DomainDE/06-VisLabelsAsXfered_DGEList.rds")
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] parallelly_1.46.1           colorout_1.3-3             
 [3] SpatialExperiment_1.20.0    edgeR_4.8.2                
 [5] dreamlet_1.8.0              SingleCellExperiment_1.32.0
 [7] SummarizedExperiment_1.40.0 Biobase_2.70.0             
 [9] GenomicRanges_1.62.1        Seqinfo_1.0.0              
[11] IRanges_2.44.0              S4Vectors_0.48.0           
[13] BiocGenerics_0.56.0         generics_0.1.4             
[15] MatrixGenerics_1.22.0       matrixStats_1.5.0          
[17] variancePartition_1.40.1    BiocParallel_1.44.0        
[19] limma_3.66.0                ggplot2_4.0.2              
[21] data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] splines_4.5.2             bitops_1.0-9              tibble_3.3.1             
  [4] graph_1.88.1              XML_3.99-0.20             lifecycle_1.0.5          
  [7] Rdpack_2.6.5              mixsqp_0.3-54             rprojroot_2.1.1          
 [10] lattice_0.22-7            MASS_7.3-65               backports_1.5.0          
 [13] magrittr_2.0.4            metafor_4.8-0             otel_0.2.0               
 [16] sessioninfo_1.2.3         DBI_1.2.3                 minqa_1.2.8              
 [19] RColorBrewer_1.1-3        abind_1.4-8               EnvStats_3.1.0           
 [22] purrr_1.2.1               rmeta_3.0                 msigdbr_25.1.1           
 [25] RCurl_1.98-1.17           ggrepel_0.9.6             pbkrtest_0.5.5           
 [28] irlba_2.3.7               annotate_1.88.0           DelayedMatrixStats_1.32.0
 [31] codetools_0.2-20          DelayedArray_0.36.0       tidyselect_1.2.1         
 [34] farver_2.1.2              lme4_1.1-38               mathjaxr_2.0-0           
 [37] iterators_1.0.14          tools_4.5.2               progress_1.2.3           
 [40] Rcpp_1.1.1                zenith_1.12.0             glue_1.8.0               
 [43] SparseArray_1.10.8        xfun_0.56                 here_1.0.2               
 [46] dplyr_1.2.0               withr_3.0.2               numDeriv_2016.8-1.1      
 [49] fastmap_1.2.0             boot_1.3-32               caTools_1.18.3           
 [52] digest_0.6.39             truncnorm_1.0-9           R6_2.6.1                 
 [55] scattermore_1.2           gtools_3.9.5              dichromat_2.0-0.1        
 [58] RSQLite_2.4.5             RhpcBLASctl_0.23-42       tidyr_1.3.2              
 [61] corpcor_1.6.10            prettyunits_1.2.0         httr_1.4.7               
 [64] S4Arrays_1.10.1           pkgconfig_2.0.3           gtable_0.3.6             
 [67] blob_1.3.0                S7_0.2.1                  XVector_0.50.0           
 [70] remaCor_0.0.20            zigg_0.0.2                GSEABase_1.72.0          
 [73] scales_1.4.0              png_0.1-8                 fANCOVA_0.6-1            
 [76] reformulas_0.4.4          ashr_2.2-63               knitr_1.51               
 [79] rstudioapi_0.18.0         rjson_0.2.23              reshape2_1.4.5           
 [82] nlme_3.1-168              curl_7.0.0                nloptr_2.2.1             
 [85] cachem_1.1.0              stringr_1.6.0             KernSmooth_2.23-26       
 [88] parallel_4.5.2            vipor_0.4.7               metadat_1.4-0            
 [91] AnnotationDbi_1.72.0      pillar_1.11.1             grid_4.5.2               
 [94] vctrs_0.7.1               gplots_3.3.0              mashr_0.2.79             
 [97] xtable_1.8-4              beeswarm_0.4.0            Rgraphviz_2.54.0         
[100] evaluate_1.0.5            KEGGgraph_1.70.0          invgamma_1.2             
[103] magick_2.9.0              mvtnorm_1.3-3             cli_3.6.5                
[106] locfit_1.5-9.12           compiler_4.5.2            rlang_1.1.7              
[109] crayon_1.5.3              SQUAREM_2021.1            plyr_1.8.9               
[112] ggbeeswarm_0.7.3          stringi_1.8.7             assertthat_0.2.1         
[115] babelgene_22.9            lmerTest_3.2-0            Biostrings_2.78.0        
[118] aod_1.3.3                 Matrix_1.7-4              hms_1.1.4                
[121] sparseMatrixStats_1.22.0  bit64_4.6.0-1             KEGGREST_1.50.0          
[124] statmod_1.5.1             rbibutils_2.4.1           Rfast_2.1.5.2            
[127] broom_1.0.12              memoise_2.0.1             RcppParallel_5.1.11-1    
[130] bit_4.6.0                 EnrichmentBrowser_2.40.0 
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-03-01
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   NA
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [1] CRAN (R 4.5.0)
 annotate               1.88.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationDbi          1.72.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 aod                    1.3.3      2023-12-13 [1] CRAN (R 4.5.0)
 ashr                   2.2-63     2023-08-21 [1] CRAN (R 4.5.0)
 assertthat             0.2.1      2019-03-21 [1] CRAN (R 4.5.0)
 babelgene              22.9       2022-09-29 [1] CRAN (R 4.5.0)
 backports              1.5.0      2024-05-23 [1] CRAN (R 4.5.0)
 beeswarm               0.4.0      2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0      2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1    2025-01-16 [1] CRAN (R 4.5.0)
 bitops                 1.0-9      2024-10-03 [1] CRAN (R 4.5.0)
 blob                   1.3.0      2026-01-14 [1] CRAN (R 4.5.2)
 boot                   1.3-32     2025-08-29 [1] CRAN (R 4.5.2)
 broom                  1.0.12     2026-01-27 [1] CRAN (R 4.5.2)
 cachem                 1.1.0      2024-05-16 [1] CRAN (R 4.5.0)
 caTools                1.18.3     2024-09-04 [1] CRAN (R 4.5.0)
 cli                    3.6.5      2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3      2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 corpcor                1.6.10     2021-09-16 [1] CRAN (R 4.5.0)
 crayon                 1.5.3      2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0      2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1   2026-01-27 [1] CRAN (R 4.5.2)
 DBI                    1.2.3      2024-06-02 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 DelayedMatrixStats     1.32.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0      2026-02-03 [1] CRAN (R 4.5.2)
 dreamlet             * 1.8.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 edgeR                * 4.8.2      2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 EnrichmentBrowser      2.40.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 EnvStats               3.1.0      2025-04-24 [1] CRAN (R 4.5.0)
 evaluate               1.0.5      2025-08-27 [1] CRAN (R 4.5.0)
 fANCOVA                0.6-1      2020-11-13 [1] CRAN (R 4.5.0)
 farver                 2.1.2      2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4      2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggbeeswarm             0.7.3      2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2      2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6      2024-09-07 [1] CRAN (R 4.5.0)
 glue                   1.8.0      2024-09-30 [1] CRAN (R 4.5.0)
 gplots                 3.3.0      2025-11-30 [1] CRAN (R 4.5.2)
 graph                  1.88.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 GSEABase               1.72.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 gtable                 0.3.6      2024-10-25 [1] CRAN (R 4.5.0)
 gtools                 3.9.5      2023-11-20 [1] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [1] CRAN (R 4.5.0)
 hms                    1.1.4      2025-10-17 [1] CRAN (R 4.5.0)
 httr                   1.4.7      2023-08-15 [1] CRAN (R 4.5.0)
 invgamma               1.2        2025-07-02 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [1] CRAN (R 4.5.2)
 iterators              1.0.14     2022-02-05 [1] CRAN (R 4.5.0)
 KEGGgraph              1.70.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KEGGREST               1.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KernSmooth             2.23-26    2025-01-01 [1] CRAN (R 4.5.2)
 knitr                  1.51       2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7     2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5      2026-01-08 [1] CRAN (R 4.5.2)
 limma                * 3.66.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 lme4                   1.1-38     2025-12-02 [1] CRAN (R 4.5.2)
 lmerTest               3.2-0      2026-01-13 [1] CRAN (R 4.5.2)
 locfit                 1.5-9.12   2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0      2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [1] CRAN (R 4.5.0)
 mashr                  0.2.79     2023-10-18 [1] CRAN (R 4.5.0)
 MASS                   7.3-65     2025-02-28 [1] CRAN (R 4.5.2)
 mathjaxr               2.0-0      2025-12-01 [1] CRAN (R 4.5.2)
 Matrix                 1.7-4      2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1      2021-11-26 [1] CRAN (R 4.5.0)
 metadat                1.4-0      2025-02-04 [1] CRAN (R 4.5.0)
 metafor                4.8-0      2025-01-28 [1] CRAN (R 4.5.0)
 minqa                  1.2.8      2024-08-17 [1] CRAN (R 4.5.0)
 mixsqp                 0.3-54     2023-12-20 [1] CRAN (R 4.5.0)
 msigdbr                25.1.1     2025-07-21 [1] CRAN (R 4.5.2)
 mvtnorm                1.3-3      2025-01-10 [1] CRAN (R 4.5.0)
 nlme                   3.1-168    2025-03-31 [1] CRAN (R 4.5.2)
 nloptr                 2.2.1      2025-03-17 [1] CRAN (R 4.5.0)
 numDeriv               2016.8-1.1 2019-06-06 [1] CRAN (R 4.5.0)
 otel                   0.2.0      2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1     2026-01-08 [1] CRAN (R 4.5.2)
 pbkrtest               0.5.5      2025-07-18 [1] CRAN (R 4.5.0)
 pillar                 1.11.1     2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [1] CRAN (R 4.5.0)
 plyr                   1.8.9      2023-10-02 [1] CRAN (R 4.5.0)
 png                    0.1-8      2022-11-29 [1] CRAN (R 4.5.0)
 prettyunits            1.2.0      2023-09-24 [1] CRAN (R 4.5.0)
 progress               1.2.3      2023-12-06 [1] CRAN (R 4.5.0)
 purrr                  1.2.1      2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1      2025-02-15 [1] CRAN (R 4.5.0)
 rbibutils              2.4.1      2026-01-21 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3      2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [1] CRAN (R 4.5.2)
 RcppParallel           5.1.11-1   2025-08-27 [1] CRAN (R 4.5.0)
 RCurl                  1.98-1.17  2025-03-22 [1] CRAN (R 4.5.0)
 Rdpack                 2.6.5      2026-01-23 [1] CRAN (R 4.5.2)
 reformulas             0.4.4      2026-02-02 [1] CRAN (R 4.5.2)
 remaCor                0.0.20     2025-08-20 [1] CRAN (R 4.5.0)
 reshape2               1.4.5      2025-11-12 [1] CRAN (R 4.5.0)
 Rfast                  2.1.5.2    2025-10-10 [1] CRAN (R 4.5.0)
 Rgraphviz              2.54.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 RhpcBLASctl            0.23-42    2023-02-11 [1] CRAN (R 4.5.0)
 rjson                  0.2.23     2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7      2026-01-09 [1] CRAN (R 4.5.2)
 rmeta                  3.0        2018-03-20 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1      2025-08-26 [1] CRAN (R 4.5.0)
 RSQLite                2.4.5      2025-11-30 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0     2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1      2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0      2025-04-24 [1] CRAN (R 4.5.0)
 scattermore            1.2        2023-06-12 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 sparseMatrixStats      1.22.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 SpatialExperiment    * 1.20.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SQUAREM                2021.1     2021-01-13 [1] CRAN (R 4.5.0)
 statmod                1.5.1      2025-10-09 [1] CRAN (R 4.5.0)
 stringi                1.8.7      2025-03-27 [1] CRAN (R 4.5.0)
 stringr                1.6.0      2025-11-04 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1      2026-01-11 [1] CRAN (R 4.5.2)
 tidyr                  1.3.2      2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [1] CRAN (R 4.5.0)
 truncnorm              1.0-9      2023-03-20 [1] CRAN (R 4.5.0)
 variancePartition    * 1.40.1     2025-12-11 [1] Bioconductor 3.22 (R 4.5.2)
 vctrs                  0.7.1      2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7      2023-12-18 [1] CRAN (R 4.5.0)
 withr                  3.0.2      2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56       2026-01-18 [1] CRAN (R 4.5.2)
 XML                    3.99-0.20  2025-11-08 [1] CRAN (R 4.5.0)
 xtable                 1.8-4      2019-04-21 [1] CRAN (R 4.5.0)
 XVector                0.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 zenith                 1.12.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 zigg                   0.0.2      2025-02-07 [1] CRAN (R 4.5.0)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────