---
title: "01-Make_VisXferDomain_DGELists"
author: "Bernie Mulvey"
date: "06-08-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(dreamlet)
library(edgeR)
library(limma)
library(SpatialExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
xlc <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")
tmpcd <- as.data.table(colData(xlc),keep.rownames=T)
tmprd <- as.data.table(rowData(xlc))

lcd <- fread("processed-data_newseg/07_visLabelTransfer/05-srtLCsmoothDoms_and_NMFpreds_lut.txt")

# if LC domain, assign LC label; else, keep nmf predicted SRT label

lcd[,domain8426:=ifelse(domain8426=="LC","LC",SRTnmfpred)]
lcd[,domain12199:=ifelse(domain12199=="LC","LC",SRTnmfpred)]
lcd[,domain16119:=ifelse(domain16119=="LC","LC",SRTnmfpred)]
lcd[,domain20079:=ifelse(domain20079=="LC","LC",SRTnmfpred)]

tmpcd <- merge.data.table(tmpcd,lcd[,.(rn,domain8426,domain12199,domain16119,domain20079)],by="rn")

tmpcd[,APOEsimple:=ifelse(APOE %in% c("E4hom","E4het"),"E4","E2")]
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(xlc),]
tmpcd$rn <- NULL

colData(xlc) <- tmpcd

rm(tmpcd)
```

### !!!! !!! !!! ####
Oligo_Complex and ACHE_SERT_LC_WM labels were only transferred to ~500 cells each (reassuringly, since these are the two very obvious multi-type spot classifications). In fact, a cell classified as these is almost certainly not one cell, because this kind of transcriptional profile corresponds explicitly to MIXED CELLS. so drop these two clusters, EVEN THE INTRA-LC DOMAIN cells thereof.

This results in removal 12 O_C and 26 A_S_L_M cells that get assigned to at least one of these domain definitions. so removing all cells effect should be negligible on LC domain itself
```{r}
keeprn <- lcd[!(SRTnmfpred %in% c("Oligo_Complex","ACHE_SERT_LC_WM")),rn]
xlc <- xlc[,colnames(xlc) %in% keeprn]

dim(xlc)
## 366 490574
```

pseudobulk
```{r}
dgels <- lapply(c("domain8426","domain12199","domain16119","domain20079"),function(d){
  ag <- dreamlet::aggregateToPseudoBulk(xlc,assay="counts",sample_id="brnum",cluster_id = d,BPPARAM = MulticoreParam(8))
  
  agcd <- as.data.table(colData(ag),keep.rownames=T)
  setnames(agcd,"rn","brnum")

  agn <- as.data.table(ag@int_colData$n_cells,keep.rownames=T)
  setnames(agn,"rn","clus")
  agn <- melt.data.table(agn,id.vars = "clus")
  setnames(agn,"variable","brnum")
  agn[,brnum:=as.character(brnum)]
  agn[,clus_brnum:=paste0(clus,"_",brnum)]
  agn <- agn[value>=10] ## these are the samples we want to keep to start: 10+ cells for domain/nmfpred type
  setnames(agn,"value","ncells")

  
  agm <- as.list(assays(ag))
  agc <- unlist(mapply(x=agm,n=names(agm),SIMPLIFY=FALSE,function(x,n){
    x <- x[,colnames(x) %in% agn[clus==n,brnum]]
    if(is.vector(x)){return(0)} else {
    sum(greaterOrEqual(colSums(x),2000))}})) ## number of pseudobulks with 10+ cells also having 2000+ counts
  ## only keep those where at least 10 samples meet that
  kpb <- names(agc[agc>=10])
  agn <- agn[clus %in% kpb]
  agm <- agm[kpb]
  
  ## now we want to make one giant sample matrix with the remaining samples
  agm.filt <- mapply(a=agm,n=names(agm),SIMPLIFY=FALSE,function(a,n){
    a <- a[,colnames(a) %in% agn[clus==n]$brnum]
    colnames(a) <- paste0(n,"_",colnames(a))
    return(a)
  })
  agm.filt <- do.call("cbind",agm.filt)
  
  # make a per cluster-donor samples table
  agn <- merge.data.table(agn,agcd[,.(brnum,sample_id,Sex,APOE,APOEsimple,YRI,Age)],by="brnum")
  stopifnot(all(agn$clus_brnum %in% colnames(agm.filt)))
  agn <- as.data.frame(agn,row.names=agn$clus_brnum)[colnames(agm.filt),]
  dge <- DGEList(counts=agm.filt, samples=agn)
  
  return(dge)
  
})
names(dgels) <- c("LCsmooth8426_orxfer","LCsmooth12199_orxfer","LCsmooth16119_orxfer","LCsmooth20079_orxfer")

### append a gene info table to each
tmprd <- tmprd[,.(ID,Symbol)]
setnames(tmprd,c("gene_id","gene_name"))
tmprd <- as.data.frame(tmprd,row.names=tmprd$gene_name)

dgels <- lapply(dgels,function(x){
  tmprd2 <- copy(tmprd)
  tmprd2 <- tmprd2[rownames(x$counts),]
  stopifnot(nrow(tmprd2)==nrow(tmprd))
  x$genes <- tmprd2
  return(x)  
})

saveRDS(dgels,"processed-data_newseg/08_VisXferDomainDE/01a-unfilt_unnorm_pbulk_LCdomain_and_nmfpred_DGElists_2kctsampmin.RDS")
```

seshinf
```{r}
sessionInfo()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] parallelly_1.47.0           colorout_1.3-3             
 [3] SpatialExperiment_1.20.0    edgeR_4.8.2                
 [5] dreamlet_1.8.0              SingleCellExperiment_1.32.0
 [7] SummarizedExperiment_1.40.0 Biobase_2.70.0             
 [9] GenomicRanges_1.62.1        Seqinfo_1.0.0              
[11] IRanges_2.44.0              S4Vectors_0.48.1           
[13] BiocGenerics_0.56.0         generics_0.1.4             
[15] MatrixGenerics_1.22.0       matrixStats_1.5.0          
[17] variancePartition_1.40.1    BiocParallel_1.44.0        
[19] limma_3.66.0                ggplot2_4.0.2              
[21] data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] splines_4.5.2             bitops_1.0-9             
  [3] tibble_3.3.1              graph_1.88.1             
  [5] XML_3.99-0.20             lifecycle_1.0.5          
  [7] Rdpack_2.6.5              mixsqp_0.3-54            
  [9] rprojroot_2.1.1           lattice_0.22-7           
 [11] MASS_7.3-65               backports_1.5.0          
 [13] magrittr_2.0.4            rmarkdown_2.30           
 [15] metafor_4.8-0             yaml_2.3.12              
 [17] otel_0.2.0                DBI_1.2.3                
 [19] minqa_1.2.8               RColorBrewer_1.1-3       
 [21] abind_1.4-8               EnvStats_3.1.0           
 [23] purrr_1.2.1               rmeta_3.0                
 [25] msigdbr_25.1.1            RCurl_1.98-1.17          
 [27] ggrepel_0.9.6             pbkrtest_0.5.5           
 [29] irlba_2.3.7               annotate_1.88.0          
 [31] DelayedMatrixStats_1.32.0 codetools_0.2-20         
 [33] DelayedArray_0.36.1       tidyselect_1.2.1         
 [35] farver_2.1.2              lme4_1.1-38              
 [37] mathjaxr_2.0-0            iterators_1.0.14         
 [39] tools_4.5.2               progress_1.2.3           
 [41] Rcpp_1.1.1                zenith_1.12.0            
 [43] glue_1.8.0                SparseArray_1.10.8       
 [45] xfun_0.56                 here_1.0.2               
 [47] dplyr_1.2.0               withr_3.0.2              
 [49] numDeriv_2016.8-1.1       fastmap_1.2.0            
 [51] boot_1.3-32               caTools_1.18.3           
 [53] digest_0.6.39             truncnorm_1.0-9          
 [55] R6_2.6.1                  scattermore_1.2          
 [57] gtools_3.9.5              dichromat_2.0-0.1        
 [59] RSQLite_2.4.5             RhpcBLASctl_0.23-42      
 [61] tidyr_1.3.2               corpcor_1.6.10           
 [63] prettyunits_1.2.0         httr_1.4.7               
 [65] S4Arrays_1.10.1           pkgconfig_2.0.3          
 [67] gtable_0.3.6              rsconnect_1.7.0          
 [69] blob_1.3.0                S7_0.2.1                 
 [71] XVector_0.50.0            remaCor_0.0.20           
 [73] htmltools_0.5.9           zigg_0.0.2               
 [75] GSEABase_1.72.0           scales_1.4.0             
 [77] png_0.1-8                 fANCOVA_0.6-1            
 [79] reformulas_0.4.4          ashr_2.2-63              
 [81] knitr_1.51                rstudioapi_0.18.0        
 [83] rjson_0.2.23              reshape2_1.4.5           
 [85] nlme_3.1-168              curl_7.0.0               
 [87] nloptr_2.2.1              cachem_1.1.0             
 [89] stringr_1.6.0             KernSmooth_2.23-26       
 [91] parallel_4.5.2            vipor_0.4.7              
 [93] metadat_1.4-0             AnnotationDbi_1.72.0     
 [95] pillar_1.11.1             grid_4.5.2               
 [97] vctrs_0.7.1               gplots_3.3.0             
 [99] mashr_0.2.79              xtable_1.8-4             
[101] beeswarm_0.4.0            Rgraphviz_2.54.0         
[103] evaluate_1.0.5            KEGGgraph_1.70.0         
[105] invgamma_1.2              magick_2.9.0             
[107] mvtnorm_1.3-3             cli_3.6.5                
[109] locfit_1.5-9.12           compiler_4.5.2           
[111] rlang_1.1.7               crayon_1.5.3             
[113] SQUAREM_2021.1            plyr_1.8.9               
[115] ggbeeswarm_0.7.3          stringi_1.8.7            
[117] assertthat_0.2.1          babelgene_22.9           
[119] lmerTest_3.2-0            Biostrings_2.78.0        
[121] aod_1.3.3                 Matrix_1.7-4             
[123] hms_1.1.4                 sparseMatrixStats_1.22.0 
[125] bit64_4.6.0-1             KEGGREST_1.50.0          
[127] statmod_1.5.1             rbibutils_2.4.1          
[129] Rfast_2.1.5.2             broom_1.0.12             
[131] memoise_2.0.1             RcppParallel_5.1.11-1    
[133] bit_4.6.0                 EnrichmentBrowser_2.40.0 