---
title: "04-OneInteractionPerRun_IntraLCVisLabels_DE"
author: "Bernie Mulvey"
date: "07-14-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(edgeR)
library(limma)
## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
dgels <- readRDS("processed-data_newseg/08_VisXferDomainDE/03a-unfilt_unnorm_pbulk_bySRTpredLabel_withinSmoothedLCOnly_DGElists_2kctsampmin.RDS")

dgels <- lapply(dgels,function(x){
  ## change metadata colnames to match previous code
  colnames(x$samples)[colnames(x$samples)=="clus"] <- "label"
  
  ## make a donor-level meta column
  x$samples$donor <- gsub(x$samples$brnum,pattern="L",replacement="")
  return(x)
})
```


##### In the next SEVERAL blocks, we will define a function that runs each DE analysis so that we can parallelize across dgelists (categorical or continuous YRI) and analysis parameters (defined functions). #####

##### MODEL 1: SEX DE (Sex_cluster "interaction" edgeR style #######
####################################################################
```{r}
sexfun <- function(dgel,n){
    dgelsex <- dgel
    dgelsex$samples$label_sex <- paste0(dgelsex$samples$label, "_", dgel$samples$Sex)
    
    # design mat
    des <- model.matrix(~0 + label_sex + APOEsimple + Age + YRI, data = dgelsex$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "label_sex", replacement = "")
    
    
    # filter using design
    dgelsex <- edgeR::normLibSizes(dgelsex)
    keep <- filterByExpr.DGEList(dgelsex, design = des)
    dgelsex <- dgelsex[keep, , keep.lib.sizes = FALSE]
    dgelsex <- edgeR::normLibSizes(dgelsex)
    
    # define contrasts to pass to makeContrasts (we have to do it this way due to environment weirdness with parallelization)    contrasts_list <- list()
    contrasts_list <- list()
    for (current_label in unique(dgelsex$samples$label)) {
        contrast_name <- paste0(current_label, "_Sex")
        contrast_formula <- paste0(current_label, "_M - ", current_label, "_F")
        
        new_contrast <- list(contrast_formula)
        names(new_contrast) <- contrast_name
        contrasts_list <- c(contrasts_list, new_contrast)
    }
    trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))

    ## fit and fit contrasts
    v.swt <- voomLmFit(dgelsex, design = des, block = as.factor(dgelsex$samples$donor), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    ## get top tables, save an RDS with the DE results and other parts
    sexlabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
    })
    names(sexlabel_detabs) <- colnames(trastmat)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelsex, trasts = trastmat, detabs = sexlabel_detabs)
    
    saveRDS(keeplist, paste0("processed-data_newseg/08_VisXferDomainDE/MODULARIZEDTMP_OUT/04bTMP_keeplist_sexlabel_noMixClusts_", n, ".RDS"))
    gc(full = TRUE)
}
```

#####################################################################
##### MODEL 2: E2/E4 carrier status "interaction" edgeR style #######
#####################################################################
```{r}
e2e4fun <- function(dgel,n){dgelE2E4 <- dgel
    dgelE2E4$samples$APOlargeGrp <- ifelse(dgelE2E4$samples$APOE %in% c("E2hom", "E2het"), yes = "E2", no = "E4")
    stopifnot(length(unique(dgelE2E4$samples$APOlargeGrp)) == 2)
    dgelE2E4$samples$label_E2E4 <- paste0(dgelE2E4$samples$label, "_", dgelE2E4$samples$APOlargeGrp)
    
    des <- model.matrix(~0 + label_E2E4 + Sex + Age + YRI, data = dgelE2E4$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "label_E2E4", replacement = "")

    # Filter using the design specified
    dgelE2E4 <- edgeR::normLibSizes(dgelE2E4)
    keep <- filterByExpr.DGEList(dgelE2E4, design = des)
    dgelE2E4 <- dgelE2E4[keep, , keep.lib.sizes = FALSE]
    dgelE2E4 <- edgeR::normLibSizes(dgelE2E4)
    
    # define contrasts to pass to makeContrasts (we have to do it this way due to environment weirdness with parallelization)
    contrasts_list <- list()
    for (current_label in unique(dgelE2E4$samples$label)) {
        contrast_name <- paste0(current_label, "_E4vE2")
        contrast_formula <- paste0(current_label, "_E4 - ", current_label, "_E2")
        
        new_contrast <- list(contrast_formula)
        names(new_contrast) <- contrast_name
        contrasts_list <- c(contrasts_list, new_contrast)
    }
    trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))

    ## fit and fit contrasts
    v.swt <- voomLmFit(dgelE2E4, design = des, block = as.factor(dgelE2E4$samples$donor), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    ## get top tables, save an RDS with the DE results and other parts
    E2E4label_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
    })
    names(E2E4label_detabs) <- colnames(trastmat)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelE2E4, trasts = trastmat, detabs = E2E4label_detabs)
                     
    saveRDS(keeplist, paste0("processed-data_newseg/08_VisXferDomainDE/MODULARIZEDTMP_OUT/04bTMP_keeplist_E2E4label_noMixClusts_", n, ".RDS"))
    gc(full = TRUE)}
```

##########################################################################
#### MODEL 3: APOE alleles (E2hom/het, 4hom/het) "intxns" edgeR style ####
##########################################################################
```{r}
apogenofun <- function(dgel,n){
    dgelApoAlleles <- dgel
    dgelApoAlleles$samples$label_apo <- paste0(dgelApoAlleles$samples$label, "_", dgelApoAlleles$samples$APOE)
    
    des <- model.matrix(~0 + label_apo + Sex + Age + YRI, data = dgelApoAlleles$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "label_apo", replacement = "")
    
    # filter using the design specified
    dgelApoAlleles <- edgeR::normLibSizes(dgelApoAlleles)
    keep <- filterByExpr.DGEList(dgelApoAlleles, design = des)
    dgelApoAlleles <- dgelApoAlleles[keep, , keep.lib.sizes = FALSE]
    dgelApoAlleles <- edgeR::normLibSizes(dgelApoAlleles)

    contrasts_list <- list()
    for (current_label in unique(dgelApoAlleles$samples$label)) {
        # Create a named vector of contrast formulas for the current label;
        # account for the fact that there are no E2hom pseudobulks for some intra-LC types
        if(!(paste0(current_label,"_E2hom") %in% levels(des))){
          formulas <- c(
              paste0(current_label, "_E2het - (0.5*", current_label, "_E4het + 0.5*", current_label, "_E4hom)"),
              paste0(current_label, "_E4hom - ", current_label, "_E2het"),
              paste0(current_label, "_E4het - ", current_label, "_E2het"),
              paste0(current_label, "_E4hom - ", current_label, "_E4het"),
              paste0(current_label, "_E4hom - ((", current_label, "_E4het + ", current_label, "_E2het)/2)"))
          names(formulas) <- paste0(current_label, c("_E2het_vs_anyE4", "_E4hom_vs_anyE2", "_E4het_vs_anyE2", "_E4hom_vs_E4het", "_E4hom_vs_others"))
        } else {
          formulas <- c(
              paste0(current_label, "_E2hom - (0.5*", current_label, "_E4het + 0.5*", current_label, "_E4hom)"),
              paste0(current_label, "_E2het - (0.5*", current_label, "_E4het + 0.5*", current_label, "_E4hom)"),
              paste0(current_label, "_E4hom - (0.5*", current_label, "_E2het + 0.5*", current_label, "_E2hom)"),
              paste0(current_label, "_E4het - (0.5*", current_label, "_E2het + 0.5*", current_label, "_E2hom)"),
              paste0(current_label, "_E4hom - ", current_label, "_E4het"),
              paste0(current_label, "_E4hom - ((", current_label, "_E4het + ", current_label, "_E2het + ", current_label, "_E2hom)/3)")
          )
          names(formulas) <- paste0(current_label, c("_E2hom_vs_anyE4", "_E2het_vs_anyE4", "_E4hom_vs_anyE2", "_E4het_vs_anyE2", "_E4hom_vs_E4het", "_E4hom_vs_others"))
        }
        contrasts_list <- c(contrasts_list, as.list(formulas))
    }
    trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))

    # fit and fit contrasts
    v.swt <- voomLmFit(dgelApoAlleles, design = des, block = as.factor(dgelApoAlleles$samples$donor), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    # get toptabs and other components, save
    ApoAllelesLabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
    })
    names(ApoAllelesLabel_detabs) <- colnames(trastmat)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelApoAlleles, trasts = trastmat, detabs = ApoAllelesLabel_detabs)
                     
    saveRDS(keeplist, paste0("processed-data_newseg/08_VisXferDomainDE/MODULARIZEDTMP_OUT/04bTMP_keeplist_ApoAllelesLabel_noMixClusts_", n, ".RDS"))
    gc(full = TRUE)
}
```

#############################################################################
#### MODEL 4: label-by-YRI (continuous) "intxns" edgeR style ####
#############################################################################
```{r}
ancesfun <- function(dgel,n){
    dgelances <- dgel
    
    # make design matrices based on YRI being continuous or categorical
    if (is.numeric(dgelances$samples$YRI)) {
        # Continuous YRI setup
        des <- model.matrix(~label:YRI + Sex + APOEsimple + Age, data = dgelances$samples)
        colnames(des) <- gsub(x = colnames(des), pattern = "^label(.*):YRI$", replacement = "\\1_by_YRI")
        trastmat <- paste0(unique(dgelances$samples$label), "_by_YRI") # These are coefficients, not contrasts
        stopifnot(all(trastmat %in% colnames(des)))
    } else {
        # Categorical YRI setup
        dgelances$samples$labAnces <- paste0(dgelances$samples$label, "_", dgelances$samples$YRI)
        des <- model.matrix(~0 + labAnces + Sex + APOEsimple + Age, data = dgelances$samples)
        
        contrasts_list <- list()
        for (current_label in unique(dgelances$samples$label)) {
            contrast_name <- paste0(current_label, "_YRI")
            contrast_formula <- paste0("labAnces", current_label, "_YRI - labAnces", current_label, "_CEU")
            
            new_contrast <- list(contrast_formula)
            names(new_contrast) <- contrast_name
            contrasts_list <- c(contrasts_list, new_contrast)
        }
        trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))
    }
    
    # Filtering using appropriate design from above
    dgelances <- edgeR::normLibSizes(dgelances)
    keep <- filterByExpr.DGEList(dgelances, design = des)
    dgelances <- dgelances[keep, , keep.lib.sizes = FALSE]
    dgelances <- edgeR::normLibSizes(dgelances)

    # Fit
    v.swt <- voomLmFit(dgelances, design = des, block = as.factor(dgelances$samples$donor), adaptive.span = TRUE, sample.weights = TRUE)
    
    # Fit contrasts or coefficients depending on YRI type
    if (is.numeric(dgelances$samples$YRI)) {
        v.swt.fit <- contrasts.fit(v.swt, coefficients = trastmat)
    } else {
        v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    }
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)
    
    # Get results and other components, save
    yrilabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
    })
    names(yrilabel_detabs) <- colnames(v.swt.fit)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelances, trasts = trastmat, detabs = yrilabel_detabs)

    saveRDS(keeplist, paste0("processed-data_newseg/08_VisXferDomainDE/MODULARIZEDTMP_OUT/04bTMP_keeplist_AncesByLabel_noMixClusts_", n, ".RDS"))
    gc(full = TRUE)
}
```

#############################################################################
####### MODEL 5: label*YRI*E4orE2 (cat-cat-cat) "intxns" edgeR style ########
#############################################################################
```{r}
ancesapofun <- function(dgel,n){
    if (is.numeric(dgel$samples$YRI)) {
        message("Subsetting to 75%+ ancestry groups", n)
      newsamp <- as.data.table(dgel$samples,keep.rownames=T)
      newct <- dgel$counts
      
      newsamp <- newsamp[YRI>0.75|YRI<0.25]
      newsamp[,YRIcat:=ifelse(YRI>0.75,"YRI","CEU")]
      newct <- newct[,colnames(newct) %in% newsamp$rn]
      newsamp <- as.data.frame(newsamp,row.names=newsamp$rn)[colnames(newct),]
      newg <- as.data.frame(dgel$genes,row.names=rownames(dgel$genes))[rownames(newct),]
      newdgel <- DGEList(counts=newct,samples=newsamp,genes=newg)
    }
  
    # for the categorical data, set it up and go
    newdgel$samples$APOlargeGrp <- ifelse(newdgel$samples$APOE %in% c("E2hom", "E2het"), yes = "E2", no = "E4")
    newdgel$samples$lab_APOE_PredomAnces <- paste0(newdgel$samples$label, "_", newdgel$samples$APOlargeGrp, "_", newdgel$samples$YRIcat)
    
    # design matrix
    des <- model.matrix(~0 + lab_APOE_PredomAnces + Sex + Age, data = newdgel$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "^lab_APOE_PredomAnces(.*)$", replacement = "\\1")
    
    # filter using the design
    newdgel <- edgeR::normLibSizes(newdgel)
    keep <- filterByExpr.DGEList(newdgel, design = des)
    newdgel <- newdgel[keep, , keep.lib.sizes = FALSE]
    newdgel <- edgeR::normLibSizes(newdgel)

    # define contrasts
    contrasts_list <- list()
    for (current_label in unique(newdgel$samples$label)) {
        formulas <- c(
            paste0(current_label, "_E4_YRI - ", current_label, "_E4_CEU"),
            paste0(current_label, "_E2_YRI - ", current_label, "_E2_CEU"),
            paste0(current_label, "_E4_CEU - ", current_label, "_E2_CEU"),
            paste0(current_label, "_E4_YRI - ", current_label, "_E2_YRI")
        )
        names(formulas) <- paste0(current_label, c("_E4YRI_vs_E4CEU", "_E2YRI_vs_E2CEU", "_E4CEU_vs_E2CEU", "_E4YRI_vs_E2YRI"))
        contrasts_list <- c(contrasts_list, as.list(formulas))
    }
    trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))

    # fit, fit contrasts
    v.swt <- voomLmFit(newdgel, design = des, block = as.factor(newdgel$samples$donor), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    # get toptabs, other results, and save
    yriapolabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        o <- as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"),keep.rownames=T)
        return(o)
    })
    names(yriapolabel_detabs) <- colnames(v.swt.fit)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = newdgel, trasts = trastmat, detabs = yriapolabel_detabs)
                     
    saveRDS(keeplist, paste0("processed-data_newseg/08_VisXferDomainDE/MODULARIZEDTMP_OUT/04bTMP_keeplist_YRIAPOlabel_noMixClusts_", n, ".RDS"))
    gc(full = TRUE)
}
```


## now run all of these simultaneously
```{r}
funs <- list(sexfun,e2e4fun,apogenofun,ancesfun,ancesapofun)

mapply(dgel=dgels,n=names(dgels),SIMPLIFY = FALSE,FUN=function(dgel,n){
    lapply(X=funs,FUN=function(f){
        f(dgel,n)
    })
})
```

## clean up, load the lists and concatenate into one giant result obj per domain definition set
```{r}
rm(list=ls())
gc(full=T)
```

## load the tmp RDSes
```{r}
tmp04 <- list.files("processed-data_newseg/08_VisXferDomainDE/MODULARIZEDTMP_OUT",full.names = TRUE)
tmp04 <- grep(tmp04,pattern="04b",value=T)
dom8426 <- list.files("processed-data_newseg/08_VisXferDomainDE/MODULARIZEDTMP_OUT",pattern="8426_orxfer.RDS",full.names = TRUE)
dom12199 <- list.files("processed-data_newseg/08_VisXferDomainDE/MODULARIZEDTMP_OUT",pattern="12199_orxfer.RDS",full.names = TRUE)
dom16119 <- list.files("processed-data_newseg/08_VisXferDomainDE/MODULARIZEDTMP_OUT",pattern="16119_orxfer.RDS",full.names = TRUE)
dom20079 <- list.files("processed-data_newseg/08_VisXferDomainDE/MODULARIZEDTMP_OUT",pattern="20079_orxfer.RDS",full.names = TRUE)


## combine into one metalist
dom8426out <- lapply(dom8426,readRDS)
names(dom8426out) <- gsub(dom8426,pattern="processed-data_newseg/08_VisXferDomainDE/MODULARIZEDTMP_OUT/04bTMP_keeplist_(.*)_noMix.*$",replacement="\\1")

dom12199out <- lapply(dom12199,readRDS)
names(dom12199out) <- gsub(dom12199,pattern="processed-data_newseg/08_VisXferDomainDE/MODULARIZEDTMP_OUT/04bTMP_keeplist_(.*)_noMix.*$",replacement="\\1")

dom16119out <- lapply(dom16119,readRDS)
names(dom16119out) <- gsub(dom16119,pattern="processed-data_newseg/08_VisXferDomainDE/MODULARIZEDTMP_OUT/04bTMP_keeplist_(.*)_noMix.*$",replacement="\\1")

dom20079out <- lapply(dom20079,readRDS)
names(dom20079out) <- gsub(dom20079,pattern="processed-data_newseg/08_VisXferDomainDE/MODULARIZEDTMP_OUT/04bTMP_keeplist_(.*)_noMix.*$",replacement="\\1")


outlist <- list(dom8426=dom8426out, dom12199=dom12199out,dom16119=dom8426out, dom20079=dom20079out)

saveRDS(outlist,"processed-data_newseg/08_VisXferDomainDE/04-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_withinLCsrtNMFpredTypes.RDS")
```
## can now delete 04bTMP RDSes.

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] parallelly_1.47.0           colorout_1.3-3              SpatialExperiment_1.20.0   
 [4] edgeR_4.8.2                 dreamlet_1.8.0              SingleCellExperiment_1.32.0
 [7] SummarizedExperiment_1.40.0 Biobase_2.70.0              GenomicRanges_1.62.1       
[10] Seqinfo_1.0.0               IRanges_2.44.0              S4Vectors_0.48.1           
[13] BiocGenerics_0.56.0         generics_0.1.4              MatrixGenerics_1.22.0      
[16] matrixStats_1.5.0           variancePartition_1.40.1    BiocParallel_1.44.0        
[19] limma_3.66.0                ggplot2_4.0.2               data.table_1.18.4          

loaded via a namespace (and not attached):
  [1] splines_4.5.2             bitops_1.0-9              tibble_3.3.1             
  [4] graph_1.88.1              XML_3.99-0.20             lifecycle_1.0.5          
  [7] Rdpack_2.6.5              mixsqp_0.3-54             rprojroot_2.1.1          
 [10] lattice_0.22-7            MASS_7.3-65               backports_1.5.0          
 [13] magrittr_2.0.4            metafor_4.8-0             rmarkdown_2.30           
 [16] yaml_2.3.12               otel_0.2.0                sessioninfo_1.2.3        
 [19] DBI_1.2.3                 minqa_1.2.8               RColorBrewer_1.1-3       
 [22] abind_1.4-8               EnvStats_3.1.0            purrr_1.2.1              
 [25] rmeta_3.0                 msigdbr_25.1.1            RCurl_1.98-1.17          
 [28] ggrepel_0.9.6             pbkrtest_0.5.5            irlba_2.3.7              
 [31] annotate_1.88.0           DelayedMatrixStats_1.32.0 codetools_0.2-20         
 [34] DelayedArray_0.36.1       tidyselect_1.2.1          farver_2.1.2             
 [37] lme4_1.1-38               mathjaxr_2.0-0            iterators_1.0.14         
 [40] tools_4.5.2               progress_1.2.3            Rcpp_1.1.1               
 [43] zenith_1.12.0             glue_1.8.0                SparseArray_1.10.8       
 [46] xfun_0.56                 here_1.0.2                dplyr_1.2.0              
 [49] withr_3.0.2               numDeriv_2016.8-1.1       fastmap_1.2.0            
 [52] boot_1.3-32               caTools_1.18.3            digest_0.6.39            
 [55] truncnorm_1.0-9           R6_2.6.1                  scattermore_1.2          
 [58] gtools_3.9.5              dichromat_2.0-0.1         RSQLite_2.4.5            
 [61] RhpcBLASctl_0.23-42       tidyr_1.3.2               corpcor_1.6.10           
 [64] prettyunits_1.2.0         httr_1.4.7                S4Arrays_1.10.1          
 [67] pkgconfig_2.0.3           gtable_0.3.6              rsconnect_1.7.0          
 [70] blob_1.3.0                S7_0.2.1                  XVector_0.50.0           
 [73] remaCor_0.0.20            htmltools_0.5.9           zigg_0.0.2               
 [76] GSEABase_1.72.0           scales_1.4.0              png_0.1-8                
 [79] fANCOVA_0.6-1             reformulas_0.4.4          ashr_2.2-63              
 [82] knitr_1.51                rstudioapi_0.18.0         rjson_0.2.23             
 [85] reshape2_1.4.5            curl_7.0.0                nlme_3.1-168             
 [88] nloptr_2.2.1              cachem_1.1.0              stringr_1.6.0            
 [91] KernSmooth_2.23-26        vipor_0.4.7               metadat_1.4-0            
 [94] AnnotationDbi_1.72.0      pillar_1.11.1             grid_4.5.2               
 [97] vctrs_0.7.1               gplots_3.3.0              mashr_0.2.79             
[100] xtable_1.8-4              beeswarm_0.4.0            Rgraphviz_2.54.0         
[103] evaluate_1.0.5            KEGGgraph_1.70.0          magick_2.9.0             
[106] invgamma_1.2              mvtnorm_1.3-3             cli_3.6.5                
[109] locfit_1.5-9.12           compiler_4.5.2            rlang_1.1.7              
[112] crayon_1.5.3              SQUAREM_2021.1            plyr_1.8.9               
[115] ggbeeswarm_0.7.3          stringi_1.8.7             assertthat_0.2.1         
[118] babelgene_22.9            lmerTest_3.2-0            Biostrings_2.78.0        
[121] aod_1.3.3                 Matrix_1.7-4              hms_1.1.4                
[124] sparseMatrixStats_1.22.0  bit64_4.6.0-1             KEGGREST_1.50.0          
[127] statmod_1.5.1             rbibutils_2.4.1           Rfast_2.1.5.2            
[130] broom_1.0.12              memoise_2.0.1             RcppParallel_5.1.11-1    
[133] bit_4.6.0                 EnrichmentBrowser_2.40.0 
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-07-14
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [1] CRAN (R 4.5.0)
 annotate               1.88.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationDbi          1.72.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 aod                    1.3.3      2023-12-13 [1] CRAN (R 4.5.0)
 ashr                   2.2-63     2023-08-21 [1] CRAN (R 4.5.0)
 assertthat             0.2.1      2019-03-21 [1] CRAN (R 4.5.0)
 babelgene              22.9       2022-09-29 [1] CRAN (R 4.5.0)
 backports              1.5.0      2024-05-23 [1] CRAN (R 4.5.0)
 beeswarm               0.4.0      2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0      2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1    2025-01-16 [1] CRAN (R 4.5.0)
 bitops                 1.0-9      2024-10-03 [1] CRAN (R 4.5.0)
 blob                   1.3.0      2026-01-14 [1] CRAN (R 4.5.2)
 boot                   1.3-32     2025-08-29 [1] CRAN (R 4.5.2)
 broom                  1.0.12     2026-01-27 [1] CRAN (R 4.5.2)
 cachem                 1.1.0      2024-05-16 [1] CRAN (R 4.5.0)
 caTools                1.18.3     2024-09-04 [1] CRAN (R 4.5.0)
 cli                    3.6.5      2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3      2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 corpcor                1.6.10     2021-09-16 [1] CRAN (R 4.5.0)
 crayon                 1.5.3      2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0      2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.4     2026-05-06 [1] CRAN (R 4.5.2)
 DBI                    1.2.3      2024-06-02 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.1     2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 DelayedMatrixStats     1.32.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0      2026-02-03 [1] CRAN (R 4.5.2)
 dreamlet             * 1.8.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 edgeR                * 4.8.2      2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 EnrichmentBrowser      2.40.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 EnvStats               3.1.0      2025-04-24 [1] CRAN (R 4.5.0)
 evaluate               1.0.5      2025-08-27 [1] CRAN (R 4.5.0)
 fANCOVA                0.6-1      2020-11-13 [1] CRAN (R 4.5.0)
 farver                 2.1.2      2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4      2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1     2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 ggbeeswarm             0.7.3      2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2      2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6      2024-09-07 [1] CRAN (R 4.5.0)
 glue                   1.8.0      2024-09-30 [1] CRAN (R 4.5.0)
 gplots                 3.3.0      2025-11-30 [1] CRAN (R 4.5.2)
 graph                  1.88.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 GSEABase               1.72.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 gtable                 0.3.6      2024-10-25 [1] CRAN (R 4.5.0)
 gtools                 3.9.5      2023-11-20 [1] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [1] CRAN (R 4.5.0)
 hms                    1.1.4      2025-10-17 [1] CRAN (R 4.5.0)
 htmltools              0.5.9      2025-12-04 [1] CRAN (R 4.5.2)
 httr                   1.4.7      2023-08-15 [1] CRAN (R 4.5.0)
 invgamma               1.2        2025-07-02 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [1] CRAN (R 4.5.2)
 iterators              1.0.14     2022-02-05 [1] CRAN (R 4.5.0)
 KEGGgraph              1.70.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KEGGREST               1.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KernSmooth             2.23-26    2025-01-01 [1] CRAN (R 4.5.2)
 knitr                  1.51       2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7     2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5      2026-01-08 [1] CRAN (R 4.5.2)
 limma                * 3.66.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 lme4                   1.1-38     2025-12-02 [1] CRAN (R 4.5.2)
 lmerTest               3.2-0      2026-01-13 [1] CRAN (R 4.5.2)
 locfit                 1.5-9.12   2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0      2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [1] CRAN (R 4.5.0)
 mashr                  0.2.79     2023-10-18 [1] CRAN (R 4.5.0)
 MASS                   7.3-65     2025-02-28 [1] CRAN (R 4.5.2)
 mathjaxr               2.0-0      2025-12-01 [1] CRAN (R 4.5.2)
 Matrix                 1.7-4      2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1      2021-11-26 [1] CRAN (R 4.5.0)
 metadat                1.4-0      2025-02-04 [1] CRAN (R 4.5.0)
 metafor                4.8-0      2025-01-28 [1] CRAN (R 4.5.0)
 minqa                  1.2.8      2024-08-17 [1] CRAN (R 4.5.0)
 mixsqp                 0.3-54     2023-12-20 [1] CRAN (R 4.5.0)
 msigdbr                25.1.1     2025-07-21 [1] CRAN (R 4.5.2)
 mvtnorm                1.3-3      2025-01-10 [1] CRAN (R 4.5.0)
 nlme                   3.1-168    2025-03-31 [1] CRAN (R 4.5.2)
 nloptr                 2.2.1      2025-03-17 [1] CRAN (R 4.5.0)
 numDeriv               2016.8-1.1 2019-06-06 [1] CRAN (R 4.5.0)
 otel                   0.2.0      2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.47.0     2026-04-17 [1] CRAN (R 4.5.2)
 pbkrtest               0.5.5      2025-07-18 [1] CRAN (R 4.5.0)
 pillar                 1.11.1     2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [1] CRAN (R 4.5.0)
 plyr                   1.8.9      2023-10-02 [1] CRAN (R 4.5.0)
 png                    0.1-8      2022-11-29 [1] CRAN (R 4.5.0)
 prettyunits            1.2.0      2023-09-24 [1] CRAN (R 4.5.0)
 progress               1.2.3      2023-12-06 [1] CRAN (R 4.5.0)
 purrr                  1.2.1      2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1      2025-02-15 [1] CRAN (R 4.5.0)
 rbibutils              2.4.1      2026-01-21 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3      2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [1] CRAN (R 4.5.2)
 RcppParallel           5.1.11-1   2025-08-27 [1] CRAN (R 4.5.0)
 RCurl                  1.98-1.17  2025-03-22 [1] CRAN (R 4.5.0)
 Rdpack                 2.6.5      2026-01-23 [1] CRAN (R 4.5.2)
 reformulas             0.4.4      2026-02-02 [1] CRAN (R 4.5.2)
 remaCor                0.0.20     2025-08-20 [1] CRAN (R 4.5.0)
 reshape2               1.4.5      2025-11-12 [1] CRAN (R 4.5.0)
 Rfast                  2.1.5.2    2025-10-10 [1] CRAN (R 4.5.0)
 Rgraphviz              2.54.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 RhpcBLASctl            0.23-42    2023-02-11 [1] CRAN (R 4.5.0)
 rjson                  0.2.23     2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7      2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30       2025-09-28 [1] CRAN (R 4.5.0)
 rmeta                  3.0        2018-03-20 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1      2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0      2025-12-06 [1] CRAN (R 4.5.2)
 RSQLite                2.4.5      2025-11-30 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0     2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1     2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1      2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0      2025-04-24 [1] CRAN (R 4.5.0)
 scattermore            1.2        2023-06-12 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 sparseMatrixStats      1.22.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 SpatialExperiment    * 1.20.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SQUAREM                2021.1     2021-01-13 [1] CRAN (R 4.5.0)
 statmod                1.5.1      2025-10-09 [1] CRAN (R 4.5.0)
 stringi                1.8.7      2025-03-27 [1] CRAN (R 4.5.0)
 stringr                1.6.0      2025-11-04 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1      2026-01-11 [1] CRAN (R 4.5.2)
 tidyr                  1.3.2      2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [1] CRAN (R 4.5.0)
 truncnorm              1.0-9      2023-03-20 [1] CRAN (R 4.5.0)
 variancePartition    * 1.40.1     2025-12-11 [1] Bioconductor 3.22 (R 4.5.2)
 vctrs                  0.7.1      2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7      2023-12-18 [1] CRAN (R 4.5.0)
 withr                  3.0.2      2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56       2026-01-18 [1] CRAN (R 4.5.2)
 XML                    3.99-0.20  2025-11-08 [1] CRAN (R 4.5.0)
 xtable                 1.8-4      2019-04-21 [1] CRAN (R 4.5.0)
 XVector                0.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12     2025-12-10 [1] CRAN (R 4.5.2)
 zenith                 1.12.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 zigg                   0.0.2      2025-02-07 [1] CRAN (R 4.5.0)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────────────────