---
title: "03c-Concat_raw_NMDAPInewseg_SPEs_Jun26BJM"
author: "Bernie Mulvey"
date: "06-03-26"
output: html_document
---

```{r}
library(data.table)
library(here)
library(SpatialExperiment)
library(SingleCellExperiment)
library(readxl)
library(tidyverse)

#files <- list.files(here('processed-data/xenium/NM_DAPI_rawSPE/'), pattern = "\\.RDS$", full.names = TRUE)
files <- list.files(here('processed-data/01_spe/NMDAPInewseg_rawSPE'), pattern = "\\.RDS$", full.names = TRUE)

# 1) Read all SPEs and tag brnum
spe_list <- lapply(files, function(f) {
  spe <- readRDS(f)
  br  <- gsub(f,pattern="^.*/01_spe/NMDAPInewseg_rawSPE/(.*)\\.RDS$",replacement="\\1")
  br <- gsub(br,pattern="-",replacement="")
  colData(spe)$brnum <- br
  colData(spe)$sample_id <- br # make the spe-recognized sample ids unique, ie. brnum
  colData(spe)$unique_cellid <- paste0(spe$brnum,"_",spe$Barcode)
  spe
})

combined_spe <- do.call(cbind, spe_list)
colnames(combined_spe) <- combined_spe$unique_cellid

## append metadata
demos <- fread(here("../LFF_spatial_LC/raw-data/demos.txt"))
demos <- unique(demos[,.(Donors,Age,Sex,Ancestry,Rin,APOE,hemisphere)])
demos[hemisphere=="L",Donors:=paste0(Donors,"L")]
demos[Donors=="Br6119(re-dis)",Donors:="Br6119"]
demos <- demos[Donors %in% unique(combined_spe$brnum)]
stopifnot(all(unique(combined_spe$brnum) %in% demos$Donors))

tmpcd <- as.data.table(colData(combined_spe),keep.rownames=T)
tmpcd <- merge.data.table(tmpcd,demos,by.x="brnum",by.y="Donors")

tmpcd[,donor:=gsub(brnum,pattern="L",replacement="")]

## append YRI/CEU/CHB estimates
flares <- fread(here("../LFF_spatial_LC/raw-data/ancestryAndHaplotypes/FLARE/global_ancestry_estimates.txt"))
flares <- flares[ID %in% tmpcd$donor]
tmpcd <- merge.data.table(tmpcd,flares,by.x="donor",by.y="ID")

stopifnot(nrow(tmpcd)==ncol(combined_spe))
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(combined_spe),]
tmpcd$rn <- NULL # redundant with unique_cellid
tmpcd$hemisphere <- NULL
tmpcd$Genotype <- NULL
colData(combined_spe) <- tmpcd

saveRDS(combined_spe, here("processed-data_newseg/01_spe/", "NMDAPI_raw_combined_spe.RDS"))
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] lubridate_1.9.4             forcats_1.0.1               stringr_1.6.0              
 [4] dplyr_1.2.0                 purrr_1.2.1                 readr_2.1.6                
 [7] tidyr_1.3.2                 tibble_3.3.1                ggplot2_4.0.2              
[10] tidyverse_2.0.0             readxl_1.4.5                SpatialExperiment_1.20.0   
[13] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0 Biobase_2.70.0             
[16] GenomicRanges_1.62.1        Seqinfo_1.0.0               IRanges_2.44.0             
[19] S4Vectors_0.48.1            BiocGenerics_0.56.0         generics_0.1.4             
[22] MatrixGenerics_1.22.0       matrixStats_1.5.0           here_1.0.2                 
[25] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] gtable_0.3.6        rjson_0.2.23        xfun_0.56           lattice_0.22-7     
 [5] tzdb_0.5.0          vctrs_0.7.1         tools_4.5.2         R.oo_1.27.1        
 [9] pkgconfig_2.0.3     Matrix_1.7-4        RColorBrewer_1.1-3  S7_0.2.1           
[13] lifecycle_1.0.5     compiler_4.5.2      farver_2.1.2        htmltools_0.5.9    
[17] yaml_2.3.12         pillar_1.11.1       R.utils_2.13.0      rsconnect_1.7.0    
[21] DelayedArray_0.36.1 sessioninfo_1.2.3   magick_2.9.0        abind_1.4-8        
[25] tidyselect_1.2.1    digest_0.6.39       stringi_1.8.7       rprojroot_2.1.1    
[29] fastmap_1.2.0       grid_4.5.2          cli_3.6.5           SparseArray_1.10.8 
[33] magrittr_2.0.4      S4Arrays_1.10.1     dichromat_2.0-0.1   withr_3.0.2        
[37] scales_1.4.0        timechange_0.4.0    rmarkdown_2.30      XVector_0.50.0     
[41] otel_0.2.0          cellranger_1.1.0    R.methodsS3_1.8.2   hms_1.1.4          
[45] evaluate_1.0.5      knitr_1.51          rlang_1.1.7         Rcpp_1.1.1         
[49] glue_1.8.0          rstudioapi_0.18.0   R6_2.6.1           
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-02
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ─────────────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cellranger             1.1.0    2016-07-27 [1] CRAN (R 4.5.0)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.1   2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                * 1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 forcats              * 1.0.1    2025-09-25 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 ggplot2              * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                 * 1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 hms                    1.1.4    2025-10-17 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 lubridate            * 1.9.4    2024-12-08 [1] CRAN (R 4.5.0)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 purrr                * 1.2.1    2026-01-09 [1] CRAN (R 4.5.2)
 R.methodsS3            1.8.2    2022-06-13 [1] CRAN (R 4.5.0)
 R.oo                   1.27.1   2025-05-02 [1] CRAN (R 4.5.0)
 R.utils                2.13.0   2025-02-24 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 readr                * 2.1.6    2025-11-14 [1] CRAN (R 4.5.2)
 readxl               * 1.4.5    2025-03-07 [1] CRAN (R 4.5.0)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1   2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 stringi                1.8.7    2025-03-27 [1] CRAN (R 4.5.0)
 stringr              * 1.6.0    2025-11-04 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble               * 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyr                * 1.3.2    2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 tidyverse            * 2.0.0    2023-02-22 [1] CRAN (R 4.5.0)
 timechange             0.4.0    2026-01-29 [1] CRAN (R 4.5.2)
 tzdb                   0.5.0    2025-03-15 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

────────────────────────────────────────────────────────────────────────────────────────────