---
title: "03-Make_XenDemos_table"
format: html
date: "06-03-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
unloadNamespace('datasets')
```

```{r}
### gah we still don't have a plain demos tsv here, need to pull from spe.
lcx <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

demos <- as.data.table(colData(lcx))
demos <- unique(demos[,.(brnum,Age,Sex,Ancestry,Rin,APOE,CEU,YRI,CHB)])
demos[,APOEcarriertype:=ifelse(APOE %in% c("E4hom","E4het"),"E4","E2")]
reordn <- names(demos)[]
demos[,donorID:=gsub(brnum,pattern="L$",replacement="")]
reordn <- c("donorID",reordn)
demos <- demos[,..reordn]
fwrite(demos,"raw-data/demos_with_hemispheres_newseg.txt",sep='\t',quote=F,row.names=F,col.names=T)
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices
[5] utils     methods   base     

other attached packages:
 [1] SpatialExperiment_1.20.0   
 [2] SingleCellExperiment_1.32.0
 [3] SummarizedExperiment_1.40.0
 [4] Biobase_2.70.0             
 [5] GenomicRanges_1.62.1       
 [6] Seqinfo_1.0.0              
 [7] IRanges_2.44.0             
 [8] S4Vectors_0.48.0           
 [9] BiocGenerics_0.56.0        
[10] generics_0.1.4             
[11] MatrixGenerics_1.22.0      
[12] matrixStats_1.5.0          
[13] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] cli_3.6.5           knitr_1.51         
 [3] magick_2.9.0        rlang_1.1.7        
 [5] xfun_0.56           otel_0.2.0         
 [7] DelayedArray_0.36.0 rjson_0.2.23       
 [9] rprojroot_2.1.1     grid_4.5.2         
[11] evaluate_1.0.5      abind_1.4-8        
[13] compiler_4.5.2      Rcpp_1.1.1         
[15] XVector_0.50.0      here_1.0.2         
[17] lattice_0.22-7      SparseArray_1.10.8 
[19] magrittr_2.0.4      Matrix_1.7-4       
[21] tools_4.5.2         S4Arrays_1.10.1    
