---
title: "05-Lowres_Xenium_HEimages_for_SPE"
author: "Bernie Mulvey"
date: "2026-06-03"
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(SpatialExperiment)
library(imager)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

library(parallelly)
library(parallel)

```

### running this all thru JHPCE conda_R/4.5 interactive command line. code placed here for concurrent markdown.
```{r}
## if necess
## setwd("/dcs05/lieber/marmaypag/LFF_spatialLC_LIBD4140/LFF_xenium_LC")

## load filtered xenium data
spe <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

## get image filepaths
brainz <- list.files("processed-data/xenium_imageProcessing_new/registrations/",pattern="_registered.png",recursive = FALSE)
brainz.dt <- as.data.table(cbind(brainz,gsub(brainz,pattern="-L",replacement="L")))
setnames(brainz.dt,c("dir","brnum"))
brainz.dt[,brnum:=gsub(brnum,pattern="^(Br.*)_registered\\.png",replacement="\\1")
brainz.dt <- brainz.dt[brnum %in% spe$brnum]

stopifnot(all(unique(spe$brnum) %in% brainz.dt$brnum))
fwrite(brainz.dt,"processed-data_newseg/02_QC/05-tmpbrainzdt.txt",sep='\t',quote=F,row.names=F,col.names=T)

```

load it
```{r}
brainz.dt <- fread("processed-data_newseg/02_QC/05-tmpbrainzdt.txt")
```

guess we can't parallelize something png-y here. so iterate...
```{r}
if(!dir.exists("processed-data/xenium_imageProcessing_new/HE_DAPIregist_lowres")){
  dir.create("processed-data/xenium_imageProcessing_new/HE_DAPIregist_lowres")  
}

b<-1

for (b in c(1:nrow(brainz.dt))){
    samp <- brainz.dt[b,brnum]
    fullres.path <- paste0("processed-data/xenium_imageProcessing_new/registrations/",brainz.dt[b,dir])
    
    ## make an output directory for our files
    if (!dir.exists(paste0("processed-data/xenium_imageProcessing_new/HE_DAPIregist_lowres/",samp))){
      dir.create(paste0("processed-data/xenium_imageProcessing_new/HE_DAPIregist_lowres/",samp))
    }
    
    outdir <- paste0("processed-data/xenium_imageProcessing_new/HE_DAPIregist_lowres/",samp,"/")
    
    # load the full res HE-DAPI registered image
    fullres <- load.image(file.path(fullres.path))
    
    ## calculate scale factors for a 800 px max-dimension image
    lowres_max_size=800
    hires_scalef=0.2125# um/px

    low_over_hi <- lowres_max_size / max(dim(fullres)[seq(2)])
    lowres_scalefact <- low_over_hi/hires_scalef
    sr_json <- list(
        tissue_hires_scalef = hires_scalef,
        tissue_lowres_scalef = lowres_scalefact,
        spot_diameter_fullres = (5/hires_scalef)) # 5um nuclei / (.2125 um/px) = px 
    
    outimg <- imager::resize(
    fullres,
    as.integer(low_over_hi * dim(fullres)[1]),
    as.integer(low_over_hi * dim(fullres)[2]))
    
    imager::save.image(outimg,paste0(outdir,"tissue_lowres_image.png"))
    
    write(
        rjson::toJSON(sr_json),
        file.path(paste0(outdir,"scalefactors_json.json")))
    
    rm(fullres,outimg,sr_json,lowres_scalefact,low_over_hi,outdir,samp,dir)
    gc(full=T)
}
```

```{r}
spes <- lapply(unique(spe$brnum),function(b){spe[,spe$brnum==b]})
names(spes) <- unique(spe$brnum)

newspes <- mcmapply(d=spes,n=names(spes),mc.cores=10,SIMPLIFY=FALSE,function(d,n){
    img <- readImgData(path=paste0("processed-data/xenium_imageProcessing_new/HE_DAPIregist_lowres/",n,"/"),sample_id = n,scaleFactors = paste0("processed-data/xenium_imageProcessing_new/HE_DAPIregist_lowres/",n,"/scalefactors_json.json"))
    
    newspe <- SpatialExperiment(assays=list(counts=counts(d)),sample_id = n,spatialCoords = spatialCoords(d),imgData = img,colData=colData(d),rowData=rowData(d))

    return(newspe)    
})
names(newspes) <- names(spes)

i<-1
for (i in c(1:length(newspes))){
    
    if(i==1){
        outspe <- newspes[[i]]
    }
    else {
        outspe <- cbind(outspe,newspes[[i]])
    }
}
stopifnot(dim(outspe)==dim(spe))
dim(outspe) ## 
length(unique(outspe$brnum)) ##
```

save
```{r}
saveRDS(outspe,"processed-data_newseg/02_QC/05a-spe_filtered_genetargonly_log-and-nonlog-norms_withLowresHEs.RDS")
```


```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.0 Patched (2025-05-21 r88220)
Platform: x86_64-conda-linux-gnu
Running under: Rocky Linux 9.4 (Blue Onyx)

Matrix products: default
BLAS:   /jhpce/shared/community/core/conda_R/4.5/R/lib64/R/lib/libRblas.so 
LAPACK: /jhpce/shared/community/core/conda_R/4.5/R/lib64/R/lib/libRlapack.so;  LAPACK version 3.12.1

locale:
 [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C              
 [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8    
 [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8   
 [7] LC_PAPER=en_US.UTF-8       LC_NAME=C                 
 [9] LC_ADDRESS=C               LC_TELEPHONE=C            
[11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C       

time zone: US/Eastern
tzcode source: system (glibc)

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices datasets  utils    
[8] methods   base     

other attached packages:
 [1] imager_1.0.8                magrittr_2.0.4             
 [3] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
 [5] SummarizedExperiment_1.40.0 Biobase_2.70.0             
 [7] GenomicRanges_1.62.1        Seqinfo_1.0.0              
 [9] IRanges_2.44.0              S4Vectors_0.48.0           
[11] BiocGenerics_0.56.0         generics_0.1.4             
[13] MatrixGenerics_1.22.0       matrixStats_1.5.0          
[15] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] Matrix_1.7-4        compiler_4.5.0      rjson_0.2.23       
 [4] Rcpp_1.1.1          stringr_1.6.0       magick_2.9.1       
 [7] bmp_0.3.1           png_0.1-8           lattice_0.22-9     
[10] XVector_0.50.0      S4Arrays_1.10.1     igraph_2.2.2       
[13] DelayedArray_0.36.0 readbitmap_0.1.5    tiff_0.1-12        
[16] rlang_1.1.7         stringi_1.8.7       SparseArray_1.10.9 
[19] cli_3.6.5           grid_4.5.0          lifecycle_1.0.5    
[22] vctrs_0.7.1         glue_1.8.0          abind_1.4-8        
[25] purrr_1.2.1         tools_4.5.0         jpeg_0.1-11        
[28] pkgconfig_2.0.3    

─ Session info ───────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.0 Patched (2025-05-21 r88220)
 os       Rocky Linux 9.4 (Blue Onyx)
 system   x86_64, linux-gnu
 ui       X11
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       US/Eastern
 date     2026-03-25
 pandoc   3.7.0.1 @ /jhpce/shared/community/core/conda_R/4.5/bin/pandoc
 quarto   NA

─ Packages ───────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [2] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.0)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.0)
 bmp                    0.3.1    2025-09-22 [1] CRAN (R 4.5.0)
 cli                    3.6.5    2025-04-23 [2] CRAN (R 4.5.0)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [2] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.0)
 glue                   1.8.0    2024-09-30 [2] CRAN (R 4.5.0)
 igraph                 2.2.2    2026-02-12 [1] CRAN (R 4.5.0)
 imager               * 1.0.8    2025-12-23 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.0)
 jpeg                   0.1-11   2025-03-21 [2] CRAN (R 4.5.0)
 lattice                0.22-9   2026-02-09 [1] CRAN (R 4.5.0)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.0)
 magick                 2.9.1    2026-02-28 [1] CRAN (R 4.5.0)
 magrittr             * 2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] CRAN (R 4.5.0)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.0)
 matrixStats          * 1.5.0    2025-01-07 [2] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [2] CRAN (R 4.5.0)
 png                    0.1-8    2022-11-29 [2] CRAN (R 4.5.0)
 purrr                  1.2.1    2026-01-09 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.0)
 readbitmap             0.1.5    2018-06-27 [1] CRAN (R 4.5.0)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.0)
 S4Arrays               1.10.1   2025-12-01 [1] Bioconductor 3.22 (R 4.5.0)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.0)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.0)
 sessioninfo            1.2.3    2025-02-05 [2] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.0)
 SparseArray            1.10.9   2026-03-06 [1] Bioconductor 3.22 (R 4.5.0)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.0)
 stringi                1.8.7    2025-03-27 [2] CRAN (R 4.5.0)
 stringr                1.6.0    2025-11-04 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.0)
 tiff                   0.1-12   2023-11-28 [2] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.0)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.0)

 [1] /users/bmulvey/R/4.5
 [2] /jhpce/shared/community/core/conda_R/4.5/R/lib64/R/site-library
 [3] /jhpce/shared/community/core/conda_R/4.5/R/lib64/R/library
 [4] /users/bmulvey/R/4.3
 * ── Packages attached to the search path.

──────────────────────────────────────────────────────────────────────────────