---
title: "04-1thresh_srtLCdomain_fineK2step_regists_and_spatialplots"
author: "Bernie Mulvey"
date: "06-05-2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(ggrastr)
library(SpatialExperiment)
library(escheR)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https:/stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallel)


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")

""
```


first let's make some plots of the relationship between visium registration scores and manual LC label inclusion
```{r}
regs <- fread("processed-data_newseg/07_visLabelTransfer/04c-ONEcond_NMF_to_KNNsmoothed_NMFsrtLCs_regist_to_manXenLC_visLC_finek2steps.txt")
sort(unique(regs$k2)) ## check that these are 3 sig digits and 
#regs[,k2:=round(as.numeric(k2),3)]



regsplt1 <- regs[Ntopdommks %in% c(10,40,70,100,130,160,190,220,240,280,300)] # visiium max overlap 240
regsplt1 <- melt.data.table(regsplt1,id.vars = c("domainDef","Ntopdommks","looseInDom","strictInDom","nmfpredInDom","k1","k2"))
```

## make the plots like we did for 07/03, plus some line plots since there's a lot of k2 values stepped thru per k1 here:

```{r}
pdf("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04a_LFFvis_regist_and_PropNMFpredLCinDom_top240mks_k2facet.pdf",height=12,width=30)
ggplot(regsplt1[Ntopdommks==240&variable=="LFF_Visium"],aes(x=nmfpredInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=10)+
  ylab("Registration to Visium LC")
dev.off()

pdf("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04a_LFFvis_regist_and_PropManLCStrict_top240mks_k2facet.pdf",height=12,width=30)
ggplot(regsplt1[Ntopdommks==240&variable=="LFF_Visium"],aes(x=strictInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=10)+
  ylab("Registration to Visium LC")
dev.off()

pdf("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04a_LFFvis_regist_and_PropNMFpredLCinDom_top40mks_k2facet.pdf",height=12,width=30)
ggplot(regsplt1[Ntopdommks==40&variable=="LFF_Visium"],aes(x=nmfpredInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=10)+
  ylab("Registration to Visium LC")
dev.off()

pdf("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04a_LFFvis_regist_and_PropManLCStrict_top40mks_k2facet.pdf",height=12,width=30)
ggplot(regsplt1[Ntopdommks==40&variable=="LFF_Visium"],aes(x=strictInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=10)+
  ylab("Registration to Visium LC")
dev.off()

###

pdf("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04b_LFFvis_regist_lines_top240mks_k2.pdf",height=5.5,width=12)
ggplot(regsplt1[Ntopdommks==240&variable=="LFF_Visium"],aes(x=k2,y=value,color=factor(k1)))+
  geom_line()+facet_wrap(~k1)+
  geom_point()+
  ylab("Registration to Visium LC")
dev.off()

pdf("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04b_LFFvis_regist_lines_top40mks_k2.pdf",height=5.5,width=12)
ggplot(regsplt1[Ntopdommks==40&variable=="LFF_Visium"],aes(x=k2,y=value,color=factor(k1)))+
  geom_line()+facet_wrap(~k1)+
  geom_point()+
  ylab("Registration to Visium LC")
dev.off()

pdf("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04b_LCstrict_regist_lines_top40mks_k2.pdf",height=5.5,width=12)
ggplot(regsplt1[Ntopdommks==40&variable=="LFF_Xenium_strict"],aes(x=k2,y=value,color=factor(k1)))+
  geom_line()+facet_wrap(~k1)+
  geom_point()+
  ylab("Registration to Strict Manual LC Cells")
dev.off()

pdf("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04b_LCstrict_regist_lines_top300mks_k2.pdf",height=5.5,width=12)
ggplot(regsplt1[Ntopdommks==300&variable=="LFF_Xenium_strict"],aes(x=k2,y=value,color=factor(k1)))+
  geom_line()+facet_wrap(~k1)+
  geom_point()+
  ylab("Registration to Strict Manual LC Cells")
dev.off()

## proportion of label-xfered LC cells in domain and nmfpred LC cells registration
pdf("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04c_srtNMFLCcells_regist_and_NMFxferLCcellProp_top300mks_k2facet.pdf",height=12,width=30)
ggplot(regsplt1[Ntopdommks==300&variable=="Xenium_All_NMFpred_LCcells"],aes(x=nmfpredInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=10)+
  ylab("Registration to Cells NMF-Xfered LFFvis LC Label")
dev.off()

pdf("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04c_srtNMFLCcells_regist_and_NMFxferLCcellProp_top40mks_k2facet.pdf",height=12,width=30)
ggplot(regsplt1[Ntopdommks==40&variable=="Xenium_All_NMFpred_LCcells"],aes(x=nmfpredInDom,y=value,color=factor(k1)))+
  geom_point()+
  facet_wrap(~k2,ncol=10)+
  ylab("Registration to Cells NMF-Xfered LFFvis LC Label")
dev.off()


## how much better does each domain definition register to Visium LC vs pseudobulking nmfpred LC cells?
pdf("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04d_VisiumLC-srtNMFprojLCcells_registdiff_top40mks_k1facet.pdf",height=5.5,width=12)
ggplot(regs[Ntopdommks==40],aes(y=(LFF_Visium-Xenium_All_NMFpred_LCcells),x=k2,col=as.factor(k1)))+
  geom_point()+
  geom_line()
dev.off()

pdf("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04d_VisiumLC-srtNMFprojLCcells_registdiff_top240mks_k1facet.pdf",height=5.5,width=12)
ggplot(regs[Ntopdommks==240],aes(y=(LFF_Visium-Xenium_All_NMFpred_LCcells),x=k2,col=as.factor(k1)))+
  geom_point()+
  geom_line()
dev.off()

pdf("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04e_VisiumLC-LCstrict_registdiff_top40mks_k1facet.pdf",height=5.5,width=12)
ggplot(regs[Ntopdommks==40],aes(y=(LFF_Visium-LFF_Xenium_strict),x=k2,col=as.factor(k1)))+
  geom_point()+
  geom_line()
dev.off()

pdf("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04e_VisiumLC-LCstrict_registdiff_top240mks_k1facet.pdf",height=5.5,width=12)
ggplot(regs[Ntopdommks==240],aes(y=(LFF_Visium-LFF_Xenium_strict),x=k2,col=as.factor(k1)))+
  geom_point()+
  geom_line()
dev.off()

rm(regs,regsplt1)
```

## get KNN smoothing and transferred labels
```{r}
dualscores <- readRDS("processed-data_newseg/07_visLabelTransfer/02b-srtXferLCknnsmoothing_step2k75.RDS")
## we only need up to k1=0.2
dualscores <- dualscores[1:5]

## we want to label individual "LC" cells that fall outside the domain to assess their boundaries, so retrieve the cellwise visium-nmf-predicted-labels. 
xferlab <- readRDS("processed-data_newseg/07_visLabelTransfer/01-VistoXen_NMFlabelxfer_tol1e8_k63_alph0.RDS")
xferlab <- xferlab$targets
visxfer <- lapply(xferlab,function(x){
  x$nmf_preds <- as.character(x$nmf_preds)
  y <- as.data.table(colData(x),keep.rownames=T)
  # y$nmf_preds <- as.character(y$nmf_preds) 
  y <- y[,.(rn,nmf_preds)]
  setnames(y,2,"nmfpred")
  return(y)
})
visxfer <- rbindlist(visxfer)

# visxfer[nmfpred=="LC",nmfpred:="nonDom_NE"] ## making this something different to keep track of where we're making this change/looking for it

## now attach
dualscores <- lapply(dualscores,function(x){
  y <- merge.data.table(x,visxfer,by="rn")
  return(y)
})


spe <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

logcounts(spe) <- NULL #dont need it

## cleanup
rm(xferlab)
gc(full=T)
```

```{r}
pal <- readRDS("../LFF_spatial_LC/plots/finalpal_somuchfornorainbows.RDS")

names(pal)
pal <- pal[c(8,1,6)] # green blue red respec.
pal <- c(pal,"#DEDEDE")

# may seem odd now but makes it easiest to see LC cells within the domain
names(pal) <- c("Domain","nonDom_NE","LC_cell","Other")
```

## run through the combos that seem promising on both visium registration and LC cell retention
# plots with one donor's xenium sample and all visium samples in a row for each threshold
```{r}
lcv <- readRDS("../LFF_spatial_LC/processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

vdom <- readRDS("../LFF_spatial_LC/processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
vdom <- vdom$HARMONYlmbna_HDG_SVG_2575
vdom <- vdom[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(vdom,2,"clusid")

vanno <- fread("../LFF_spatial_LC/processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")
vdom <- merge.data.table(vdom,vanno[,.(clusid,anno)],by="clusid")
vdom[,anno:=ifelse(anno=="LC","Domain","Other")] ## for pal consistency with how we're marking up the xenium plots
vdom <- DataFrame(vdom,row.names=vdom$rn)[colnames(lcv),]

colLabels(lcv) <- vdom$anno

rm(vdom,vanno)

lcv$brnum[lcv$brnum=="Br6119(re-dis)"] <- "Br6119"
## drop donors with limited LC
lcv <- lcv[,!(lcv$brnum %in% paste0("Br",c(5517,5276,5712)))]

## we also want to match left/right sections here so change brnum to reflect the hemi of origin
lcv$brnum <- paste0(lcv$brnum,lcv$hemisphere)
lcv$brnum <- gsub(lcv$brnum,pattern="(Br.*)R",replacement="\\1")
sum(unique(lcv$brnum) %in% unique(spe$brnum))==length(unique(spe$brnum)) # k we good
```

### one plot file per top candidate threshold set and incorporating visium samples.

### LEFT OFF HERE

### prioritization strategy: look at the line plots for stable regions of 2+ points w/r/t registration values; take notes on top peak registration k1, k2 pairs for each of 
-240 marker Vis LFF regs
0.08: 0.426, 0.439, 0.452
0.12: 0.199, .212, .226
0.16: cant tell
0.2: cant tell


-40 marker vis lff regs not used--too variable step to step
0.08: cant tell
0.12: cant tell
0.16: .052, .066, .079, .119, .132, .146
0.2: .079, .092, .106, .119

-240 marker (visium domain MINUS SRT NMF-xfer labeled LC cells).
0.08: cant tell
0.12: cant tell
0.16: cant tell
0.2: .079 .092 .106

-40 marker (visium domain MINUS SRT NMF-xfer labeled LC cells).
0.08: cant tell
0.12: .199, .212
0.16: .119, .132, .146, .159
0.2: .079 .092 .106

300 marker, all manual LC strict cells

0.08: .346, .359, .372, .426, .439
0.12: 0.199 / can't tell
0.16: 0.092, 0.106
0.2: .066, .079

40 marker, all manual LC strict cells
0.08: .426, .439
0.12: .212, .226, .239
0.16: .092, .106, .119, .132, .146, .159
0.2: .066 .079 .092 .106

240 marker ∆visium, all strict
0.08: .426/cant tell
0.12: .212, .226
0.16: cant tell
0.2: .079, .092, .106

40 marker ∆visium, all strict
0.08: can't tell
0.12: .199, .212
0.16: .119, .132, .146, .159
0.2: .066 .079, .092, .106

## get recurring sets from the above...the following pairs emerge as consistently prioritized (note these k2 values were already shifted by -0.001 to avoid rounding and machine precision mistakes.):
```{r}
priort <- list(
  "0.08"=c(0.426, 0.439, 0.452, 0.346, 0.359, 0.372, 0.426, 0.439, 0.426, 0.439, 0.426),
  "0.12"=c(0.199, 0.212, 0.226, 0.199, 0.212, 0.199, 0.212, 0.226, 0.239, 0.212, 0.226, 0.199, 0.212, 0.119),
  "0.16"=c(0.052, 0.066, 0.079, 0.119, 0.132, 0.146, 0.119, 0.132, 0.146, 0.159, 0.092, 0.106, 0.092, 0.106, 0.119, 0.132, 0.146, 0.159, 0.119, 0.132, 0.146, 0.159),
  "0.20"=c(0.079, 0.092, 0.106, 0.119, 0.079, 0.092, 0.106, 0.079, 0.092, 0.106, 0.066, 0.079, 0.066, 0.079, 0.092, 0.106, 0., 0.079, 0.092, 0.106, 0.066, 0.079, 0.092, 0.106))
priort <- lapply(priort,as.data.table)
priort <- rbindlist(priort,idcol="k1")
setnames(priort,"V1","k2thresh")
priort[,k2thresh:=round(k2thresh,3)]
priort[,k1:=round(as.numeric(k1),2)]
priort <- unique(priort[,Noccur:=.N,by=c("k1","k2thresh")][greaterOrEqual(Noccur,3)])
#         k1 k2thresh     N
#     <char>    <num> <int>
#  1:   0.08    0.426     4
#  2:   0.08    0.439     3
#  3:   0.12    0.199     4
#  4:   0.12    0.212     5
#  5:   0.12    0.226     3
#  6:   0.16    0.119     4
#  7:   0.16    0.132     4
#  8:   0.16    0.146     4
#  9:   0.16    0.159     3
# 10:   0.20    0.079     7
# 11:   0.20    0.092     6
# 12:   0.20    0.106     6
# 13:   0.20    0.066     3
```


Plot:
.08 / .426, .439
.12 / .199, .212
.16 / .119, .132, .146
.2 / 0.079, .092, .106

## QC visualization: for comparing Xenium and Visium domain structures. we can focus on a smaller subset with the cell type plots (subsequent block) once we have some good structural matches from here.
```{r}
kpairs <- priort[(k1==0.08 & k2thresh>0.42 & k2thresh<0.44)|(k1==0.12 & k2thresh>0.19 & k2thresh<0.22)|(k1==0.16 & k2thresh>0.11 & k2thresh<.15)|(greaterOrEqual(k1,0.19) & Noccur>4)] 


setDTthreads(1,restore_after_fork = FALSE)
lapply(c(1:nrow(kpairs)),function(i){
  k1 <- kpairs[i,k1]
  k2 <- kpairs[i,k2thresh]

  n <- paste0("k1thresh_",k1)
  d <- copy(dualscores[[n]]) # single brackets don't return a d.t. i guess, but double brackets do

  d <- as.data.table(d)
  d[,isLC:=ifelse(
    #greaterOrEqual(lcpred.smooth1.k25,k1) & 
    greaterOrEqual(lcpred.smooth2.k75,k2),"Domain","Other")]

  
  ## we just want domain/other for these visualizations.

  d2 <- DataFrame(d,row.names=d$rn)[colnames(spe),]
  spe$pltclus <- d2$isLC
  
  plts <- list()
  j<-1
  for (j in c(1:length(unique(spe$brnum)))){
    s <- unique(spe$brnum)[j]
    sublist <- list()
    px <- make_escheR(spe[,spe$brnum==s], y_reverse=FALSE)
    px <- rasterize(add_fill(p = px,var="pltclus",point_size=0.75),layers="Points",dpi=300,dev="cairo_png")
    px <- px+scale_fill_manual(values = pal)
    px <- px+guides(fill="none",color="none")
    px <- px+ggtitle(s)

    sublist[[1]] <- px

    i<-1
    for (i in c(1:length(unique(lcv[,lcv$brnum==s]$sample_id)))){
      curs <- unique(lcv[,lcv$brnum==s]$sample_id)[i]
      curd <- lcv[,lcv$brnum==s & lcv$sample_id==curs]
      pv <- make_escheR(curd, y_reverse=FALSE)
      pv <- rasterize(add_fill(p=pv,var="label",size=1.25,point_size=1.25),layers="Points",dpi=300,dev="cairo_png")
      pv <- pv+scale_fill_manual(
        values = pal)+guides(color="none",fill="none")
      pv <- pv+ggtitle(curs)
      sublist[[(i+1)]] <- pv
    }
    retobj <- gridExtra::arrangeGrob(grobs = sublist, ncol = length(sublist))
    plts[[j]] <- retobj
  }
  dev.off()

  png(paste0("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04f-NMFtransferLC_smoothedDom_knn25_k1_",k1,"_knn75_k2_",k2,"_VisCompare.png"),height=4*length(plts),width=14,units = "in",res=150)
  grid.arrange(grobs=plts,ncol=1)
  dev.off()
})
```

### QC visualization: emphasizing LC cells within and outside domain for top four as sanity check before finalizing:
Plot:
.08 / .426
0.12 / 0.199
0.16 / 0.119
.2 / .079


```{r}
k1thresh <- round(c(0.08,0.12,0.16,0.20),2)
setDTthreads(1,restore_after_fork = FALSE)
lapply(k1thresh,function(k1){
  if (k1==0.08){
    k2s <- round(0.426,3)
  } else if (k1==0.12){
    k2s <- round(0.199,3)
  } else if (k1==0.16){
    k2s <- round(0.119,3)
  } else if (k1==0.20){
    k2s <- round(0.079,3)
  }
  n <- paste0("k1thresh_",k1)
  d <- copy(dualscores[[n]]) # single brackets don't return a d.t. i guess, but double brackets do

  lapply(k2s,function(k2){
    d <- as.data.table(d)
    d[,isLC:=ifelse(
      #greaterOrEqual(lcpred.smooth1.k25,k1) & 
      greaterOrEqual(lcpred.smooth2.k75,k2),"Domain","Other")]

    ## mark cells outside the defined domain that nmf labeled LC.
    d[nmfpred=="LC",isLC:=ifelse(isLC=="Domain","LC_cell","nonDom_NE")]

    d2 <- DataFrame(d,row.names=d$rn)[colnames(spe),]
    spe$pltclus <- d2$isLC
    
    plts <- list()
    j<-1
    for (j in c(1:length(unique(spe$brnum)))){
      s <- unique(spe$brnum)[j]
      sublist <- list()
      px <- make_escheR(spe[,spe$brnum==s], y_reverse=FALSE)
      px <- rasterize(add_fill(p = px,var="pltclus",point_size=0.75),layers="Points",dpi=300,dev="cairo_png")
      px <- px+scale_fill_manual(values = pal)
      
      # since escheR is making the size aesthetic inaccessible, override it:
      px@layers[[1]]$aes_params$size <- NULL
      ## if we specify a conditional to aes, we can then just set a true/false value for the scale values, but right now i want to see both LC cell types, especially nondomain
      px <- px+aes(size=pltclus)+# %in% c("nonDom_NE"))+
      scale_size_manual(values = c("Domain" = 0.75, "LC_cell" = 1.5, "nonDom_NE"=2.25,"Other" = 0.75))
      ## carry on
      
      px <- px+guides(fill="none",color="none",size="none")
      px <- px+ggtitle(s)

      sublist[[1]] <- px

      i<-1
      for (i in c(1:length(unique(lcv[,lcv$brnum==s]$sample_id)))){
        curs <- unique(lcv[,lcv$brnum==s]$sample_id)[i]
        curd <- lcv[,lcv$brnum==s & lcv$sample_id==curs]
        pv <- make_escheR(curd, y_reverse=FALSE)
        pv <- rasterize(add_fill(p=pv,var="label",size=1.25,point_size=1.25),layers="Points",dpi=300,dev="cairo_png")
        pv <- pv+scale_fill_manual(
          values = pal)+guides(color="none",fill="none")
        pv <- pv+ggtitle(curs)
        sublist[[(i+1)]] <- pv
      }
      retobj <- gridExtra::arrangeGrob(grobs = sublist, ncol = length(sublist))
      plts[[j]] <- retobj
    }
    
    png(paste0("plots_newseg/07_visLabelTransfer/04_ONEcond_visXferLC_smooth_QC_and_spotplots_finesteps/04g-NMFtransferLC_smoothedDom_knn25_k1_",k1,"_knn75_k2_",k2,"_CellsInOut.png"),height=4*length(plts),width=14,units = "in",res=150)
    grid.arrange(grobs=plts,ncol=1)
    dev.off()
  })
})

```

all look fairly similar except for ± small neighborhoods per donor; Br5854 has no domain at 20079 but that's okay--it should be at most very small and the others are a little more permissive about what goes in. use all 4 of these parameter sets.
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] parallelly_1.47.0           colorout_1.3-3              escheR_1.10.0              
 [4] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
 [7] Biobase_2.70.0              GenomicRanges_1.62.1        Seqinfo_1.0.0              
[10] IRanges_2.44.0              S4Vectors_0.48.1            BiocGenerics_0.56.0        
[13] generics_0.1.4              MatrixGenerics_1.22.0       matrixStats_1.5.0          
[16] ggrastr_1.0.2               gridExtra_2.3               ggtext_0.1.2               
[19] ggplot2_4.0.2               data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] gtable_0.3.6        beeswarm_0.4.0      rjson_0.2.23        xfun_0.56          
 [5] lattice_0.22-7      vctrs_0.7.1         tools_4.5.2         tibble_3.3.1       
 [9] pkgconfig_2.0.3     Matrix_1.7-4        RColorBrewer_1.1-3  S7_0.2.1           
[13] lifecycle_1.0.5     compiler_4.5.2      farver_2.1.2        vipor_0.4.7        
[17] htmltools_0.5.9     yaml_2.3.12         pillar_1.11.1       rsconnect_1.7.0    
[21] DelayedArray_0.36.1 magick_2.9.0        abind_1.4-8         tidyselect_1.2.1   
[25] digest_0.6.39       dplyr_1.2.0         labeling_0.4.3      rprojroot_2.1.1    
[29] fastmap_1.2.0       grid_4.5.2          here_1.0.2          cli_3.6.5          
[33] SparseArray_1.10.8  magrittr_2.0.4      S4Arrays_1.10.1     dichromat_2.0-0.1  
[37] withr_3.0.2         scales_1.4.0        ggbeeswarm_0.7.3    rmarkdown_2.30     
[41] XVector_0.50.0      otel_0.2.0          evaluate_1.0.5      knitr_1.51         
[45] rlang_1.1.7         gridtext_0.1.5      Rcpp_1.1.1          glue_1.8.0         
[49] xml2_1.5.2          rstudioapi_0.18.0   R6_2.6.1           
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-03
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ─────────────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.1   2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 escheR               * 1.10.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 ggbeeswarm             0.7.3    2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggrastr              * 1.0.2    2023-06-01 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 labeling               0.4.3    2023-08-29 [1] CRAN (R 4.5.0)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.47.0   2026-04-17 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1   2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.5.0)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

────────────────────────────────────────────────────────────────────────────────────────────