---
title: "01-Concat_NM_measures"
author: "Bernie Mulvey"
date: "06-10-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

## Maddy's original copies of these per-sample tables are found in processed-data/01_spe/NMDAPInewseg_rawSPE. Those (date created/modified June 10 1208p EST) were copied over into processed-data_newseg for clarity, and we'll load them from there.
```{r}
nmcsv <- list.files("processed-data_newseg/01_spe/NM_metrics_and_matched_cells",full.names=T)

nmstats <- lapply(nmcsv,fread)
# get brnum with hemisphere where applicable
names(nmstats) <- gsub(nmcsv,pattern="^.*_and_matched_cells/(Br.*)_NM_xenium_cell_matches.csv",replacement="\\1")
## for consistency with SPE
names(nmstats) <- gsub(names(nmstats),pattern="-L",replacement="L")

## the output CSVs have the within-sample cell ids, but not the globally unique ones. easiest way to handle is recreate those global ids (brnum_id) in each table before concatenating.

## for some reason the names on these tables are not uniformly ordered, so to bind em all together also reord in each table first, using first slot's naming order
nameord <- names(nmstats[[1]])

nmstats <- mapply(s=nmstats,n=names(nmstats),SIMPLIFY=FALSE,function(s,n){
  ## get rid of NM segmentations that were not assigned a cell
  s <- s[!is.na(cell_id)]
  s[,cell_id:=paste0(n,"_",cell_id)]
  s <- s[,..nameord]
  return(s)
})

nmstats2 <- rbindlist(nmstats)
```

## load SPE and check that these cell identifiers all exist(ed at one point)
```{r}
xlc.filt <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")
# how many NM segs were assigned to a cell we still have?
sum(unique(nmstats2$cell_id) %in% colnames(xlc.filt)) #14233

xlc.unfilt <- readRDS("processed-data_newseg/01_spe/NMDAPI_raw_combined_spe.RDS")
## were all these cells in the SPE before QC?
stopifnot(sum(unique(nmstats2$cell_id) %in% colnames(xlc.unfilt))==length(unique(nmstats2$cell_id)))
## ok, good.
```

## there are also 3 cases where 2 NM segmentations were assigned to a cell. since there's so few of those, easier to just drop them than write a methods paragraph about what we did about it.
```{r}
dropids <- nmstats2[,.N,by="cell_id"][N>1,cell_id]
nmstats2 <- nmstats2[!(cell_id %in% dropids)]

## subset to cells in the filtered datset
nmstats2 <- nmstats2[cell_id %in% colnames(xlc)]

## make the nmstats2 brain identifier match everything else we use (no dash in Br####-L)
nmstats2[,brain:=gsub(brain,pattern="-L",replacement="\\1")]

fwrite(nmstats2,"processed-data_newseg/10_NManalyses/01-NMmetrics_per_QCretained_cell_noMultiNMsegCells.txt",sep='\t',quote=F,row.names=F)
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0         parallelly_1.47.0          
 [3] colorout_1.3-3              SpatialExperiment_1.20.0   
 [5] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
 [7] Biobase_2.70.0              GenomicRanges_1.62.1       
 [9] Seqinfo_1.0.0               IRanges_2.44.0             
[11] S4Vectors_0.48.1            BiocGenerics_0.56.0        
[13] generics_0.1.4              MatrixGenerics_1.22.0      
[15] matrixStats_1.5.0           gridExtra_2.3              
[17] ggtext_0.1.2                ggplot2_4.0.2              
[19] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] SparseArray_1.10.8  xml2_1.5.2          lattice_0.22-7     
 [4] digest_0.6.39       magrittr_2.0.4      evaluate_1.0.5     
 [7] grid_4.5.2          RColorBrewer_1.1-3  fastmap_1.2.0      
[10] Matrix_1.7-4        rprojroot_2.1.1     scales_1.4.0       
[13] codetools_0.2-20    abind_1.4-8         cli_3.6.5          
[16] rlang_1.1.7         XVector_0.50.0      DelayedArray_0.36.1
[19] withr_3.0.2         yaml_2.3.12         otel_0.2.0         
[22] S4Arrays_1.10.1     parallel_4.5.2      tools_4.5.2        
[25] dplyr_1.2.0         here_1.0.2          vctrs_0.7.1        
[28] R6_2.6.1            magick_2.9.0        lifecycle_1.0.5    
[31] pkgconfig_2.0.3     pillar_1.11.1       gtable_0.3.6       
[34] rsconnect_1.7.0     glue_1.8.0          Rcpp_1.1.1         
[37] xfun_0.56           tibble_3.3.1        tidyselect_1.2.1   
[40] rstudioapi_0.18.0   knitr_1.51          dichromat_2.0-0.1  
[43] rjson_0.2.23        farver_2.1.2        htmltools_0.5.9    
[46] rmarkdown_2.30      compiler_4.5.2      S7_0.2.1           
[49] gridtext_0.1.5     
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-10
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.1   2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 ggplot2              * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.47.0   2026-04-17 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1   2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────