---
title: "04-LMM_wdLC-srtLC-orManualLC-annotdCells_allNMsegs_logcounts_toNMmetrics_keepranfx"
author: "Bernie Mulvey"
date: "06-15-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(lme4)
library(car)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

for plotting residualized values, we want to have the random effects as well -- which the original version of this code didn't include in the returned output. so we'll re-run, now outputting a random effect coef table and a main effect coef table per gene per APOE variable encoding and NM metric.

read in spe, append YRI genomic proportions, and slide info. this time, subset to NM+ segmentations that mapped to a CELL with an LC label by any approach used (manual expression-based definitions, weber-divecha label transfer, or SRT label transfer).
```{r}
lcx <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")

## label transfers and manual lcs
manlc <- fread("processed-data_newseg/03_ManualLC/01-XprBasedLClabels_and_ManLCcellsWithSegsGainedorAlteredByNM.txt")
wdx <- fread("processed-data_newseg/04_snSeqLabelTransfer/06-wdLCsmoothings_LUT.txt")
srt <- fread("processed-data_newseg/07_visLabelTransfer/05-srtLCsmoothDoms_and_NMFpreds_lut.txt")

lccells <- merge.data.table(manlc,wdx,by="rn")
lccells <- merge.data.table(lccells,srt,by="rn")

lccells <- lccells[LCloose=="LC"|LCstrict=="LC"|SRTnmfpred=="LC"|nmfpred=="NE",rn] # 16.5k cells across all definitions, before considering NM seg regist.

## read in nm measures-retain all for the moment
lcxnm <- fread("processed-data_newseg/10_NManalyses/01-NMmetrics_per_QCretained_cell.txt")
lcxnm <- lcxnm[cell_id %in% lccells,
              .(cell_id,NM_mean_intensity,NM_median_intensity)] 
# ^ 3453 LC cells ever annotated as LC-NE AND registered to an NM seg

lcnm <- DataFrame(lcxnm,row.names=lcxnm$cell_id)[colnames(lcx),]

## assign
colData(lcx)$NM_mean_intensity <- lcnm$NM_mean_intensity
colData(lcx)$NM_median_intensity <- lcnm$NM_median_intensity
# subset to NM+
lcx <- lcx[,!is.na(lcx$NM_mean_intensity)]
stopifnot(ncol(lcx)==nrow(lcxnm))

### get metadata -- only intensity is continuous and appropriate for an LMM. we also need to know the slide for each sample.
flares <- fread("../LFF_spatial_LC/raw-data/ancestryAndHaplotypes/FLARE/global_ancestry_estimates.txt")

lcnm.cd <- as.data.table(colData(lcx),keep.rownames=T)
lcnm.cd <- lcnm.cd[,.(rn,NM_mean_intensity,NM_median_intensity,Age,Sex,APOE,donor,brnum)] # brnum = sample (donor with hemisphere); donor entries do not have hemisphere
setnames(lcnm.cd,"APOE","APOEgeno")

### append YRI
stopifnot(all(lcnm.cd$donor%in%flares$ID))
lcnm.cd <- merge.data.table(lcnm.cd,flares[,.(ID,YRI)],by.x="donor",by.y="ID")

### xenium slide info
techmet <- fread("raw-data/donor_sample_slide_kit.txt")
stopifnot(all(lcnm.cd$brnum %in% techmet$Donor_sample))
lcnm.cd <- merge.data.table(lcnm.cd,techmet[,.(Donor_sample,Xenium_slide)],by.x="brnum",by.y="Donor_sample")

## we'll use two models with different approaches for APOE: E4+/- or individual genotypes (already in APOEgeno as E4hom/E4het/E2hom/E2het)
lcnm.cd[,E4carrier:="No"]
lcnm.cd[APOEgeno %in% c("E4hom","E4het"),E4carrier:="Yes"]

lcnm.lct <- logcounts(lcx)
lcnm.lct <- as.data.frame(as.matrix(lcnm.lct))
```

## clean up for modeling runs
```{r}
rm(lcx,lcxnm,lcnm,techmet,manlc, srt, wdx, flares)
gc(full=T)
```

iterate over each gene and metric and APOE modeling specification
## note that, unlike modeling gene upon NM (expr ~ NM + covars), these models (NM ~ expr + covars) do NOT return "singular fit" warnings en masse.
nested rfx: none for donor (2 samples max per donor, mostly 1); sample (brnum) within slide 

```{r}
setDTthreads(1,restore_after_fork=FALSE)

nmcor_apoes <- lapply(c("E4carrier","APOEgeno"),function(apo){
    
    nmcors <- lapply(c("NM_mean_intensity","NM_median_intensity"),function(nm){
        results <- mclapply(c(1:nrow(lcnm.lct)),mc.cores=8,FUN=function(i){

            setDTthreads(1,restore_after_fork=FALSE)
            xpr <- lcnm.lct[i,]
            xpr <- t(xpr)
            xpr <- as.data.table(xpr,keep.rownames=T)
            xpr <- merge.data.table(xpr,lcnm.cd,by="rn")
            
            curg <- colnames(xpr)[2]
            
            setnames(xpr,2,"val")
            setnames(xpr,nm,"NMvar")
            # fit lmms
            f <- formula(paste0("NMvar ~ val + Age + Sex + YRI +", apo," + (1|Xenium_slide) + (1|Xenium_slide:brnum)"))
            f2 <- formula(paste0("NMvar ~ Age + Sex + YRI + ", apo," + (1|Xenium_slide) + (1|Xenium_slide:brnum)")) # no term for gene
            
            lmmnm <- lmer(f,data=xpr)
            
            ## to perform LRT test, we shouldn't use REML, so run a second non REML
            lmmnm.compare <- lmer(f,data=xpr,REML = FALSE)
            lmmno <- lmer(f2,data=xpr,REML=FALSE)
            
            ## get LRT
            lrt <- stats::anova(lmmnm.compare,lmmno)
            
            ## get anova for model coeffs in the REML NM model (car::Anova)
            an <- as.data.table(Anova(lmmnm),keep.rownames=T)
            
            ## get model coefficients
            fixfx <- as.data.table(fixef(lmmnm),keep.rownames=T)
            setnames(fixfx,c("variable","coef"))
            ## split Sex and APO into a variable column and group column for consistency with the anova table (e.g., SexM -> Sex , M) comes out as Sex here but "SexM" for the fixef. so to make sure we keep the sex coefficients, change SexM to Sex here
            fixfx[variable=="SexM",var_group:="M"]
            fixfx[variable=="SexM",variable:="Sex"]
            # pattern to fix the APOE fixed effects will begin with the variable we're testing
            fixpat <- paste0(apo,"(.*)")
            fixfx[variable %in% grep(variable,pattern=apo,value=T),var_group:=gsub(variable,pattern=fixpat,replacement="\\1")]
            fixfx[variable %in% grep(variable,pattern=apo,value=T),variable:=apo]
            
            ## make cleaner intercept rows
            fixfx[variable=="(Intercept)",variable:="Intercept"]
            ## append model coefficients
            
            an <- merge.data.table(fixfx,an,by.x="variable",by.y="rn")
            an[variable=="val",variable:=curg]
            
            ## append LRT test
            lrt <- as.data.table(lrt,keep.rownames=T)[rn=="lmmnm.compare",.(rn,Chisq,Df,`Pr(>Chisq)`,logLik)]
            ## change 'loglik' to coef; change rn to "LRTvs_no_gene_coef"
            lrt[,rn:="LRTvs_no_gene_coef"]
            setnames(lrt,"rn","variable")
            
            an <- rbind(an,lrt,fill=T)
            
            ## add the NMvar being fitted
            an[,NMvar:=nm]
            
            ## get the ranfx table
            ranfxtab <- as.data.table(ranef(lmmnm))
            ranfxtab[,gene_name:=curg]
            ranfxtab[,NMvar:=nm]
            
            retlist <- list(an,ranfxtab)
            names(retlist) <- c("fixed","ranfx")
            return(retlist)
        })
        names(results) <- rownames(lcnm.lct)
        return(results)
    })
    names(nmcors) <- c("NM_mean_intensity","NM_median_intensity")
    return(nmcors)

})
names(nmcor_apoes) <- c("E4carrierYN","PerGeno")


### a handful of these return error messages indicating one of the variables is constant. https://stats.stackexchange.com/questions/366784/how-to-get-around-the-glmer-warning-downdated-vtv-is-not-positive-definite
    
### only retain entries that are tabular results; concat those tabs for fixed and rand fx
nmcor_apoe_fixfxtabs <- lapply(nmcor_apoes,function(x){
    fixlist <- lapply(x,function(y){
        lapply(y, function(z){
            z[["fixed"]]
        })
    })
    fixlist2 <- lapply(fixlist,FUN=function(x){
        ret <- x[unlist(lapply(x,is.data.table))]
        ret <- rbindlist(ret,idcol="gene_name")
        return(ret)
    })
    return(fixlist2)
})

nmcor_apoe_ranfxtabs <- lapply(nmcor_apoes,function(x){
    ranlist <- lapply(x,function(y){
        lapply(y, function(z){
            z[["ranfx"]]
        })
    })
    ranlist2 <- lapply(ranlist,FUN=function(x){
        ret <- x[unlist(lapply(x,is.data.table))]
        ret <- rbindlist(ret,idcol="gene_name")
        return(ret)
    })
    return(ranlist2)
})


## spot check that we have random effects for all genes that we got fixed effect results for
stopifnot(all(unique(nmcor_apoe_fixfxtabs$E4carrierYN$NM_mean_intensity$gene_name) %in% unique(nmcor_apoe_ranfxtabs$E4carrierYN$NM_mean_intensity$gene_name)))

### final list should rearrange these back together now
finalout <- list(
    E4carrierYN = list(
        NM_mean_intensity = list(
            fixed = nmcor_apoe_fixfxtabs$E4carrierYN$NM_mean_intensity,
            ranfx = nmcor_apoe_ranfxtabs$E4carrierYN$NM_mean_intensity
        ),
        NM_median_intensity = list(
            fixed = nmcor_apoe_fixfxtabs$E4carrierYN$NM_median_intensity,
            ranfx = nmcor_apoe_ranfxtabs$E4carrierYN$NM_median_intensity
        )
    ),
    PerGeno = list(
        NM_mean_intensity = list(
            fixed = nmcor_apoe_fixfxtabs$PerGeno$NM_mean_intensity,
            ranfx = nmcor_apoe_ranfxtabs$PerGeno$NM_mean_intensity
        ),
        NM_median_intensity = list(
            fixed = nmcor_apoe_fixfxtabs$PerGeno$NM_median_intensity,
            ranfx = nmcor_apoe_ranfxtabs$PerGeno$NM_median_intensity
        )
))

# save it!
saveRDS(finalout,"processed-data_newseg/10_NManalyses/04-Manual-wdXfer-orSRTxfer_LClabeled_AllNMsegMapped_meanMedIntensity_LMMed_by_geneLogcounts_2apoeFactorizations_withRanEfFits.RDS")
```

seshinf
```{r}
sessioninfo::session_info()
sessionInfo()
```
─ Session info ──────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-15
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 boot                   1.3-32   2025-08-29 [1] CRAN (R 4.5.2)
 car                  * 3.1-5    2026-02-03 [1] CRAN (R 4.5.2)
 carData              * 3.0-6    2026-01-30 [1] CRAN (R 4.5.2)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.1   2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 Formula                1.2-5    2023-02-24 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 ggplot2              * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 lme4                 * 1.1-38   2025-12-02 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 MASS                   7.3-65   2025-02-28 [1] CRAN (R 4.5.2)
 Matrix               * 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 minqa                  1.2.8    2024-08-17 [1] CRAN (R 4.5.0)
 nlme                   3.1-168  2025-03-31 [1] CRAN (R 4.5.2)
 nloptr                 2.2.1    2025-03-17 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.47.0   2026-04-17 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 rbibutils              2.4.1    2026-01-21 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 Rdpack                 2.6.5    2026-01-23 [1] CRAN (R 4.5.2)
 reformulas             0.4.4    2026-02-02 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1   2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────
> sessionInfo()
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods  
[8] base     

other attached packages:
 [1] car_3.1-5                   carData_3.0-6              
 [3] BiocParallel_1.44.0         parallelly_1.47.0          
 [5] colorout_1.3-3              lme4_1.1-38                
 [7] Matrix_1.7-4                SpatialExperiment_1.20.0   
 [9] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
[11] Biobase_2.70.0              GenomicRanges_1.62.1       
[13] Seqinfo_1.0.0               IRanges_2.44.0             
[15] S4Vectors_0.48.1            BiocGenerics_0.56.0        
[17] generics_0.1.4              MatrixGenerics_1.22.0      
[19] matrixStats_1.5.0           gridExtra_2.3              
[21] ggtext_0.1.2                ggplot2_4.0.2              
[23] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] gtable_0.3.6        rjson_0.2.23        xfun_0.56          
 [4] lattice_0.22-7      vctrs_0.7.1         tools_4.5.2        
 [7] Rdpack_2.6.5        tibble_3.3.1        pkgconfig_2.0.3    
[10] RColorBrewer_1.1-3  S7_0.2.1            lifecycle_1.0.5    
[13] compiler_4.5.2      farver_2.1.2        codetools_0.2-20   
[16] htmltools_0.5.9     yaml_2.3.12         Formula_1.2-5      
[19] nloptr_2.2.1        pillar_1.11.1       MASS_7.3-65        
[22] rsconnect_1.7.0     DelayedArray_0.36.1 sessioninfo_1.2.3  
[25] reformulas_0.4.4    magick_2.9.0        boot_1.3-32        
[28] abind_1.4-8         nlme_3.1-168        tidyselect_1.2.1   
[31] digest_0.6.39       dplyr_1.2.0         splines_4.5.2      
[34] rprojroot_2.1.1     fastmap_1.2.0       grid_4.5.2         
[37] here_1.0.2          cli_3.6.5           SparseArray_1.10.8 
[40] magrittr_2.0.4      S4Arrays_1.10.1     dichromat_2.0-0.1  
[43] withr_3.0.2         scales_1.4.0        rmarkdown_2.30     
[46] XVector_0.50.0      otel_0.2.0          evaluate_1.0.5     
[49] knitr_1.51          rbibutils_2.4.1     rlang_1.1.7        
[52] gridtext_0.1.5      Rcpp_1.1.1          glue_1.8.0         
[55] xml2_1.5.2          rstudioapi_0.18.0   minqa_1.2.8        
[58] R6_2.6.1           