---
title: "06-CompareNMassocs05_to_Visium"
format: html
date: "06-29-26"
author: "Bernie Mulvey"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
# library(SpatialFeatureExperiment)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
precision <- sqrt(.Machine$double.eps)
(x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
precision <- sqrt(.Machine$double.eps)
(x <= y) | (abs(x-y) <= precision)}

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost='localhost')


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = '#000000'), axis.title.x = element_text(size = 9,color = '#000000'), axis.text.y = element_text(size = 8,color = '#000000'), axis.title.y = element_text(size =9,color = '#000000'), plot.title = element_markdown(size = 10,hjust=0.5,color = '#000000'), strip.text = element_text(size=11), legend.text = element_text(size=7,color = '#000000'), legend.title = element_text(size=8,hjust=0.5,color = '#000000'),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position='plot'))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

```{r}
vislmm <- readRDS("../LFF_spatial_LC/processed-data/12_DEanalyses_removedsampsAndFinalNMseg/11-LCNMspots_NMmetric_LMMed_by_geneLogcounts_2apoeFactorizations_withRanEfFits.RDS")

# get fixed effects tables for pixel intensity (no NM prop in xenium), with gene effects only
vislmm <- lapply(vislmm,function(v){
    y <- lapply(v,function(x){
        setnames(x$fixed,"Pr(>Chisq)","pval_vis")
        setnames(x$fixed,"coef","coef_vis")
        x$fixed <- x$fixed[variable==gene_name]
        x$fixed[,fdr_vis:=p.adjust(pval_vis,method="BH")]
        x$fixed[,bonf_vis:=pval_vis/length(unique(x$fixed$gene_name))]
        return(x$fixed)
    })
    return(y)
})
vislmm2 <- list(E4carrierYN=vislmm$E4carrierYN$Intensity_NM,PerGeno=vislmm$PerGeno$Intensity_NM)
rm(vislmm)

## remember, the sign convention flipped in xenium--positive means more NM with higher xpr in xenium
xenlmm <- readRDS("processed-data_newseg/10_NManalyses/05-EverInLCdomain_wdXfer-orSRTxfer_LClabeled_NonMultimapNM_meanMedIntensity_LMMed_by_geneLogcounts_2apoeFactorizations_withRanEfFits.RDS")

xenlmm <- lapply(xenlmm,function(v){
    y <- lapply(v,function(x){
        setnames(x$fixed,"Pr(>Chisq)","pval_xen")
        setnames(x$fixed,"coef","coef_xen")
        ## get coefs in same sign convention
        x$fixed[,coef_xen:=-1*coef_xen]
        x$fixed <- x$fixed[variable==gene_name]
        x$fixed[,fdr_xen:=p.adjust(pval_xen,method="BH")]
        x$fixed[,bonf_xen:=pval_xen/length(unique(x$fixed$gene_name))]
        return(x$fixed)
    })
    return(y)
})
```

### combine
```{r}
mods <- c("E4carrierYN","PerGeno")

reslist <- list()

for (mod in mods){
    xenvals <- names(xenlmm[[mod]])
    for (xenval in xenvals){
        curxen <- xenlmm[[mod]][[xenval]]
        curvis <- vislmm2[[mod]]
        cur <- merge.data.table(curxen[,.(gene_name,coef_xen,pval_xen,fdr_xen,bonf_xen)],curvis[,.(gene_name,coef_vis,fdr_vis,bonf_vis)],by="gene_name")
        cur[,mod:=mod]
        reslist[[(length(reslist)+1)]] <- cur
        names(reslist)[length(reslist)] <- paste0(mod,"_",xenval)
    }
}

ans <- lapply(reslist,function(x){
    conc <- x[lessOrEqual(bonf_vis,0.05)&lessOrEqual(pval_xen,0.05)&(coef_vis*coef_xen)>0]
    disconc <- x[lessOrEqual(bonf_vis,0.05)&lessOrEqual(pval_xen,0.05)&(coef_vis*coef_xen)<0]
    anslist <- list()
    anslist[[1]] <- conc
    anslist[[2]] <- paste0(length(conc$gene_name)," concordant; ",length(disconc$gene_name)," discordant; ",nrow(x[lessOrEqual(bonf_vis,0.05)&lessOrEqual(pval_xen,0.05)])," total")
    return(anslist)
})
```
all comparisons: 163-164 concordant / 37-39 discordant / 200-203 total vis bonf sig + xen nom sig

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils    
[5] methods   base     

other attached packages:
[1] BiocParallel_1.44.0 parallelly_1.47.0  
[3] gridExtra_2.3       ggtext_0.1.2       
[5] ggplot2_4.0.2       data.table_1.18.4  

loaded via a namespace (and not attached):
 [1] gtable_0.3.6       dplyr_1.2.0       
 [3] compiler_4.5.2     tidyselect_1.2.1  
 [5] Rcpp_1.1.1         xml2_1.5.2        
 [7] parallel_4.5.2     dichromat_2.0-0.1 
 [9] scales_1.4.0       here_1.0.2        
[11] R6_2.6.1           generics_0.1.4    
[13] knitr_1.51         tibble_3.3.1      
[15] rprojroot_2.1.1    pillar_1.11.1     
[17] RColorBrewer_1.1-3 rlang_1.1.7       
[19] xfun_0.56          S7_0.2.1          
[21] otel_0.2.0         cli_3.6.5         
[23] withr_3.0.2        magrittr_2.0.4    
[25] grid_4.5.2         gridtext_0.1.5    
[27] lifecycle_1.0.5    vctrs_0.7.1       
[29] evaluate_1.0.5     glue_1.8.0        
[31] farver_2.1.2       codetools_0.2-20  
[33] tools_4.5.2        pkgconfig_2.0.3   
> sessioninfo::session_info()
─ Session info ──────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       Positron
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-29
 pandoc   NA
 quarto   1.9.38 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────
 package      * version date (UTC) lib source
 BiocParallel * 1.44.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli            3.6.5   2025-04-23 [1] CRAN (R 4.5.0)
 codetools      0.2-20  2024-03-31 [1] CRAN (R 4.5.2)
 data.table   * 1.18.4  2026-05-06 [1] CRAN (R 4.5.2)
 dichromat      2.0-0.1 2022-05-02 [1] CRAN (R 4.5.0)
 dplyr          1.2.0   2026-02-03 [1] CRAN (R 4.5.2)
 evaluate       1.0.5   2025-08-27 [1] CRAN (R 4.5.0)
 farver         2.1.2   2024-05-13 [1] CRAN (R 4.5.0)
 generics       0.1.4   2025-05-09 [1] CRAN (R 4.5.0)
 ggplot2      * 4.0.2   2026-02-03 [1] CRAN (R 4.5.2)
 ggtext       * 0.1.2   2022-09-16 [1] CRAN (R 4.5.0)
 glue           1.8.0   2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra    * 2.3     2017-09-09 [1] CRAN (R 4.5.0)
 gridtext       0.1.5   2022-09-16 [1] CRAN (R 4.5.0)
 gtable         0.3.6   2024-10-25 [1] CRAN (R 4.5.0)
 here           1.0.2   2025-09-15 [1] CRAN (R 4.5.0)
 knitr          1.51    2025-12-20 [1] CRAN (R 4.5.2)
 lifecycle      1.0.5   2026-01-08 [1] CRAN (R 4.5.2)
 magrittr       2.0.4   2025-09-12 [1] CRAN (R 4.5.0)
 otel           0.2.0   2025-08-29 [1] CRAN (R 4.5.0)
 parallelly   * 1.47.0  2026-04-17 [1] CRAN (R 4.5.2)
 pillar         1.11.1  2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig      2.0.3   2019-09-22 [1] CRAN (R 4.5.0)
 R6             2.6.1   2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer   1.1-3   2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp           1.1.1   2026-01-10 [1] CRAN (R 4.5.2)
 rlang          1.1.7   2026-01-09 [1] CRAN (R 4.5.2)
 rprojroot      2.1.1   2025-08-26 [1] CRAN (R 4.5.0)
 S7             0.2.1   2025-11-14 [1] CRAN (R 4.5.2)
 scales         1.4.0   2025-04-24 [1] CRAN (R 4.5.0)
 sessioninfo    1.2.3   2025-02-05 [1] CRAN (R 4.5.0)
 tibble         3.3.1   2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect     1.2.1   2024-03-11 [1] CRAN (R 4.5.0)
 vctrs          0.7.1   2026-01-23 [1] CRAN (R 4.5.2)
 withr          3.0.2   2024-10-28 [1] CRAN (R 4.5.0)
 xfun           0.56    2026-01-18 [1] CRAN (R 4.5.2)
 xml2           1.5.2   2026-01-17 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────