---
title: "04a-Write_transcriptsCSVgz_files_forMolExp"
author: "Bernie Mulvey"
date: "04-15-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(arrow) ## v22 will not work here as 10x isn't actually adhering to parquet, see arrow https://github.com/apache/arrow/issues/48066

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

### due to JHPCE package availability/uninstallability by version, this starts to get a little messy to implement. so, locally, create a temp directory of xenium outputs copied from JHPCE -- we need transcripts and nuclear boundaries.

### MoleculeExperiment, despite still being in Bioconductor, is out of date with regards to how Xenium data is stored, namely that transcripts data is no longer output in a csv.gz (which is what ME will look for when we read in transcript level data later). So we need to create this file from the formats that are available--we'll use the parquet files.

First, pull down the files listed in code/04/00 into a temporary directory organized by brnum. (Done)

Now we need to go through all of those and write a transcripts.csv.gz to each dir. ME uses data.table internally to read in the .gz tables, so hopefully writing it out with fwrite should give us something compatible on its way back in.
```{r}
tsparq <- list.files("code/01_spe/04_NMDAPI_nuclearCount_spe/xentmp",pattern="transcripts.parquet",recursive=T,full.names = T)

lapply(tsparq,function(x){
  curparq <- read_parquet(x,as_data_frame = T)
  opath <- gsub(x,pattern=".parquet",replacement=".csv.gz")
  fwrite(curparq,opath,sep=",",compress = "gzip")
})

## test run worked. nice
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] colorout_1.3-3      arrow_23.0.1.2      data.table_1.18.2.1

loaded via a namespace (and not attached):
 [1] vctrs_0.7.1       cli_3.6.5         knitr_1.51        rlang_1.1.7.9000 
 [5] xfun_0.56         otel_0.2.0        purrr_1.2.1       assertthat_0.2.1 
 [9] glue_1.8.0        bit_4.6.0         rprojroot_2.1.1   htmltools_0.5.9  
[13] rmarkdown_2.30    evaluate_1.0.5    fastmap_1.2.0     yaml_2.3.12      
[17] lifecycle_1.0.5   compiler_4.5.2    here_1.0.2        rstudioapi_0.18.0
[21] digest_0.6.39     R6_2.6.1          tidyselect_1.2.1  magrittr_2.0.4   
[25] tools_4.5.2       bit64_4.6.0-1    
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-04-15
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────────────
 package     * version    date (UTC) lib source
 arrow       * 23.0.1.2   2026-03-25 [1] CRAN (R 4.5.2)
 assertthat    0.2.1      2019-03-21 [2] CRAN (R 4.5.0)
 bit           4.6.0      2025-03-06 [2] CRAN (R 4.5.0)
 bit64         4.6.0-1    2025-01-16 [2] CRAN (R 4.5.0)
 cli           3.6.5      2025-04-23 [2] CRAN (R 4.5.0)
 colorout    * 1.3-3      2026-01-31 [2] Github (jalvesaq/colorout@64863bb)
 data.table  * 1.18.2.1   2026-01-27 [2] CRAN (R 4.5.2)
 digest        0.6.39     2025-11-19 [2] CRAN (R 4.5.2)
 evaluate      1.0.5      2025-08-27 [2] CRAN (R 4.5.0)
 fastmap       1.2.0      2024-05-15 [2] CRAN (R 4.5.0)
 glue          1.8.0      2024-09-30 [2] CRAN (R 4.5.0)
 here          1.0.2      2025-09-15 [2] CRAN (R 4.5.0)
 htmltools     0.5.9      2025-12-04 [2] CRAN (R 4.5.2)
 knitr         1.51       2025-12-20 [2] CRAN (R 4.5.2)
 lifecycle     1.0.5      2026-01-08 [2] CRAN (R 4.5.2)
 magrittr      2.0.4      2025-09-12 [2] CRAN (R 4.5.0)
 otel          0.2.0      2025-08-29 [2] CRAN (R 4.5.0)
 purrr         1.2.1      2026-01-09 [2] CRAN (R 4.5.2)
 R6            2.6.1      2025-02-15 [2] CRAN (R 4.5.0)
 rlang         1.1.7.9000 2026-02-04 [1] Github (r-lib/rlang@bae1fe1)
 rmarkdown     2.30       2025-09-28 [2] CRAN (R 4.5.0)
 rprojroot     2.1.1      2025-08-26 [2] CRAN (R 4.5.0)
 rstudioapi    0.18.0     2026-01-16 [2] CRAN (R 4.5.2)
 sessioninfo   1.2.3      2025-02-05 [2] CRAN (R 4.5.0)
 tidyselect    1.2.1      2024-03-11 [2] CRAN (R 4.5.0)
 vctrs         0.7.1      2026-01-23 [2] CRAN (R 4.5.2)
 xfun          0.56       2026-01-18 [2] CRAN (R 4.5.2)
 yaml          2.3.12     2025-12-10 [2] CRAN (R 4.5.2)

 [1] /Users/bmulvey/Library/R/arm64/4.5/library
 [2] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────────────