---
title: "SF-AllISSdomains_with_SRTs_select_donors"
author: "Bernie Mulvey"
date: "07-14-2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(ggrastr)
library(SpatialExperiment)
library(escheR)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https:/stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallel)


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")

""
```

## get KNN smoothing lookup tables
```{r}
srtdom <- fread("processed-data_newseg/07_visLabelTransfer/05-srtLCsmoothDoms_and_NMFpreds_lut.txt")
sndom <- fread("processed-data_newseg/04_snSeqLabelTransfer/06-wdLCsmoothings_LUT.txt")

## get Xenium obj
spe <- readRDS("processed-data_newseg/02_QC/02b-spe_filtered_genetargonly_log-and-nonlog-norms.RDS")
logcounts(spe) <- NULL #dont need it

xendoms <- merge.data.table(sndom,srtdom,by="rn")
xendoms <- DataFrame(xendoms,row.names=xendoms$rn)[colnames(spe),]
xendoms$rn <- NULL
colData(spe) <- cbind(colData(spe),xendoms)

## get Visium with annots
lcv <- readRDS("../LFF_spatial_LC/processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

vdom <- readRDS("../LFF_spatial_LC/processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
vdom <- vdom$HARMONYlmbna_HDG_SVG_2575
vdom <- vdom[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(vdom,2,"clusid")

vanno <- fread("../LFF_spatial_LC/processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")
vdom <- merge.data.table(vdom,vanno[,.(clusid,anno)],by="clusid")
vdom[,anno:=ifelse(anno=="LC","LC","Other")] ## for pal consistency with how we're marking up the xenium plots
vdom <- DataFrame(vdom,row.names=vdom$rn)[colnames(lcv),]

colLabels(lcv) <- vdom$anno

## tweak SPE to be able to match visium-xenium samples by hemisphere
lcv$brnum[lcv$brnum=="Br6119(re-dis)"] <- "Br6119" # and fix a separate annoyance
## drop donors with limited LC
lcv <- lcv[,!(lcv$brnum %in% paste0("Br",c(5517,5276,5712)))]

## we also want to match left/right sections here so change brnum to reflect the hemi of origin
lcv$brnum <- paste0(lcv$brnum,lcv$hemisphere)
lcv$brnum <- gsub(lcv$brnum,pattern="(Br.*)R",replacement="\\1")
sum(unique(lcv$brnum) %in% unique(spe$brnum))==length(unique(spe$brnum)) # k good

rm(vdom,vanno,srtdom,sndom)
```

## LC/Other palette
```{r}
pal <- readRDS("../LFF_spatial_LC/plots/finalpal_somuchfornorainbows.RDS")
pal <- c(pal[8],"#DEDEDE") # for same green and gray as other supp. with domains)
names(pal) <- c("LC","Other")
```

## plot the SRT samples and corresp. 6 xenium domains   for select donors
# plots with one donor's xenium sample and all visium samples in a column for each threshold
# exemplars with 2 visium sections:
Br6476
Br5426
Br3974
(Br1706, but has 4 visium sections)
```{r}
dons <- paste0("Br",c(6476,5426,3974))
setDTthreads(1,restore_after_fork = FALSE)
plts <- lapply(dons,function(d){
  domdefs <- paste0("domain",c(12412,8492,8426,12199,16119,20079))
  
  sublist <- list()
  i<-1
  for (i in c(1:length(unique(lcv[,lcv$brnum==d]$sample_id)))){
    curs <- unique(lcv[,lcv$brnum==d]$sample_id)[i]
    curd <- lcv[,lcv$brnum==d & lcv$sample_id==curs]
    pv <- make_escheR(curd, y_reverse=FALSE)
    pv <- rasterize(add_fill(p=pv,var="label",size=1.25,point_size=1.25),layers="Points",dpi=300,dev="cairo_png")
    pv <- pv+scale_fill_manual(
      values = pal)+guides(color="none",fill="none")
    pv <- pv+ggtitle(paste0(curs,"_",d))
    sublist[[(length(sublist)+1)]] <- pv
  }
  j<-1
  for (j in c(1:length(domdefs))){
    curdom <- domdefs[j]
    px <- make_escheR(spe[,spe$brnum==d], y_reverse=FALSE)
    px <- rasterize(add_fill(p = px,var=curdom,point_size=0.75),layers="Points",dpi=300,dev="cairo_png")
    px <- px+scale_fill_manual(values = pal)
    px <- px+guides(fill="none",color="none")
    px <- px+ggtitle(curdom)
    # px <- px+ggtitle(s)

    sublist[[(length(sublist)+1)]] <- px
  }

  retobj <- gridExtra::arrangeGrob(grobs = sublist, nrow = length(sublist))
  return(retobj)
})
  dev.off()

pdf("plots_newseg/manuscript/SuppFigs/SF-All_ISS_LCdomains_vs_SRTs.pdf",height=10.5,width=8)
grid.arrange(grobs=plts,ncol=3)
dev.off()

```
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices
[5] utils     datasets  methods   base     

other attached packages:
 [1] colorout_1.3-3             
 [2] escheR_1.10.0              
 [3] SpatialExperiment_1.20.0   
 [4] SingleCellExperiment_1.32.0
 [5] SummarizedExperiment_1.40.0
 [6] Biobase_2.70.0             
 [7] GenomicRanges_1.62.1       
 [8] Seqinfo_1.0.0              
 [9] IRanges_2.44.0             
[10] S4Vectors_0.48.1           
[11] BiocGenerics_0.56.0        
[12] generics_0.1.4             
[13] MatrixGenerics_1.22.0      
[14] matrixStats_1.5.0          
[15] ggrastr_1.0.2              
[16] gridExtra_2.3              
[17] ggtext_0.1.2               
[18] ggplot2_4.0.2              
[19] data.table_1.18.4          

loaded via a namespace (and not attached):
 [1] SparseArray_1.10.8  xml2_1.5.2         
 [3] lattice_0.22-7      magrittr_2.0.4     
 [5] evaluate_1.0.5      grid_4.5.2         
 [7] RColorBrewer_1.1-3  Matrix_1.7-4       
 [9] rprojroot_2.1.1     scales_1.4.0       
[11] textshaping_1.0.4   abind_1.4-8        
[13] cli_3.6.5           rlang_1.1.7        
[15] XVector_0.50.0      DelayedArray_0.36.1
[17] withr_3.0.2         otel_0.2.0         
[19] S4Arrays_1.10.1     ggbeeswarm_0.7.3   
[21] tools_4.5.2         dplyr_1.2.0        
[23] here_1.0.2          vctrs_0.7.1        
[25] R6_2.6.1            magick_2.9.0       
[27] lifecycle_1.0.5     vipor_0.4.7        
[29] ragg_1.5.0          pkgconfig_2.0.3    
[31] beeswarm_0.4.0      pillar_1.11.1      
[33] gtable_0.3.6        glue_1.8.0         
[35] Rcpp_1.1.1          systemfonts_1.3.1  
[37] xfun_0.56           tibble_3.3.1       
[39] tidyselect_1.2.1    knitr_1.51         
[41] dichromat_2.0-0.1   rjson_0.2.23       
[43] farver_2.1.2        labeling_0.4.3     
[45] compiler_4.5.2      S7_0.2.1           
[47] gridtext_0.1.5     
> sessioninfo::session_info()
─ Session info ──────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       Positron
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-07-14
 pandoc   NA
 quarto   1.9.38 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────
 package              * version date (UTC) lib source
 abind                  1.4-8   2024-09-12 [1] CRAN (R 4.5.0)
 beeswarm               0.4.0   2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5   2025-04-23 [1] CRAN (R 4.5.0)
 colorout             * 1.3-3   2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.4  2026-05-06 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.1  2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1 2022-05-02 [1] CRAN (R 4.5.0)
 dplyr                  1.2.0   2026-02-03 [1] CRAN (R 4.5.2)
 escheR               * 1.10.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 evaluate               1.0.5   2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2   2024-05-13 [1] CRAN (R 4.5.0)
 generics             * 0.1.4   2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1  2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 ggbeeswarm             0.7.3   2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2   2026-02-03 [1] CRAN (R 4.5.2)
 ggrastr              * 1.0.2   2023-06-01 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2   2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0   2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3     2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5   2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6   2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2   2025-09-15 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51    2025-12-20 [1] CRAN (R 4.5.2)
 labeling               0.4.3   2023-08-29 [1] CRAN (R 4.5.0)
 lattice                0.22-7  2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5   2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0   2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4   2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4   2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0   2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0   2025-08-29 [1] CRAN (R 4.5.0)
 pillar                 1.11.1  2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.5.0)
 R6                     2.6.1   2025-02-15 [1] CRAN (R 4.5.0)
 ragg                   1.5.0   2025-09-02 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3   2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1   2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23  2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7   2026-01-09 [1] CRAN (R 4.5.2)
 rprojroot              2.1.1   2025-08-26 [1] CRAN (R 4.5.0)
 S4Arrays               1.10.1  2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1  2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1   2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0   2025-04-24 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3   2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0  2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8  2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0  2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0  2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 systemfonts            1.3.1   2025-10-01 [1] CRAN (R 4.5.0)
 textshaping            1.0.4   2025-10-10 [1] CRAN (R 4.5.0)
 tibble                 3.3.1   2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1   2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7   2023-12-18 [1] CRAN (R 4.5.0)
 withr                  3.0.2   2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56    2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2   2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0  2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────